### 2020-03-24 README.Rmd
initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# kwlburials

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/LiYingWang/kwlburials/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, kwlburials, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("LiYingWang/kwlburials")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2020-03-26 analysis/paper/paper.Rmd
move paper rmd from dissertation proj to here

---
title: "Bayesian exponential random graph modelling of KWL Burial network"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)
```

# Introduction

# Background

# Methods

## ERGM model

Exponential Random Graph Models (ERGMs) are a family of statistical model for network analysis to 
describe the formation of ties in a network. In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately. In contrast to traditional model where variables not interacting with each other, ERGMs assumes the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes and ties (Robins et al. 2007). In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized (Morries et al. 2008). This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data (Amati et al. 2019).

some concepts:
- endogenous statistics that ties are modelled based on the existence of other ties
- exogenous statistics that ties are formed according to monadic or dyadic node attributes

toturial: 
https://www.r-bloggers.com/ergm-tutorial/

## statistics
We want to know

ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles

We used ergm package in R for the model specification. Every term in an ERGM must have an associated algorithm for computing its value for network (Morris et al. 2008). The terms we selected for nwtwork specification include:

1. edges: ties, a measure of density, equal to kstar(1) for undirected networks

2. gwesp (geometrically weighted edgewise shared partners): a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node. It is a measure of the tendency for two actors who are tied to have x partners in common.

3. gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once. The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts

# Results

```{r model-ergm-1}
model.ergm <- burial_network_pre ~ 
  edges + # ties, a measure of density, equal to kstar(1) for undirected networks
  density +
  gwdegree(0.5, fixed = TRUE)  +
  triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
summary(model.ergm)
```

```{r model-ergm-2}
model.2 <- burial_network_pre ~ edges + # density
  gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
  # triad-based clustering according to dyad-based configurations instead of node-based
  # number means weight parameter alpha, which controls the rate of declining marginal returns
  # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
  # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
  # ergm can estimate the parameter from the data by using fixed=FALSE
  # http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
  gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
  # tendency of being in contact with multiple partners, measures of centralisation
  # distribution of node-based edge counts, each node counts only once
  # number means weight parameter decay
  # Morris et al. (2008)
summary(model.2)
```

```{r model-ergm-3}
model.3 <- burial_network_pre ~ edges +  # the overall density of the network
  nodematch('quantity') +    # quantity-based homophily, categorical nodal attribute
  # the similarity of connected nodes
  gwesp(0.2, fixed = TRUE) +    # transitivity
  gwdegree(0.8, fixed = TRUE)   # popularity
summary(model.3)
```

```{r model-bergm}
# Specify a prior distribution: normal distribution (low density and high transitivity)
prior.mean <- c(-3, 0, 1, 0) # prior mean corresponds to mean for each parameter
# follow Alberto Caimo et al. (2015) hospital example
prior.sigma <- diag(3, 4, 4) # covariance matrix structure
# normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
# where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

# Estimated posterior means, medians and 95% credible intervals for Models.3
# bergmM: Bayesian exponential random graphs models under missing data using the approximate exchange algorithm
parpost <- bergm(model.3,
                  prior.mean  = prior.mean,
                  prior.sigma = prior.sigma,
                  burn.in     = 200, # burn-in iterations for every chain of the populationm, drops the first 200
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
                  nchains     = 8, # number of chains of the population MCMC
                  gamma       = 0.7) # scalar; parallel adaptive direction sampling move factor, acceptance rate

summary(parpost) 
```

The summary presents the statistics that each θ corresponds to the parameter specified in ERGM previously. In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.

θ1 = number of ties
θ2 = individuals with the same abundance of burial goods
θ3 = gwesp is negative that rejects the assumption that actors with multiple partners in common are more likely to be directed connected
θ4 =

# some exploration
1. there is different statistics between weighted and unweighed ties

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-03-26 Note.Rmd
add note

# Note for the paper

ERGM toturial: 
https://www.r-bloggers.com/ergm-tutorial/

Bergm toturial:
http://acaimo.github.io/Bergm/tutorials/Dec_18.html

# Explore different combination of variables for Bergm
1. there is difference in statistics (mean, sd) between weighed and unweighed ties
2. difference in statistics when simplifying variables to agate and golden beads, plus a few types of prestige goods such as gold foil and fish-like shaped necklace. The mean looks fine
3. did not work well when there is too many ties in the network
4. difference in the results of using BergmM and Bergm, BergmM is for missing data
5. add the spacial information?

### 2020-03-26 analysis/paper/paper.Rmd
add note

@@
 
-## ERGM model
+## Exponential Random Graph Models
 
@@
 2007).
-In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized (Morries et al.
-2008).
-This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data (Amati et al.
-2019).
+In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized [@Morris2008].
+This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data [@Amati2019].
 
@@
 - exogenous statistics that ties are formed according to monadic or dyadic node attributes
-
-toturial: 
-https://www.r-bloggers.com/ergm-tutorial/
 
@@
 We used ergm package in R for the model specification.
-Every term in an ERGM must have an associated algorithm for computing its value for network (Morris et al.
-2008).
+Every term in an ERGM must have an associated algorithm for computing its value for network [@Morris2008].
 The terms we selected for nwtwork specification include:
@@
 
-# some exploration
-1. there is different statistics between weighted and unweighed ties
-
-Here is an example of inline code `r x` in the middle of a sentence.
+inline code `r x` example
 

### 2020-03-27 analysis/paper/paper.Rmd
paste some paragraphs from proposal and reorganize as a start

@@
 
-# Background
+Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity (Ehrhardt 2009; Gamble and Zepeda 2002; Janes 2013; Johansen and Bauer 2015).
+Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting (Coward 2013: 252).
+One way to study the relationships between individual is network analysis, where burials represent actors, nodes in the network, linked by sharing the same attributes such as the same prestige goods.
+
+Social Network Analysis (SNA) is a tool for understanding network structure and its effects on every part within the network (Freeman 2004; Salvini 2010).
+A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
+Their relationships are flows of resources that reflect relations of control, dependence, and cooperation.
+In other words, actors are viewed as ‘nodes’ in a wider network, in which nodes are connected by multiple relationships, or ‘edges’, to other nodes (Coward 2013; Mizoguchi 2013; Potgieter, et al.
+2008).
+SNA provides a way to visualize, quantify, and statistically test hypotheses about the relationships between actors.
+At Kiwulan, I will use individual burials as actors or nodes in the social network, and use SNA methods to test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions.
+Network ties will be drawn between burials with similar variables that can represent the same ‘community’ or social group.
+To measure change in social inequality, I will use the binary presence or absence of burial goods, the amount of prestige goods, grave dimensional attributes, and ordinal variables of age and sex categories to determine which actors have connections in the network (Sosna, et al.
+2012).
+Once the relationships between the actors have been described, I will compute statistical measures such as network degree centrality, graph density, and graph centralization.
+These measures are helpful to identify individual’s role and trends of network relationship, and I will use them to formally test hypotheses about the structure of the social network (Brughmans 2010: 280; Gjesfjeld 2014: 192).
+I predict that if there is a change from corporate to network mode following the European contact, then I expect the social network will show statistically significantly higher degrees of centrality and a higher density in the post-European period compared to pre-European period.
+This is because a social network graph that is highly centralized and very sparse indicates high concentrations of wealth, and higher social inequality.
+
+Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
+We ask whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society.
+We use social network analysis (SNA) where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
+Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
+
+Exponential random graph models (ERGMs) are an important family of statistical models for analyzing network data and evaluating models of network structure.
+However, ERGMs are difficult to compute because their normalizing constant, which depends on model parameters, is intractable.
+A Bayesian framework allows for parameter inference using MCMC strategies which avoid the need for computationally intensive calculations of the normalizing constants.
+Using the R language, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan.
+We expect the network to present a higher degree of clustering (e.g.
+the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period.
+This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on indigenous groups.
+A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.
+
+# Archaeological background
+
+Based on the quantity and quality of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
 
@@
 
-Exponential Random Graph Models (ERGMs) are a family of statistical model for network analysis to 
-describe the formation of ties in a network.
+Exponential Random Graph Models (ERGMs) are a family of statistical model for network analysis to describe the formation of ties in a network.
 In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately.

### 2020-03-29 analysis/paper/paper.Rmd
add implication of gwdegree

@@
 gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once.
-The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts
+The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts.
+A measure of the tendency of the number of edges into a node.
+gwdegree estimates the change in tie likelihood given the degree of the nodes involved, but with marginally decreasing weighting as degree increases(https://www.r-bloggers.com/ergm-tutorial/).
+This statistic captures the tendency towards centralisation in the degree distribution of the network.
 
@@
   # Morris et al. (2008)
+  # The closer decay is to zero, the more gwdegree considers low degree nodes relative to high degree nodes
 summary(model.2)
@@
 θ2 = individuals with the same abundance of burial goods
-θ3 = gwesp is negative that rejects the assumption that actors with multiple partners in common are more likely to be directed connected
-θ4 =
+θ3 = gwesp, a positive sign suggests a tendency of burials in our sample to group together in closed transitive structures.
+In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organised in closed structures.
+Burials with multiple prestige goods in common are more likely to be directed connected.
+θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.
 

### 2020-03-30 Note.Rmd
working on introduction

@@
 5. add the spacial information?
+
+# some paragraphs from proposal 
+At Kiwulan, we view individual burials as actors or nodes in the social network, and use SNA methods to test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions.
+Network ties will be drawn between burials with similar variables that can represent the same ‘community’ or social group.
+To measure change in social inequality, I will use the binary presence or absence of burial goods, the amount of prestige goods, grave dimensional attributes, and ordinal variables of age and sex categories to determine which actors have connections in the network (Sosna, et al.
+2012).
+
+Once the relationships between the actors have been described, I will compute statistical measures such as network degree centrality, graph density, and graph centralization.
+These measures are helpful to identify individual’s role and trends of network relationship, and I will use them to formally test hypotheses about the structure of the social network (Brughmans 2010: 280; Gjesfjeld 2014: 192).
+I predict that if there is a change from corporate to network mode following the European contact, then I expect the social network will show statistically significantly higher degrees of centrality and a higher density in the post-European period compared to pre-European period.
+This is because a social network graph that is highly centralized and very sparse indicates high concentrations of wealth, and higher social inequality.

### 2020-03-30 analysis/paper/paper.Rmd
working on introduction

@@
 Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting (Coward 2013: 252).
-One way to study the relationships between individual is network analysis, where burials represent actors, nodes in the network, linked by sharing the same attributes such as the same prestige goods.
-
-Social Network Analysis (SNA) is a tool for understanding network structure and its effects on every part within the network (Freeman 2004; Salvini 2010).
+One way to study the relationships between individual is social network analysis (SNA), which can visualize, quantify, and statistically test hypotheses about the relationships to understand network structure and its effects on every part within the network (Freeman 2004; Salvini 2010).
 A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
 Their relationships are flows of resources that reflect relations of control, dependence, and cooperation.
-In other words, actors are viewed as ‘nodes’ in a wider network, in which nodes are connected by multiple relationships, or ‘edges’, to other nodes (Coward 2013; Mizoguchi 2013; Potgieter, et al.
+In other words, actors are viewed as "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes (Coward 2013; Mizoguchi 2013; Potgieter, et al.
 2008).
-SNA provides a way to visualize, quantify, and statistically test hypotheses about the relationships between actors.
-At Kiwulan, I will use individual burials as actors or nodes in the social network, and use SNA methods to test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions.
-Network ties will be drawn between burials with similar variables that can represent the same ‘community’ or social group.
-To measure change in social inequality, I will use the binary presence or absence of burial goods, the amount of prestige goods, grave dimensional attributes, and ordinal variables of age and sex categories to determine which actors have connections in the network (Sosna, et al.
-2012).
-Once the relationships between the actors have been described, I will compute statistical measures such as network degree centrality, graph density, and graph centralization.
-These measures are helpful to identify individual’s role and trends of network relationship, and I will use them to formally test hypotheses about the structure of the social network (Brughmans 2010: 280; Gjesfjeld 2014: 192).
-I predict that if there is a change from corporate to network mode following the European contact, then I expect the social network will show statistically significantly higher degrees of centrality and a higher density in the post-European period compared to pre-European period.
-This is because a social network graph that is highly centralized and very sparse indicates high concentrations of wealth, and higher social inequality.
 
-Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
-We ask whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society.
-We use social network analysis (SNA) where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
-Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
-
-Exponential random graph models (ERGMs) are an important family of statistical models for analyzing network data and evaluating models of network structure.
+Exponential Random Graph Models (ERGMs) are an important family of statistical model for network analysis to describe the formation of ties in a network.
+In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately.
+In contrast to traditional model where variables not interacting with each other, ERGMs assumes the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes and ties (Robins et al.
+2007).
+In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized [@Morris2008].
+This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data [@Amati2019].
 However, ERGMs are difficult to compute because their normalizing constant, which depends on model parameters, is intractable.
 A Bayesian framework allows for parameter inference using MCMC strategies which avoid the need for computationally intensive calculations of the normalizing constants.
+
 Using the R language, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan.
+We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group.
+We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions.
+We ask the questions: 1.
+Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
+2.
+Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society can be detected by burials networks.
 We expect the network to present a higher degree of clustering (e.g.
@@
 
+We build network where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
+
 # Methods
@@
 ## Exponential Random Graph Models
-
-Exponential Random Graph Models (ERGMs) are a family of statistical model for network analysis to describe the formation of ties in a network.
-In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately.
-In contrast to traditional model where variables not interacting with each other, ERGMs assumes the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes and ties (Robins et al.
-2007).
-In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized [@Morris2008].
-This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data [@Amati2019].
 

### 2020-04-01 analysis/paper/paper.Rmd
add sentences to the background of burials

@@
 
+The burials we analyzed in this study are from Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
+The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data and diagnostic porcelain as time indicators.
+
 Based on the quantity and quality of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
-
 We build network where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
@@
 
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis (Caimo et al, 2015).
+We specified the prior based on normal distribution that assumes a low density and high transitivity network.
+
 ```{r model-bergm}
@@
 prior.sigma <- diag(3, 4, 4) # covariance matrix structure
-# normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-04-01 analysis/paper/experiment.R

@@ -1,0 +1,61 @@
+library(network)
+
+burial_network_pre <-
+  network(edges_for_network_pre, # the network object
+          vertex.attr = nodes_pre, # node list
+          directed = FALSE, # specify whether the network is directed
+          ignore.eval = FALSE, # FALSE = weighted
+          loops = FALSE, # do we allow self ties (should not allow them)
+          matrix.type = "edgelist") # the type of input
+
+# plot
+plot(burial_network_pre, vertex.cex = 1) # It seems the last two nodes are missing?? only 27 nodes
+
+
+library(tidygraph)
+library(ggraph)
+
+relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
+                               edges = edges_for_network_pre,
+                               directed = FALSE)
+
+relation_tidy_pre %>%
+  activate(edges) %>%
+  arrange(desc(common_counts))
+
+ggraph(relation_tidy_pre) +
+  geom_edge_link() +
+  geom_node_point() +
+  theme_graph()
+
+ggraph(relation_tidy_pre, layout = "graphopt") +
+  geom_node_point() +
+  geom_edge_link(alpha = 0.8) +
+  scale_edge_width(range = c(0.2, 2)) +
+  labs(edge_width = "common item") +
+  theme_graph()
+
+
+rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia", "Emily"))
+rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
+                          to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
+graph_output <- tbl_graph(nodes = rstat_nodes, edges = rstat_edges)
+ggraph(graph_output, layout = "graphopt") +
+  geom_node_point() +
+  geom_edge_link(alpha = 0.8) +
+  scale_edge_width(range = c(0.2, 2)) +
+  labs(edge_width = "common item") +
+  theme_graph()
+
+#------network
+rstat_nodes$id <- 1:nrow(rstat_nodes)
+example_network <-
+  network(rstat_edges, # the network object
+          vertex.attr = tibble(rstat_nodes), # node list
+          directed = FALSE, # specify whether the network is directed
+          ignore.eval = FALSE, # FALSE = weighted
+          loops = FALSE, # do we allow self ties (should not allow them)
+          matrix.type = "edgelist") # the type of input
+
+# plot
+plot(example_network, vertex.cex = 1) # It seems the last two nodes are missing?? only 27 nodes

### 2020-04-01 Note.Rmd
paste some paragraphs from dissertation proposal

@@
 This is because a social network graph that is highly centralized and very sparse indicates high concentrations of wealth, and higher social inequality.
+
+# Testing the predictions of the models
+
+My hypothesis is that if European colonization created monopolies for trade between a small number of indigenous people, and this resulted in the shift from a more corporate to a more network organization, then I would expect to see changes in mortuary practices, food consumption and feasting behaviors, and social network organization at Kiwulan.
+
+Before European contact, households consisted of single and multi-family units and were organized in a corporate mode with limited economic inequality primarily based on age, gender, or different corporate groups.
+I expect that burial goods will reveal a relatively even distribution of the quantity and quality of items.
+Social network analysis will indicate multiple small groups of similar size.
+If there is unequal distribution of prestige or high quality goods, then it will be restricted to group who share similar demographic characteristics such as age.
+Production and consumption of ceramics will be based on corporate groups.
+Thus, diversity of sources and style of pottery within a family unit will be more similar than between family units.
+Activity areas of family units will reflect more even distribution of high ranked subsistence resources, and cache pits will present less variability in size and contents.
+
+After European contact, there will be more obvious inequalities in the distribution of prestige materials in burials and settlement areas, indicative of an increase in the concentration of wealth at Kiwulan.
+Prestige goods will not only appear in adult burials but also child burials, which reflect wealth accumulation within a family.
+Higher quality imported goods will mainly appear in a few but not all individual burials because of control of long distance trade.
+Social network analysis of burials will present a higher density and higher degree of centrality based on individual power.
+Production and consumption of ceramics will extend beyond the corporate groups and present more standardized sizes and shapes, which show an emphasis on production for feasting, and control of production by a small number of individuals.
+Activity areas will reflect uneven distribution of high ranked subsistence resources, and cache pits will present higher variability in size and contents.

### 2020-04-01 analysis/paper/paper.Rmd
paste some paragraphs from dissertation proposal

@@
 
-The burials we analyzed in this study are from Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
-The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data and diagnostic porcelain as time indicators.
+The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
+The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans.
+The burials were located in the north margin of the AD section which is the largest open area at Kiwulan site.
+
+The archaeological evidence in the Yilan plain indicates some elements of social complexity around 17th century, including increasingly sedentary settlements, cultivation of rice, complex inter- and intra-trade networks, the use of prestige goods, and differential burial practices (Chen 2004; Cheng 2008; Li 2014).
+Previous studies of social organization of the Kiwulan site suggest two different interpretations of this evidence.
+Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
+However, Hsieh (2012) suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
+She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status.
+The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese (Borao Mateo 2009; Li and Wu 2006).
+If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived.
+In this case, pre-existing inequalities might have become more accentuated as a response to European contact.
+Did the degree of economic inequality change over time, and how did European contact impact indigenous social organization?
+These questions motivate this project.
+
+The archaeological materials include around 139,900 pottery sherds, 12,500 imported ceramics and stonewares, 1,100 wooden artifacts, 730 stone tools, 425 metal artifacts, 7,400 glass beads and agates beads, 100 pipes, several ornaments made of animal bones, shells, and wood, and a faunal assemblage (Chen 2007).
+In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated (Figure 5).
+These postholes were found aligned in a north-south direction with construction marks, which were interpreted as the remains of house structures.
+Moreover, the area with burials is oriented in an east-west direction located on the north side of the house structures (Chen 2007).
+Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society (Cheng 2008; Hsieh 2009a).
+In the following section I explain the model I will use to evaluate these claims for social inequality, and how I will determine the likely causes of any changes over time that I identify.
+
+The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials (Chen 2005; Cheng 2008), which might indicate the pursuit of prestige or wealth.
+In addition to prestige, European colonial powers could have provided military support to indigenous allies.
+For example, Voss’s (2005) study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people.
+A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
+In this study, I will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.
 

### 2020-04-01 analysis/paper/paper.Rmd
update the description for terms in ERGMs

@@
 2.
-gwesp (geometrically weighted edgewise shared partners): a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node.
+gwesp (geometrically weighted edgewise shared partners): triad-based clustering based on a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node.
 It is a measure of the tendency for two actors who are tied to have x partners in common.
+In orther words, a tendency for those with shared partners to become tied, or tendency of ties to cluster together.
 
@@
   gwdegree(0.5, fixed = TRUE) +
-  triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
+  triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network, can be replaced by "gwdegree"
 summary(model.ergm)
@@
 ```{r model-ergm-2}
+# check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
+# also check here Morris et al. (2008)
 model.2 <- burial_network_pre ~ edges + # density
-  gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
-  # triad-based clustering according to dyad-based configurations instead of node-based
+  gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle)
   # number means weight parameter alpha, which controls the rate of declining marginal returns
@@
   # ergm can estimate the parameter from the data by using fixed=FALSE
-  # http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
+
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
@@
   # number means weight parameter decay
-  # Morris et al. (2008)
   # The closer decay is to zero, the more gwdegree considers low degree nodes relative to high degree nodes

### 2020-04-01 analysis/paper/paper.Rmd
add references

@@
 
-Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity (Ehrhardt 2009; Gamble and Zepeda 2002; Janes 2013; Johansen and Bauer 2015).
-Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting (Coward 2013: 252).
-One way to study the relationships between individual is social network analysis (SNA), which can visualize, quantify, and statistically test hypotheses about the relationships to understand network structure and its effects on every part within the network (Freeman 2004; Salvini 2010).
+Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
+Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting [@Coward2013, pp.
+252].
+One way to study the relationships between individual is social network analysis (SNA), which can visualize, quantify, and statistically test hypotheses about the relationships to understand network structure and its effects on every part within the network [@Freeman2004; @Salvini2010].
 A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
 Their relationships are flows of resources that reflect relations of control, dependence, and cooperation.
-In other words, actors are viewed as "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes (Coward 2013; Mizoguchi 2013; Potgieter, et al.
-2008).
+In other words, actors are viewed as "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes [@Coward2013; @Mizoguchi2013; @Potgieter2008].
 
@@
 
-Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis (Caimo et al, 2015).
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
 We specified the prior based on normal distribution that assumes a low density and high transitivity network.

### 2020-04-02 analysis/paper/paper.Rmd
working on archaeological background

@@
 
-The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan that experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
-The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans.
-The burials were located in the north margin of the AD section which is the largest open area at Kiwulan site.
-
-The archaeological evidence in the Yilan plain indicates some elements of social complexity around 17th century, including increasingly sedentary settlements, cultivation of rice, complex inter- and intra-trade networks, the use of prestige goods, and differential burial practices (Chen 2004; Cheng 2008; Li 2014).
-Previous studies of social organization of the Kiwulan site suggest two different interpretations of this evidence.
+The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
+The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation].
+The excavation revealed abundent pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [\@Chen2007].
+In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
+The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for comparison.
+Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and in-situ wooden posts.
+Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society [\@Cheng 2008; \@Hsieh 2009a].
 Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
@@
 These questions motivate this project.
-
-The archaeological materials include around 139,900 pottery sherds, 12,500 imported ceramics and stonewares, 1,100 wooden artifacts, 730 stone tools, 425 metal artifacts, 7,400 glass beads and agates beads, 100 pipes, several ornaments made of animal bones, shells, and wood, and a faunal assemblage (Chen 2007).
-In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated (Figure 5).
-These postholes were found aligned in a north-south direction with construction marks, which were interpreted as the remains of house structures.
-Moreover, the area with burials is oriented in an east-west direction located on the north side of the house structures (Chen 2007).
-Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society (Cheng 2008; Hsieh 2009a).
-In the following section I explain the model I will use to evaluate these claims for social inequality, and how I will determine the likely causes of any changes over time that I identify.
 

### 2020-04-02 analysis/paper/paper.Rmd
add references

@@
 The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation].
-The excavation revealed abundent pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [\@Chen2007].
+The excavation revealed abundent pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
@@
 Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and in-situ wooden posts.
-Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society [\@Cheng 2008; \@Hsieh 2009a].
-Cheng (2008) interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
-However, Hsieh (2012) suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
+Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society [@Cheng2008; @Hsieh2012].
+@Cheng2008 interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
+However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
 She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status.
-The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese (Borao Mateo 2009; Li and Wu 2006).
+The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; Li and Wu 2006).
 If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived.

### 2020-04-03 analysis/paper/KWL_burial_post.R
run Bergm code for post-European successfully, takes ~1 hr

@@ -194,10 +194,12 @@ bergmC(formula,
        rm.alpha = 0, n.aux.draws = 400, aux.thin = 50,
        estimate = c("MLE", "CD"), ...)
 
-
 # Model assessment, Bayesian goodness of fit diagnostics:
 bgof(parpost,
      aux.iters = 10000,
      n.deg     = 14,
      n.dist    = 15,
      n.esp     = 9)
+
+
+

### 2020-04-03 analysis/paper/paper.Rmd
run Bergm code for post-European successfully, takes ~1 hr
working on archaeological background

@@
 The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
-The burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation].
-The excavation revealed abundent pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
+The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
@@
 The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; Li and Wu 2006).
-If Cheng is correct and socio-economic inequality was present, it might be rooted in the use of prestige goods traded in the area before the Europeans arrived.
-In this case, pre-existing inequalities might have become more accentuated as a response to European contact.
-Did the degree of economic inequality change over time, and how did European contact impact indigenous social organization?
-These questions motivate this project.
+Analysing burial data will give an insight into the social reactions of Kiwulan residents and provide a way to examine and explain the changes in social inequality at Kiwulan before and after European contact.
+
+Burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation].
+The burials that were disturbed severely due to modern construction are excluded due to incomplete data.
+Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, accoutring to the total number of burial goods to indicate the index of wealth.
 
@@
 
-## statistics
+## Statistics
 We want to know

### 2020-04-03 analysis/paper/paper.Rmd
work on method section

@@
 
+We assume burial network represents the relations among people and social structure based on the argument that 
 # Methods
 
+Our purpose is to understand the characteristics of burial network and the mechanism of the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
+Previous studies have shown that there are some evidence of social inequality observed on this site but the degree and causes remains unclear.
+To answer our question about the shifts in social inequality during and after the arrival of foreign presence, we hypothesize:
+
+• Hypothesis 1: There are significant reciprocal structure effects in burial network before the contact
+• Hypothesis 2: There are significant transitivity structure effects in burial network after the contact
+• Hypothesis 3: There are significant popularity structure effects in in burial network after the contact
+
 ## Exponential Random Graph Models
+
+We want to know
 
@@
 ## Statistics
-We want to know
-
 ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles
@@
 
-```{r model-ergm-1}
+```{r pre-model-ergm-1}
 model.ergm <- burial_network_pre ~ 
@@
 
-```{r model-ergm-2}
+```{r pre-model-ergm-2}
 # check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
@@
 
-```{r model-ergm-3}
+```{r pre-model-ergm-3}
 model.3 <- burial_network_pre ~ edges +  # the overall density of the network
@@
 
-```{r model-bergm}
+```{r pre-model-bergm}
 # Specify a prior distribution: normal distribution (low density and high transitivity)

### 2020-04-03 analysis/paper/KWL_burial_pre.R

@@ -257,7 +257,7 @@ summary(parpost) # Each θ corresponds to the parameter specified in ERGM previo
 # there is different statistics between weighted and unweighted ties
 # θ1 = number of ties, θ2 = individuals with the same abundance of burial goods
 # θ3 = gwesp is negative that rejects the assumption that actors with multiple partners in common are more likely to be directed connected
-# θ4 =
+# θ4 = gwdegree, negative estimates means strong edges are not necessarily centralised or dispersed in the degree distribution.
 
 plot(parpost)
 

### 2020-04-04 analysis/paper/paper.Rmd
work on method section and add references

@@
 Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society can be detected by burials networks.
+3.
+How the relations among burials may affect the quantity of burial goods?
 We expect the network to present a higher degree of clustering (e.g.
@@
 We assume burial network represents the relations among people and social structure based on the argument that 
+## Bayesian inference for Exponential Random Graph Models
+
+Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
+This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
+
 # Methods
@@
 Our purpose is to understand the characteristics of burial network and the mechanism of the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
-Previous studies have shown that there are some evidence of social inequality observed on this site but the degree and causes remains unclear.
-To answer our question about the shifts in social inequality during and after the arrival of foreign presence, we hypothesize:
+Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
+To answer our question for the shifts in social inequality during and after the arrival of the foreign presences, we hypothesize:
 
@@
 
-## Exponential Random Graph Models
+The computational processes in this paper were performed by the ergm package for simulation and model specification and then bergm package using R programming [@Hunter2008; @Morris2008; @Caimo2014].
+We set the number of chains to eight.
+For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
+We set the number of iterations used to simulate a network y′ at each iteration to 10000.
 
-We want to know
+## Statistics
 
@@
 
-## Statistics
 ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles

### 2020-04-04 analysis/paper/paper.Rmd
add sentences about goodness-of-fit

@@
 We set the number of iterations used to simulate a network y′ at each iteration to 10000.
+The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with a set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
+This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
 
@@
 
+Figure shows the Bayesian goodness of fit diagnostics plots where the red lines represent the distributions of the observed data and the box plots represent the GOF distributions of 10000 network graphs simulated from ERGMs based on the estimated posterior distribution.
+The grey lines mark the 95% intervals.
+
 inline code `r x` example

### 2020-04-04 analysis/paper/KWL_burial_pre.R

@@ -297,6 +297,9 @@ bgof(parpost,
      n.dist    = 15,
      n.esp     = 9)
 
+# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
+# a result that is very difficult to obtain in practice (Caimo 2017)
+
 #-------------------------set different parameters------------------------------
 # three competing models based on ERGM formulas to fit the data using gwesp and gwdegree
 

### 2020-04-05 analysis/paper/paper.Rmd
Explore age as additional node attribute, the result doesn't change a lot. Add sentences about chronology

@@
 
-Burials were assigned to pre-European contact period, post-European contact period, and Chinese period based on the stratigraphic data, radiocarbon dates, and diagnostic porcelain such as white-glazed jars as a time indicator of the arrival of the Europeans[citation].
-The burials that were disturbed severely due to modern construction are excluded due to incomplete data.
-Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, accoutring to the total number of burial goods to indicate the index of wealth.
+Burials were assigned to the pre-European period, European and post-European period, and Chinese period for network comparison.
+Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
+The chronology is then evaluated by the burials with light grey glazed jars, also called An-ping jars produced most likely in China and circulated in Southeast Asia from the late 16th to 18th century but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
+In addition to the diagnostic porcelain, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007].
+Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting from incomplete data.
+Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, according to the total number of burial goods as an index of wealth.
 
@@
 
-Based on the quantity and quality of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
+Based on the quantity of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
 We build network where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
@@
 We assume burial network represents the relations among people and social structure based on the argument that
+
 ## Bayesian inference for Exponential Random Graph Models

### 2020-04-05 analysis/paper/burials_by_age.R

@@ -2,160 +2,252 @@ library(readxl)
 library(tidyverse)
 library(here)
 
-burial <- read_excel(here("index", "data", "burials", "Kiwulan_Burials.xlsx"))
+burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
 
-# combine some age column 
-burial_age <- 
-  burial %>% 
-  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>% 
-  filter(!is.na(Phase), !is.na(Age)) %>% 
+# burial data with combined age and three phases
+burial_three_period_age_tidy <-
+  burial %>%
+  rename(burial_label = ID) %>%
+  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
+  filter(!is.na(Phase)) %>%
   mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
          Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
-         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>% 
-  mutate(`Indo-Pacific_bead` = as.numeric(`Indo-Pacific_bead`),
-         `Gold_leaf` = as.numeric(`Gold_leaf`),
-         Stoneware = as.numeric(Stoneware)) %>% 
-  mutate(Age_com = case_when(
+         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
+  mutate_at(21:ncol(.), as.numeric) %>%
+  janitor::remove_empty(which = "cols") %>%
+  mutate(total = rowSums(.[c(21:48, 50, 55, 56)], na.rm = TRUE)) %>% #prestige goods
+  mutate(Porcelain = rowSums(.[c(40:48, 55, 56)], na.rm = TRUE)) %>% #B&W, porcelain, Anping, kendi
+  mutate(Porcelain = ifelse(Porcelain == 0, NA, Porcelain)) %>%
+  mutate(quantity = case_when(
+    total == 0 ~ "none",
+    total > 0 & total <= 7 ~ "low",
+    total > 7 & total < 100 ~ "medium",
+    total >= 100 ~ "high",
+    TRUE ~ "")) %>% # the classification is based on the result of histogram
[…]

### 2020-04-06 analysis/paper/paper.Rmd
minor edits for intro and method section, change the boxplot number for bgof

@@
 This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data [@Amati2019].
-However, ERGMs are difficult to compute because their normalizing constant, which depends on model parameters, is intractable.
-A Bayesian framework allows for parameter inference using MCMC strategies which avoid the need for computationally intensive calculations of the normalizing constants.
+However, ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
+A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
 
@@
 Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting from incomplete data.
-Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, according to the total number of burial goods as an index of wealth.
 
@@
 
+We use burial data from different time periods at Kiwulan and consider the effects of both exogenous and endogenous factors that may provide meaningful explanation of the tendency of burials correlate with each other and the underlying mechnism.
+
+n concerning connectivity, inter-cultural contacts and exchange among those sites.
 Our purpose is to understand the characteristics of burial network and the mechanism of the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
@@
 This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
+
+Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, according to the total number of burial goods as an index of wealth.
 

### 2020-04-06 analysis/paper/burials_by_age.R

@@ -248,6 +248,6 @@ plot(parpost)
 # Model assessment, Bayesian goodness of fit diagnostics:
 bgof(parpost,
      aux.iters = 10000,
-     n.deg     = 14,
+     n.deg     = 20,
      n.dist    = 15,
-     n.esp     = 9)
+     n.esp     = 15)

### 2020-04-07 analysis/paper/paper.Rmd
add code to the rmd file and edit the method section

@@
 
+```{r read-data}
+library(readxl)
+library(tidyverse)
+library(here)
+
+# read data
+burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
+```
+
+```{r tidy-data}
+# combine some periods
+burial_three_period_tidy <-
+  burial %>%
+  rename(burial_label = ID) %>%
+  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
+  filter(!is.na(Phase)) %>%
+  mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
+         Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
+         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
+  mutate_at(21:ncol(.), as.numeric) %>%
+  janitor::remove_empty(which = "cols") %>%
+  mutate(total = rowSums(.[c(21:48, 50, 55, 56)], na.rm = TRUE)) %>% #prestige goods
+  mutate(Porcelain = rowSums(.[c(40:48, 55, 56)], na.rm = TRUE)) %>% #B&W, porcelain, Anping, kendi
+  mutate(Porcelain = ifelse(Porcelain == 0, NA, Porcelain)) %>%
+  mutate(quantity = case_when(
+    total == 0 ~ "none",
+    total > 0 & total <= 7 ~ "low",
+    total > 7 & total < 100 ~ "medium",
+    total >= 100 ~ "high",
+    TRUE ~ "other")) %>% # the classification is based on the result of histogram
+  mutate(Age_scale = case_when(
+    `Age` %in% c("1","2") ~ "0-12",
+    `Age` == "3" ~ "12~20",
+    `Age` %in% c("4","5","6","7","8") ~ "+20",
+    TRUE ~ "NA")) %>%
+  mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
+         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
+         Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
+         Agate_bead_low = ifelse(Agate_bead == 1, 1, NA),
+         Agate_bead_med = ifelse(Agate_bead > 1 & Agate_bead <10, 1, NA),
+         Agate_bead_high = ifelse(Agate_bead > 10, 1, NA),
+         `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` < 100, 1, NA),
+         `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
+         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% 
+  # the low-med-high index is determined by interpretation for histograms for those burial good types
+  select(burial_label,
+         Phase,
+         Age_scale,
+         Gold_bead_low,
+         Gold_bead_med,
+         Gold_bead_high,
+         Agate_bead_low,
+         Agate_bead_med,
+         Agate_bead_high,
+         #Agate_bead, #female burials
+         #Golden_bead,
+         Porcelain,
+         Gold_leaf, #prestige good
+         fish_shape_knit, #prestige good
+         #Bell, #children's burials
+         quantity)
+         #total) # select specific variable to drop columns (uninformative variables)
+
+# number of each phase
+burial_three_period_number <-
+  burial_three_period_tidy %>%
+  count(Phase)
+```
+
+```{r network-object-pre}
+# filter pre burials
+burial_pre <-
+  burial_three_period_tidy %>%
+  filter(Phase == "pre") %>%
+  janitor::remove_empty(which = "cols")
+
+# create node list using burial index by phase
+#--------------------pre--------------------------------
+nodes_pre <-
+  burial_three_period_tidy %>%
+  filter(Phase == "pre") %>%
+  select(burial_label) %>%
+  rowid_to_column("id")
+
+# pair wise combinations for burials as index for later map function
+library(gtools)
+burial_comb_pre = combinations(length(nodes_pre$burial_label), #t(combn(nodes_pre$burial_label, 2))
+                               2, nodes_pre$burial_label,
+                               repeats = TRUE)
+colnames(burial_comb_pre) = c("burial_1", "burial_2")
+burial_comb_pre = as_tibble(burial_comb_pre)
+
+# create list for each burial that contains the burial good types and their counts
+edge_list_pre <-
+  burial_three_period_tidy %>%
+  select(burial_label, 4:12) %>% # need to change for each exploration
+  pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
+  #mutate(burial_connection = rep(unique(burial_label), length.out = length(burial_label)))
+  group_by(burial_label) %>%
+  nest()
+
+# pair-wise list
+common_counts_lst_pre <-
+map2(burial_comb_pre$burial_1,
+     burial_comb_pre$burial_2,
+     ~bind_cols(
+       edge_list_pre %>%
+       filter(burial_label == .x) %>%
+       unnest(data),
+       edge_list_pre %>%
+         filter(burial_label == .y) %>%
+         unnest(data)) %>%
+       rowwise() %>%
+       mutate(common_counts = sum(count, count1)))
+
+# count of ornament types in common between each pair of burials
+common_counts_vct_pre <- map_int(common_counts_lst_pre, ~sum(!is.na(.x$common_counts)))
+
+burial_comb_with_common_counts_pre <-
+burial_comb_pre %>%
+  mutate(common_counts = common_counts_vct_pre) %>%
+  mutate(common_counts = ifelse(burial_1 == burial_2, 0, common_counts)) # use 0 for self loop
+
+# change label to ids for node linking
+edges_pre <-
+  burial_comb_with_common_counts_pre %>%
+  left_join(nodes_pre, by = c("burial_1" = "burial_label")) %>%
+  rename(from = id) %>%
+  left_join(nodes_pre, by = c("burial_2" = "burial_label")) %>%
+  rename(to = id)
+
+edges_for_network_pre <-
+  select(edges_pre, from, to, common_counts) %>%
+  filter (!common_counts == 0) # remove rows with no goods in common
+ # %>% mutate(common_counts = ifelse(common_counts > 1, 1, common_counts)) # for unweighted network
+```
+
 The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
@@
 
+some concepts:
+- endogenous statistics that ties are modelled based on the existence of other ties
+- exogenous statistics that ties are formed according to monadic or dyadic node attributes
+
 # Methods
 
-We use burial data from different time periods at Kiwulan and consider the effects of both exogenous and endogenous factors that may provide meaningful explanation of the tendency of burials correlate with each other and the underlying mechnism.
-
-n concerning connectivity, inter-cultural contacts and exchange among those sites.
-Our purpose is to understand the characteristics of burial network and the mechanism of the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
-Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
-To answer our question for the shifts in social inequality during and after the arrival of the foreign presences, we hypothesize:
+We use burial data from different time periods at Kiwulan and consider the effects of both exogenous and endogenous factors that could provide meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
+To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
+We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
+To answer our question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
 
@@
 
-some concepts:
-- endogenous statistics that ties are modelled based on the existence of other ties
-- exogenous statistics that ties are formed according to monadic or dyadic node attributes
-
 ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles
@@
 It is a measure of the tendency for two actors who are tied to have x partners in common.
-In orther words, a tendency for those with shared partners to become tied, or tendency of ties to cluster together.
+In other words, a tendency for those with shared partners to become tied, or tendency of ties to cluster together.
 
@@
 A measure of the tendency of the number of edges into a node.
-gwdegree estimates the change in tie likelihood given the degree of the nodes involved, but with marginally decreasing weighting as degree increases(https://www.r-bloggers.com/ergm-tutorial/).
 This statistic captures the tendency towards centralisation in the degree distribution of the network.
+
+```{r network-ggraph}
+# network analysis using ggraph pkg
+library(tidygraph)
+library(ggraph)
+
+relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
+                               edges = edges_for_network_pre,
+                               directed = FALSE)
+
+relation_tidy_pre %>%
+  activate(edges) %>%
+  arrange(desc(common_counts))
+
+ggraph(relation_tidy_pre) +
+  geom_edge_link() +
+  geom_node_point() +
+  theme_graph()
+
+ggraph(relation_tidy_pre, layout = "graphopt") +
+  geom_node_point() +
+  geom_edge_link(aes(width = common_counts), alpha = 0.8) +
+  scale_edge_width(range = c(0.2, 2)) +
+  geom_node_text(aes(label = burial_label), repel = TRUE) +
+  labs(edge_width = "common item") +
+  theme_graph()
+```
+
+```{r network-network-pkg}
+# network analysis using network pkg
+library(network)
+
+burial_network_pre <-
+  network(edges_for_network_pre, # the network object
+          vertex.attr = nodes_pre, # node list
+          directed = FALSE, # specify whether the network is directed
+          ignore.eval = FALSE, # FALSE = weighted
+          loops = FALSE, # do we allow self ties (should not allow them)
+          matrix.type = "edgelist") # the type of input
+
+# plot
+plot(burial_network_pre, vertex.cex = 1) # last two nodes are missing?? only 27 nodes
+```
+
+```{r add-node-attribute-color}
+# prepare notwork objects for Bayesian ERGMs 
+library(statnet)
+library(Bergm)
+
+# Add attributes to the network object burial_network_pre
+set.vertex.attribute(burial_network_pre, "quantity", burial_pre$quantity)
+set.vertex.attribute(burial_network_pre, "age", burial_pre$Age_scale)
+
+# plot
+set.seed(30)
+quantity <- get.vertex.attribute(burial_network_pre, "quantity")
+age <- get.vertex.attribute(burial_network_pre, "age")
+
+ID <- get.vertex.attribute(burial_network_pre, "burial_label") # not sure how to get id on the network plot
+
+plot(burial_network_pre,
+     vertex.col = "quantity", # age
+     vertex.cex = 1.5)
+
+legend("topleft",
+       col = c(2, 3, 1, 4), # need to adjust each time
+       pch    = 20,
+       legend = unique(quantity), # age
+       title  = 'Burial good counts')
+
+#? can't get the items in legend in order, tried as.factor and set their level but does not work
+#? can't match color with categories, using an odd method at the moment
+```
 
@@
 
-```{r pre-model-ergm-1}
+```{r network-ERGMs}
+# model 1
 model.ergm <- burial_network_pre ~
@@
   density +
-  gwdegree(0.5, fixed = TRUE) +
   triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network, can be replaced by "gwdegree"
 summary(model.ergm)
-```
 
-```{r pre-model-ergm-2}
+# model 2, check here Morris et al. (2008) 
 # check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
-# also check here Morris et al. (2008)
 model.2 <- burial_network_pre ~ edges + # density
@@
 summary(model.2)
-```
 
-```{r pre-model-ergm-3}
+# model 3 with node attributes
 model.3 <- burial_network_pre ~ edges +  # the overall density of the network
-  nodematch('quantity') + # quantity-based homophily, categorical nodal attribute
-  # the similarity of connected nodes
+  nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
+  nodematch('age') +
   gwesp(0.2, fixed = TRUE) +    # transitivity
@@
 
-Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
-We specified the prior based on normal distribution that assumes a low density and high transitivity network.
-
-```{r pre-model-bergm}
+```{r Bayesian-ERGMs}
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 0, 1, 0) # prior mean corresponds to mean for each parameter
+prior.mean <- c(-3, 0, 0, 1, 0) # prior mean corresponds to mean for each parameter
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 4, 4) # covariance matrix structure
+prior.sigma <- diag(3, 5, 5) # covariance matrix structure
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
@@
 # Estimated posterior means, medians and 95% credible intervals for Models.3
-# bergmM: Bayesian exponential random graphs models under missing data using the approximate exchange algorithm
-parpost <- bergm(model.3,
+# bergmM: Bayesian exponential random graphs models under missing data using the approximate exchange algorithm. Use this because of missing data for age variable
+parpost <- bergmM(model.3,
                   prior.mean  = prior.mean,
                   prior.sigma = prior.sigma,
-                  burn.in     = 200, # burn-in iterations for every chain of the populationm, drops the first 200
+                  burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
                   main.iters  = 2000, # iterations for every chain of the population
@@
 
+# summary of parameter estimation
 summary(parpost)
+
+# plot the parameter posterior distribution
+plot(parpost)
 ```
+
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
+We specified the prior based on normal distribution that assumes a low density and high transitivity network.
 
@@
 θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.
+
+```{r Bayesian-gof}
+# Model assessment, Bayesian goodness of fit diagnostics:
+bgof(parpost,
+     aux.iters = 10000,
+     n.deg     = 14,
+     n.dist    = 15,
+     n.esp     = 9)
+
+# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
+# a result that is very difficult to obtain in practice (Caimo 2017)
+```
 

### 2020-04-08 analysis/paper/KWL_burial_pre.R
try different prior but not much difference, add sentences to statistics

@@ -302,36 +302,3 @@ bgof(parpost,
 # An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
 # a result that is very difficult to obtain in practice (Caimo 2017)
 
-#-------------------------set different parameters------------------------------
-# three competing models based on ERGM formulas to fit the data using gwesp and gwdegree
-
-mod.1 <- burial_network_pre ~ edges +
-  gwesp(0.2, fixed = TRUE)
-summary(mod.1)
-
-mod.2 <- burial_network_pre ~ edges +
-  gwdegree(0.8, fixed = TRUE)
-summary(mod.2)
-
-mod.3 <- burial_network_pre ~ edges +
-  gwesp(0.2, fixed = TRUE) +
-  gwdegree(0.8, fixed = TRUE)
-summary(mod.3)
-
-mod.sel <- bergm(mod.1,
-                 prior.mean  = NULL,
-                 prior.sigma = NULL,
-                 aux.iters = 100, # auxiliary iterations used for network simulation
-                 main.iters = 20, # iterations for every chain of the population
-                 burn.in = 200, # burn-in iterations for every chain of the population
-                 nchains = 8, # number of chains of the population MCMC
-                 gammas = 0.5)
-
-summary(parpost)
-plot(parpost)
-
-mod.sel <- bergm(mod.1, iters = 25000,
-                 aux.iters = 2000,
-                 main.iters = rep(700, 3),
-                 burn.ins = rep(100, 3),
-                 gammas = c(1, 1, 0.8))

### 2020-04-08 analysis/paper/paper.Rmd
try different prior but not much difference, add sentences to statistics

@@
 output: 
-    bookdown::html_document2:
+    bookdown::word_document2:
       fig_caption: yes
@@
 To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
-We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
+We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.
+Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
 To answer our question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
@@
 Every term in an ERGM must have an associated algorithm for computing its value for network [@Morris2008].
-The terms we selected for nwtwork specification include:
+The terms we selected for network specification include:
 
@@
 
+4. nodematch: uniform homophily that refers to the similarity of connected nodes. A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics.
+
 ```{r network-ggraph}
@@
 
-ggraph(relation_tidy_pre) +
-  geom_edge_link() +
-  geom_node_point() +
-  theme_graph()
-
+# plot
 ggraph(relation_tidy_pre, layout = "graphopt") +
@@
 
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
+We specified the prior based on normal distribution that assumes a low density and high transitivity network.
+
 ```{r Bayesian-ERGMs}
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 0, 0, 1, 0) # prior mean corresponds to mean for each parameter
+# need to observe the graph and select suitable prior 
+prior.mean <- c(1, 0, 0, 3, 0) # prior mean corresponds to mean for each parameter
 # follow Alberto Caimo et al. (2015) hospital example
@@
 
-Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
-We specified the prior based on normal distribution that assumes a low density and high transitivity network.
-
 The summary presents the statistics that each θ corresponds to the parameter specified in ERGM previously.
@@
 θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.
+
+The plots from left to right show estimated marginal posterior densities, trace plots, and autocorrelation plots.
 
@@
 The grey lines mark the 95% intervals.
-
-inline code `r x` example
 

### 2020-04-08 analysis/paper/paper.Rmd
merge

@@
 To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
-We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.
-Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
+We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
 To answer our question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
@@
 # plot
+ggraph(relation_tidy_pre, layout = "graphopt") +
+  geom_node_point() +
+  geom_edge_link(aes(width = common_counts), alpha = 0.8) +
+  scale_edge_width(range = c(0.2, 2)) +
+  geom_node_text(aes(label = burial_label), repel = TRUE) +
+  labs(edge_width = "common item") +
+  theme_graph()
+```
+
+```{r network-ggraph}
+# network analysis using ggraph pkg
+library(tidygraph)
+library(ggraph)
+
+relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
+                               edges = edges_for_network_pre,
+                               directed = FALSE)
+
+relation_tidy_pre %>%
+  activate(edges) %>%
+  arrange(desc(common_counts))
+
+ggraph(relation_tidy_pre) +
+  geom_edge_link() +
+  geom_node_point() +
+  theme_graph()
+
 ggraph(relation_tidy_pre, layout = "graphopt") +
@@
 
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
+We specified the prior based on normal distribution that assumes a low density and high transitivity network.
+
 The summary presents the statistics that each θ corresponds to the parameter specified in ERGM previously.
@@
 The plots from left to right show estimated marginal posterior densities, trace plots, and autocorrelation plots.
+
+```{r Bayesian-gof}
+# Model assessment, Bayesian goodness of fit diagnostics:
+bgof(parpost,
+     aux.iters = 10000,
+     n.deg     = 14,
+     n.dist    = 15,
+     n.esp     = 9)
+
+# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
+# a result that is very difficult to obtain in practice (Caimo 2017)
+```
+
+Figure shows the Bayesian goodness of fit diagnostics plots where the red lines represent the distributions of the observed data and the box plots represent the GOF distributions of 10000 network graphs simulated from ERGMs based on the estimated posterior distribution.
+The grey lines mark the 95% intervals.
 

### 2020-04-08 analysis/paper/KWL_burial_pre.R
add label and other effects to network plot

@@ -192,8 +192,8 @@ legend("topleft",
 # predict ties, as function of network structures,
 # endogenous statistics that ties are modelled based on the existence of other ties
 # exogenous statistics that ties are formed according to monadic or dyadic node attributes
-# Kiwulan case
 # https://www.r-bloggers.com/ergm-tutorial/
+# https://rdrr.io/cran/ergm/man/ergm-terms.html
 
 # every term in an ERGM must have an associated algorithm for computing its value for network
 # Morris et al. 2008
@@ -229,6 +229,7 @@ model.2 <- burial_network_pre ~ edges + # density
 summary(model.2)
 
 #--------------------Bayesian inference for ERGMs-------------------------
+# https://rdrr.io/cran/Bergm/man/bergmM.html
 model.3 <- burial_network_pre ~ edges +  # the overall density of the network
   nodematch('quantity') +    # quantity-based homophily, categorical nodal attribute
   # the similarity of connected nodes

### 2020-04-08 analysis/paper/burials_by_age.R

@@ -181,8 +181,11 @@ age <- get.vertex.attribute(burial_network_pre, "age")
 
 ID <- get.vertex.attribute(burial_network_pre, "burial_label") # not sure how to get id on the network plot
 plot(burial_network_pre,
+     displaylabels = TRUE,
      vertex.col = "age", # quantity
-     vertex.cex = 1.5)
+     vertex.cex = degree(burial_network_pre, cmode = 'indegree') / 5, # size nodes to their in-degree
+     #vertex.sides = ifelse(burial_network_pre %v% "", 4, 50),
+     pad = 1) #protects the labels from getting clipped
 
 legend("topleft",
        col = c(3, 1, 2, 4), # need to adjust each time
@@ -248,7 +251,7 @@ plot(parpost)
 # try 4 competing models
 m1 <- burial_network_pre ~ edges +
   nodematch("age") +
-  nodematch("quantity")+
+  nodematch("quantity") +
   gwesp(0.8, fixed = TRUE) +
   gwdegree(0.5, fixed = TRUE)
 

### 2020-04-09 analysis/paper/KWL_burial_pre.R
increase the prior number for gwesp to 5 and the gof improves a bit

@@ -216,7 +216,8 @@ model.2 <- burial_network_pre ~ edges + # density
   # triad-based clustering according to dyad-based configurations instead of node-based
   # number means weight parameter alpha, which controls the rate of declining marginal returns
   # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
-  # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
+  # not much difference in a range of 0-1.5, the lower the value, the less likely the model is to be degenerate
+  # start close to zero and move up, looking for how well it matches the count of triangles
   # ergm can estimate the parameter from the data by using fixed=FALSE
   # http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees

### 2020-04-09 analysis/paper/burials_by_age.R

@@ -252,7 +252,7 @@ plot(parpost)
 m1 <- burial_network_pre ~ edges +
   nodematch("age") +
   nodematch("quantity") +
-  gwesp(0.8, fixed = TRUE) +
+  gwesp(0.8, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
   gwdegree(0.5, fixed = TRUE)
 
 m2 <- burial_network_pre ~ edges +
@@ -270,7 +270,7 @@ m4 <- burial_network_pre ~ edges +
   gwdegree(0.8, fixed = TRUE)
 
 mod <- bergmM(m1,
-             prior.mean  = c(1, 0, 0, 3, 0),
+             prior.mean  = c(1, 0, 0, 5, 0),
              prior.sigma = diag(3, 5, 5),
              burn.in     = 200,
              main.iters  = 2000,
@@ -283,6 +283,6 @@ summary(mod)
 # Model assessment, Bayesian goodness of fit diagnostics:
 bgof(mod,
      aux.iters = 10000,
-     n.deg     = 14,
+     n.deg     = 15,
      n.dist    = 15,
-     n.esp     = 9)
+     n.esp     = 10)

### 2020-04-10 analysis/paper/paper.Rmd
add label and other effects to network plot
increase the prior number for gwesp to 5 and the gof improves a bit
add a reference and minor update

--- 2020-04-08 analysis/paper/paper.Rmd
@@
 She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status.
-The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; Li and Wu 2006).
+The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; @LiandWu2006).
 Analysing burial data will give an insight into the social reactions of Kiwulan residents and provide a way to examine and explain the changes in social inequality at Kiwulan before and after European contact.
@@
 
-The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials (Chen 2005; Cheng 2008), which might indicate the pursuit of prestige or wealth.
+The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials [@Cheng2008], which might indicate the pursuit of prestige or wealth.
 In addition to prestige, European colonial powers could have provided military support to indigenous allies.
@@
 4. nodematch: uniform homophily that refers to the similarity of connected nodes. A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics.
+
+```{r network-ggraph}
+# network analysis using ggraph pkg
+library(tidygraph)
+library(ggraph)
+
+relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
+                               edges = edges_for_network_pre,
+                               directed = FALSE)
+
+relation_tidy_pre %>%
+  activate(edges) %>%
+  arrange(desc(common_counts))
+
+# plot
+ggraph(relation_tidy_pre, layout = "graphopt") +
+  geom_node_point() +
+  geom_edge_link(aes(width = common_counts), alpha = 0.8) +
+  scale_edge_width(range = c(0.2, 2)) +
+  geom_node_text(aes(label = burial_label), repel = TRUE) +
+  labs(edge_width = "common item") +
+  theme_graph()
+```
 
@@
 
+```{r Bayesian-gof}
+# Model assessment, Bayesian goodness of fit diagnostics:
+bgof(parpost,
+     aux.iters = 10000,
+     n.deg     = 14,
+     n.dist    = 15,
+     n.esp     = 9)
+
+# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
+# a result that is very difficult to obtain in practice (Caimo 2017)
+```
+
+Figure shows the Bayesian goodness of fit diagnostics plots where the red lines represent the distributions of the observed data and the box plots represent the GOF distributions of 10000 network graphs simulated from ERGMs based on the estimated posterior distribution.
+The grey lines mark the 95% intervals.
+X-axis means the proportion of nodes
+
 # Discussion

### 2020-04-10 analysis/paper/burials_by_age.R

@@ -179,11 +179,10 @@ set.seed(30)
 quantity <- get.vertex.attribute(burial_network_pre, "quantity")
 age <- get.vertex.attribute(burial_network_pre, "age")
 
-ID <- get.vertex.attribute(burial_network_pre, "burial_label") # not sure how to get id on the network plot
-plot(burial_network_pre,
+ID <- get.vertex.attribute(burial_network_pre, "burial_label")
      displaylabels = TRUE,
      vertex.col = "age", # quantity
-     vertex.cex = degree(burial_network_pre, cmode = 'indegree') / 5, # size nodes to their in-degree
+     vertex.cex = degree(burial_network_pre, cmode = 'indegree') / 5, #size nodes to their in-degree
      #vertex.sides = ifelse(burial_network_pre %v% "", 4, 50),
      pad = 1) #protects the labels from getting clipped
 

### 2020-04-10 analysis/paper/paper.Rmd
add sentences to Bayesian section
update the description for prior in Rmd

@@
 This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
+For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
 
@@
 
-ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles
-
 We used ergm package in R for the model specification.
@@
 
-4. nodematch: uniform homophily that refers to the similarity of connected nodes. A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics.
+4.
+nodematch: uniform homophily between nodes that refers to the similarity of connected nodes.
+A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics.
+We use wealth index and age as covariants that reflects the similarity between nodes.
 
@@
 
-Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis [@Caimo2017].
-We specified the prior based on normal distribution that assumes a low density and high transitivity network.
+Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis if there is no previous observation for a network[@Caimo2017].
+Since our original data presents a pattern of dense ties, we specified the prior based on our observation that assumes a high density and high transitivity network.
 
@@
 ```
+
+## Structure Effects
 

### 2020-04-12 analysis/paper/paper.Rmd
reorganize some paragraphs and add some subheadings

@@
 ---
-title: "Bayesian exponential random graph modelling of KWL Burial network"
+title: "Bayesian exponential random graph modeling of KWL Burial network"
 author:
@@
 
+<!--burial study- network analysis-->
 Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
@@
 
+<!--current approaches to network analysis-->
 Exponential Random Graph Models (ERGMs) are an important family of statistical model for network analysis to describe the formation of ties in a network.
@@
 A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
+
+## Bayesian modeling 
+<!--Bayesian inference for Exponential Random Graph Models--> 
+Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
+This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
+For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
+- weighted data
 
@@
 
+<!--archaeological setting-->
 The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
@@
 
+<!--chronology-->
 Burials were assigned to the pre-European period, European and post-European period, and Chinese period for network comparison.
@@
 
+<!--unanswered issues-->
 The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials [@Cheng2008], which might indicate the pursuit of prestige or wealth.
@@
 
-## Bayesian inference for Exponential Random Graph Models
-
-Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
-This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
-For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
-
-some concepts:
-- endogenous statistics that ties are modelled based on the existence of other ties
-- exogenous statistics that ties are formed according to monadic or dyadic node attributes
-
 # Methods
 
-We use burial data from different time periods at Kiwulan and consider the effects of both exogenous and endogenous factors that could provide meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
+<!--procesures-->
+
+## data
+
+We use burial data collected based on published excavation report, and original excavation records with detailed stratigraphy and context description.
+Burial data provides information about direction, size, depth, age, gender, classes and quantity of grave goods, and spatial relationship between them.
+We consider the effects of both exogenous and endogenous factors to generate meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
 To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
@@
 
-## Statistics
+## Model specification and estimation
+<!--Network construction-->
+
+some concepts:
+- endogenous statistics that ties are modelled based on the existence of other ties
+- exogenous statistics that ties are formed according to monadic or dyadic node attributes
 
@@
 We use wealth index and age as covariants that reflects the similarity between nodes.
+
+<!--Model fitting and comparison to alternative approaches-->
+
+<!-- Goodness-of-fit assessment-->
+
+# Results
 
@@
 
-# Results
+## ERGM parameter estimation
+<!--interpretation of network plots across phases-->
 
@@
 ```
+
+<!--interpretation of Bayesian modeling-->
 
@@
 
-# Acknowledgements
+# Acknowledgments
 

### 2020-04-12 analysis/paper/burials_by_age.R

@@ -275,7 +275,7 @@ m1 <- burial_network_pre ~ edges +
   nodematch("quantity") +
   nodematch("gender") +
   gwesp(0.8, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
-  gwdegree(0.5, fixed = TRUE)
+  gwdegree(0.2, fixed = TRUE)
 
 m2 <- burial_network_pre ~ edges +
   nodematch("age") +
@@ -292,7 +292,7 @@ m4 <- burial_network_pre ~ edges +
   gwdegree(0.8, fixed = TRUE)
 
 mod <- bergmM(m1,
-             prior.mean  = c(1, 0, 0, 0, 5, 0),
+             prior.mean  = c(1, 0, 0, 0, 10, 0),
              prior.sigma = diag(3, 6, 6),
              burn.in     = 200,
              main.iters  = 2000,
@@ -307,4 +307,4 @@ bgof(mod,
      aux.iters = 10000,
      n.deg     = 15,
      n.dist    = 15,
-     n.esp     = 10)
+     n.esp     = 15)

### 2020-04-13 analysis/paper/paper.Rmd
edit the ERGM para and add references

@@
 <!--current approaches to network analysis-->
-Exponential Random Graph Models (ERGMs) are an important family of statistical model for network analysis to describe the formation of ties in a network.
-In traditional statistical model for network, a dyad, the pair of nodes, is measured by the presence of a tie according to the attributes of the nodes or ties, where each predictor variable for forming a tie is measured separately.
-In contrast to traditional model where variables not interacting with each other, ERGMs assumes the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes and ties (Robins et al.
-2007).
-In a network using ERGMs, predictor variables, called "network statistics", are direct functions of the ties themselves that the probability of occurrence of the configuration of ties can be hypothesized [@Morris2008].
-This approach enables networks to be constructed in a variety of assumption and contexts to model social network in the past using archaeological data [@Amati2019].
-However, ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
+Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007].
+In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measure separately by the presence of a tie without considering interactions among ties [@Morris2008].
+By contrast, statistical models assumes social networks as dependent variables that can represent the network dependencies between tie variables [@Snijders2011].
+In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes/ties.
+Predictor variables, called network statistics, such as reciprocity, transitivity, homophily, are direct functions of the ties themselves where the probability of occurrence of the configuration of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+With the development of Markov chain Monte Carlo (MCMC) algorithms that produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, ERGMs can deal with complicated dependence patterns observed at one moment in time [@Snijders2006].
+The advantage of networks specification under assumptions and contexts can be useful to explain archaeological network in a given time.
+@Amati2019 demonstrate the use of ERGMs to reconstruct networks consist of a set of sites in the Caribbean during the period AD 100–400.
+They found out the structure of network can be efficiently explained by multiple interdependent mechanisms instead of hypotheses of distance and cultural homophily exclusively.
+However, they also point out some limitation of ERGMs
+
+However, two major limitations ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
 A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].

### 2020-04-13 analysis/paper/KWL_burial_pre.R

@@ -219,7 +219,7 @@ model.2 <- burial_network_pre ~ edges + # density
   # number means weight parameter alpha, which controls the rate of declining marginal returns
   # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
   # not much difference in a range of 0-1.5, the lower the value, the less likely the model is to be degenerate
-  # start close to zero and move up, looking for how well it matches the count of triangles
+  # start close to zero and move up, looking for how well it matches the count of triangles, https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2443947/
   # ergm can estimate the parameter from the data by using fixed=FALSE
   # http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees

### 2020-04-14 Note.Rmd
update the description for prior in Rmd
edit intro and download hergm

@@
 http://acaimo.github.io/Bergm/tutorials/Dec_18.html
+
+# Ergms
+ERGM- consider transitivity by taking into account high-order k-triangles. Triangles and higher order cycles
 
@@
 Activity areas will reflect uneven distribution of high ranked subsistence resources, and cache pits will present higher variability in size and contents.
+
+# acceptance rate
+https://www.ibm.com/support/pages/acceptance-rate-amos-bayesian-sem
+The acceptance rate is the proportion of times the Metropolis sampling algorithm generates a sample of parameter values that differs from the previous sample.
+For example, an acceptance rate of .2 means that, in generating a sequence of parameter values, the algorithm generates new parameter values 20% of the time and repeats the previous parameter values 80% of the time.
+
+One rule of thumb is that the MCMC algorithm is most effective when the acceptance rate is between .2 and .5.
+If you see that the acceptance rate is outside this range, it may be worthwhile to try adjusting the tuning parameter.
+If the acceptance rate is less than .2, try making the tuning parameter smaller.
+If the acceptance rate is greater than .5, try making the tuning parameter larger.

### 2020-04-14 analysis/paper/paper.Rmd
edit intro and download hergm

@@
 They found out the structure of network can be efficiently explained by multiple interdependent mechanisms instead of hypotheses of distance and cultural homophily exclusively.
-However, they also point out some limitation of ERGMs
-
-However, two major limitations ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
+However, they also point out some limitations of application such as static outcome and sensitive to missing data.
+The two major limitations ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
 A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
@@
 <!--Network construction-->
+
+it can accomodate binary network statistics to describe the weighted network topology [@Snijders2006]
 

### 2020-04-14 analysis/paper/burials_by_age.R

@@ -184,8 +184,8 @@ set.vertex.attribute(burial_network_pre, "gender", burial_pre$gender)
 set.seed(30)
 quantity <- get.vertex.attribute(burial_network_pre, "quantity")
 age <- get.vertex.attribute(burial_network_pre, "age")
-
 ID <- get.vertex.attribute(burial_network_pre, "burial_label")
+
 plot(burial_network_pre,
      displaylabels = TRUE,
      vertex.col = "quantity",
@@ -207,7 +207,7 @@ legend("topleft",
 
 # ? can't get the items in legend in order
 # ? tried as.factor and set their level but does not work
-# ? can't match color with categories, using an odd method at the moment
+# ? can't match color with categories, adjust manually every time
 
 #------------------creating ERGM model-------------------------------------
 # model 1 considers density and triad relations (for cluster)
@@ -293,7 +293,7 @@ m4 <- burial_network_pre ~ edges +
   gwdegree(0.8, fixed = TRUE)
 
 mod <- bergmM(m1,
-             prior.mean  = c(1, 0, 0, 0, 5, 0, 0),
+             prior.mean  = c(2, 0, 0, 0, 5, 0, 0),
              prior.sigma = diag(3, 7, 7),
              burn.in     = 200,
              main.iters  = 2000,
@@ -309,4 +309,9 @@ bgof(mod,
      n.deg     = 15,
      n.dist    = 15,
      n.esp     = 15)
+
+#-------------------------using hergm--------------------------
+
+
+
 

### 2020-04-14 analysis/paper/experiment.R

@@ -59,3 +59,10 @@ example_network <-
 
 # plot
 plot(example_network, vertex.cex = 1) # It seems the last two nodes are missing?? only 27 nodes
+
+# try hergm
+library("hergm")
+
+# Download package from CRAN archive
+require(devtools)
+install_version("hergm", version = "4.1-6", repos = "http://cran.us.r-project.org")

### 2020-04-15 analysis/paper/paper.Rmd
add code of post-Euro to rmd file and knit it

@@
     TRUE ~ "NA")) %>%
+    mutate(gender = case_when(
+    `Gender` %in% c("1","2") ~ "male",
+    `Gender` %in% c("3","4") ~ "female",
+    TRUE ~ "NA")) %>%
   mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
@@
          Age_scale,
+         gender,
          Gold_bead_low,
@@
 
-```{r network-object-pre}
+```{r prep-network-object-pre}
 # filter pre burials
@@
   burial_three_period_tidy %>%
-  select(burial_label, 4:12) %>% # need to change for each exploration
+  select(burial_label, 5:13) %>% # need to change for each exploration
   pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
@@
  # %>% mutate(common_counts = ifelse(common_counts > 1, 1, common_counts)) # for unweighted network
+```
+
+```{r prep-network-object-post}
+# filter post burials
+burial_post <-
+  burial_three_period_tidy %>%
+  filter(Phase == "post") %>%
+  janitor::remove_empty(which = "cols")
+
+# create node list
+nodes_post <-
+  burial_three_period_tidy %>%
+  filter(Phase == "post") %>%
+  select(burial_label) %>%
+  rowid_to_column("id")
+
+# pair wise combinations for burials as index for later map function
+library(gtools)
+burial_comb_post = combinations(length(nodes_post$burial_label),
+                               2, nodes_post$burial_label,
+                               repeats = TRUE)
+colnames(burial_comb_post) = c("burial_1", "burial_2")
+burial_comb_post = as_tibble(burial_comb_post)
+
+# create list for each burial that contains the burial good types and their counts
+edge_list_post <-
+  burial_three_period_tidy %>%
+  select(burial_label, 5:13) %>% # need to change for each exploration
+  pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
+  group_by(burial_label) %>%
+  nest()
+
+# pair-wise list
+common_counts_lst_post <-
+  map2(burial_comb_post$burial_1,
+       burial_comb_post$burial_2,
+       ~bind_cols(
+         edge_list_post %>%
+           filter(burial_label == .x) %>%
+           unnest(data),
+         edge_list_post %>%
+           filter(burial_label == .y) %>%
+           unnest(data)) %>%
+         rowwise() %>%
+         mutate(common_counts = sum(count, count1)))
+
+# count of ornament types in common between each pair of burials
+common_counts_vct_post <- map_int(common_counts_lst_post, ~sum(!is.na(.x$common_counts)))
+
+burial_comb_with_common_counts_post <-
+  burial_comb_post %>%
+  mutate(common_counts = common_counts_vct_post) %>%
+  mutate(common_counts = ifelse(burial_1 == burial_2, 0, common_counts)) # use 0 for self loop
+
+# change label to ids for node linking
+edges_post <-
+  burial_comb_with_common_counts_post %>%
+  left_join(nodes_post, by = c("burial_1" = "burial_label")) %>%
+  rename(from = id) %>%
+  left_join(nodes_post, by = c("burial_2" = "burial_label")) %>%
+  rename(to = id)
+
+edges_for_network_post <-
+  select(edges_post, from, to, common_counts) %>%
+  filter (!common_counts == 0) # remove rows with no goods in common
 ```
@@
 library(ggraph)
-
+# pre-Euro
 relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
@@
   theme_graph()
-```
 
-```{r network-ggraph}
-# network analysis using ggraph pkg
-library(tidygraph)
-library(ggraph)
-
-relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
-                               edges = edges_for_network_pre,
+# post-Euro
+relation_tidy_post <- tbl_graph(nodes = nodes_post,
+                               edges = edges_for_network_post,
                                directed = FALSE)
 
-relation_tidy_pre %>%
+relation_tidy_post %>%
   activate(edges) %>%
@@
 # plot
-ggraph(relation_tidy_pre, layout = "graphopt") +
+ggraph(relation_tidy_post, layout = "graphopt") +
   geom_node_point() +
@@
 
-```{r network-network-pkg}
-# network analysis using network pkg
+```{r network-object}
 library(network)
@@
 
-# plot
-plot(burial_network_pre, vertex.cex = 1) # last two nodes are missing?? only 27 nodes
+burial_network_post <-
+  network(edges_for_network_post,
+          vertex.attr = nodes_post,
+          directed = FALSE,
+          ignore.eval = FALSE, 
+          loops = FALSE, 
+          matrix.type = "edgelist") 
 ```
@@
 set.vertex.attribute(burial_network_pre, "age", burial_pre$Age_scale)
+set.vertex.attribute(burial_network_pre, "gender", burial_pre$gender)
 
@@
 
-ID <- get.vertex.attribute(burial_network_pre, "burial_label") # not sure how to get id on the network plot
-
 plot(burial_network_pre,
-     vertex.col = "quantity", # age
-     vertex.cex = 1.5)
+     displaylabels = TRUE,
+     vertex.col = "quantity",
+     vertex.cex = degree(burial_network_pre, cmode = 'indegree')/5, #size nodes
+     pad = 1)
 
@@
 #? can't match color with categories, using an odd method at the moment
+
+# Add attributes to the network object burial_network_post
+set.vertex.attribute(burial_network_post, "quantity", burial_post$quantity)
+set.vertex.attribute(burial_network_post, "age", burial_post$Age_scale)
+set.vertex.attribute(burial_network_post, "gender", burial_post$gender)
+
+# plot
+set.seed(30)
+quantity <- get.vertex.attribute(burial_network_post, "quantity")
+age <- get.vertex.attribute(burial_network_post, "age")
+ID <- get.vertex.attribute(burial_network_post, "burial_label") # not sure how to get id on the network plot
+
+plot(burial_network_post,
+     displaylabels = TRUE,
+     vertex.col = "quantity",
+     vertex.cex = degree(burial_network_post, cmode = 'indegree') / 13, #size nodes to their in-degree
+     #vertex.sides = ifelse(burial_network_pre %v% "", 4, 50),
+     pad = 1)
+
+legend("topleft",
+       col = c(4, 3, 2, 1), # need to adjust each time
+       pch    = 20,
+       legend = unique(quantity),
+       title  = 'Burial good counts')
 ```
@@
 model.3 <- burial_network_pre ~ edges +  # the overall density of the network
-  nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
+  nodematch('quantity') + # quantity-based homophily
   nodematch('age') +
-  gwesp(0.2, fixed = TRUE) +    # transitivity
-  gwdegree(0.8, fixed = TRUE)   # popularity
+  nodematch('gender') +
+  gwesp(0.75, fixed = TRUE) +    # transitivity
+  gwnsp(0.75, fixed = TRUE) +
+  gwdegree(0.2, fixed = TRUE) # popularity
 summary(model.3)
@@
 # need to observe the graph and select suitable prior 
-prior.mean <- c(1, 0, 0, 3, 0) # prior mean corresponds to mean for each parameter
+prior.mean <- c(2, 0, 0, 0, 5, 0, 0) 
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 5, 5) # covariance matrix structure
+prior.sigma <- diag(3, 7) # covariance matrix structure
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
@@
                   nchains     = 8, # number of chains of the population MCMC
-                  gamma       = 0.7) # scalar; parallel adaptive direction sampling move factor, acceptance rate
+                  gamma       = 0.2) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
@@
      aux.iters = 10000,
-     n.deg     = 14,
+     n.deg     = 15,
      n.dist    = 15,
-     n.esp     = 9)
-
-# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
-# a result that is very difficult to obtain in practice (Caimo 2017)
-```
-
-Figure shows the Bayesian goodness of fit diagnostics plots where the red lines represent the distributions of the observed data and the box plots represent the GOF distributions of 10000 network graphs simulated from ERGMs based on the estimated posterior distribution.
-The grey lines mark the 95% intervals.
-
-```{r Bayesian-gof}
-# Model assessment, Bayesian goodness of fit diagnostics:
-bgof(parpost,
-     aux.iters = 10000,
-     n.deg     = 14,
-     n.dist    = 15,
-     n.esp     = 9)
+     n.esp     = 15)
 

### 2020-04-15 analysis/paper/paper.Rmd
add some more ergm terms into model specification section

@@
 
-<!--procesures-->
+<!--procedures-->
 
@@
 3.
+gwnsp (geometrically weighted non-edgewise shared partner statistics): a function of the non-edgewise shared partner statistics.
+The edgewise and non-edgewise shared partner statistics reflect tendencies toward transitivity abd compared together would give a more comprehensive result.
+
+4.
 gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once.
@@
 
-4.
+5.
 nodematch: uniform homophily between nodes that refers to the similarity of connected nodes.
@@
 We use wealth index and age as covariants that reflects the similarity between nodes.
+Here we consider age, gender, and index of wealth as node attributes covariants.
 

### 2020-04-15 analysis/paper/burials_by_age.R

@@ -293,7 +293,7 @@ m4 <- burial_network_pre ~ edges +
   gwdegree(0.8, fixed = TRUE)
 
 mod <- bergmM(m1,
-             prior.mean  = c(2, 0, 0, 0, 5, -1, 0),
+             prior.mean  = c(2, 0, 0, 0, 4, -1, 0),
              prior.sigma = diag(3, 7, 7),
              burn.in     = 200,
              main.iters  = 2000,

### 2020-04-16 analysis/paper/paper.Rmd
add description to the pre-Euro network graph

@@
 
+The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle.
+Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods is limited to burials with rich goods.
+
 ```{r add-node-attribute-color}

### 2020-05-28 analysis/paper/paper.Rmd
work on the background section

@@
 <!--archaeological setting-->
-The burial data we analyzed in this paper are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
+The burial data we analyzed are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
 The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
-The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for comparison.
-Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and in-situ wooden posts.
-Previous research indicates the presence of economic inequality between burials, but there is debate about whether these represent an egalitarian or stratified society [@Cheng2008; @Hsieh2012].
+The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for temporal comparison.
+Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and `in-situ` wooden posts.
+
+Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials, accompanied with a debate over the differentiation represent an economic inequality supporting an egalitarian society or political inequality indicating a stratified society [@Cheng2008; @Hsieh2012].
 @Cheng2008 interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
@@
 <!--chronology-->
-Burials were assigned to the pre-European period, European and post-European period, and Chinese period for network comparison.
+Burials were primarily assigned to the pre-European period, European and post-European period, and Chinese period for network comparison.
 Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
@@
 
-We use burial data collected based on published excavation report, and original excavation records with detailed stratigraphy and context description.
+We use burial data collected from published excavation report, and original excavation records with detailed stratigraphy and context description.
+Given that burials in Chinese phase are far less than previous two phases, we only describe our observations for those burials for comparison.
 Burial data provides information about direction, size, depth, age, gender, classes and quantity of grave goods, and spatial relationship between them.
@@
 We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.
-Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but the degree over time and causes remain unclear.
-To answer our question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
+Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but how does the degree changes over time and what variables decide the changes remain unclear.
+To answer the research question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
 
@@
 
-Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, low, and none, according to the total number of burial goods as an index of wealth.
+Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, and none as an index of wealth.
+In addition to wealth, we also consider another two node variables, age and gender, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
 
@@
 We used ergm package in R for the model specification.
-Every term in an ERGM must have an associated algorithm for computing its value for network [@Morris2008].
+Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
 The terms we selected for network specification include:
@@
 
-The summary presents the statistics that each θ corresponds to the parameter specified in ERGM previously.
+The network statistics presents that each θ corresponds to the parameter specified in ERGM previously.
 In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.

### 2020-05-30 analysis/paper/paper.Rmd
udpate background section

@@
 <!--archaeological setting-->
-The burial data we analyzed are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century.
+We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007].
 The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
-The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for temporal comparison.
-Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and `in-situ` wooden posts.
+The burials are mostly located in the AD section, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
+Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
 
-Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials, accompanied with a debate over the differentiation represent an economic inequality supporting an egalitarian society or political inequality indicating a stratified society [@Cheng2008; @Hsieh2012].
-@Cheng2008 interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy.
-However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods.
-She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status.
-The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; @LiandWu2006).
-Analysing burial data will give an insight into the social reactions of Kiwulan residents and provide a way to examine and explain the changes in social inequality at Kiwulan before and after European contact.
+Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials.
+However, there is a debate over whether the differentiation between burials hints an economic inequality supporting an egalitarian society [@Hsieh2012], or political inequality indicating a stratified society [@Cheng2008].
+@Cheng2008 interprets the unequal distribution of prestige materials, especially the gold-foil beads between burials, as evidence for hierarchy.
+However, @Hsieh2012 suggests a relatively egalitarian society based on the comparative analysis of frequency and proportion of all burial goods.
+She found that the burials with rich goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
+The use of prestige goods in burials might be traced back before the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006).
+For example, the Spanish document by Spanish Priest documented that the Spanish soldier used agate beads to exchange materials from local indigenous people since they were valuable in their culture[Citation].
+Also, there was a custom that people would bury imported materials with deceased to display their status and influence [Ciation].
+Despite we observed the economic differentiation in burial goods and the value of imported goods according to historical documents, the extent of economic differentiation and how it changed over time remain unclear.
+Analyzing all burial data using network analysis approaches and compare the networks before and after the foreign presences will give an insight into the degree and changes in social inequality.
+This is important to understand the social reactions of Kiwulan residents in the context of foreign contact.
 
 <!--chronology-->
-Burials were primarily assigned to the pre-European period, European and post-European period, and Chinese period for network comparison.
+We assigned burials to the pre-European period, European and post-European period for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
 Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
-The chronology is then evaluated by the burials with light grey glazed jars, also called An-ping jars produced most likely in China and circulated in Southeast Asia from the late 16th to 18th century but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
-In addition to the diagnostic porcelain, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007].
-Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting from incomplete data.
+The chronology is then evaluated by the burials with time marker, light gray glazed jars, also called An-ping jars.
+The jars were believed to be produced in Southeast China and circulated in Southeast Asia from the late 16th to 18th century, but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
+In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007].
+Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data.
 
 <!--unanswered issues-->
-The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials [@Cheng2008], which might indicate the pursuit of prestige or wealth.
+The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
 In addition to prestige, European colonial powers could have provided military support to indigenous allies.
@@
 A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
-In this study, I will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.
+In this study, We will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.
 
@@
 
-The network statistics presents that each θ corresponds to the parameter specified in ERGM previously.
+The network statistics presents each θ that corresponds to the parameter specified in ERGM previously.
 In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.

### 2020-06-03 analysis/paper/paper.Rmd
udpate method section

@@
 
-We use burial data collected from published excavation report, and original excavation records with detailed stratigraphy and context description.
-Given that burials in Chinese phase are far less than previous two phases, we only describe our observations for those burials for comparison.
-Burial data provides information about direction, size, depth, age, gender, classes and quantity of grave goods, and spatial relationship between them.
-We consider the effects of both exogenous and endogenous factors to generate meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
-To be specific, our purpose is to understand the underlying mechanisms of the network structures and how the relationship influence on the pattern of accumulation of wealth during different foreign contact periods.
-We focus on connectivity, clusters, centralization, and homophily among those burials that could reflect the social relations between individuals guided by social structures.
-Previous studies have shown that there are some evidence of social inequality observed from Kiwulan but how does the degree changes over time and what variables decide the changes remain unclear.
-To answer the research question for the shifts in social inequality to a more wealth differentiation during and after the arrival of the foreign presences, we hypothesize:
+We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy.
+Given that burials in Chinese phase are less than previous two phases, we only compare these burials by their grave goods and consitions without further network analysis.
+Burial data provides information about the location, orientation, size, depth, age, gender, quantity of grave goods, and spatial relationship among them.
+We consider the effects of both exogenous and endogenous factors for meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
+To be specific, our purpose is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods.
+We focus on connectivity, clusters, centralization, and homophily those properties of network that can reflect the social relations between individuals guided by social structures.
+Previous studies have shown that there is evidence of economic inequality between Kiwualn burials.
+However, the degree of inequality over time, and which variable decides the changes remain unclear.
 
-• Hypothesis 1: There are significant reciprocal structure effects in burial network before the contact
-• Hypothesis 2: There are significant transitivity structure effects in burial network after the contact
-• Hypothesis 3: There are significant popularity structure effects in in burial network after the contact
+We hypothesize that after the arrival of the European presence in the 17th century, a group of ambitious individuals competed with each other for trading with the Europeans due to the high value of trade goods in their society, which led to the accumulation of high quality trade goods.
+Social inequality gradually increased among Kiwulan society, which can be observed from the burials from a more equal distribution of grave goods to a more differentiation in goods accumulation.
+We hypothesize that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence.
 
-The computational processes in this paper were performed by the ergm package for simulation and model specification and then bergm package using R programming [@Hunter2008; @Morris2008; @Caimo2014].
+We performed network analysis with Bayesian inference with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling using R programming [@Caimo2014].
 We set the number of chains to eight.

### 2020-06-09 analysis/paper/paper.Rmd
edit archaeology background

@@
 In other words, actors are viewed as "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes [@Coward2013; @Mizoguchi2013; @Potgieter2008].
+
+<!--archaeology case-->
+The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact in the 17th century[@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
+In addition to prestige, European colonial powers could have provided military support to indigenous allies.
+For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people.
+A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
+In this study, we will use this model of competition to explain the changes in social inequality at Kiwulan, an Iron Age site in northesatern Taiwan, which covers the presence of the Europeans, inculding the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
 
@@
 The burials are mostly located in the AD section, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
-Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
+Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized space arrangement.
 
@@
 She found that the burials with rich goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
-The use of prestige goods in burials might be traced back before the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006).
-For example, the Spanish document by Spanish Priest documented that the Spanish soldier used agate beads to exchange materials from local indigenous people since they were valuable in their culture[Citation].
-Also, there was a custom that people would bury imported materials with deceased to display their status and influence [Ciation].
-Despite we observed the economic differentiation in burial goods and the value of imported goods according to historical documents, the extent of economic differentiation and how it changed over time remain unclear.
-Analyzing all burial data using network analysis approaches and compare the networks before and after the foreign presences will give an insight into the degree and changes in social inequality.
-This is important to understand the social reactions of Kiwulan residents in the context of foreign contact.
+The earliest record of the use of prestige goods in burials can be traced back around the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006].
+For example, the Spanish document by a Spanish Priest described that the Spanish soldier used agate beads to exchange materials from local indigenous people since they were valuable in their culture [@LiandWu2006].
+In addition, there was a custom that people would bury imported materials with deceased to display their status and influence [@LiandWu2006].
+Despite we observed the economic differentiation in burial goods and the high value of imported goods according to historical documents, the extent of economic differentiation and how it changed over time remain unclear.
+Analyzing and comparing burial data from different phases using network analysis approaches will give an insight into the degree and changes in social inequality.
+This is important to understand the social reactions of Kiwulan residents in the context of the foreign contact.
 
@@
 Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data.
-
-<!--unanswered issues-->
-The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
-In addition to prestige, European colonial powers could have provided military support to indigenous allies.
-For example, Voss’s (2005) study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people.
-A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
-In this study, We will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.
-
-Based on the quantity of grave goods among burials, we can reconstruct social networks from the distribution of prestige goods, including imported ceramic and beads based on the principle that prestige materials usually relate to social status.
-We build network where burials represent actors (nodes in the network) linked by sharing the same prestige goods.
-
-We assume burial network represents the relations among people and social structure based on the argument that
+With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar variables.
+This is based on the principle that prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations].
+Therefore, we build network where burials represent actors (nodes in the network) that are linked when they have th same prestige goods in common.
+The prestige goods we identified including agate beads, gold-foil beads, and imported Chinese porcelains, according to previous burials analysis [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
 

### 2020-06-16 analysis/paper/paper.Rmd
work on method and reorganize paragraphs

@@
 Predictor variables, called network statistics, such as reciprocity, transitivity, homophily, are direct functions of the ties themselves where the probability of occurrence of the configuration of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+The advantage of networks specification under assumptions and contexts can be useful to explain archaeological network in a given time.
+For example, @Amati2019 demonstrate the use of ERGMs to reconstruct networks consist of a set of sites in the Caribbean during the period AD 100–400.
+They found the structure of network can be efficiently explained by multiple interdependent mechanisms instead of hypotheses of distance and cultural homophily exclusively.
+However, they also point out some limitations of application such as static outcome and sensitivity to missing data.
+Another limitation is that ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
 With the development of Markov chain Monte Carlo (MCMC) algorithms that produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, ERGMs can deal with complicated dependence patterns observed at one moment in time [@Snijders2006].
-The advantage of networks specification under assumptions and contexts can be useful to explain archaeological network in a given time.
-@Amati2019 demonstrate the use of ERGMs to reconstruct networks consist of a set of sites in the Caribbean during the period AD 100–400.
-They found out the structure of network can be efficiently explained by multiple interdependent mechanisms instead of hypotheses of distance and cultural homophily exclusively.
-However, they also point out some limitations of application such as static outcome and sensitive to missing data.
-The two major limitations ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
 A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
@@
 In addition, there was a custom that people would bury imported materials with deceased to display their status and influence [@LiandWu2006].
-Despite we observed the economic differentiation in burial goods and the high value of imported goods according to historical documents, the extent of economic differentiation and how it changed over time remain unclear.
+Despite we observed the economic differentiation in burial goods and the high value of imported goods according to historical documents, the degree of inequality over time, and which variable decides the changes remain unclear.
 Analyzing and comparing burial data from different phases using network analysis approaches will give an insight into the degree and changes in social inequality.
@@
 
-<!--procedures-->
+## data and hypoethesis
 
-## data
+We analyze archaeological burial networks from a Iron Age site, Kiwulan, in northeastern Taiwan.
+We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy.
+Given that burials in Chinese phase are less than previous two phases, we only compare the quantity and types of burial goods in burials in Chinese phase rather than network analysis.
+For the quantity of burial goods, we assigned the burials into three classes, high, medium, and low as an index of wealth as a variable to indicate the extent of the accumulation of wealth.
+In addition to wealth, we also consider another two social variables, age and gender, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
 
-We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy.
-Given that burials in Chinese phase are less than previous two phases, we only compare these burials by their grave goods and consitions without further network analysis.
-Burial data provides information about the location, orientation, size, depth, age, gender, quantity of grave goods, and spatial relationship among them.
-We consider the effects of both exogenous and endogenous factors for meaningful explanation of the characteristics of burial network and tendency of burials correlate with each other.
-To be specific, our purpose is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods.
-We focus on connectivity, clusters, centralization, and homophily those properties of network that can reflect the social relations between individuals guided by social structures.
-Previous studies have shown that there is evidence of economic inequality between Kiwualn burials.
-However, the degree of inequality over time, and which variable decides the changes remain unclear.
-
+The purpose of our analysis is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods.
 We hypothesize that after the arrival of the European presence in the 17th century, a group of ambitious individuals competed with each other for trading with the Europeans due to the high value of trade goods in their society, which led to the accumulation of high quality trade goods.
 Social inequality gradually increased among Kiwulan society, which can be observed from the burials from a more equal distribution of grave goods to a more differentiation in goods accumulation.
-We hypothesize that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence.
+Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence compared to the network before the European presence.
 
-We performed network analysis with Bayesian inference with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling using R programming [@Caimo2014].
-We set the number of chains to eight.
+## Model specification and estimation
+<!--Network construction-->
+
+We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.
+ERGMs assume that each network tie is a random variable which is dependent to each other somehow through the shared attributes or a reciprocal relationship [@Robins2007].
+The dependence of ties can be hypothesized by parameters correspond to configurations in the network, such as a single tie, or a transitive triad.
+A model is built up by a distribution of random graphs according to these configurations that are also called network statistics.
+According to the formation of ties, network statistics can be divided into two ways, endogenous statistics and exogenous statistics [@Amati2019].
+In endogenous model, the dependence of ties is decided by the presence of other ties, while in exogenous model, the dependence of ties is determined by node attributes.
+We use both exogenous and endogenous statistics to understand the characteristics of burial network and tendency of burials correlate with each other.
+To describe the relationship linked by shared burial goods better, the network is weighted to capture the multiple effects [@Snijders2006].
+The network variables we examine include connectivity, clusters, centralization, and homophily in a undirected relations to reflect the social relations between individuals guided by social structures.
+Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations.
+Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
+
+```{r ergm-terms-table}
+ergm_terms <- readxl::read_excel(here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
+kableExtra::kable(ergm_terms,
+             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
+```
+
+We conduct network analysis using R programming with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling [@Caimo2014].
+We set the number of chains to six.
 For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
@@
 This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
-
-Since the accumulation of imported prestige goods in burials is commonly observed, we assign the burials into four classes, high, medium, and none as an index of wealth.
-In addition to wealth, we also consider another two node variables, age and gender, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
-
-## Model specification and estimation
-<!--Network construction-->
-
-it can accomodate binary network statistics to describe the weighted network topology [@Snijders2006]
-
-some concepts:
-- endogenous statistics that ties are modelled based on the existence of other ties
-- exogenous statistics that ties are formed according to monadic or dyadic node attributes
-
-We used ergm package in R for the model specification.
-Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
-The terms we selected for network specification include:
-
-1. edges: ties, a measure of density, equal to kstar(1) for undirected networks
-
-2.
-gwesp (geometrically weighted edgewise shared partners): triad-based clustering based on a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node.
-It is a measure of the tendency for two actors who are tied to have x partners in common.
-In other words, a tendency for those with shared partners to become tied, or tendency of ties to cluster together.
-
-3.
-gwnsp (geometrically weighted non-edgewise shared partner statistics): a function of the non-edgewise shared partner statistics.
-The edgewise and non-edgewise shared partner statistics reflect tendencies toward transitivity abd compared together would give a more comprehensive result.
-
-4.
-gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once.
-The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts.
-A measure of the tendency of the number of edges into a node.
-This statistic captures the tendency towards centralisation in the degree distribution of the network.
-
-5.
-nodematch: uniform homophily between nodes that refers to the similarity of connected nodes.
-A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics.
-We use wealth index and age as covariants that reflects the similarity between nodes.
-Here we consider age, gender, and index of wealth as node attributes covariants.
 

### 2020-06-16 R/paper_note.R

@@ -1,0 +1,9 @@
+1. edges: ties, a measure of density, equal to kstar(1) for undirected networks
+
+2. gwesp (geometrically weighted edgewise shared partners): triad-based clustering based on a dyad-based configurations (not node-based) that counts the number of times that both nodes have a tie to a third node. It is a measure of the tendency for two actors who are tied to have x partners in common. In other words, a tendency for those with shared partners to become tied, or tendency of ties to cluster together.
+
+3. gwnsp (geometrically weighted non-edgewise shared partner statistics): a function of the non-edgewise shared partner statistics. The edgewise and non-edgewise shared partner statistics reflect tendencies toward transitivity and compared together would give a more comprehensive result.
+
+4. gwdegree (geometrically weighted degree distribution): a parametric form that represents the frequency distribution for nodal degree, each node counts only once. The local configuration is degrees and stars that are equivalent representations for the distribution of node-based edge counts. A measure of the tendency of the number of edges into a node. This statistic captures the tendency towards centralisation in the degree distribution of the network.
+
+5. nodematch: uniform homophily between nodes that refers to the similarity of connected nodes. A statistics that counts the number of edges between nodes sharing the same covariate value or characteristics. We use wealth index and age as covariants that reflects the similarity between nodes. Here we consider age, gender, and index of wealth as node attributes covariants.

### 2020-06-16 analysis/paper/paper.Rmd
some edits and feedback on methods section

@@
 ## Bayesian modeling
+
 <!--Bayesian inference for Exponential Random Graph Models--> 
@@
 
-We analyze archaeological burial networks from a Iron Age site, Kiwulan, in northeastern Taiwan.
-We use burial data collected from the published excavation report, and original fieldwork note with detailed descriptions about context and stratigraphy.
-Given that burials in Chinese phase are less than previous two phases, we only compare the quantity and types of burial goods in burials in Chinese phase rather than network analysis.
-For the quantity of burial goods, we assigned the burials into three classes, high, medium, and low as an index of wealth as a variable to indicate the extent of the accumulation of wealth.
-In addition to wealth, we also consider another two social variables, age and gender, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
+We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan to generate networks.
+We used burial data collected from the published excavation report, and original fieldwork notes with detailed descriptions about context and stratigraphy.
+We do not include the burials in Chinese phase in our network analysis because of the small sample size.
+<!-- but we will briefly discuss these in the discussion section --> We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
+In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
 
-The purpose of our analysis is to understand the underlying mechanisms of the network structures and how their relationships have an influence on the accumulation of wealth between different foreign contact periods.
-We hypothesize that after the arrival of the European presence in the 17th century, a group of ambitious individuals competed with each other for trading with the Europeans due to the high value of trade goods in their society, which led to the accumulation of high quality trade goods.
-Social inequality gradually increased among Kiwulan society, which can be observed from the burials from a more equal distribution of grave goods to a more differentiation in goods accumulation.
-Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects in burial network after the European presence compared to the network before the European presence.
+The purpose of our analysis is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
+We hypothesize that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
+This led to the accumulation of high quality trade goods among a small numbers of individuals.
+Social inequality gradually increased in Kiwulan, and we predict this can be observed from the burials.
+We hypothesize a change from a more equal distribution of grave goods to a more differentiated distribution of goods.
+Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects <!-- the reader needs to be introduced to these terms before you can use them like this...
+instead, need to foreground or foreshadow these concepts...
+"We used network analysis statistics to quantify the relationships and formally test our hypothesis.
+"--> in burial network after the European presence, compared to the network before European presence.
 
@@
 
-We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.
-ERGMs assume that each network tie is a random variable which is dependent to each other somehow through the shared attributes or a reciprocal relationship [@Robins2007].
-The dependence of ties can be hypothesized by parameters correspond to configurations in the network, such as a single tie, or a transitive triad.
+We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.<!-- seems like a long sentence --> ERGMs assume that each network tie <!-- does the reader know what this is?--> is a random variable which is dependent on each other somehow through the shared attributes or a reciprocal relationship [@Robins2007].
+The dependence of ties can be hypothesized by parameters corresponding to configurations in the network, such as a single tie, or a transitive triad <!-- does the reader know what is 'transitive?
+--> .
 A model is built up by a distribution of random graphs according to these configurations that are also called network statistics.
-According to the formation of ties, network statistics can be divided into two ways, endogenous statistics and exogenous statistics [@Amati2019].
-In endogenous model, the dependence of ties is decided by the presence of other ties, while in exogenous model, the dependence of ties is determined by node attributes.
-We use both exogenous and endogenous statistics to understand the characteristics of burial network and tendency of burials correlate with each other.
-To describe the relationship linked by shared burial goods better, the network is weighted to capture the multiple effects [@Snijders2006].
-The network variables we examine include connectivity, clusters, centralization, and homophily in a undirected relations to reflect the social relations between individuals guided by social structures.
-Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations.
+According to the formation of ties, network statistics can be divided into two types, endogenous statistics and exogenous statistics [@Amati2019].
+In endogenous models <!--statistic or model?
+what exactly?
+these seem like two different things that are getting mixed up.
+Choose one way to describe and stick to it to avoid confusing the reader -->, the dependence of ties is decided by the presence of other ties, while in an exogenous model, the dependence of ties is determined by node attributes <!-- node?
+-->.
+We use both exogenous and endogenous statistics to understand the characteristics of the burial network at Kiwulan and the correlate of burials.
+To describe the relationship linked by shared burial goods, the network is weighted to capture the multiple effects <!-- of what?
+the reader doesn't know about any effects yet --> [@Snijders2006].
+The network variables we examine include connectivity, clusters, centralization, and homophily in undirected relations <!-- something wrong here, 'relations' seems wrong here --> to reflect the social relations <!-- avoid repeating the same work in a sentence --> between individuals guided by social structures.
+Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations <!-- put this earlier in the paragraph --> .
 Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
@@
 
-We conduct network analysis using R programming with the ergm package for simulation and model specification [@Hunter2008; @Morris2008], and then bergm package for Bayesian modeling [@Caimo2014].
-We set the number of chains to six.
+We conducted a network analysis using the R programming language with the ergm package for simulation and model specification [@Hunter2008; @Morris2008].
+We used the bergm package for Bayesian modeling [@Caimo2014].
+<!-- why did we use two packages?
+why not just one?
+this needs to be expanded a bit to make it clear to the reader --> We set the number of chains to six.
 For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
-We set the number of iterations used to simulate a network y′ at each iteration to 10000.
-The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with a set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
+We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
+The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
 This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
@@
 <!-- Goodness-of-fit assessment-->
+
+<!-- "We test our hypothesis by measuring reciprocality,  transitivity, and  popularity structure effects and expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects" This is where you can connect the anthropological question, the archaeological hypothesis and the statistical method for testing the hypothesis  -->
 

### 2020-06-21 analysis/paper/paper.Rmd
edit method section

@@
 
-We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan to generate networks.
-We used burial data collected from the published excavation report, and original fieldwork notes with detailed descriptions about context and stratigraphy.
+We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks.
+The data includes information of gender, age, burial rites, and types of burial goods.
 We do not include the burials in Chinese phase in our network analysis because of the small sample size.
-<!-- but we will briefly discuss these in the discussion section --> We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
-In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
+A network is a structure consists of a set of nodes (actors) and edges (ties) connecting them representing relationships [@Brughmans2018]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
+This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
+252].
+Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
+In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts.
 
-The purpose of our analysis is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
-We hypothesize that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
-This led to the accumulation of high quality trade goods among a small numbers of individuals.
-Social inequality gradually increased in Kiwulan, and we predict this can be observed from the burials.
-We hypothesize a change from a more equal distribution of grave goods to a more differentiated distribution of goods.
-Thus, we predict that the burial network will present less reciprocality, less transitivity, but higher popularity structure effects <!-- the reader needs to be introduced to these terms before you can use them like this...
-instead, need to foreground or foreshadow these concepts...
-"We used network analysis statistics to quantify the relationships and formally test our hypothesis.
-"--> in burial network after the European presence, compared to the network before European presence.
+We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
+The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
+In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals to built their personal status and power.
+This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with social differentiation.
+If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by the social processes.
+This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.
 
@@
 
-We use Bayesian reference on Exponential Random Graph Models (ERGMs) to conduct statistical network analysis to test what variables may have contributed to the economic inequality observed between burials over time under foreign contact conditions.<!-- seems like a long sentence --> ERGMs assume that each network tie <!-- does the reader know what this is?--> is a random variable which is dependent on each other somehow through the shared attributes or a reciprocal relationship [@Robins2007].
-The dependence of ties can be hypothesized by parameters corresponding to configurations in the network, such as a single tie, or a transitive triad <!-- does the reader know what is 'transitive?
---> .
-A model is built up by a distribution of random graphs according to these configurations that are also called network statistics.
-According to the formation of ties, network statistics can be divided into two types, endogenous statistics and exogenous statistics [@Amati2019].
-In endogenous models <!--statistic or model?
-what exactly?
-these seem like two different things that are getting mixed up.
-Choose one way to describe and stick to it to avoid confusing the reader -->, the dependence of ties is decided by the presence of other ties, while in an exogenous model, the dependence of ties is determined by node attributes <!-- node?
--->.
-We use both exogenous and endogenous statistics to understand the characteristics of the burial network at Kiwulan and the correlate of burials.
-To describe the relationship linked by shared burial goods, the network is weighted to capture the multiple effects <!-- of what?
-the reader doesn't know about any effects yet --> [@Snijders2006].
-The network variables we examine include connectivity, clusters, centralization, and homophily in undirected relations <!-- something wrong here, 'relations' seems wrong here --> to reflect the social relations <!-- avoid repeating the same work in a sentence --> between individuals guided by social structures.
-Table \@ref(tab:ergm-terms-table) lists the network statistics we used in the model construction for burials with archaeological interpretations <!-- put this earlier in the paragraph --> .
+We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis.
+ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+A number of parameters used for modeling the dependencies expresses the structure or property of a network, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006].
+With a research question for a certain social process, we can build a hypothesis-based model to generate a distribution of random networks according to the parameters we specified, also called configurations or network statistics.
+This helps to understand the underlying processes of an observed network by comparing it with a large set of possible networks generated by the model.
+
+Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations.
 Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
+Based on our hypothesis, the burial network after the European presence will present less reciprocity, less transitivity, but higher popularity structure effects compared to the network before the European presence.
+Since burials tend to have multiple types of burial goods in common, such as personal ornaments, imported porcelains, and gold foils, the network ties are weighted instead of a binary data expressed by 0 and 1 [@Snijders2006].
+We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
+In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
 
@@
 
-The plots from left to right show estimated marginal posterior densities, trace plots, and autocorrelation plots.
+The plots from left to right show estimated marginal posterior densities, trace plots, and auto correlation plots.
 
@@
 # Discussion
+
+<!-- briefly discuss Chinese burials -->
 

### 2020-06-22 analysis/paper/paper.Rmd
edit method section and add references

@@
 We do not include the burials in Chinese phase in our network analysis because of the small sample size.
-A network is a structure consists of a set of nodes (actors) and edges (ties) connecting them representing relationships [@Brughmans2018]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
+A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
 This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
 252].
+Since burials tend to have multiple prestige goods in common, such as personal ornaments and imported porcelains, the network ties are weighted instead of binary variables, where the value 1 represents a tie between two nodes and the value 0 otherwise [@Snijders2011].
+For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
+Our networks are non-directed, which means ties have no orientation forming the relationship between actors.
+
 Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
 In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts.
-
 We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
@@
 
-## Model specification and estimation
+## Model specification 
 <!--Network construction-->
@@
 ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-A number of parameters used for modeling the dependencies expresses the structure or property of a network, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006].
-With a research question for a certain social process, we can build a hypothesis-based model to generate a distribution of random networks according to the parameters we specified, also called configurations or network statistics.
-This helps to understand the underlying processes of an observed network by comparing it with a large set of possible networks generated by the model.
+In ERGMs, ties form a small structure in a network called graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations are the structure or property of a network and can be expressed as network statistics.
+By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks.
+This distribution consists of a large number of possible networks that provides a statistical inference to an observed network [@Robins2007].
+This helps to understand whether our observed network shows significantly more or less property of interest than the large set of random networks we modeled.
 
 Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations.
-Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008].
-Based on our hypothesis, the burial network after the European presence will present less reciprocity, less transitivity, but higher popularity structure effects compared to the network before the European presence.
-Since burials tend to have multiple types of burial goods in common, such as personal ornaments, imported porcelains, and gold foils, the network ties are weighted instead of a binary data expressed by 0 and 1 [@Snijders2006].
+Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
+We test our hypothesis by measuring reciprocity, transitivity, popularity, and homophily structure effects.
+We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence.
 We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
@@
 
-We conducted a network analysis using the R programming language with the ergm package for simulation and model specification [@Hunter2008; @Morris2008].
-We used the bergm package for Bayesian modeling [@Caimo2014].
-<!-- why did we use two packages?
-why not just one?
-this needs to be expanded a bit to make it clear to the reader --> We set the number of chains to six.
+We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014].
+Bergm uses ergm packages for simulation and model specification, and provides a framework for Bayesian parameter estimation on the result of exponential random graph models [@Hunter2008; @Morris2008; @Handcock2008].
+We set the number of chains to six.
 For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
@@
 <!-- Goodness-of-fit assessment-->
-
-<!-- "We test our hypothesis by measuring reciprocality,  transitivity, and  popularity structure effects and expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects" This is where you can connect the anthropological question, the archaeological hypothesis and the statistical method for testing the hypothesis  -->
 

### 2020-06-24 analysis/paper/paper.Rmd
expand and edit method section

@@
 <!--chronology-->
-We assigned burials to the pre-European period, European and post-European period for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
+We assigned burials to the pre-European period (n = 26), European and post-European period (n = 45) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
 Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
@@
 Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data.
-With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar variables.
-This is based on the principle that prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations].
+With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar prestige goods.
+This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
+252].
+Also, prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations].
 Therefore, we build network where burials represent actors (nodes in the network) that are linked when they have th same prestige goods in common.
-The prestige goods we identified including agate beads, gold-foil beads, and imported Chinese porcelains, according to previous burials analysis [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
+The prestige goods we identified including agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments, according to previous research [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
 
@@
 
-We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks.
-The data includes information of gender, age, burial rites, and types of burial goods.
+We analyzed archaeological burials from an Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks.
 We do not include the burials in Chinese phase in our network analysis because of the small sample size.
-A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
-This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
-252].
-Since burials tend to have multiple prestige goods in common, such as personal ornaments and imported porcelains, the network ties are weighted instead of binary variables, where the value 1 represents a tie between two nodes and the value 0 otherwise [@Snijders2011].
+A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007].
+In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
+Due to gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels based on the distribution of their counts to indicate quantity, high, medium, and low.
+That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them.
+For prestige goods not commonly found across burials, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess the good in common.
+Node attributes here include osteological data and mortuary practices, such as age, gender, burial ritual, and wealth level.
+Burial ritual is identified by locally made ceramics on the top of grave that suggests the presence of funeral feasting.
+Wealth level is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
+Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
@@
 
-Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
-In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts.
 We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
 The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
-In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals to built their personal status and power.
-This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with social differentiation.
-If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by the social processes.
+In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power.
+This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with more social inequality.
+If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by those social processes.
+Network analysis provides a way to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
+Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches.
 This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.
@@
 ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In ERGMs, ties form a small structure in a network called graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations are the structure or property of a network and can be expressed as network statistics.
+In ERGMs, ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
 By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks.
-This distribution consists of a large number of possible networks that provides a statistical inference to an observed network [@Robins2007].
-This helps to understand whether our observed network shows significantly more or less property of interest than the large set of random networks we modeled.
+Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007].
+This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.
 
-Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations.
+Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations.
 Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
@@
 We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence.
-We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
-In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].
+This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality.
+We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
+This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any.
 
@@
 We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014].
-Bergm uses ergm packages for simulation and model specification, and provides a framework for Bayesian parameter estimation on the result of exponential random graph models [@Hunter2008; @Morris2008; @Handcock2008].
-We set the number of chains to six.
+Bergm provides a framework for Bayesian parameter estimation on the result of exponential random graph models, which depends on ergm packages for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
+Despite exponential random graph models can better describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
+A normalizing constant is a step to make probability distribution to integrate to one, which becomes hard to compute for larger set of networks [@Caimo2017bayesian].
+Another problem is the degeneracy that means the probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
+To overcome these limitations, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the exchange algorithm, which solves computational problems in ERGMs [@Caimo2011].
+
+MCMC simulation can estimate the posterior distribution of interest by of random sampling without the need of normal distribution for the prior and likelihood.
+MCMC is a random probability distribution that describes a sequence of events or transitions (chain) from one state to another state following probabilistic rules.
+In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within that chosen set [@Caimo2011].
+In our case, we set the number of chains to six.
 For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
@@
 <!-- briefly discuss Chinese burials -->
+Age and gender attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].
 

### 2020-07-03 analysis/paper/paper.Rmd
edit method section

@@
 
+<!--Bayesian approaches-->
+
 We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014].
@@
 To overcome these limitations, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the exchange algorithm, which solves computational problems in ERGMs [@Caimo2011].
+
+<!--Model fitting and comparison-->
 
@@
 This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
-
-<!--Model fitting and comparison to alternative approaches-->
+Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
+Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with higher transitivity.
 
@@
 
-Normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) is viewed as a suitable prior model for the model parameters of interests in network analysis if there is no previous observation for a network[@Caimo2017].
-Since our original data presents a pattern of dense ties, we specified the prior based on our observation that assumes a high density and high transitivity network.
-
 ```{r Bayesian-ERGMs}

### 2020-07-17 analysis/paper/paper.Rmd
update code and edit the method section

@@
 # combine some periods
-burial_three_period_tidy <-
+burial_three_period_age_tidy <-
   burial %>%
@@
     total >= 100 ~ "high",
-    TRUE ~ "other")) %>% # the classification is based on the result of histogram
+    TRUE ~ "")) %>% # the classification is based on the result of histogram
   mutate(Age_scale = case_when(
@@
     TRUE ~ "NA")) %>%
+  mutate(ritual = case_when(
+    `Stamped_ceramic` == "2" ~ "two pots",
+    `Stamped_ceramic` == "1" ~ "one pot",
+    TRUE ~ "NA")) %>%
   mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
@@
          `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
-         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% 
-  # the low-med-high index is determined by interpretation for histograms for those burial good types
+         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% #based on the result of histogram
   select(burial_label,
@@
          gender,
+         ritual, # consider to remove if not very informative
          Gold_bead_low,
@@
          Agate_bead_high,
-         #Agate_bead, #female burials
-         #Golden_bead,
-         Porcelain,
+         Porcelain, #prestige good
          Gold_leaf, #prestige good
          fish_shape_knit, #prestige good
-         #Bell, #children's burials
          quantity)
-         #total) # select specific variable to drop columns (uninformative variables)
 
@@
   count(Phase)
+
+pre_E <- burial_three_period_number$Phase[4]
+post_E <- burial_three_period_number$Phase[3]
+
 ```
@@
 <!--chronology-->
-We assigned burials to the pre-European period (n = 26), European and post-European period (n = 45) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
+We assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
 Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
@@
 
-We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
+Our hypothesis is that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
 The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
@@
 
-MCMC simulation can estimate the posterior distribution of interest by of random sampling without the need of normal distribution for the prior and likelihood.
-MCMC is a random probability distribution that describes a sequence of events or transitions (chain) from one state to another state following probabilistic rules.
+Markov chain Monte Carlo (MCMC) is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
+We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms.
+This enables a random or stochastic simulation in a long run without depending on the previous move.
+More chains ensures a more desirable distribution close to the target distribution that we seek, which is called the convergence of the Markov chains.
 In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within that chosen set [@Caimo2011].
@@
 The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
-This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
+This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
 Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
-Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with higher transitivity.
+Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
 

### 2020-07-17 analysis/paper/paper.Rmd
update code and edit the method section

@@
          Agate_bead_high,
+         #Agate_bead, #female burials
+         #Golden_bead,
          Porcelain, #prestige good
@@
          fish_shape_knit, #prestige good
+         #Bell, #children's burials
          quantity)
@@
 # number of each phase
-burial_three_period_number <-
-  burial_three_period_tidy %>%
+burial_three_period_age_number <-
+  burial_three_period_age_tidy %>%
   count(Phase)
 
-pre_E <- burial_three_period_number$Phase[4]
-post_E <- burial_three_period_number$Phase[3]
+pre_E <- burial_three_period_age_number$Phase[4]
+post_E <- burial_three_period_age_number$Phase[3]
 
@@
 burial_pre <-
-  burial_three_period_tidy %>%
+  burial_three_period_age_tidy %>%
   filter(Phase == "pre") %>%
@@
 nodes_pre <-
-  burial_three_period_tidy %>%
+  burial_three_period_age_tidy %>%
   filter(Phase == "pre") %>%
@@
 edge_list_pre <-
-  burial_three_period_tidy %>%
-  select(burial_label, 5:13) %>% # need to change for each exploration
+  burial_three_period_age_tidy %>%
+  select(burial_label, 6:14) %>% # need to change for each exploration
   pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
@@
          rowwise() %>%
-       mutate(common_counts = sum(count, count1)))
+         mutate(common_counts = sum(count...3, count...6)))
 
@@
 burial_post <-
-  burial_three_period_tidy %>%
+  burial_three_period_age_tidy %>%
   filter(Phase == "post") %>%
-  janitor::remove_empty(which = "cols")
+  janitor::remove_empty(which = "cols") %>%
+  filter(quantity != "none") # remove burial without burial goods
 
@@
 nodes_post <-
-  burial_three_period_tidy %>%
+  burial_three_period_age_tidy %>%
   filter(Phase == "post") %>%
+  filter(quantity != "none") %>%
   select(burial_label) %>%
@@
 edge_list_post <-
-  burial_three_period_tidy %>%
-  select(burial_label, 5:13) %>% # need to change for each exploration
+  burial_three_period_age_tidy %>%
+  select(burial_label, 6:14) %>% # need to change for each exploration
   pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
@@
          rowwise() %>%
-         mutate(common_counts = sum(count, count1)))
+         mutate(common_counts = sum(count...3, count...6)))
 
@@
 Bergm provides a framework for Bayesian parameter estimation on the result of exponential random graph models, which depends on ergm packages for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
-Despite exponential random graph models can better describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
+Although exponential random graph models can better describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
 A normalizing constant is a step to make probability distribution to integrate to one, which becomes hard to compute for larger set of networks [@Caimo2017bayesian].
@@
 This enables a random or stochastic simulation in a long run without depending on the previous move.
-More chains ensures a more desirable distribution close to the target distribution that we seek, which is called the convergence of the Markov chains.
-In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within that chosen set [@Caimo2011].
+More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains.
+In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011].
 In our case, we set the number of chains to six.
@@
 The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle.
-Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods is limited to burials with rich goods.
+Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods only for burials with rich goods.
 

### 2020-07-17 analysis/paper/002-burials-pre-network-modelling.R

@@ -4,15 +4,15 @@
 burial_pre <-
   burial_three_period_age_tidy %>%
   filter(Phase == "pre") %>%
-  janitor::remove_empty(which = "cols") %>%
-  filter(quantity != "none") # remove burial without burial goods
+  janitor::remove_empty(which = "cols")
+  #filter(quantity != "none") # remove burial without burial goods
 
 # create node list using burial index by phase
 #--------------------pre--------------------------------
 nodes_pre <-
   burial_three_period_age_tidy %>%
   filter(Phase == "pre") %>%
-  filter(quantity != "none") %>% # remove burial without burial goods
+  #filter(quantity != "none") %>% # remove burial without burial goods
   select(burial_label) %>%
   rowid_to_column("id")
 

### 2020-08-20 analysis/paper/paper.Rmd

@@
          Agate_bead_high,
-         #Agate_bead, #female burials
-         #Golden_bead,
          Porcelain, #prestige good
@@
          fish_shape_knit, #prestige good
-         #Bell, #children's burials
          quantity)
@@
 post_E <- burial_three_period_age_number$Phase[3]
-
 ```

### 2020-08-25 analysis/paper/paper.Rmd
minor edit for intro and add a reference

@@
 <!--burial study- network analysis-->
-Burials provide valuable information about social structures that reflect not only the personal life and status of the deceased, but also the expressions of their descendants, from which we are able to detect the social differences that can relate to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
+Burials provide valuable information about social structures that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods [@Pearson1999].
+This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
 Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting [@Coward2013, pp.
@@
 
-Using the R language, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan.
+Using the R software, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan.
 We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group.
@@
 <!--archaeological setting-->
-We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007].
+We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (1350-1850 AD) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007].
 The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
-The burials are mostly located in the AD section, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
+The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
 Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized space arrangement.

### 2020-09-06 analysis/paper/paper.Rmd
tidy code and add summary for gwd and esp

@@
 A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
-In this study, we will use this model of competition to explain the changes in social inequality at Kiwulan, an Iron Age site in northesatern Taiwan, which covers the presence of the Europeans, inculding the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
+In this study, we will use this model of competition to explain the changes in social inequality at Kiwulan, an Iron Age site in northesatern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
 
@@
 model.ergm <- burial_network_pre ~
-  edges + # ties, a measure of density, equal to kstar(1) for undirected networks
+  edges + 
   density +
-  triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network, can be replaced by "gwdegree"
+  triangle 
 summary(model.ergm)
 
-# model 2, check here Morris et al. (2008) 
-# check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
-model.2 <- burial_network_pre ~ edges + # density
+# model 2 considers cluster and degree
+model.2 <- burial_network_pre ~ 
+  edges + # density
   gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle)
-  # number means weight parameter alpha, which controls the rate of declining marginal returns
-  # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
-  # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
-  # ergm can estimate the parameter from the data by using fixed=FALSE
-
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
-  # tendency of being in contact with multiple partners, measures of centralisation
-  # distribution of node-based edge counts, each node counts only once
-  # number means weight parameter decay
-  # The closer decay is to zero, the more gwdegree considers low degree nodes relative to high degree nodes
 summary(model.2)

### 2020-09-06 analysis/paper/002-burials-pre-network-modelling.R

@@ -150,10 +150,24 @@ model.ergm <- burial_network_pre ~
   triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
 summary(model.ergm)
 
-# model 2 considers cluster and degree
+# Edgewise Shared Partners
+summary(burial_network_pre ~ esp(0:10))
+summary(burial_network_pre ~ gwdegree(0:10))
+
+# model 2 considers cluster and degree, Morris et al. (2008)
+# check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
 model.2 <- burial_network_pre ~ edges + # density
-  gwesp(0.75, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
+  gwesp(0.75, fixed = TRUE) +
+  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
+  # number means weight parameter alpha, which controls the rate of declining marginal returns
+  # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
+  # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
+  # ergm can estimate the parameter from the data by using fixed=FALSE
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
+  # tendency of being in contact with multiple partners, measures of centralisation
+  # distribution of node-based edge counts, each node counts only once
+  # number means weight parameter decay
+  # The closer decay is to zero, the more gwdegree considers low degree nodes relative to high degree nodes
 summary(model.2)
 
 # model 3 considers cluster, degree, and node attributes

### 2020-09-20 analysis/paper/paper.Rmd
minor updates for method
edit the main text

@@
 
-Using the R software, we use Bayesian SNA to study burials from Kiwulan, an Iron Age site in northeast Taiwan.
+Using the R software, we apply Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan.
 We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group.
@@
 
+## Bayesian modelling
+
 ```{r ergm-terms-table}
@@
 We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
+Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
 The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
@@
 Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
-Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
 
@@
 Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods only for burials with rich goods.
+The post-European network also presented subgroups with more edges connecting burials.
 

### 2020-09-20 analysis/paper/002-burials-pre-network-modelling.R

@@ -133,7 +133,7 @@ plot(burial_network_pre,
      pad = 1) #protects the labels from getting clipped
 
 legend("topleft",
-       col = c(2, 3, 1), # need to adjust each time
+       col = c(2, 3, 1, 4), # need to adjust each time
        pch    = 20,
        legend = unique(quantity),# quantity
        title  = 'Burial good counts')
@@ -176,7 +176,6 @@ model.3 <- burial_network_pre ~ edges +  # the overall density of the network
   nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
   nodematch('age') +
   nodematch('gender') +
-  nodematch('ritual') +
   absdiff('total') +
   gwesp(0.75, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(0.75, fixed = TRUE) + #prior = -1
@@ -185,9 +184,9 @@ summary(model.3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-1, 0, 0, 0, 0, 0, 3, -1, 0) # positive prior number for edge means high density
+prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(5, 9, 9) # covariance matrix structure
+prior.sigma <- diag(5, 8, 8) # covariance matrix structure
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
 

### 2020-09-28 analysis/paper/paper.Rmd
add sentences to method section

@@
 
-We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014].
-Bergm provides a framework for Bayesian parameter estimation on the result of exponential random graph models, which depends on ergm packages for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
-Although exponential random graph models can better describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
-A normalizing constant is a step to make probability distribution to integrate to one, which becomes hard to compute for larger set of networks [@Caimo2017bayesian].
-Another problem is the degeneracy that means the probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-To overcome these limitations, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the exchange algorithm, which solves computational problems in ERGMs [@Caimo2011].
+We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm for model specification [@Hunter2008; @Morris2008; @Handcock2008].
+Although exponential random graph models can describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
+A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
+Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
+To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
+This alleviate above computational problems.
+By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
+Posterior probability enables us to estimate the effect of ERGM parameter by observing the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
+The criteria is that the odds ratio grater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
 
@@
 
-Markov chain Monte Carlo (MCMC) is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
+Markov chain Monte Carlo (MCMC) algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
 We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms.

### 2020-09-29 analysis/paper/paper.Rmd
edit ERGMs section

@@
 
+# Exponential random graph models
+
 <!--current approaches to network analysis-->
@@
 In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measure separately by the presence of a tie without considering interactions among ties [@Morris2008].
-By contrast, statistical models assumes social networks as dependent variables that can represent the network dependencies between tie variables [@Snijders2011].
+By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011].
 In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes/ties.
-Predictor variables, called network statistics, such as reciprocity, transitivity, homophily, are direct functions of the ties themselves where the probability of occurrence of the configuration of ties can be specified and hypothesized [@Morris2008; @Robins2007].
-The advantage of networks specification under assumptions and contexts can be useful to explain archaeological network in a given time.
-For example, @Amati2019 demonstrate the use of ERGMs to reconstruct networks consist of a set of sites in the Caribbean during the period AD 100–400.
-They found the structure of network can be efficiently explained by multiple interdependent mechanisms instead of hypotheses of distance and cultural homophily exclusively.
-However, they also point out some limitations of application such as static outcome and sensitivity to missing data.
-Another limitation is that ERGMs are difficult to compute because their normalizing constant depending on model parameters is intractable.
-With the development of Markov chain Monte Carlo (MCMC) algorithms that produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, ERGMs can deal with complicated dependence patterns observed at one moment in time [@Snijders2006].
-A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) strategies which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
+One can model network dependencies according to a common experience or research questions to represent relationships between actors through network variables.
+For example, reciprocity is a variable describing two actors' relationship, and transitivity describes two actors' relationship via a shared third actor.
+Those predictor variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time.
+For instance, @Amati2019 demonstrated the use of ERGMs to reconstruct three different network for 15 sites in the Caribbean between AD 100 to 400 to explore the interaction mechanisms.
+Those models test the hypotheses for the relations between sites in terms of proximity, inter-cultural, and exchange network to explain the presence of hub sites.
+They found that the observed presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of the hypothesis of proximity or cultural homophily exclusively.
+However, they also point out some limitations of application, such as static outcome and sensitivity to missing data.
+In addition, it could be difficult to work on ERGMs because their normalizing constant depending on model parameters is intractable.
+A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
+MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns [@Snijders2006].
 
@@
 
-We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm for model specification [@Hunter2008; @Morris2008; @Handcock2008].
+We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
 Although exponential random graph models can describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].

### 2020-09-29 analysis/paper/paper.Rmd
making burial map

@@
 However, they also point out some limitations of application, such as static outcome and sensitivity to missing data.
-In addition, it could be difficult to work on ERGMs because their normalizing constant depending on model parameters is intractable.
+In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits ERGMs applications.
 A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
-MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns [@Snijders2006].
+MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
 

### 2020-09-29 analysis/paper/000-burial-location.R

@@ -1,0 +1,29 @@
+library(sf)
+library(ggsn)
+# read in the spatial data
+burial_shape <- invisible(st_read(here("analysis", "data", "raw_data", "shapefiles", "AD_burial.shp"), quiet = TRUE))
+post <- invisible(st_read(here("analysis", "data", "raw_data", "shapefiles", "AD_postholes.shp"), quiet = TRUE))
+AD <- invisible(st_read(here("analysis", "data", "raw_data", "shapefiles", "AD_zone.shp"), quiet = TRUE))
+AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "shapefiles", "AD_withdata.shp"), quiet = TRUE))
+location <- invisible(st_read(here("analysis", "data", "raw_data", "shapefiles", "location.shp"), quiet = TRUE))
+
+# join and find centroid point of burial
+burial_three_period_age_tidy_geo <-
+  burial_shape %>%
+  left_join(burial_three_period_age_tidy,
+            by = c("No_burial" = "burial_label")) %>%
+  mutate(burial_cent = st_centroid(geometry))
+
+# mapping burials over the AD section
+KWL_burial_map <-
+  burial_three_period_age_tidy_geo %>%
+  filter(!is.na(geometry)) %>%
+  ggplot() +
+  geom_sf(aes(fill = Phase.y)) +
+  geom_sf(data = post, color = 'blue', size = 0.2, alpha = 0.4) +
+  geom_sf(data = AD_data, fill = NA) +
+  #geom_sf_text(data = AD_data, aes(label = `坑號`), size = 2.5,
+               #alpha = 0.3, fontface = "bold") +
+  theme_minimal() +
+  blank() +
+  north(AD_data, scale = 0.2, symbol = 3)

### 2020-10-03 analysis/paper/paper.Rmd
edit Intro

@@
 <!--burial study- network analysis-->
-Burials provide valuable information about social structures that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods [@Pearson1999].
+Burials provide valuable information about past human behavior that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods and arrangements [@Pearson1999].
+Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures.
 This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
-Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow us to infer social structures, for examples, grave goods usually relate to social practices of trade, exchange, and gifting [@Coward2013, pp.
+For examples, grave goods can reflect consumption patterns and social practices in a broader of cultural contexts involves personal identity or social relation through trade, exchange, and gifting [@Coward2013, pp.
 252].
-One way to study the relationships between individual is social network analysis (SNA), which can visualize, quantify, and statistically test hypotheses about the relationships to understand network structure and its effects on every part within the network [@Freeman2004; @Salvini2010].
+The correlation between burials could be viewed as an indicator of social relationship in a society.
+One way to study the relationships between individual is social network analysis (SNA).
 A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
-Their relationships are flows of resources that reflect relations of control, dependence, and cooperation.
-In other words, actors are viewed as "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes [@Coward2013; @Mizoguchi2013; @Potgieter2008].
+SNA can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the structure of the network [@Freeman2004; @Salvini2010].
+To be specific, actors are "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes [@Coward2013; @Mizoguchi2013; @Potgieter2008].
+Relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relation in a given society.
 

### 2020-10-06 analysis/paper/paper.Rmd
Get first three moments of GOF distributions

@@
 Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation with the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
-This alleviate above computational problems.
+To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
+This alleviate above computational problems and gives a better convergence results.
 By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
-Posterior probability enables us to estimate the effect of ERGM parameter by observing the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
-The criteria is that the odds ratio grater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
+Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
+The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
 

### 2020-10-08 analysis/paper/paper.Rmd
tried other parameters and changed object names

@@
 This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
-For examples, grave goods can reflect consumption patterns and social practices in a broader of cultural contexts involves personal identity or social relation through trade, exchange, and gifting [@Coward2013, pp.
+For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom.
+Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp.
 252].
-The correlation between burials could be viewed as an indicator of social relationship in a society.
-One way to study the relationships between individual is social network analysis (SNA).
+Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization.
+One way to study relationships between individuals is social network analysis (SNA).
 A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
@@
 <!--archaeology case-->
-The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact in the 17th century[@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
+The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
 In addition to prestige, European colonial powers could have provided military support to indigenous allies.

### 2020-10-08 analysis/paper/003-burials-pre-network-modelling.R

@@ -144,12 +144,12 @@ legend("topleft",
 
 #------------------creating ERGM model-------------------------------------
 # model 1 considers density and triad relations (for cluster)
-model.ergm <- burial_network_pre ~
+model_pre_1 <- burial_network_pre ~
   edges + # ties, a measure of density, equal to kstar(1) for undirected networks
   density +
   gwdegree(0.5, fixed = TRUE)  +
   triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
-summary(model.ergm)
+summary(model_pre_1)
 
 # Edgewise Shared Partners
 summary(burial_network_pre ~ esp(0:10))
@@ -157,7 +157,7 @@ summary(burial_network_pre ~ gwdegree(0:10))
 
 # model 2 considers cluster and degree, Morris et al. (2008)
 # check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
-model.2 <- burial_network_pre ~ edges + # density
+model_pre_2 <- burial_network_pre ~ edges + # density
   gwesp(0.75, fixed = TRUE) +
   # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
   # number means weight parameter alpha, which controls the rate of declining marginal returns
@@ -169,30 +169,31 @@ model.2 <- burial_network_pre ~ edges + # density
   # distribution of node-based edge counts, each node counts only once
   # number means weight parameter decay
   # The closer decay is to zero, the more gwdegree considers low degree nodes relative to high degree nodes
-summary(model.2)
+summary(model_pre_2)
 
 # model 3 considers cluster, degree, and node attributes
-model.3 <- burial_network_pre ~ edges +  # the overall density of the network
+model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the network
   nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
   nodematch('age') +
   nodematch('gender') +
+  nodematch('ritual') +
   absdiff('total') +
   gwesp(0.75, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(0.75, fixed = TRUE) + #prior = -1
   gwdegree(0.8, fixed = TRUE)
-summary(model.3)
+summary(model_pre_3)
 
[…]

### 2020-10-08 analysis/paper/004-burials-post-network-modelling.R

@@ -113,6 +113,7 @@ set.vertex.attribute(burial_network_post, "quantity", burial_post$quantity)
 set.vertex.attribute(burial_network_post, "age", burial_post$Age_scale)
 set.vertex.attribute(burial_network_post, "gender", burial_post$gender)
 set.vertex.attribute(burial_network_post, "ritual", burial_post$ritual)
+set.vertex.attribute(burial_network_post, "total", burial_post$total)
 
 # plot
 set.seed(30)
@@ -154,14 +155,15 @@ model.3 <- burial_network_post ~ edges +  # the overall density of the network
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
+  absdiff('total') +
   gwesp(1.8, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(1.8, fixed = TRUE) +
   gwdegree(0.8, fixed = TRUE)   # popularity
 summary(model.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(5, 8, 8) # covariance matrix structure
+prior.mean <- c(-1, 0, 0, 0, 0, 0, 3, -1, 0) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(5, 9, 9) # covariance matrix structure
 
 parpost <- bergm(model.3,
                  prior.mean  = prior.mean,
@@ -177,8 +179,56 @@ summary(parpost)
 plot(parpost)
 
 # Model assessment, Bayesian goodness of fit diagnostics:
-bgof(parpost,
-     aux.iters = 10000,
-     n.deg     = 20,
-     n.dist    = 15,
-     n.esp     = 25)
+bgof_post <-
+  bgof(parpost,
+       aux.iters = 10000,
+       n.deg     = 20,
+       n.dist    = 15,
+       n.esp     = 25)
+
+summary(bgof_post)
+
+# calculate moments for observed and simulated GOF distribution
+obs_degree_post <- describe(bgof_post$obs.degree * 0:(length(bgof_post$obs.degree)-1))
+obs_dist_post <- describe(bgof_post$obs.dist * 1:length(bgof_post$obs.dist))
+obs_esp_post <- describe(bgof_post$obs.esp * 0:(length(bgof_post$obs.esp)-1))
+
[…]

### 2020-10-08 analysis/paper/paper.Rmd
edited method section and added references

@@
 Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization.
-One way to study relationships between individuals is social network analysis (SNA).
-A social network contains a set of actors and a set of relationships connecting these actors, which could be groups, organizations or persons.
-SNA can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the structure of the network [@Freeman2004; @Salvini2010].
-To be specific, actors are "nodes" in a wider network, in which nodes are connected by multiple relationships, or "edges", to other nodes [@Coward2013; @Mizoguchi2013; @Potgieter2008].
-Relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relation in a given society.
+A common way to study connections between individuals is social network analysis (SNA), which has been applied increasingly to archaeological studies [@Mills2017].
+A network contains a set of actors and a set of relationships connecting these actors, which could be persons, groups, or organizations.
+In other words, actors are "nodes", in which nodes are connected by multiple relationships, or "edges", to other nodes.
+The basic unit for SNA analysis is dyad relationship, which is defined as a connection between two distinct actors [@Lee2016].
+Relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relation in a given society [@Marin2011].
+Using SNA, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the structure of the network [@Freeman2004; @Salvini2010].
+The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]
 
 <!--archaeology case-->
-The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
+The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and accumulation of them in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
+Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups.
 In addition to prestige, European colonial powers could have provided military support to indigenous allies.
@@
 Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007].
-In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measure separately by the presence of a tie without considering interactions among ties [@Morris2008].
+In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measured separately by the presence of a tie without consideration of interactions among ties [@Morris2008].
 By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011].
-In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes/ties.
-One can model network dependencies according to a common experience or research questions to represent relationships between actors through network variables.
-For example, reciprocity is a variable describing two actors' relationship, and transitivity describes two actors' relationship via a shared third actor.
-Those predictor variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or the attributes of nodes/ties, called covariates.
+One can model network variables, such as dependencies, nodal attributes, and dyadic attributes, to generate edges between nodes based on a common experience for observed relationships.
+For example, common network dependencies include reciprocity that describes two actors' relationship, and transitivity that describes two actors' relationship via a shared third actor.
+Those variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
 The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time.

### 2020-10-08 analysis/paper/003-burials-pre-network-modelling.R

@@ -177,7 +177,7 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
-  absdiff('total') +
+  #absdiff('total') +
   gwesp(0.75, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(0.75, fixed = TRUE) + #prior = -1
   gwdegree(0.8, fixed = TRUE)
@@ -185,9 +185,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-1, 0, 0, 0, 0, 0, 3, -1, 0) # positive prior number for edge means high density
+prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(5, 9, 9) # covariance matrix structure
+prior.sigma <- diag(5, 8, 8) # covariance matrix structure
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
 
@@ -200,7 +200,7 @@ pre_bergm <- bergmM(model_pre_3,
                   main.iters  = 2000, # iterations for every chain of the population
                   aux.iters   = 10000, # MCMC steps used for network simulation
                   nchains     = 6, # number of chains of the population MCMC
-                  gamma       = 0.5) # scalar; parallel adaptive direction sampling move factor, acceptance rate
+                  gamma       = 0.2) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
 summary(pre_bergm)
 

### 2020-10-08 analysis/paper/004-burials-post-network-modelling.R

@@ -155,15 +155,15 @@ model.3 <- burial_network_post ~ edges +  # the overall density of the network
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
-  absdiff('total') +
+  #absdiff('total') +
   gwesp(1.8, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(1.8, fixed = TRUE) +
   gwdegree(0.8, fixed = TRUE)   # popularity
 summary(model.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-1, 0, 0, 0, 0, 0, 3, -1, 0) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(5, 9, 9) # covariance matrix structure
+prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(5, 8, 8) # covariance matrix structure
 
 parpost <- bergm(model.3,
                  prior.mean  = prior.mean,
@@ -172,7 +172,7 @@ parpost <- bergm(model.3,
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
                  nchains     = 6, # number of chains of the population MCMC
-                 gamma       = 0.5) # scalar; parallel adaptive direction sampling move factor, acceptance rate
+                 gamma       = 0.2) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
 summary(parpost)
 

### 2020-10-10 analysis/paper/paper.Rmd
edit and fix the node number for pre-E

@@
 Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization.
-A common way to study connections between individuals is social network analysis (SNA), which has been applied increasingly to archaeological studies [@Mills2017].
-A network contains a set of actors and a set of relationships connecting these actors, which could be persons, groups, or organizations.
-In other words, actors are "nodes", in which nodes are connected by multiple relationships, or "edges", to other nodes.
-The basic unit for SNA analysis is dyad relationship, which is defined as a connection between two distinct actors [@Lee2016].
-Relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relation in a given society [@Marin2011].
-Using SNA, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the structure of the network [@Freeman2004; @Salvini2010].
+A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017].
+Networks are represented by graphs that consists of a set of actors, such as people, groups, or organizations, and a set of links connecting them.
+Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
+The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016].
+In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011].
+With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010].
 The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]
@@
 <!--current approaches to network analysis-->
-Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007].
-In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measured separately by the presence of a tie without consideration of interactions among ties [@Morris2008].
-By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011].
-In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or the attributes of nodes/ties, called covariates.
-One can model network variables, such as dependencies, nodal attributes, and dyadic attributes, to generate edges between nodes based on a common experience for observed relationships.
-For example, common network dependencies include reciprocity that describes two actors' relationship, and transitivity that describes two actors' relationship via a shared third actor.
-Those variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
-The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time.
-For instance, @Amati2019 demonstrated the use of ERGMs to reconstruct three different network for 15 sites in the Caribbean between AD 100 to 400 to explore the interaction mechanisms.
-Those models test the hypotheses for the relations between sites in terms of proximity, inter-cultural, and exchange network to explain the presence of hub sites.
-They found that the observed presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of the hypothesis of proximity or cultural homophily exclusively.
-However, they also point out some limitations of application, such as static outcome and sensitivity to missing data.
-In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits ERGMs applications.
-A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
+Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures by directly modeling the formation of edges, or ties, between nodes [@Robins2007].
+The assumption is that an observed network is randomly generated depending on the presence or absence of other edges.
+ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008].
+In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011].
+For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby.
+Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common.
+We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+Network specification under assumptions and contexts is useful for exploring archaeological network.
+For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs.
+Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed.
+They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively.
+However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data.
+In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs.
+Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
 MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].

### 2020-10-10 analysis/paper/002-burial-location.R

@@ -27,7 +27,8 @@ df_pre_distance <-
   arrange(Id)
 
 pre_distance <-
-  st_distance(df_pre_distance$burial_cent)
+  st_distance(df_pre_distance$burial_cent) %>%
+  filter()
 
 # mapping burials over the AD section
 KWL_burial_map <-

### 2020-10-10 analysis/paper/003-burials-pre-network-modelling.R

@@ -66,7 +66,7 @@ edges_pre <-
 edges_for_network_pre <-
   select(edges_pre, from, to, common_counts) %>%
   filter (!common_counts == 0) # remove rows with no goods in common
-# %>% mutate(common_counts = ifelse(common_counts > 1, 1, common_counts)) # for unweighted network
+  #mutate(common_counts = ifelse(common_counts > 1, 1, common_counts)) # for unweighted network
 
 #1--------------------network analysis using ggraph pkg------------------------------
 library(tidygraph)
@@ -96,6 +96,7 @@ ggraph(relation_tidy_pre, layout = "graphopt") +
 #2-------------------network analysis using network pkg-------------------------------
 library(network)
 
+attr(edges_for_network_pre, "n") = 29
 burial_network_pre <-
   network(edges_for_network_pre, # the network object
           vertex.attr = nodes_pre, # node list
@@ -105,9 +106,9 @@ burial_network_pre <-
           matrix.type = "edgelist") # the type of input
 
 # plot
-plot(burial_network_pre, vertex.cex = 1) # It seems the last two nodes are missing?? only 27 nodes
+plot(burial_network_pre, vertex.cex = 1) # exclude the last two nodes since no edges, 27 left
 
-#-----------------------Bayesian ERGMs ------------------------------
+#-----------------------attach attributes and make graphs------------------------------
 library(statnet)
 library(Bergm)
 
@@ -117,6 +118,15 @@ set.vertex.attribute(burial_network_pre, "age", burial_pre$Age_scale)
 set.vertex.attribute(burial_network_pre, "gender", burial_pre$gender)
 set.vertex.attribute(burial_network_pre, "ritual", burial_pre$ritual)
 set.vertex.attribute(burial_network_pre, "total", burial_pre$total)
+
+#get distance matrix, need to run 002 code first
+pre_distance_n <- network(pre_distance, directed = F)
+set.edge.attribute(pre_distance_n, "dist", pre_distance_n)
+
+test_1 <- burial_network_pre ~
+  edges +
+  edgecov(pre_distance_n, "dist")
+summary(test_1)
 
 # plot
 set.seed(30)

### 2020-10-10 analysis/paper/004-burials-post-network-modelling.R

@@ -168,7 +168,7 @@ prior.sigma <- diag(5, 8, 8) # covariance matrix structure
 parpost <- bergm(model.3,
                  prior.mean  = prior.mean,
                  prior.sigma = prior.sigma,
-                 burn.in     = 200, # burn-in iterations for every chain of the populationm, drops the first 200
+                 burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
                  nchains     = 6, # number of chains of the population MCMC

### 2020-10-14 analysis/paper/paper.Rmd
edit Intro

@@
 <!--burial study- network analysis-->
-Burials provide valuable information about past human behavior that can reflect not only the personal life and status of the deceased, but also the expressions of their descendants through grave goods and arrangements [@Pearson1999].
+Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole.
+With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013].
 Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures.
@@
 Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups.
-In addition to prestige, European colonial powers could have provided military support to indigenous allies.
-For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people.
+In addition to prestige, European colonial powers could have provided military support to Indigenous allies.
+For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people.
 A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
@@
 2.
-Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an indigenous society can be detected by burials networks.
+Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks.
 3.
@@
 the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period.
-This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on indigenous groups.
+This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
 A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.
@@
 The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
-Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized space arrangement.
+Those burials are oriented in an east-west direction on the north side of the house structures indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement.
 
@@
 The earliest record of the use of prestige goods in burials can be traced back around the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006].
-For example, the Spanish document by a Spanish Priest described that the Spanish soldier used agate beads to exchange materials from local indigenous people since they were valuable in their culture [@LiandWu2006].
+For example, the Spanish document by a Spanish Priest described that the Spanish soldier used agate beads to exchange materials from local Indigenous people since they were valuable in their culture [@LiandWu2006].
 In addition, there was a custom that people would bury imported materials with deceased to display their status and influence [@LiandWu2006].

### 2020-10-15 analysis/paper/paper.Rmd
edit method section and work on egodata

@@
 A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017].
-Networks are represented by graphs that consists of a set of actors, such as people, groups, or organizations, and a set of links connecting them.
+Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007].
 Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
@@
 <!--current approaches to network analysis-->
-Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures by directly modeling the formation of edges, or ties, between nodes [@Robins2007].
+Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007].
 The assumption is that an observed network is randomly generated depending on the presence or absence of other edges.
@@
 
-## data and hypoethesis
+## data and hypothesis
 
-We analyzed archaeological burials from an Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks.
-We do not include the burials in Chinese phase in our network analysis because of the small sample size.
-A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007].
+We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan.
+We do not include the burials of the Chinese phase in our network analysis due to small sample size.
 In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
-Due to gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels based on the distribution of their counts to indicate quantity, high, medium, and low.
+Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
 That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them.
-For prestige goods not commonly found across burials, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess the good in common.
-Node attributes here include osteological data and mortuary practices, such as age, gender, burial ritual, and wealth level.
-Burial ritual is identified by locally made ceramics on the top of grave that suggests the presence of funeral feasting.
-Wealth level is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods.
+For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common.
+Node attributes here include osteological data and mortuary practices, such as age, gender, presence of burial rituals, and overall burial goods quantity index.
+Burial ritual is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records.
+Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
-Our networks are non-directed, which means ties have no orientation forming the relationship between actors.
+Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
 

### 2020-10-16 analysis/paper/paper.Rmd
reorganized paras

@@
 For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people.
-A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals.
-In this study, we will use this model of competition to explain the changes in social inequality at Kiwulan, an Iron Age site in northesatern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
+A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inquqality due to competition among individuals.
+In this study, we use this pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
+Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival.
+This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.
 
-# Exponential random graph models
-
-<!--current approaches to network analysis-->
-Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007].
-The assumption is that an observed network is randomly generated depending on the presence or absence of other edges.
-ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008].
-In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011].
-For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby.
-Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common.
-We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
-Network specification under assumptions and contexts is useful for exploring archaeological network.
-For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs.
-Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed.
-They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively.
-However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data.
-In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs.
-Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011].
-MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
-
-## Bayesian modeling
-
-<!--Bayesian inference for Exponential Random Graph Models--> 
-Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
-This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
-For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
-- weighted data
-
-Using the R software, we apply Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan.
+We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan.
 We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group.
@@
 
-# Archaeological background
+# Exponential random graph models in a Bayseisan framework
+
+<!--current approaches to network analysis-->
+Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007].
+The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008].
+In ERGMs, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011].
+For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
+We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
+By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks.
+Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007].
+This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.
+
+ERGMs can describe a network by capturing the dependency structure by specifying parameters under assumptions and contexts, which is useful for exploring archaeological network.
+For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs.
+Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed.
+They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively.
+However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data.
+In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
+A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
+Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
+Those complex computational processes in ERGMs can be also resolved by implementing ERGMs in a Bayesian framework.
+
+<!--Bayesian inference for Exponential Random Graph Models--> 
+Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
+This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
+For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced with Bayesian modeling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
+To overcome these limitations in ERGMs, Bayesian modeling applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
+MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
+This alleviate above computational problems and gives a better convergence results.
+By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
+Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
+The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
+
+# Kiwulan in northeastern Taiwan
 
@@
 The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
-Those burials are oriented in an east-west direction on the north side of the house structures indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement.
+Those burials are oriented in an east-west direction on the north side of the residential area indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement.
+Previous studies point out a presence of uneven distribution of prestige goods across burials without an agreement that whether this uneven distribution hints vertical social differences.
+@Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society.
+However, @Hsieh2012 suggested a relatively egalitarian based on the comparative analysis of frequency and proportion of all burial goods.
+She found that the burials with high valued burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
 
-Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials.
-However, there is a debate over whether the differentiation between burials hints an economic inequality supporting an egalitarian society [@Hsieh2012], or political inequality indicating a stratified society [@Cheng2008].
-@Cheng2008 interprets the unequal distribution of prestige materials, especially the gold-foil beads between burials, as evidence for hierarchy.
-However, @Hsieh2012 suggests a relatively egalitarian society based on the comparative analysis of frequency and proportion of all burial goods.
-She found that the burials with rich goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
-The earliest record of the use of prestige goods in burials can be traced back around the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006].
-For example, the Spanish document by a Spanish Priest described that the Spanish soldier used agate beads to exchange materials from local Indigenous people since they were valuable in their culture [@LiandWu2006].
-In addition, there was a custom that people would bury imported materials with deceased to display their status and influence [@LiandWu2006].
-Despite we observed the economic differentiation in burial goods and the high value of imported goods according to historical documents, the degree of inequality over time, and which variable decides the changes remain unclear.
-Analyzing and comparing burial data from different phases using network analysis approaches will give an insight into the degree and changes in social inequality.
-This is important to understand the social reactions of Kiwulan residents in the context of the foreign contact.
+The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006].
+For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006].
+In addition, there was a custom that people would bury imported beads and ceramics with deceased to display their status and influence [@LiandWu2006].
+Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
+The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well identified.
+That is, the observed uneven distribution may be explained as result of social change from less social differences to more social differences when Indigenous societies involved in the complex trade network stimulated by the Europeans.
+We assume that the changes would be supported by chronological differences in the structure of burial networks.
+Analyzing and comparing networks between different phases relative to cultural contact will give an insight into the degree and changes in social differences.
+This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
 
@@
 We do not include the burials of the Chinese phase in our network analysis due to small sample size.
-In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common.
+We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here.
+In our case, burials are nodes that are linked when they have goods in common.
 Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
@@
 For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common.
-Node attributes here include osteological data and mortuary practices, such as age, gender, presence of burial rituals, and overall burial goods quantity index.
-Burial ritual is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records.
+Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index.
+Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records.
 Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods.
@@
 
-Our hypothesis is that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society.
+Our pericolonial impact hypothesis is that restricted grave goods or accumulation of goods present in a few individuals after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access high value goods.
 The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
-In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power.
-This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with more social inequality.
-If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by those social processes.
-Network analysis provides a way to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships.
-Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches.
-This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.
+In addition to the interest of burial goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power.
+If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
+Network analysis provides a way to examine the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.
 
-## Model specification 
-<!--Network construction-->
-
-We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis.
-ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In ERGMs, ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks.
-Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007].
-This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.
-
-Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations.
-Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
-We test our hypothesis by measuring reciprocity, transitivity, popularity, and homophily structure effects.
-We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence.
-This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality.
-We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
-This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any.
-
-## Bayesian modelling
-
+## Model specification in a Bayesian framework
 ```{r ergm-terms-table}
@@
 
-<!--Bayesian approaches-->
+<!--Network construction and specification-->
+We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis of social change using the R programming language with bergm package [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
+Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations.
+Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
+Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects.
+We modeled several node attributes for homophily effects, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
+We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried close to each other.
+Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality.
+We expect that observed network after the European arrival will have a better match with our model compared with the network before European contact.
 
-We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
-Although exponential random graph models can describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
-A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
-Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
-This alleviate above computational problems and gives a better convergence results.
-By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
-Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
-The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
+<!---Based on our hypothesis, we model a network using parameters including reciprocity, transitivity, popularity, and homophily structure effects.
+We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence.
+This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality.
+We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
+This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any.
+--->
 
-<!--Model fitting and comparison-->
-
-Markov chain Monte Carlo (MCMC) algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
+<!--Model fitting and prior setting--> 
+After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm.
+MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
 We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms.
@@
 We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
-Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
+
+<!-- Goodness-of-fit assessment-->
 The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
@@
 
-<!-- Goodness-of-fit assessment-->
+<!--Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
+The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
+This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
+Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017] -->
 

### 2020-10-20 analysis/paper/paper.Rmd
adjusting settings for different prior

@@
 With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013].
-Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals and allow archaeologists to infer social structures.
+Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures.
 This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
@@
 
-We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan.
+We analyzed burial data from published excavation reports, and original fieldwork notes of Kiwulan.
 We do not include the burials of the Chinese phase in our network analysis due to small sample size.
@@
 
-<!---Based on our hypothesis, we model a network using parameters including reciprocity, transitivity, popularity, and homophily structure effects.
-We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence.
-This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality.
-We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
-This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any.
---->
-
 <!--Model fitting and prior setting--> 
@@
 Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
-
-<!--Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density.
-The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
-This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
-Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017] -->
 
@@
 θ3 = gwesp, a positive sign suggests a tendency of burials in our sample to group together in closed transitive structures.
-In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organised in closed structures.
+In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organized in closed structures.
 Burials with multiple prestige goods in common are more likely to be directed connected.

### 2020-10-20 analysis/paper/003-burials-pre-network-modelling.R

@@ -188,16 +188,16 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   nodematch('ritual') +
   #absdiff('total') +
   gwesp(0.75, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
-  gwnsp(0.75, fixed = TRUE) + #prior = -1
+  gwnsp(0.75, fixed = TRUE) + #0.75, #prior = -1
   gwdegree(0.8, fixed = TRUE) +
   edgecov(pre_distance_n, "dist")
 summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0, -1) # positive prior number for edge means high density
+prior.mean <- c(-1, 1, -1, 0, 0, 3, -1, 1, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(5, 9, 9) # covariance matrix structure
+prior.sigma <- diag(3, 9, 9) # covariance matrix structure
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
 
@@ -248,7 +248,7 @@ sim_esp_pre <-
                                        /length(bgof_pre$obs.esp)))
 
 # make dataframes for distribution stats
-distribution_degree <-
+distribution_degree_pre <-
   data.frame("moments" = c("Mean", "Variance", "Skewness"),
              "Observed" = round(c(obs_degree_pre$mean,
                                   obs_degree_pre$sd,
@@ -256,7 +256,7 @@ distribution_degree <-
              "Model" = round(c(mean(sim_degree_pre$mean),
                                mean(sim_degree_pre$sd),
                                mean(sim_degree_pre$skew)),2))
-distribution_dist <-
+distribution_dist_pre <-
   data.frame("moments" = c("Mean", "Variance", "Skewness"),
              "Observed" = round(c(obs_dist_pre$mean,
                                   obs_dist_pre$sd,
@@ -264,7 +264,7 @@ distribution_dist <-
              "Model" = round(c(mean(sim_dist_pre$mean),
[…]

### 2020-10-20 analysis/paper/004-burials-post-network-modelling.R

@@ -172,8 +172,8 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-1, 0, 0, 0, 0, 3, -1, 0, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(5, 9, 9) # covariance matrix structure
+prior.mean <- c(-1, 1, -1, 0, 0, 3, -1, 1, -1) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(3, 9, 9) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,
@@ -192,9 +192,9 @@ plot(post_bergm)
 bgof_post <-
   bgof(post_bergm,
        aux.iters = 10000,
-       n.deg     = 20,
+       n.deg     = 40,
        n.dist    = 15,
-       n.esp     = 25)
+       n.esp     = 35)
 
 summary(bgof_post)
 

### 2020-10-22 analysis/paper/paper.Rmd
edit code and method section

@@
 
-# Kiwulan in northeastern Taiwan
+# Burial data from Kiwulan, NE Taiwan
 
@@
 post_E <- burial_three_period_age_number$Phase[3]
+chi <- burial_three_period_age_number$Phase[1]
 ```
@@
 Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
-The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well identified.
-That is, the observed uneven distribution may be explained as result of social change from less social differences to more social differences when Indigenous societies involved in the complex trade network stimulated by the Europeans.
+The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.
+That is, the observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans.
 We assume that the changes would be supported by chronological differences in the structure of burial networks.
-Analyzing and comparing networks between different phases relative to cultural contact will give an insight into the degree and changes in social differences.
+Analyzing and comparing networks between different phases relative to cultural contact will give an insight into changes in social differences.
 This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
@@
 <!--chronology-->
-We assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
+To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
+We excluded the burials from the Chinese phase (n = `r chi`) for analysis due to much smaller sample size.
 Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
-The chronology is then evaluated by the burials with time marker, light gray glazed jars, also called An-ping jars.
-The jars were believed to be produced in Southeast China and circulated in Southeast Asia from the late 16th to 18th century, but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
+The chronology is then evaluated by the burials with light gray glazed jars, also called An-ping jars that were believed to be produced in Southeast China.
+They circulated in Southeast Asia from the late 16th to 18th century, and often found at sites with the European occupation in the early 17th century in many Aisan places, including Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
+
 In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007].
@@
 
-We analyzed burial data from published excavation reports, and original fieldwork notes of Kiwulan.
-We do not include the burials of the Chinese phase in our network analysis due to small sample size.
+We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
 We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here.
-In our case, burials are nodes that are linked when they have goods in common.
-Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
+We treat the presence of relations between burials based on number or types of  prestige items from regional trade network.
+where a few prestige goods controlled by a small number of individuals.
+This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals.
+Based on this assumption, we treat the presence of relations between burials through the burial goods in comoon.
+Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classifed those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
 That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them.
-For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common.
+For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common.
 Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index.

### 2020-10-22 analysis/paper/003-burials-pre-network-modelling.R

@@ -198,6 +198,7 @@ summary(model_pre_3)
 prior.mean <- c(-1, 1, -1, 0, 0, 3, -1, 1, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
 prior.sigma <- diag(3, 9, 9) # covariance matrix structure
+
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
 

### 2020-10-22 analysis/paper/004-burials-post-network-modelling.R

@@ -165,8 +165,8 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
   nodematch('gender') +
   nodematch('ritual') +
   #absdiff('total') +
-  gwesp(1.8, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
-  gwnsp(1.8, fixed = TRUE) +
+  gwesp(0.75, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
+  gwnsp(0.75, fixed = TRUE) + # original 1.8
   gwdegree(0.8, fixed = TRUE) +
   edgecov(post_distance_n, "dist")
 summary(model.post.3)

### 2020-10-22 analysis/paper/005-bergm-distribution-stats.R

@@ -22,7 +22,7 @@ edges_value <-
 
 trans_value <-
   ggplot(bergm_two_phases,
-       aes(x = gwesp.fixed.1.8, fill = phase)) +
+       aes(x = gwesp.fixed.0.75, fill = phase)) +
   geom_histogram(position = "identity") +
   theme_minimal()
 

### 2020-10-23 analysis/paper/paper.Rmd
add code for burial value, run model, and edit Intro

@@
 <!--burial study- network analysis-->
-Burials contains valuable information about individual mortuary practices and inherited structure of a society as a whole.
-With the development of network science and graph theory, it is promising to study relation among burials and explore its social implication [@Sosna2013].
+Burials contains valuable information of both individual mortuary practices and the structure of a society inherited in burial relations.
+With the development of network science and graph theory, studying relations among burials and exploring its social implication become promising [@Sosna2013].
 Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures.
@@
 For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom.
-Burials with grave goods in common often indicates underlying social relation connected through trade, exchange, and gifting [@Coward2013, pp.
+Burials with grave goods in common often indicate some underlying social relations connected through trade, exchange, and gifting [@Coward2013, pp.
 252].
-Identifying the structure of relationships among burials allows a further inferences on a structure of a social organization.
-A common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies recently [@Mills2017].
+Identifying the structure of relationships among burials can give an insight to the structure of a social organization.
+One common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies in recent years [@Mills2017].
 Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007].
@@
 The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016].
-In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, reflecting social relations in societies [@Marin2011].
-With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affects the structure [@Freeman2004; @Salvini2010].
-The application in archaeology is widely used for [@Coward2013; @Mizoguchi2013; @Potgieter2008]
+In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, indicating social relations in societies [@Marin2011].
+With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
+The application in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]
 
 <!--archaeology case-->
-The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and accumulation of them in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society.
-Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups.
-In addition to prestige, European colonial powers could have provided military support to Indigenous allies.
-For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people.
-A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inquqality due to competition among individuals.
-In this study, we use this pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
+The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in those societies.
+The prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups.
+The concept of prestige goods may develop earlier before the European arrival, but could be amplified during the European colonial period.
+For example, a person with more imported goods may be treated as higher status.
+A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals.
+In this study, we use a pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
 Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival.
@@
 We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan.
-We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group.
-We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions.
+We viewed individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same burial goods that can represent the same social group with similar economic or political status.
+We construct a network that represents higher social inequality and tested hypotheses that burial relations becomes more hierarchical after the European arrival.
 We ask the questions: 1.
-Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities?
+Do the observed burial data after the European presence shows more similar network quality as our hypothesized model?
 2.
@@
 3.
-How the relations among burials may affect the quantity of burial goods?
-We expect the network to present a higher degree of clustering (e.g.
-the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period.
+What is the major variables affecting or forming the higher degree of popularity (e.g.
+a few node has more relationship with other nodes) in the European contact period that might hint at social heterogeneity during this period.
 This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
@@
 This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals.
-Based on this assumption, we treat the presence of relations between burials through the burial goods in comoon.
-Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classifed those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
+Based on this assumption, we treat the presence of relations between burials through the burial goods in common.
+Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classified those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
 That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them.

### 2020-10-23 analysis/paper/000-value-burials.R

@@ -4,15 +4,15 @@ library(here)
 
 burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
 
-# run 000-value-burial first before the following code
 # burial data with combined age and three phases
-burial_three_period_age_tidy <-
+burial_values <-
   burial %>%
   rename(burial_label = ID) %>%
   mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
   filter(!is.na(Phase)) %>%
   mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
          Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
+         Glass_bead = ifelse(Glass_bead == "shatter", "1", Glass_bead),
          Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
   mutate_at(21:ncol(.), as.numeric) %>%
   janitor::remove_empty(which = "cols") %>%
@@ -21,6 +21,8 @@ burial_three_period_age_tidy <-
   mutate(Porcelain = ifelse(Porcelain == 0, NA, Porcelain)) %>%
   mutate(Stonewares = rowSums(.[c(50, 55)], na.rm = TRUE)) %>% #stoneware, Anping jars
   mutate(Stonewares = ifelse(Stonewares == 0, NA, Stonewares)) %>%
+  mutate(Metal_bangles = rowSums(.[c(26, 28, 29)], na.rm = TRUE)) %>% #different bangles
+  mutate(Metal_bangles = ifelse(Metal_bangles == 0, NA, Porcelain)) %>%
   mutate(quantity = case_when(
     total == 0 ~ "none",
     total > 0 & total <= 7 ~ "low",
@@ -36,43 +38,41 @@ burial_three_period_age_tidy <-
     `Gender` %in% c("1","2") ~ "male",
     `Gender` %in% c("3","4") ~ "female",
     TRUE ~ "NA")) %>%
-  mutate(ritual = case_when(
-    `Stamped_ceramic` == "2" ~ "two pots",
-    `Stamped_ceramic` == "1" ~ "one pot",
-    TRUE ~ "NA")) %>%
-  mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
-         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
[…]

### 2020-10-23 analysis/paper/003-burials-pre-network-modelling.R

@@ -117,16 +117,11 @@ set.vertex.attribute(burial_network_pre, "quantity", burial_pre$quantity)
 set.vertex.attribute(burial_network_pre, "age", burial_pre$Age_scale)
 set.vertex.attribute(burial_network_pre, "gender", burial_pre$gender)
 set.vertex.attribute(burial_network_pre, "ritual", burial_pre$ritual)
-set.vertex.attribute(burial_network_pre, "total", burial_pre$total)
+set.vertex.attribute(burial_network_pre, "burial_value", burial_pre$burial_value)
 
 #get distance matrix, need to run 002 code first
 pre_distance_n <- network(pre_distance, directed = F)
 set.edge.attribute(pre_distance_n, "dist", pre_distance_n)
-
-test_1 <- burial_network_pre ~
-  edges +
-  edgecov(pre_distance_n, "dist")
-summary(test_1)
 
 # plot
 set.seed(30)
@@ -167,7 +162,7 @@ summary(burial_network_pre ~ gwdegree(0:10))
 # model 2 considers cluster and degree, Morris et al. (2008)
 # check out the terms: http://mailman13.u.washington.edu/pipermail/statnet_help/2010/000575.html
 model_pre_2 <- burial_network_pre ~ edges + # density
-  gwesp(0.75, fixed = TRUE) +
+  gwesp(0.8, fixed = TRUE) +
   # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
   # number means weight parameter alpha, which controls the rate of declining marginal returns
   # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
@@ -186,7 +181,7 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
-  #absdiff('total') +
+  absdiff('burial_value') +
   gwesp(0.75, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   gwnsp(0.75, fixed = TRUE) + #0.75, #prior = -1
   gwdegree(0.8, fixed = TRUE) +
@@ -195,9 +190,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
[…]

### 2020-10-23 analysis/paper/004-burials-post-network-modelling.R

@@ -217,28 +217,25 @@ sim_esp_post <-
                                     /length(bgof_post$obs.esp)))
 
 # make dataframes for distribution stats
-distribution_degree_post <-
+posterior_distribution_post <-
   data.frame("moments" = c("Mean", "Variance", "Skewness"),
-             "Observed" = round(c(obs_degree_post$mean,
+             "Observed.degree" = round(c(obs_degree_post$mean,
                                   obs_degree_post$sd,
                                   obs_degree_post$skew), 2),
-             "Model" = round(c(mean(sim_degree_post$mean),
+             "Model.degree" = round(c(mean(sim_degree_post$mean),
                                mean(sim_degree_post$sd),
-                               mean(sim_degree_post$skew)),2))
-distribution_dist_post <-
-  data.frame("moments" = c("Mean", "Variance", "Skewness"),
-             "Observed" = round(c(obs_dist_post$mean,
-                                  obs_dist_post$sd,
-                                  obs_dist_post$skew), 2),
-             "Model" = round(c(mean(sim_dist_post$mean),
-                               mean(sim_dist_post$sd),
-                               mean(sim_dist_post$skew)),2))
-distribution_esp_post <-
-  data.frame("moments" = c("Mean", "Variance", "Skewness"),
-             "Observed" = round(c(obs_esp_post$mean,
-                                  obs_esp_post$sd,
-                                  obs_esp_post$skew), 2),
-             "Model" = round(c(mean(sim_esp_post$mean),
-                               mean(sim_esp_post$sd),
-                               mean(sim_esp_post$skew)),2))
+                               mean(sim_degree_post$skew)),2),
+             "Observed.distance" = round(c(obs_dist_post$mean,
+                                           obs_dist_post$sd,
+                                           obs_dist_post$skew), 2),
+             "Model.distance" = round(c(mean(sim_dist_post$mean),
+                                        mean(sim_dist_post$sd),
+                                        mean(sim_dist_post$skew)),2),
+             "Observed.esp" = round(c(obs_esp_post$mean,
+                                      obs_esp_post$sd,
+                                      obs_esp_post$skew), 2),
+             "Model.esp" = round(c(mean(sim_esp_post$mean),
+                                   mean(sim_esp_post$sd),
+                                   mean(sim_esp_post$skew)),2),
+             "Phase" = "post-European")
 

### 2020-10-23 analysis/paper/005-bergm-distribution-stats.R

@@ -14,6 +14,7 @@ post_bergm_stats <-
 bergm_two_phases <-
   rbind(pre_bergm_stats, post_bergm_stats)
 
+# making posterior distribution plot for both phases
 edges_value <-
   ggplot(bergm_two_phases,
        aes(x = edges, fill = phase)) +
@@ -38,8 +39,26 @@ dist_value <-
   geom_histogram(alpha = 0.8, position = "identity") +
   theme_minimal()
 
+# plot them together
 library(cowplot)
 plot_grid(edges_value,
           trans_value,
           degree_value,
           dist_value)
+
+# distribution stats for both phases
+distribution_two_phases <-
+  rbind(posterior_distribution_pre,
+        posterior_distribution_post)
+
+distribution_two_phases_longer <-
+  distribution_two_phases %>%
+  pivot_longer(cols = starts_with(c("Observed", "Model")),
+               names_to = "parameter",
+               values_to = "number") %>%
+  separate(parameter, c("value", "parameter"))
+
+ggplot(distribution_two_phases_longer,
+       aes(number, parameter)) +
+  geom_point(aes(color = Phase, shape = value), size = 3) +
+  facet_wrap(~moments, scales = "free")

### 2020-10-23 analysis/paper/paper.Rmd
edits to intro

@@
 <!--burial study- network analysis-->
-Burials contains valuable information of both individual mortuary practices and the structure of a society inherited in burial relations.
-With the development of network science and graph theory, studying relations among burials and exploring its social implication become promising [@Sosna2013].
-Mortuary practices such as burial forms, grave goods, and related ritual practices could represent the social relations between individuals, allowing the inference for social structures.
-This enables the study of social differences based on burials to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
-For examples, the use of burial goods can reflect social practices in a broader of cultural contexts to present personal wealth, social roles, or a traditional custom.
-Burials with grave goods in common often indicate some underlying social relations connected through trade, exchange, and gifting [@Coward2013, pp.
+Burials contain valuable information about both individual mortuary practices and the structure of a society.
+The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration their social implications [@Sosna2013].
+Mortuary practices such as burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [citation needed].
+This means we can use data from burials to study social differences to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
+For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs.
+Burials with grave goods in common often indicate some similar underlying social relations, for example, individuals that share similar access to trade, exchange, and gifting networks [@Coward2013, pp.
 252].
-Identifying the structure of relationships among burials can give an insight to the structure of a social organization.
-One common way to study connections between individuals is network analysis, which has been applied increasingly to archaeological studies in recent years [@Mills2017].
+Identifying the structure of relationships among burials can give insights into the structure of social organization.
+
+One common way to study connections between individuals is network analysis, which has been increasingly applied by archaeologists in recent years [@Mills2017].
 Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007].
-Technically speaking, actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
-The basic unit of relationship is called dyad, defined as a connection between two distinct actors [@Lee2016].
+<!-- define 'links' a bit more --> Actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
+The basic unit of relationship is called a dyad, defined as a connection between two distinct actors [@Lee2016].
 In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, indicating social relations in societies [@Marin2011].
-With network analysis, we can quantify, visualize, and statistically test hypotheses of the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-The application in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]
+With network analysis, we can quantify, visualize, and statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
+Network analysis in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]
 

### 2020-10-26 analysis/paper/paper.Rmd
delete code chunks in the paper rmd file

@@
 ```{r tidy-data}
-# combine some periods
-burial_three_period_age_tidy <-
+# number of each phase
+burial_phases_sample_size <-
   burial %>%
-  rename(burial_label = ID) %>%
   mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
   filter(!is.na(Phase)) %>% 
-  mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
-         Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
-         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
-  mutate_at(21:ncol(.), as.numeric) %>%
-  janitor::remove_empty(which = "cols") %>%
-  mutate(total = rowSums(.[c(21:48, 50, 55, 56)], na.rm = TRUE)) %>% #prestige goods
-  mutate(Porcelain = rowSums(.[c(40:48, 55, 56)], na.rm = TRUE)) %>% #B&W, porcelain, Anping, kendi
-  mutate(Porcelain = ifelse(Porcelain == 0, NA, Porcelain)) %>%
-  mutate(quantity = case_when(
-    total == 0 ~ "none",
-    total > 0 & total <= 7 ~ "low",
-    total > 7 & total < 100 ~ "medium",
-    total >= 100 ~ "high",
-    TRUE ~ "")) %>% # the classification is based on the result of histogram
-  mutate(Age_scale = case_when(
-    `Age` %in% c("1","2") ~ "0-12",
-    `Age` == "3" ~ "12~20",
-    `Age` %in% c("4","5","6","7","8") ~ "+20",
-    TRUE ~ "NA")) %>%
-  mutate(gender = case_when(
-    `Gender` %in% c("1","2") ~ "male",
-    `Gender` %in% c("3","4") ~ "female",
-    TRUE ~ "NA")) %>%
-  mutate(ritual = case_when(
-    `Stamped_ceramic` == "2" ~ "two pots",
-    `Stamped_ceramic` == "1" ~ "one pot",
-    TRUE ~ "NA")) %>%
-  mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
-         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
-         Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
-         Agate_bead_low = ifelse(Agate_bead == 1, 1, NA),
-         Agate_bead_med = ifelse(Agate_bead > 1 & Agate_bead <10, 1, NA),
-         Agate_bead_high = ifelse(Agate_bead > 10, 1, NA),
-         `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` < 100, 1, NA),
-         `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
-         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% #based on the result of histogram
-  select(burial_label,
-         Phase,
-         Age_scale,
-         gender,
-         ritual, # consider to remove if not very informative
-         Gold_bead_low,
-         Gold_bead_med,
-         Gold_bead_high,
-         Agate_bead_low,
-         Agate_bead_med,
-         Agate_bead_high,
-         Porcelain, #prestige good
-         Gold_leaf, #prestige good
-         fish_shape_knit, #prestige good
-         quantity)
-
-# number of each phase
-burial_three_period_age_number <-
-  burial_three_period_age_tidy %>%
   count(Phase)
 
-pre_E <- burial_three_period_age_number$Phase[4]
-post_E <- burial_three_period_age_number$Phase[3]
-chi <- burial_three_period_age_number$Phase[1]
-```
-
-```{r prep-network-object-pre}
-# filter pre burials
-burial_pre <-
-  burial_three_period_age_tidy %>%
-  filter(Phase == "pre") %>%
-  janitor::remove_empty(which = "cols")
-
-# create node list using burial index by phase
-#--------------------pre--------------------------------
-nodes_pre <-
-  burial_three_period_age_tidy %>%
-  filter(Phase == "pre") %>%
-  select(burial_label) %>%
-  rowid_to_column("id")
-
-# pair wise combinations for burials as index for later map function
-library(gtools)
-burial_comb_pre = combinations(length(nodes_pre$burial_label), #t(combn(nodes_pre$burial_label, 2))
-                               2, nodes_pre$burial_label,
-                               repeats = TRUE)
-colnames(burial_comb_pre) = c("burial_1", "burial_2")
-burial_comb_pre = as_tibble(burial_comb_pre)
-
-# create list for each burial that contains the burial good types and their counts
-edge_list_pre <-
-  burial_three_period_age_tidy %>%
-  select(burial_label, 6:14) %>% # need to change for each exploration
-  pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
-  #mutate(burial_connection = rep(unique(burial_label), length.out = length(burial_label)))
-  group_by(burial_label) %>%
-  nest()
-
-# pair-wise list
-common_counts_lst_pre <-
-  map2(burial_comb_pre$burial_1,
-       burial_comb_pre$burial_2,
-       ~bind_cols(
-         edge_list_pre %>%
-           filter(burial_label == .x) %>%
-           unnest(data),
-         edge_list_pre %>%
-           filter(burial_label == .y) %>%
-           unnest(data)) %>%
-         rowwise() %>%
-         mutate(common_counts = sum(count...3, count...6)))
-
-# count of ornament types in common between each pair of burials
-common_counts_vct_pre <- map_int(common_counts_lst_pre, ~sum(!is.na(.x$common_counts)))
-
-burial_comb_with_common_counts_pre <-
-  burial_comb_pre %>%
-  mutate(common_counts = common_counts_vct_pre) %>%
-  mutate(common_counts = ifelse(burial_1 == burial_2, 0, common_counts)) # use 0 for self loop
-
-# change label to ids for node linking
-edges_pre <-
-  burial_comb_with_common_counts_pre %>%
-  left_join(nodes_pre, by = c("burial_1" = "burial_label")) %>%
-  rename(from = id) %>%
-  left_join(nodes_pre, by = c("burial_2" = "burial_label")) %>%
-  rename(to = id)
-
-edges_for_network_pre <-
-  select(edges_pre, from, to, common_counts) %>%
-  filter (!common_counts == 0) # remove rows with no goods in common
-# %>% mutate(common_counts = ifelse(common_counts > 1, 1, common_counts)) # for unweighted network
-```
-
-```{r prep-network-object-post}
-# filter post burials
-burial_post <-
-  burial_three_period_age_tidy %>%
-  filter(Phase == "post") %>%
-  janitor::remove_empty(which = "cols") %>%
-  filter(quantity != "none") # remove burial without burial goods
-
-# create node list
-nodes_post <-
-  burial_three_period_age_tidy %>%
-  filter(Phase == "post") %>%
-  filter(quantity != "none") %>%
-  select(burial_label) %>%
-  rowid_to_column("id")
-
-# pair wise combinations for burials as index for later map function
-library(gtools)
-burial_comb_post = combinations(length(nodes_post$burial_label),
-                               2, nodes_post$burial_label,
-                               repeats = TRUE)
-colnames(burial_comb_post) = c("burial_1", "burial_2")
-burial_comb_post = as_tibble(burial_comb_post)
-
-# create list for each burial that contains the burial good types and their counts
-edge_list_post <-
-  burial_three_period_age_tidy %>%
-  select(burial_label, 6:14) %>% # need to change for each exploration
-  pivot_longer(-burial_label, names_to = "goods", values_to = "count") %>%
-  group_by(burial_label) %>%
-  nest()
-
-# pair-wise list
-common_counts_lst_post <-
-  map2(burial_comb_post$burial_1,
-       burial_comb_post$burial_2,
-       ~bind_cols(
-         edge_list_post %>%
-           filter(burial_label == .x) %>%
-           unnest(data),
-         edge_list_post %>%
-           filter(burial_label == .y) %>%
-           unnest(data)) %>%
-         rowwise() %>%
-         mutate(common_counts = sum(count...3, count...6)))
-
-# count of ornament types in common between each pair of burials
-common_counts_vct_post <- map_int(common_counts_lst_post, ~sum(!is.na(.x$common_counts)))
-
-burial_comb_with_common_counts_post <-
-  burial_comb_post %>%
-  mutate(common_counts = common_counts_vct_post) %>%
-  mutate(common_counts = ifelse(burial_1 == burial_2, 0, common_counts)) # use 0 for self loop
-
-# change label to ids for node linking
-edges_post <-
-  burial_comb_with_common_counts_post %>%
-  left_join(nodes_post, by = c("burial_1" = "burial_label")) %>%
-  rename(from = id) %>%
-  left_join(nodes_post, by = c("burial_2" = "burial_label")) %>%
-  rename(to = id)
-
-edges_for_network_post <-
-  select(edges_post, from, to, common_counts) %>%
-  filter (!common_counts == 0) # remove rows with no goods in common
+pre_E <- burial_phases_sample_size %>% filter(Phase == "pre") %>% pull()
+post_E <- burial_phases_sample_size %>% filter(Phase == "post") %>% pull()
+chi <- burial_phases_sample_size %>% filter(Phase == "chi") %>% pull()
 ```
@@
 
-```{r network-ggraph}
-# network analysis using ggraph pkg
-library(tidygraph)
-library(ggraph)
-# pre-Euro
-relation_tidy_pre <- tbl_graph(nodes = nodes_pre,
-                               edges = edges_for_network_pre,
-                               directed = FALSE)
-
-relation_tidy_pre %>%
-  activate(edges) %>%
-  arrange(desc(common_counts))
-
-# plot
-ggraph(relation_tidy_pre, layout = "graphopt") +
-  geom_node_point() +
-  geom_edge_link(aes(width = common_counts), alpha = 0.8) +
-  scale_edge_width(range = c(0.2, 2)) +
-  geom_node_text(aes(label = burial_label), repel = TRUE) +
-  labs(edge_width = "common item") +
-  theme_graph()
-
-# post-Euro
-relation_tidy_post <- tbl_graph(nodes = nodes_post,
-                               edges = edges_for_network_post,
-                               directed = FALSE)
-
-relation_tidy_post %>%
-  activate(edges) %>%
-  arrange(desc(common_counts))
-
-# plot
-ggraph(relation_tidy_post, layout = "graphopt") +
-  geom_node_point() +
-  geom_edge_link(aes(width = common_counts), alpha = 0.8) +
-  scale_edge_width(range = c(0.2, 2)) +
-  geom_node_text(aes(label = burial_label), repel = TRUE) +
-  labs(edge_width = "common item") +
-  theme_graph()
-```
-
-```{r network-object}
-library(network)
-
-burial_network_pre <-
-  network(edges_for_network_pre, # the network object
-          vertex.attr = nodes_pre, # node list
-          directed = FALSE, # specify whether the network is directed
-          ignore.eval = FALSE, # FALSE = weighted
-          loops = FALSE, # do we allow self ties (should not allow them)
-          matrix.type = "edgelist") # the type of input
-
-burial_network_post <-
-  network(edges_for_network_post,
-          vertex.attr = nodes_post,
-          directed = FALSE,
-          ignore.eval = FALSE, 
-          loops = FALSE, 
-          matrix.type = "edgelist") 
-```
-
 The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle.
@@
 The post-European network also presented subgroups with more edges connecting burials.
-
-```{r add-node-attribute-color}
-# prepare notwork objects for Bayesian ERGMs 
-library(statnet)
-library(Bergm)
-
-# Add attributes to the network object burial_network_pre
-set.vertex.attribute(burial_network_pre, "quantity", burial_pre$quantity)
-set.vertex.attribute(burial_network_pre, "age", burial_pre$Age_scale)
-set.vertex.attribute(burial_network_pre, "gender", burial_pre$gender)
-
-# plot
-set.seed(30)
-quantity <- get.vertex.attribute(burial_network_pre, "quantity")
-age <- get.vertex.attribute(burial_network_pre, "age")
-
-plot(burial_network_pre,
-     displaylabels = TRUE,
-     vertex.col = "quantity",
-     vertex.cex = degree(burial_network_pre, cmode = 'indegree')/5, #size nodes
-     pad = 1)
-
-legend("topleft",
-       col = c(2, 3, 1, 4), # need to adjust each time
-       pch    = 20,
-       legend = unique(quantity), # age
-       title  = 'Burial good counts')
-
-#? can't get the items in legend in order, tried as.factor and set their level but does not work
-#? can't match color with categories, using an odd method at the moment
-
-# Add attributes to the network object burial_network_post
-set.vertex.attribute(burial_network_post, "quantity", burial_post$quantity)
-set.vertex.attribute(burial_network_post, "age", burial_post$Age_scale)
-set.vertex.attribute(burial_network_post, "gender", burial_post$gender)
-
-# plot
-set.seed(30)
-quantity <- get.vertex.attribute(burial_network_post, "quantity")
-age <- get.vertex.attribute(burial_network_post, "age")
-ID <- get.vertex.attribute(burial_network_post, "burial_label") # not sure how to get id on the network plot
-
-plot(burial_network_post,
-     displaylabels = TRUE,
-     vertex.col = "quantity",
-     vertex.cex = degree(burial_network_post, cmode = 'indegree') / 13, #size nodes to their in-degree
-     #vertex.sides = ifelse(burial_network_pre %v% "", 4, 50),
-     pad = 1)
-
-legend("topleft",
-       col = c(4, 3, 2, 1), # need to adjust each time
-       pch    = 20,
-       legend = unique(quantity),
-       title  = 'Burial good counts')
-```
 
@@
 ```{r network-ERGMs}
-# model 1
-model.ergm <- burial_network_pre ~
-  edges + 
-  density +
-  triangle 
-summary(model.ergm)
-
-# model 2 considers cluster and degree
-model.2 <- burial_network_pre ~ 
-  edges + # density
-  gwesp(0.2, fixed = TRUE) +  # transitivity(cohesion; triangle)
-  gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
-summary(model.2)
-
-# model 3 with node attributes
-model.3 <- burial_network_pre ~ edges +  # the overall density of the network
-  nodematch('quantity') + # quantity-based homophily
-  nodematch('age') +
-  nodematch('gender') +
-  gwesp(0.75, fixed = TRUE) +    # transitivity
-  gwnsp(0.75, fixed = TRUE) +
-  gwdegree(0.2, fixed = TRUE) # popularity
-summary(model.3)
 ```
@@
 ```{r Bayesian-ERGMs}
-# Specify a prior distribution: normal distribution (low density and high transitivity)
-# need to observe the graph and select suitable prior 
-prior.mean <- c(2, 0, 0, 0, 5, 0, 0) 
-# follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 7) # covariance matrix structure
-# where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
-
-# Estimated posterior means, medians and 95% credible intervals for Models.3
-# bergmM: Bayesian exponential random graphs models under missing data using the approximate exchange algorithm. Use this because of missing data for age variable
-parpost <- bergmM(model.3,
-                  prior.mean  = prior.mean,
-                  prior.sigma = prior.sigma,
-                  burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
-                  main.iters  = 2000, # iterations for every chain of the population
-                  aux.iters   = 10000, # MCMC steps used for network simulation
-                  nchains     = 8, # number of chains of the population MCMC
-                  gamma       = 0.2) # scalar; parallel adaptive direction sampling move factor, acceptance rate
-
-# summary of parameter estimation
-summary(parpost)
-
-# plot the parameter posterior distribution
-plot(parpost)
 ```
@@
 ```{r Bayesian-gof}
-# Model assessment, Bayesian goodness of fit diagnostics:
-bgof(parpost,
-     aux.iters = 10000,
-     n.deg     = 15,
-     n.dist    = 15,
-     n.esp     = 15)
-
-# An estimated ERGM is fitting perfectly a certain observed network if the red line falls inside this interval
-# a result that is very difficult to obtain in practice (Caimo 2017)
 ```

### 2020-10-28 analysis/paper/paper.Rmd
adjust gamma and add orientation variable

@@
 <!--burial study- network analysis-->
-Burials contain valuable information about both individual mortuary practices and the structure of a society.
+Burials contain valuable information about both individual mortuary practices and the structure of a society .
 The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration their social implications [@Sosna2013].
@@
 For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
 Those configurations represent the structure or the property of a network and can be expressed by network statistics.
@@
 <!--Network construction and specification-->
-We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis of social change using the R programming language with bergm package [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008].
+We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language with the bergm package [@Caimo2014], which built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
 Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations.
@@
 We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
+For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].
+To
 
@@
 The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
-This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.
-Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].
+The Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+This provides a statistical approach to check how well the estimated posterior parameter distribution based on our hypotheses can reproduce networks similar to the general structural features of the observed network.
+We then compared the distribution of our observed networks, pre-European and post-European networks, with the distribution of our hypothesized model with higher popularity.
+We expect to see the network after the European presence has a better fit with the hypothesis, indicative of a structure with higher social differentiation.
 

### 2020-10-28 analysis/paper/000-goods-class-plot.R

@@ -58,3 +58,12 @@ burial_three_period_tidy_explore %>%
   ggplot(aes(`Indo-Pacific_bead`)) +
   geom_histogram() +
   scale_x_log10()
+
+# orientation
+burial %>%
+  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
+  filter(!is.na(Phase)) %>%
+  ggplot(aes(x = Degree_axis, fill = Phase)) +
+  geom_histogram()+
+  coord_polar() +
+  scale_x_continuous(limits = c(0,360))

### 2020-10-28 analysis/paper/000-value-burials.R

@@ -76,6 +76,7 @@ burial_with_type_value <-
   mutate(burial_value = rowSums(across(where(is.numeric)))) %>%
   select(burial_label, burial_value)
 
+
 # distribution plot of burial value
 mean(burial_with_type_value$burial_value)
 quantile(burial_with_type_value$burial_value)

### 2020-10-28 analysis/paper/001-data-tidy.R

@@ -1,9 +1,3 @@
-library(readxl)
-library(tidyverse)
-library(here)
-
-burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
-
 # run 000-value-burial first before the following code
 # burial data with combined age and three phases
 burial_three_period_age_tidy <-
@@ -40,6 +34,10 @@ burial_three_period_age_tidy <-
     `Stamped_ceramic` == "2" ~ "two pots",
     `Stamped_ceramic` == "1" ~ "one pot",
     TRUE ~ "NA")) %>%
+  mutate(orientation = case_when(
+    `Degree_axis` %in% c(275:330) ~ "northwest",
+    `Degree_axis` %in% c(0:90) ~ "northeast",
+    TRUE ~ "NA")) %>%
   mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
          Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
          Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
@@ -71,7 +69,8 @@ burial_three_period_age_tidy <-
          quantity,
          total,
          burial_value,
-         value_class) # select specific variable to drop columns (uninformative variables)
+         value_class,
+         orientation) # select specific variable to drop columns (uninformative variables)
 
 # number of each phase
 burial_three_period_age_number <-

### 2020-10-28 analysis/paper/002-burial-location.R

@@ -44,7 +44,7 @@ df_pre_distance <-
   arrange(Id)
 
 pre_distance <-
-  st_distance(df_pre_distance$burial_cent)
+  st_distance(df_pre_distance$burial_cent) # will be ran in 003
 
 # get distance between post-E burials
 df_post_distance <-
@@ -55,7 +55,7 @@ df_post_distance <-
   arrange(Id)
 
 post_distance <-
-  st_distance(df_post_distance$burial_cent)
+  st_distance(df_post_distance$burial_cent) # will be ran in 004
 
 # mapping burials over the AD section
 KWL_burial_map <-

### 2020-10-29 analysis/paper/paper.Rmd
edit Intro, adjust model, and add references

@@
 <!--burial study- network analysis-->
-Burials contain valuable information about both individual mortuary practices and the structure of a society .
+Burial analysis has long developed in archaeology to study societies through mortuary practices, such as cultural and biological records in burial contexts, reflecting the social status or identity of the deceased [@Binford1971; @Saxe1970].
+This approach is criticized for its simple connection between real life in society and death expression without considering the manipulation of the living or distortion [@Hodder1980; @Pearson1982].
+However, quantitative analysis for burial data is still an effective way to answer questions related to social representations with both social representation for death and life in mind [@Chapman2003].
+Mortuary practices include burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
 The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration their social implications [@Sosna2013].
-Mortuary practices such as burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [citation needed].
 This means we can use data from burials to study social differences to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
@@
 One common way to study connections between individuals is network analysis, which has been increasingly applied by archaeologists in recent years [@Mills2017].
-Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007].
-<!-- define 'links' a bit more --> Actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
+Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links representing relationship, such as friendships, connecting them [@Brughmans2018; @Robins2007].
+Actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
 The basic unit of relationship is called a dyad, defined as a connection between two distinct actors [@Lee2016].
@@
 With network analysis, we can quantify, visualize, and statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Network analysis in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]
+Network analysis in archaeology is widely used for different topics, such as exploring hunter-gatherer exchange network in the Kuril Islands based on ceramic compositions [@Erik2015]; [@Coward2013; @Mizoguchi2013; @Potgieter2008].
+
+ore specifically, this research constructs network models from the compositional analysis of ceramic artifacts recovered in the Kuril Islands of northeast Asia and evaluates network centrality measures using bootstrap simulation and sensitivity analysis.
+Results of this research suggest that while archaeologists may never fully identify the "true" network of past relationships, network models that approximate "true" network structure can provide useful metrics in exploring the behavior of past hunter-gatherer populations.
+Overall, given the challenges associated with hunter-gatherer archaeological data, it is argued here that critically evaluating the structure and stability of network models is an essential first step in developing an archaeological network analysis that is relevant and informative to research on past small-scale societies.
 

### 2020-10-29 analysis/paper/003-burials-pre-network-modelling.R

@@ -122,7 +122,7 @@ set.vertex.attribute(burial_network_pre, "burial_value", burial_pre$burial_value
 set.vertex.attribute(burial_network_pre, "orientation", burial_pre$orientation)
 
 #get distance matrix, need to run 002 code first
-pre_distance_n <- network(pre_distance, directed = F)
+pre_distance_n <- network(pre_distance, matrix.type = "adjacency", directed = F)
 set.edge.attribute(pre_distance_n, "dist", pre_distance_n)
 
 # plot
@@ -179,24 +179,24 @@ summary(model_pre_2)
 
 # model 3 considers cluster, degree, and node attributes
 model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the network
-  nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
+  #nodematch('quantity') + # quantity-based homophily, categorical nodal attribute, the similarity of connected nodes
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
   nodematch('value_class') +
   nodematch('orientation') +
-  absdiff('burial_value') +
+  #absdiff('burial_value') +
   gwesp(0.8, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
-  gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
+  #gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
   gwdegree(0.8, fixed = TRUE) +
-  edgecov(pre_distance_n, "dist")
+  dyadcov(pre_distance_n, "dist")
 summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, 1, -1, 0, 0, 1, 0, 1, 1, -1, 3, -1) # positive prior number for edge means high density
+prior.mean <- c(-3, -1, 0, 0, 1, 0, 1, 3, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 12, 12) # covariance matrix structure
+prior.sigma <- diag(3, 9, 9) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
[…]

### 2020-10-30 analysis/paper/paper.Rmd
adjust models and edit Intro

--- 2020-10-28 analysis/paper/paper.Rmd
@@
 <!--burial study- network analysis-->
-Burials contain valuable information about both individual mortuary practices and the structure of a society .
-The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration their social implications [@Sosna2013].
-Mortuary practices such as burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [citation needed].
+Burial analysis has long developed in archaeology to study societies through mortuary practices, such as cultural and biological records in burial contexts, reflecting the social status or identity of the deceased [@Binford1971; @Saxe1970].
+Despite the criticism for not considering the burial manipulation by the living or situations of less connection to real life [@Hodder1980; @Pearson1982], burial analysis with a quantitative approach is still effective to answer questions related to social representations with both archaeological contexts of death and life in mind [@Chapman2003].
+Mortuary practices include burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
+The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration of their social implications [@Sosna2013].
 This means we can use data from burials to study social differences to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
 For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs.
-Burials with grave goods in common often indicate some similar underlying social relations, for example, individuals that share similar access to trade, exchange, and gifting networks [@Coward2013, pp.
+Burials with the same mortuary practices may indicate some similar underlying social relations, for example, individuals with grave goods in common that share similar access to trade, exchange, and gifting networks [eg.
+@Coward2013, pp.
 252].
@@
 
-One common way to study connections between individuals is network analysis, which has been increasingly applied by archaeologists in recent years [@Mills2017].
-Networks are represented as graphs consisting of a set of actors, such as people, groups, or organizations, and a set of links connecting them [@Brughmans2018; @Robins2007].
-<!-- define 'links' a bit more --> Actors are "nodes" connected by multiple "edges", representing relationships, between nodes.
+To measure relations and patterns underlying the relations, social network analysis (SNA), is a useful method that has been increasingly applied by archaeologists to study relationship between humans, or sites in past societies [@Mills2017].
+A social network can be expressed as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, for example, friendship, collaborations, information flow, trade ties, or any forms of connection of interest [@Marin2011; @Wasserman1994].
 The basic unit of relationship is called a dyad, defined as a connection between two distinct actors [@Lee2016].
-In archaeology, relationships can be viewed as flows of resources that reflect relations of control, dependence, and cooperation, indicating social relations in societies [@Marin2011].
-With network analysis, we can quantify, visualize, and statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Network analysis in archaeology is widely used for different topics including distribution of archaeological sites, grave goods, and social complexity [@Coward2013; @Mizoguchi2013; @Potgieter2008]
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on archaeological theories for relationships [@Mills2017; @Brughmans2018].
+This allows answering questions related to interaction and exchange, identities, and power relationships between social groups or archaeological sites by quantifying and visualizing social relations from archaeological data.
+There are two common approaches of archaeological applications, one is analyzing structures of networks built on archaeological data and finding patterns, and another is modeling theory to represent and test dependence of relationships [@Brughmans2018].
+The modeling approach is useful to statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
+Those approaches are widely used for different topics, such as exploring hunter-gatherer exchange network in the Kuril Islands based on ceramic compositions [@Erik2015]; long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013]; hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013].
 

### 2020-10-30 analysis/paper/003-burials-pre-network-modelling.R

@@ -107,6 +107,8 @@ burial_network_pre <-
           matrix.type = "edgelist") # the type of input
 
 plot(burial_network_pre, vertex.cex = 1)
+network.density(burial_network_pre)
+network.dyadcount(burial_network_pre, na.omit = TRUE)
 
 #-----------------------attach attributes and make graphs------------------------------
 library(statnet)
@@ -184,7 +186,7 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   nodematch('gender') +
   nodematch('ritual') +
   nodematch('value_class') +
-  nodematch('orientation') +
+  #nodematch('orientation') +
   #absdiff('burial_value') +
   gwesp(0.8, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   #gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
@@ -194,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, -1, 0, 0, 1, 0, 1, 3, -1) # positive prior number for edge means high density
+prior.mean <- c(-3, -1, 0, 0, 1, 1, 3, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 9, 9) # covariance matrix structure
+prior.sigma <- diag(3, 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
@@ -209,7 +211,7 @@ pre_bergm <- bergmM(model_pre_3,
                   burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
                   main.iters  = 2000, # iterations for every chain of the population
                   aux.iters   = 10000, # MCMC steps used for network simulation
-                  nchains     = 6, # number of chains of the population MCMC
+                  nchains     = 16, # number of chains of the population MCMC
                   gamma       = 0) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
 summary(pre_bergm)

### 2020-10-30 analysis/paper/004-burials-post-network-modelling.R

@@ -103,7 +103,7 @@ burial_network_post <-
 
 # plot
 plot(burial_network_post, vertex.cex = 1)
-
+network.density(burial_network_post)
 #-----------------------Bayesian ERGMs------------------------------
 library(statnet)
 library(Bergm)
@@ -115,9 +115,10 @@ set.vertex.attribute(burial_network_post, "gender", burial_post$gender)
 set.vertex.attribute(burial_network_post, "ritual", burial_post$ritual)
 set.vertex.attribute(burial_network_post, "value_class", burial_post$value_class)
 set.vertex.attribute(burial_network_post, "burial_value", burial_post$burial_value)
+set.vertex.attribute(burial_network_post, "orientation", burial_post$orientation)
 
 #get distance matrix, need to run 002 code first
-post_distance_n <- network(post_distance, directed = F)
+post_distance_n <- network(post_distance, matrix.type = "adjacency", directed = F)
 set.edge.attribute(post_distance_n, "dist", post_distance_n)
 
 # plot
@@ -155,21 +156,22 @@ summary(model.2)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 model.post.3 <- burial_network_post ~ edges +  # the overall density of the network
-  nodematch('quantity') +  # quantity-based homophily, the similarity of connected nodes
+  #nodematch('quantity') +  # quantity-based homophily, the similarity of connected nodes
   nodematch('age') +
   nodematch('gender') +
   nodematch('ritual') +
   nodematch('value_class') +
-  absdiff('burial_value') +
+  #nodematch('orientation') +
+  #absdiff('burial_value') +
   gwesp(1.7, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
-  gwnsp(1.7, fixed = TRUE) + # original 1.8
+  #gwnsp(1.7, fixed = TRUE) + # original 1.8
   gwdegree(0.8, fixed = TRUE) +
-  edgecov(post_distance_n, "dist")
+  dyadcov(post_distance_n, "dist")
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 1, -1, 0, 0, 1, 1, 1, -1, 3, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(3, 11, 11) # covariance matrix structure
[…]

### 2020-11-02 analysis/paper/paper.Rmd
edit chronology section and adject bootstrap parameters

@@
 <!--chronology-->
-To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies.
-We excluded the burials from the Chinese phase (n = `r chi`) for analysis due to much smaller sample size.
-Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology.
-The chronology is then evaluated by the burials with light gray glazed jars, also called An-ping jars that were believed to be produced in Southeast China.
-They circulated in Southeast Asia from the late 16th to 18th century, and often found at sites with the European occupation in the early 17th century in many Aisan places, including Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011].
-
-In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007].
-Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data.
-With the information of burials, including quantity of grave goods, types of grave goods, burial identities, we can reconstruct social relations between individuals by linking them according to similar prestige goods.
+To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison based on established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon dates [@LWandBM2020ornament; @Chen2007].
+We excluded the burials from the Chinese phase (n = `r chi`) for analysis due to the smaller sample size.
+There are eight burials heavily disturbed by modern construction that are excluded to avoid misinterpreting due to distorted chronology.
+We reconstructed social relations between individuals by linking them according to similar prestige goods.
 This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
 252].
-Also, prestige materials usually relate to a representation of social status, where the same economic or social class, tend to use similar goods as a status indicators [citations].
-Therefore, we build network where burials represent actors (nodes in the network) that are linked when they have th same prestige goods in common.
-The prestige goods we identified including agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments, according to previous research [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
+Also, prestige materials usually relate to a representation of social status because of their limited accessibility that can be used as status indicators [citations].
+Therefore, we build networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
+The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments that are considered as high-value items by studies across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
 

### 2020-11-04 analysis/paper/paper.Rmd
adjust the parameter for ritual

@@
 <!--burial study- network analysis-->
-Burial analysis has long developed in archaeology to study societies through mortuary practices, such as cultural and biological records in burial contexts, reflecting the social status or identity of the deceased [@Binford1971; @Saxe1970].
-Despite the criticism for not considering the burial manipulation by the living or situations of less connection to real life [@Hodder1980; @Pearson1982], burial analysis with a quantitative approach is still effective to answer questions related to social representations with both archaeological contexts of death and life in mind [@Chapman2003].
-Mortuary practices include burial forms, grave goods, and related ritual practices that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
-The development of network science and graph theory provides many new tools for the study of the relations between burials and exploration of their social implications [@Sosna2013].
-This means we can use data from burials to study social differences to answer questions related to social inequality and complexity [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
+Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent social status or identity of the deceased [@Binford1971; @Saxe1970].
+Despite the criticism for overlooking the burial manipulation by the living that does not necessarily connect to the real life [@Hodder1980; @Pearson1982], with comparisons between different archaeological contexts of death and life in mind, burials can still provide valuable information to infer past societies [@Chapman2003].
+For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
+To study social relations, the recent development of network science and graph theory provides many new tools for the exploration and analyzing the structure of relational data [@Brandes2013].
+This means we can answer anthropological questions about social differentiation, and social inequality or complexity based on our assumptions of archaeological data [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
 For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs.
-Burials with the same mortuary practices may indicate some similar underlying social relations, for example, individuals with grave goods in common that share similar access to trade, exchange, and gifting networks [eg.
+The flow of material goods  Burials with the same goods may indicate some similar underlying social relations where individuals share similar access to trade, exchange, and gifting networks [eg.
 @Coward2013, pp.

### 2020-11-04 analysis/paper/001-data-tidy.R

@@ -30,10 +30,9 @@ burial_three_period_age_tidy <-
     `Gender` %in% c("1","2") ~ "male",
     `Gender` %in% c("3","4") ~ "female",
     TRUE ~ "NA")) %>%
-  mutate(ritual = case_when(
-    `Stamped_ceramic` == "2" ~ "two pots",
-    `Stamped_ceramic` == "1" ~ "one pot",
-    TRUE ~ "none")) %>%
+  mutate(ritual_pottery = case_when(
+    `Stamped_ceramic` %in% c("1", "2") ~ "presence",
+    TRUE ~ "absense")) %>%
   mutate(orientation = case_when(
     `Degree_axis` %in% c(275:330) ~ "northwest",
     `Degree_axis` %in% c(0:90) ~ "northeast",
@@ -52,7 +51,7 @@ burial_three_period_age_tidy <-
          Phase,
          Age_scale,
          gender,
-         ritual, # consider to remove if not very informative
+         ritual_pottery, # consider to remove if not very informative
          Gold_bead_low,
          Gold_bead_med,
          Gold_bead_high,
@@ -65,7 +64,7 @@ burial_three_period_age_tidy <-
          Stonewares, #prestige good
          Gold_leaf, #prestige good
          fish_shape_knit, #prestige good
-         #Bell, #children's burials
+         #Bell, #children burials
          quantity,
          total,
          burial_value,

### 2020-11-04 analysis/paper/004-burials-post-network-modelling.R

@@ -112,7 +112,7 @@ library(Bergm)
 set.vertex.attribute(burial_network_post, "quantity", burial_post$quantity)
 set.vertex.attribute(burial_network_post, "age", burial_post$Age_scale)
 set.vertex.attribute(burial_network_post, "gender", burial_post$gender)
-set.vertex.attribute(burial_network_post, "ritual", burial_post$ritual)
+set.vertex.attribute(burial_network_post, "ritual_pottery", burial_post$ritual_pottery)
 set.vertex.attribute(burial_network_post, "value_class", burial_post$value_class)
 set.vertex.attribute(burial_network_post, "burial_value", burial_post$burial_value)
 set.vertex.attribute(burial_network_post, "orientation", burial_post$orientation)
@@ -159,7 +159,7 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
   #nodematch('quantity') +  # quantity-based homophily, the similarity of connected nodes
   nodematch('age') +
   nodematch('gender') +
-  nodematch('ritual') +
+  nodematch('ritual_pottery') +
   nodematch('value_class') +
   #nodematch('orientation') +
   #absdiff('burial_value') +
@@ -170,7 +170,7 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, -1, 0, 0, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
+prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
 prior.sigma <- diag(3, 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
@@ -197,6 +197,7 @@ bgof_post <-
 summary(bgof_post)
 
 # calculate moments for observed and simulated GOF distribution
+library(psych)
 obs_degree_post <- describe(bgof_post$obs.degree * 0:(length(bgof_post$obs.degree)-1))
 obs_dist_post <- describe(bgof_post$obs.dist * 1:length(bgof_post$obs.dist))
 obs_esp_post <- describe(bgof_post$obs.esp * 0:(length(bgof_post$obs.esp)-1))

### 2020-11-05 analysis/paper/paper.Rmd
add more citations and adjust parameters

@@
 For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
-To study social relations, the recent development of network science and graph theory provides many new tools for the exploration and analyzing the structure of relational data [@Brandes2013].
-This means we can answer anthropological questions about social differentiation, and social inequality or complexity based on our assumptions of archaeological data [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
-For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs.
-The flow of material goods  Burials with the same goods may indicate some similar underlying social relations where individuals share similar access to trade, exchange, and gifting networks [eg.
+To study the relations among burials, the recent development of network science provides many new tools to visualize and analyze relationships in archaeological data.
+Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013].
+A basic network is viewed as a aggregation of pattern that includes individual elements, pair-wise relationships, and overall structure showing global patterning that are represented in data.
+This provide a useful framework for dealing with relational data in archaeology, such as past trade is abstracted as a network of individual entities with the flow of goods between them to represent their interaction [@Collar2015, pp.
+4].
+Similar conceptual concepts can be applied to other cases to explore a wide variety of topics in terms of archaeological data.
+For burials, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs that can infer social differentiation or complexity[@Gamble2002; @Janes2013].
+Burials with the same goods may indicate some similar underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
 @Coward2013, pp.
 252].
-Identifying the structure of relationships among burials can give insights into the structure of social organization.
+This enables the exploration of the structure of the past social organization through identifying the relationships among burials.
 

### 2020-11-05 analysis/paper/003-burials-pre-network-modelling.R

@@ -190,7 +190,7 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   #absdiff('burial_value') +
   gwesp(0.8, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   #gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
-  gwdegree(0.8, fixed = TRUE) +
+  gwdegree(0.8, fixed = TRUE) + # prior = 3
   dyadcov(pre_distance_n, "dist")
 summary(model_pre_3)
 
@@ -198,7 +198,7 @@ summary(model_pre_3)
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
 prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(3, 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(10,1,1,1,1,10,10,1), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-05 analysis/paper/004-burials-post-network-modelling.R

@@ -157,8 +157,8 @@ summary(model.2)
 #--------------------Bayesian inference for ERGMs-------------------------
 model.post.3 <- burial_network_post ~ edges +  # the overall density of the network
   #nodematch('quantity') +  # quantity-based homophily, the similarity of connected nodes
-  nodematch('age') +
-  nodematch('gender') +
+  #nodematch('age') +
+  #nodematch('gender') +
   nodematch('ritual_pottery') +
   nodematch('value_class') +
   #nodematch('orientation') +
@@ -170,8 +170,8 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(3, 8, 8) # covariance matrix structure
+prior.mean <- c(-3, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(3, 6, 6) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,
@@ -179,7 +179,7 @@ post_bergm <- bergmM(model.post.3,
                  burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
-                 nchains     = 16, # number of chains of the population MCMC
+                 nchains     = 12, # number of chains of the population MCMC
                  gamma       = 0) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
 summary(post_bergm)

### 2020-11-06 analysis/paper/paper.Rmd
add citations to Intro

--- 2020-11-04 analysis/paper/paper.Rmd
@@
 <!--burial study- network analysis-->
-Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent social status or identity of the deceased [@Binford1971; @Saxe1970].
+Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
 Despite the criticism for overlooking the burial manipulation by the living that does not necessarily connect to the real life [@Hodder1980; @Pearson1982], with comparisons between different archaeological contexts of death and life in mind, burials can still provide valuable information to infer past societies [@Chapman2003].
 For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
-To study social relations, the recent development of network science and graph theory provides many new tools for the exploration and analyzing the structure of relational data [@Brandes2013].
-This means we can answer anthropological questions about social differentiation, and social inequality or complexity based on our assumptions of archaeological data [@Ehrhardt2009; @Gamble2002; @Janes2013; @Johansen2015].
-For example, burial goods can reflect social practices in broader cultural contexts to represent personal wealth, social roles, or traditional customs.
-The flow of material goods  Burials with the same goods may indicate some similar underlying social relations where individuals share similar access to trade, exchange, and gifting networks [eg.
+To study the relations among burials, the recent development of network science provides many new tools to visualize and analyze relationships.
+Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013].
+That is, a network can be viewed as a patterned aggregation that includes individual elements, pair-wise relationships (the dyads), and overall structure showing global patterning represented in data.
+
+A network is generally visualized as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on theories for relationships [@Mills2017; @Brughmans2018].
+For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp.
+4].
+Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial analysis.
+Burial goods, especially for high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status that can infer social differentiation or complexity [@Gamble2002; @Janes2013].
+Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
 @Coward2013, pp.
 252].
-Identifying the structure of relationships among burials can give insights into the structure of social organization.
+This enables the exploration of the structure of the past social organization by identifying the relationships among burials.
 
-To measure relations and patterns underlying the relations, social network analysis (SNA), is a useful method that has been increasingly applied by archaeologists to study relationship between humans, or sites in past societies [@Mills2017].
-A social network can be expressed as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, for example, friendship, collaborations, information flow, trade ties, or any forms of connection of interest [@Marin2011; @Wasserman1994].
-The basic unit of relationship is called a dyad, defined as a connection between two distinct actors [@Lee2016].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on archaeological theories for relationships [@Mills2017; @Brughmans2018].
-This allows answering questions related to interaction and exchange, identities, and power relationships between social groups or archaeological sites by quantifying and visualizing social relations from archaeological data.
-There are two common approaches of archaeological applications, one is analyzing structures of networks built on archaeological data and finding patterns, and another is modeling theory to represent and test dependence of relationships [@Brughmans2018].
-The modeling approach is useful to statistically test hypotheses about the relationships between actors to examine the structure of network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Those approaches are widely used for different topics, such as exploring hunter-gatherer exchange network in the Kuril Islands based on ceramic compositions [@Erik2015]; long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013]; hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013].
+Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanism.
+There are two common methodological approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017]; another one is modeling theory to represent and test dependence of relationships [@Brughmans2018].
+Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns.
+Recent applications include long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013], and hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013].
+For modeling approaches, they provide a robust way to statistically test hypotheses about the relationships between actors to examine the structure of the network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
+Examples include testing hunter-gatherer exchange network across the Kuril Islands based on ceramic compositions [@Erik2015], and modeling and comparing chronological and social hypotheses to explain the differences between burials [@Sosna2012].
+This paper will use modeling approaches for the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts.
 
@@
 
-# Exponential random graph models in a Bayseisan framework
+# Exponential random graph models in a Baysesian framework
 
@@
 For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].
-To
 

### 2020-11-06 analysis/paper/004-burials-post-network-modelling.R

@@ -157,8 +157,8 @@ summary(model.2)
 #--------------------Bayesian inference for ERGMs-------------------------
 model.post.3 <- burial_network_post ~ edges +  # the overall density of the network
   #nodematch('quantity') +  # quantity-based homophily, the similarity of connected nodes
-  #nodematch('age') +
-  #nodematch('gender') +
+  nodematch('age') +
+  nodematch('gender') +
   nodematch('ritual_pottery') +
   nodematch('value_class') +
   #nodematch('orientation') +
@@ -170,8 +170,8 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(3, 6, 6) # covariance matrix structure
+prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(c(10, 3, 5, 3, 3, 3, 10, 3), 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,
@@ -179,7 +179,7 @@ post_bergm <- bergmM(model.post.3,
                  burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
                  main.iters  = 2000, # iterations for every chain of the population
                  aux.iters   = 10000, # MCMC steps used for network simulation
-                 nchains     = 12, # number of chains of the population MCMC
+                 nchains     = 16, # number of chains of the population MCMC
                  gamma       = 0) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 
 summary(post_bergm)

### 2020-11-06 analysis/paper/paper.Rmd
edits on intro and methods

@@
 <!--burial study- network analysis-->
-Burials analysis has long developed in archaeology as an approach to study past social structure through mortuary practices, such as cultural and biological records in burial contexts, that can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
-Despite the criticism for overlooking the burial manipulation by the living that does not necessarily connect to the real life [@Hodder1980; @Pearson1982], with comparisons between different archaeological contexts of death and life in mind, burials can still provide valuable information to infer past societies [@Chapman2003].
-For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference for social structures [@Byrd1995; @Seikel2011].
-To study the relations among burials, the recent development of network science provides many new tools to visualize and analyze relationships.
+Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
+The material cultures and biological records of burial behaviours can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
+Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past social organisation [@Chapman2003].
+For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
+The recent development of network science provides many new tools to visualize and analyze relationships among ancient burials.
 Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013].
-That is, a network can be viewed as a patterned aggregation that includes individual elements, pair-wise relationships (the dyads), and overall structure showing global patterning represented in data.
+That is, a network can be viewed as a patterned aggregation that includes individual elements (i.e.
+individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
 
+<!-- intro to networks -->
 A network is generally visualized as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on theories for relationships [@Mills2017; @Brughmans2018].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on theories of relationships [@Mills2017; @Brughmans2018].
 For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp.
 4].
-Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial analysis.
-Burial goods, especially for high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status that can infer social differentiation or complexity [@Gamble2002; @Janes2013].
+Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
+Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
 Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
@@
 252].
-This enables the exploration of the structure of the past social organization by identifying the relationships among burials.
+This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
 
-Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanism.
-There are two common methodological approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017]; another one is modeling theory to represent and test dependence of relationships [@Brughmans2018].
+<!-- two types of network analysis -->
+Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
+There are two common approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017].
+Another approach is using models and simulations to represent and test dependence of relationships [@Brughmans2018].
 Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns.
-Recent applications include long term inter-site relationship from the Epipalaeolithic to the early Neolithic in Near East according to trade items [@Coward2013], and hierarchical communication systems in the Kofun period in Japan through prestige goods [@Mizoguchi2013].
-For modeling approaches, they provide a robust way to statistically test hypotheses about the relationships between actors to examine the structure of the network and major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Examples include testing hunter-gatherer exchange network across the Kuril Islands based on ceramic compositions [@Erik2015], and modeling and comparing chronological and social hypotheses to explain the differences between burials [@Sosna2012].
-This paper will use modeling approaches for the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts.
+Recent applications include the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Similarly, @Mizoguchi2013 analysed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods.
+For the second approach, researchers typically statistically test hypotheses about the relationships between actors to examine the structure of the network and infer major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
+Examples include testing hunter-gatherer exchange networks across the Kuril Islands based on ceramic composition [@Erik2015].
+<!-- better to have a specific case study by a scholar who is likely to peer review this paper, try to find some recent publication --> .
+This paper will use a novel Bayesian modeling approach to investigate the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organisation.
 
-<!--archaeology case-->
-The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in those societies.
-The prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups.
-The concept of prestige goods may develop earlier before the European arrival, but could be amplified during the European colonial period.
-For example, a person with more imported goods may be treated as higher status.
-A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals.
-In this study, we use a pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.
-Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival.
-This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.
+<!-- archaeological background -->
+Archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007].
+This might indicate the pursuit of prestige or wealth in those Indigenous societies.
+Prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks that involved the Dutch, Spanish, Han Chinese traders, and local Indigenous groups.
+The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period <!-- citation -->.
+For example, an Indigenous person possessing more imported goods may have been recognised as having higher status by their community <!-- citation to some historical record -->.
+A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks <!-- citation -->.
 
-We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan.
-We viewed individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same burial goods that can represent the same social group with similar economic or political status.
-We construct a network that represents higher social inequality and tested hypotheses that burial relations becomes more hierarchical after the European arrival.
-We ask the questions: 1.
-Do the observed burial data after the European presence shows more similar network quality as our hypothesized model?
-2.
-Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks.
-3.
-What is the major variables affecting or forming the higher degree of popularity (e.g.
-a few node has more relationship with other nodes) in the European contact period that might hint at social heterogeneity during this period.
-This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
-A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.
+<!-- archaeological specific case study -->
+In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century <!-- citation -->.
+Our purpose is to examine the network structure of burials before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
+We ask: (1) does the observed burial data after the start of European presence resemble the of network our hypothesized model?
+(2) did European colonial activities in 17th century Taiwan result in the emergence of social inequality in an Indigenous society in ways that can be detected by burials networks?
+(3) what are the major variables affecting or forming the higher degree of popularity (e.g.
+a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period?
+By answering these questions, this study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
 
@@
 
-<!--current approaches to network analysis-->
-Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007].
+<!-- current approaches to network analysis: important to show how what we are doing is amazing and new and not done before by other archaeologists.
+We need a sentence or two to say what do archaeologists usually do for network statistics...
+and make it clear that we are doing something new -->
+Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
 The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008].
+ERGMs differs from traditional network approaches in that edges are only formed according to their corresponding predictors separately without taking into account the interactions among edges [@Morris2008].
 In ERGMs, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011].
-For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
 Those configurations represent the structure or the property of a network and can be expressed by network statistics.
 We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
-By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks.
-Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007].
-This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.
+By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks that represent our model.
+Such distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
+This helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
 
-ERGMs can describe a network by capturing the dependency structure by specifying parameters under assumptions and contexts, which is useful for exploring archaeological network.
-For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs.
-Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed.
-They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively.
-However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data.
-In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013].
-A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
+<!-- ERGM in archaeology, and limitations --> 
+To date, ERGM has only been used once in archaeology, by @Amati2019 who reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms.
+These network models evaluated different hypotheses about the importance of three variables, proximity, inter-cultural items, and pottery types, on the presence of hub sites they observed.
+They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
+They also pointed out some limitations of ERGMs, such as static outcomes and sensitivity to missing data.
+In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
+A normalizing constant is an important step to make probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
 Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-Those complex computational processes in ERGMs can be also resolved by implementing ERGMs in a Bayesian framework.
+One solution to these limitations is available by implementing ERGMs in a Bayesian framework.
 
-<!--Bayesian inference for Exponential Random Graph Models--> 
-Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019].
+<!-- Bayesian inference for Exponential Random Graph Models --> 
+Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporates prior information about the network configurations into the model and offers uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
 This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
-For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced with Bayesian modeling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
-To overcome these limitations in ERGMs, Bayesian modeling applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
-MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
-This alleviate above computational problems and gives a better convergence results.
-By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
-Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian].
-The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.
+For archaeological studies, missing data is often a problem leading to misinterpretation.
+@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling to provides an average of 80% correct predictions for the ties, and allow 25% false predictions with a third of data missing [@Koskinen2010] <!-- maybe a little more detail and clarification needed here -->.
+The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
+MCMC algorithms can produce approximate maximum likelihood estimators and network statistics for further specification of transitivity and heterogeneity.
+The use of MCMC enables ERGMs to deal with complicated dependence patterns with ease [@Snijders2006].
+This alleviates the above computational problems and gives a better convergence results.
+By fitting an ERMG with an MCMC algorithm, a Bayesian approach provides posterior probabilities that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
+Posterior probability enables the estimation the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
+The typical criteria for interpreting the posteriors is that an odds ratio greater than one means a positive effect of a parameter, while odds ratios less than one represents negative effect.
 
-# Burial data from Kiwulan, NE Taiwan
+# Burial data from Kiwulan, Northeastern Taiwan
 
@@
 <!--archaeological setting-->
-We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (1350-1850 AD) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007].
-The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
-In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated.
-The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison.
-Those burials are oriented in an east-west direction on the north side of the residential area indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement.
-Previous studies point out a presence of uneven distribution of prestige goods across burials without an agreement that whether this uneven distribution hints vertical social differences.
-@Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society.
-However, @Hsieh2012 suggested a relatively egalitarian based on the comparative analysis of frequency and proportion of all burial goods.
-She found that the burials with high valued burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
+We analyzed burial data collected from the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD) located in northeastern Taiwan, which experienced European colonial impacts in the early 17th century and a large wave of Chinese immigrants in the 19th century [@Chen2007].
+Excavations revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
+In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with _in-situ_ posts were also excavated.
+The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison <!-- fig reference to a map of burials at the site-->.
+Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
+Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
+For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society.
+However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods.
+She found that the burials with high value burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
+One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organiation with post-European social organisation.
+Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period.
 
+<!-- take the citations from this paragraph and add them to one of the intro paras, them delete this para -->
 The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006].
@@
 <!--chronology-->
-To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`) for network comparison based on established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon dates [@LWandBM2020ornament; @Chen2007].
-We excluded the burials from the Chinese phase (n = `r chi`) for analysis due to the smaller sample size.
-There are eight burials heavily disturbed by modern construction that are excluded to avoid misinterpreting due to distorted chronology.
-We reconstructed social relations between individuals by linking them according to similar prestige goods.
-This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp.
-252].
-Also, prestige materials usually relate to a representation of social status because of their limited accessibility that can be used as status indicators [citations].
-Therefore, we build networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
-The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments that are considered as high-value items by studies across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011] and the description in historical records [@Borao2009; @LiandWu2006].
+To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
+Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [@LWandBM2020ornament; @Chen2007].
+We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
+There are eight burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology.
+We reconstructed social relations between individuals by linking them according to similar prestige goods [cf.
+@Coward2013, pp.
+252].Therefore, we built networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
+The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments.
+Previous studies indicate that these items are considered as high-value across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+Historical accounts also support these items as prestige goods [@Borao2009; @LiandWu2006].
 
@@
 
-## data and hypothesis
+## Data and Hypothesis
 
+<!-- we should tell the reader what does it mean high-medium-low --> 
 We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
-We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here.
-We treat the presence of relations between burials based on number or types of  prestige items from regional trade network.
-where a few prestige goods controlled by a small number of individuals.
-This is because prestige items could represent a relationship, where a few prestige goods controlled by a small number of individuals.
-Based on this assumption, we treat the presence of relations between burials through the burial goods in common.
-Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we classified those beads into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts.
-That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them.
-For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common.
-Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index.
-Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods.
+Gold-foil beads and agate beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high (50-1000?), medium (20-50?), and low (1-10?), of these beads.
+That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e.
+presence or absence).
+Node attributes here include osteological data, such as age, sex, and ritual pottery, and an overall burial goods values...
+index.
+Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
+The quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
+
+<!-- stopped editing here -->
 

### 2020-11-09 analysis/paper/paper.Rmd
edit Bayesian method section and reduce iterations for modeling

@@
 Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
-The material cultures and biological records of burial behaviours can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
-Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past social organisation [@Chapman2003].
+The material cultures and biological records of burial behaviors can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
+Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
 For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
@@
 They also pointed out some limitations of ERGMs, such as static outcomes and sensitivity to missing data.
-In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
-A normalizing constant is an important step to make probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
-Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
+In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
+A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
+This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
+Model degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
 One solution to these limitations is available by implementing ERGMs in a Bayesian framework.
@@
 Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporates prior information about the network configurations into the model and offers uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
-This provides better estimations for complex social network models with heterogeneous data [@Caimo2017].
-For archaeological studies, missing data is often a problem leading to misinterpretation.
-@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling to provides an average of 80% correct predictions for the ties, and allow 25% false predictions with a third of data missing [@Koskinen2010] <!-- maybe a little more detail and clarification needed here -->.
-The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011].
-MCMC algorithms can produce approximate maximum likelihood estimators and network statistics for further specification of transitivity and heterogeneity.
-The use of MCMC enables ERGMs to deal with complicated dependence patterns with ease [@Snijders2006].
-This alleviates the above computational problems and gives a better convergence results.
-By fitting an ERMG with an MCMC algorithm, a Bayesian approach provides posterior probabilities that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
-Posterior probability enables the estimation the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
+The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
+MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
+This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
+By fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
+Posterior probability estimates the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
 The typical criteria for interpreting the posteriors is that an odds ratio greater than one means a positive effect of a parameter, while odds ratios less than one represents negative effect.
+In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
+@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data missing.
 

### 2020-11-09 analysis/paper/003-burials-pre-network-modelling.R

@@ -109,6 +109,7 @@ burial_network_pre <-
 plot(burial_network_pre, vertex.cex = 1)
 network.density(burial_network_pre)
 network.dyadcount(burial_network_pre, na.omit = TRUE)
+transitivity(burial_network_pre)
 
 #-----------------------attach attributes and make graphs------------------------------
 library(statnet)
@@ -196,9 +197,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # positive prior number for edge means high density
+prior.mean <- c(-2, 0, 0, 0, 0, 2, 2, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(3, 3, 5, 3, 3, 3, 3, 3), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(1, 3, 3, 3, 3, 1, 1, 1), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-09 analysis/paper/004-burials-post-network-modelling.R

@@ -176,9 +176,9 @@ prior.sigma <- diag(c(10, 3, 5, 3, 3, 3, 10, 3), 8, 8) # covariance matrix struc
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,
                  prior.sigma = prior.sigma,
-                 burn.in     = 200, # burn-in iterations for every chain of the population, drops the first 200
-                 main.iters  = 2000, # iterations for every chain of the population
-                 aux.iters   = 10000, # MCMC steps used for network simulation
+                 burn.in     = 100, # burn-in iterations for every chain of the population, drops the first 200
+                 main.iters  = 1000, # iterations for every chain of the population
+                 aux.iters   = 3000, # MCMC steps used for network simulation
                  nchains     = 16, # number of chains of the population MCMC
                  gamma       = 0) # scalar; parallel adaptive direction sampling move factor, acceptance rate
 

### 2020-11-09 analysis/paper/006-bergm-bootstrap.R

@@ -3,12 +3,13 @@
 
 #----------------------pre-E burials example----------------------------
 # make igraph object for pre-E
+library(igraph)
 pre_E_igraph <- graph_from_data_frame(d = edges_for_network_pre,
                                       vertices = nodes_pre,
                                       directed = FALSE)
 pre_density_obs <- graph.density(pre_E_igraph)
 pre_mean_star_obs <- mean(degree(pre_E_igraph))
-pre_trans_obs <- transitivity(pre_E_igraph)
+pre_trans_obs <- transitivity(pre_E_igraph) # clustering coefficient
 pre_all_obs <- rbind(pre_density_obs,
                      pre_mean_star_obs,
                      pre_trans_obs)

### 2020-11-11 analysis/paper/paper.Rmd
add references and citations

@@
 Recent applications include the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Similarly, @Mizoguchi2013 analysed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods.
+Similarly, @Mizoguchi2013 analyzed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods.
+Regarding burials, @Sosna examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological differences between burials.
+They found that the observed concentric pattern does not match the network results projected into geographic space, but contemporary subsections that indicates different corporate groups.
 For the second approach, researchers typically statistically test hypotheses about the relationships between actors to examine the structure of the network and infer major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Examples include testing hunter-gatherer exchange networks across the Kuril Islands based on ceramic composition [@Erik2015].
-<!-- better to have a specific case study by a scholar who is likely to peer review this paper, try to find some recent publication --> .
+Examples include accessing hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+Another example is to simulate several modes of food exchange for Ancestral Pueblos and compare the results with the social networks build on the aggregation of households in the American Southwest at a regional scale [@Crabtree2015].
 This paper will use a novel Bayesian modeling approach to investigate the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organisation.
@@
 Prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks that involved the Dutch, Spanish, Han Chinese traders, and local Indigenous groups.
-The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period <!-- citation -->.
-For example, an Indigenous person possessing more imported goods may have been recognised as having higher status by their community <!-- citation to some historical record -->.
+The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period [@Borao2009; @LiandWu2006; @LWandBM2020ornament].
+For example, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
 A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks <!-- citation -->.

### 2020-11-11 analysis/paper/003-burials-pre-network-modelling.R

@@ -109,7 +109,6 @@ burial_network_pre <-
 plot(burial_network_pre, vertex.cex = 1)
 network.density(burial_network_pre)
 network.dyadcount(burial_network_pre, na.omit = TRUE)
-transitivity(burial_network_pre)
 
 #-----------------------attach attributes and make graphs------------------------------
 library(statnet)
@@ -197,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-2, 0, 0, 0, 0, 2, 2, -1) # positive prior number for edge means high density
+prior.mean <- c(-3, 0, 0, 0, 0, 5, -3, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(1, 3, 3, 3, 3, 1, 1, 1), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(10, 10, 10, 10, 10, 5, 5, 5), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-12 analysis/paper/paper.Rmd
add references and edit Intro

@@
 For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
-The recent development of network science provides many new tools to visualize and analyze relationships among ancient burials.
-Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013].
+Based on network science and graph theories, the recent development of social network analysis provides many new tools to visualize and analyze relationships for archaeological data [@Borgatti2009; @Brughmans2013].
+Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013; @Collar2015].
 That is, a network can be viewed as a patterned aggregation that includes individual elements (i.e.
@@
 <!-- intro to networks -->
-A network is generally visualized as a graph consisting of a set of socially-relevant nodes, or actors, connected by edges representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, and co-presence of material culture based on theories of relationships [@Mills2017; @Brughmans2018].
+A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
+The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples].
 For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp.
@@
 Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
-There are two common approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017].
-Another approach is using models and simulations to represent and test dependence of relationships [@Brughmans2018].
-Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns.
-Recent applications include the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Similarly, @Mizoguchi2013 analyzed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods.
-Regarding burials, @Sosna examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological differences between burials.
-They found that the observed concentric pattern does not match the network results projected into geographic space, but contemporary subsections that indicates different corporate groups.
-For the second approach, researchers typically statistically test hypotheses about the relationships between actors to examine the structure of the network and infer major factors that affect the formation of the structure [@Freeman2004; @Salvini2010].
-Examples include accessing hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-Another example is to simulate several modes of food exchange for Ancestral Pueblos and compare the results with the social networks build on the aggregation of households in the American Southwest at a regional scale [@Crabtree2015].
-This paper will use a novel Bayesian modeling approach to investigate the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organisation.
+There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples].
+Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples].
+By quantifying those network properties, archaeologists can answer a wide range of research questions.
+Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
+Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
+Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+This paper will use a novel Bayesian modeling approach to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
@@
 For example, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
-A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks <!-- citation -->.
+A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks [@Brumfiel1994; @Clark1994].
 
 <!-- archaeological specific case study -->
-In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century <!-- citation -->.
+In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
 Our purpose is to examine the network structure of burials before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.

### 2020-11-12 analysis/paper/003-burials-pre-network-modelling.R

@@ -196,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, 0, 0, 0, 0, 5, -3, -1) # positive prior number for edge means high density
+prior.mean <- c(-2, 0, 0, 0, 0, 5, 3, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(10, 10, 10, 10, 10, 5, 5, 5), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(7, 10, 10, 10, 10, 7, 7, 5), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-13 analysis/paper/paper.Rmd
edit background section

@@
 <!-- archaeological background -->
-Archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007].
-This might indicate the pursuit of prestige or wealth in those Indigenous societies.
-Prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks that involved the Dutch, Spanish, Han Chinese traders, and local Indigenous groups.
-The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period [@Borao2009; @LiandWu2006; @LWandBM2020ornament].
-For example, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
+Archaeological sites in northeastern Taiwan show evidence of imported prestige goods, such as ornaments and porcelains, in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007; @LWandBM2020ornament].
+The use of imported goods is also mentioned in historical records from the Spanish, indicative of the pursuit of prestige or wealth in those Indigenous societies [@Borao2009; @LiandWu2006].
+For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006].
+In addition, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
+Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
+The introduction of prestige goods to local Indigenous communities might occur earlier before a European presence became established, but was amplified during the European colonial period [@Borao2009; @LiandWu2006; @LWandBM2020ornament].
 A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks [@Brumfiel1994; @Clark1994].
+The observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans.
 
@@
 In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
-Our purpose is to examine the network structure of burials before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
+We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
+We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
 We ask: (1) does the observed burial data after the start of European presence resemble the of network our hypothesized model?
@@
 By answering these questions, this study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
-
+This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
 # Exponential random graph models in a Baysesian framework
@@
 She found that the burials with high value burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
-One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organiation with post-European social organisation.
+One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization.
 Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period.
-
-<!-- take the citations from this paragraph and add them to one of the intro paras, them delete this para -->
-The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006].
-For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006].
-In addition, there was a custom that people would bury imported beads and ceramics with deceased to display their status and influence [@LiandWu2006].
-Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
 The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.
-That is, the observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans.
-We assume that the changes would be supported by chronological differences in the structure of burial networks.
-Analyzing and comparing networks between different phases relative to cultural contact will give an insight into changes in social differences.
-This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
 
@@
 We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
-Gold-foil beads and agate beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high (50-1000?), medium (20-50?), and low (1-10?), of these beads.
+Gold-foil and agate beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high (> 10), medium (> 1 and < 10), and low (= 1), of these beads.
+Glass beads are in levels of high (> 100), medium (> 99 and < 900), and low (< 100).
 That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 presence or absence).
-Node attributes here include osteological data, such as age, sex, and ritual pottery, and an overall burial goods values...
-index.
+Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index.
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium(>11 and <30), and low(<12), as an index of wealth.
+The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].

### 2020-11-13 analysis/paper/003-burials-pre-network-modelling.R

@@ -196,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-2, 0, 0, 0, 0, 5, 3, -1) # positive prior number for edge means high density
+prior.mean <- c(-17, 0, 0, 2, 0, 5, 4, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(7, 10, 10, 10, 10, 7, 7, 5), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(7, 10, 10, 5, 10, 8, 5, 5), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
@@ -208,7 +208,7 @@ prior.sigma <- diag(c(7, 10, 10, 10, 10, 7, 7, 5), 8, 8) # covariance matrix str
 pre_bergm <- bergmM(model_pre_3,
                   prior.mean  = prior.mean,
                   prior.sigma = prior.sigma,
-                  burn.in     = 100, # burn-in iterations for every chain of the population, drops the first 200
+                  burn.in     = 100, # drop first 100 for every chain of the population
                   main.iters  = 1000, # iterations for every chain of the population
                   aux.iters   = 3000, # MCMC steps used for network simulation
                   nchains     = 16, # number of chains of the population MCMC

### 2020-11-13 analysis/paper/004-burials-post-network-modelling.R

@@ -167,11 +167,12 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
   #gwnsp(1.7, fixed = TRUE) + # original 1.8
   gwdegree(0.8, fixed = TRUE) +
   dyadcov(post_distance_n, "dist")
+
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, -1, 0, 1, 1, 1, 3, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(c(10, 3, 5, 3, 3, 3, 10, 3), 8, 8) # covariance matrix structure
+prior.mean <- c(-20, 0, 0, 0, 0, 4, 8, -1) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(c(6, 8, 8, 8, 8, 8, 8, 5), 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,

### 2020-11-13 analysis/paper/paper.Rmd
add a reference for MCMC

@@
 If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
-Network analysis provides a way to examine the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.
+ERGM in a Bayesian framework provides a way to examine the formation of network ties and its underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.
 
@@
 Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects.
-We modeled several node attributes for homophily effects, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
-We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried close to each other.
+We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
+We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@@LiandWu2006].
 Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality.
@@
 After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm.
-MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
+MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution[@Hamra2013].
 We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms.

### 2020-11-13 analysis/paper/paper.Rmd
editing the text together

--- 2020-11-13 analysis/paper/paper.Rmd
@@
 Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporates prior information about the network configurations into the model and offers uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
-The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
+<!-- add brief sentence to define posterior and prior to help the reader --> The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
 MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
@@
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium(>11 and <30), and low(<12), as an index of wealth.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth.
 The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
@@
 
-<!-- stopped editing here -->
-
-Our pericolonial impact hypothesis is that restricted grave goods or accumulation of goods present in a few individuals after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access high value goods.
-The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
-In addition to the interest of burial goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power.
-If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
-Network analysis provides a way to examine the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.
+<!-- our hypothesis --> 
+Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods.
+We can identify these high value trade goods had been from historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
+A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power.
+If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
+To test this prediction, we use ERGM in a Bayesian framework to examine the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
+By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
 
@@
 ```{r ergm-terms-table}
-ergm_terms <- readxl::read_excel(here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
+ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
 kableExtra::kable(ergm_terms,
@@
 
-<!--Network construction and specification-->
-We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language with the bergm package [@Caimo2014], which built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
-Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations.
+<!-- Network construction and specification -->
+We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language <!-- cite R --> with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
+Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
 Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
-Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects.
-We modeled several node attributes for homophily effects, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall.
-We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried close to each other.
-Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality.
-We expect that observed network after the European arrival will have a better match with our model compared with the network before European contact.
+Based on our hypothesis, we model a network with increased social inequality that is represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects.
+We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes to form relationships and the network ties overall.
+We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
+Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
+This would be indicated by high popularity or degree values in the network statistics.
+According to our hypothesis, the burial evidence from after the European arrival will have a better fit with our model compared with the burial evidence from before European contact.
+
+```{r}
+burn_in <- 200
+...
+```
 
 <!--Model fitting and prior setting--> 
-After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm.
-MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution.
-We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms.
-This enables a random or stochastic simulation in a long run without depending on the previous move.
-More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains.
-In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011].
+After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
+MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
+We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm.
+This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.
+More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
+In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
 In our case, we set the number of chains to six.
-For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000.
+For each chain, the number of burn-in iterations was 200 `r burn_in` and the number of iterations after the burn-in was 2000.
 We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
-For prior, the normal distribution is agreed as a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly found in real world [@Caimo2017].
+We use the normal distribution for the priors, as is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
+
+<!-- add text about anthropological assumptions of our prior values -->
+
+<!-- stopped editing here -->
 
 <!-- Goodness-of-fit assessment-->
-The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017].
-The Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+The output from our two were models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of the model [@Caimo2011; @Caimo2017].
+Our Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
 This provides a statistical approach to check how well the estimated posterior parameter distribution based on our hypotheses can reproduce networks similar to the general structural features of the observed network.

### 2020-11-14 analysis/paper/paper.Rmd
add citation, inline code, and reproducibility section

@@
 <!-- Network construction and specification -->
-We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language <!-- cite R --> with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
+We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
 Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
@@
 burn_in <- 200
+main_iter <- 1000
+aux_iter <- 10000
 ...
@@
 In our case, we set the number of chains to six.
-For each chain, the number of burn-in iterations was 200 `r burn_in` and the number of iterations after the burn-in was 2000.
-We set the number of iterations used to simulate a network y′ at each iteration to 10,000.
+For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`.
+We set the number of iterations used to simulate a network y′ at each iteration to `r aux_iter`.
 We use the normal distribution for the priors, as is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
@@
 We expect to see the network after the European presence has a better fit with the hypothesis, indicative of a structure with higher social differentiation.
+
+# Reproducibility and open source materials
+
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
+All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
+The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.
 

### 2020-11-15 analysis/paper/paper.Rmd
work on para of prior specification

@@
 Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
-Based on our hypothesis, we model a network with increased social inequality that is represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects.
-We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes to form relationships and the network ties overall.
+Based on our hypothesis, we model a network with increased social inequality that is represented by low transitivity and high popularity dyadic effects.
+We include nodal covariate effects for homophily, including age, gender, ritual activity, and the degree of wealth, to test whether nodes with similar attributes tend to form relationships presented by the network ties overall.
 We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
 Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics.
-According to our hypothesis, the burial evidence from after the European arrival will have a better fit with our model compared with the burial evidence from before European contact.
+This would be indicated by high popularity or degree values in the network statistics, with covariates showing the underlying mechanisms.
+According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
 
@@
 
-<!--Model fitting and prior setting--> 
+<!--Model fitting--> 
 After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
@@
 We set the number of iterations used to simulate a network y′ at each iteration to `r aux_iter`.
-We use the normal distribution for the priors, as is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-
-<!-- add text about anthropological assumptions of our prior values -->
 
@@
 
+<!-- add text about prior setting and anthropological assumptions of our prior values --> 
+Normal distribution for the priors is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
+Thus, we specified the priors as low density for both network models.
+For covariate based on nodal attributes, such as age, gender, and wealth index, we set the priors to follow a normal distribution with mean 0 and standard deviation 5 (i.e.
+N(0, 5)) for both models.
+We set a negative covariate effect based on physical distance between burials (N(-1, 5)).
+To evaluate our anthropological assumption of increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for examining social inequality, especially for gwesp and gwedegree.
+For example, we set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on the ritual (N(2, 5)) for network before European contact to indicate less social inequality.
+On the contrary, we set the prior for the network after the European arrival to lower transitivity (N(4, 8)), and higher popularity (N(8, 8)) to illustrate an increased social inequality.
+This prior information derives from the theory about the social structure of horizontal hierarchy, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate model at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
+The priors for density were also set differently with N(-17, 7) for earlier network and N(-20, 6) for later network.
+We should note that the mean priors and standard deviation priors indicate relative differences for comparison.
+
 <!-- Goodness-of-fit assessment-->
-The output from our two were models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of the model [@Caimo2011; @Caimo2017].
+The output from our two  models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of the model [@Caimo2011; @Caimo2017].
 Our Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.

### 2020-11-15 R/000-trial-resampling-network.R

@@ -1,5 +1,5 @@
 library(ergm.ego)
-
 # create edodata object
 as.egodata(burial_network_pre)
 
+install.packages("~/Downloads/hergm_4.1-6 (1).tar.gz", repos = NULL, type = "source")

### 2020-11-15 analysis/paper/003-burials-pre-network-modelling.R

@@ -198,7 +198,7 @@ summary(model_pre_3)
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
 prior.mean <- c(-17, 0, 0, 2, 0, 5, 4, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(8, 9, 9, 5, 7, 8, 5, 5), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(7, 7, 7, 5, 7, 7, 7, 5), 8, 8) # covariance matrix structure
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) as a suitable prior model for the model parameters of interests
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-16 analysis/paper/paper.Rmd
edit assessments for MCMC and work on para for results

@@
 Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
-The material cultures and biological records of burial behaviors can represent the social status or identity of the deceased [@Binford1971; @Saxe1970].
+The material cultures and biological records of burial behaviors can represent the social ranking or identity of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
 Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
@@
 This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
+
 # Exponential random graph models in a Baysesian framework
@@
 Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
-Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008].
+Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
 Based on our hypothesis, we model a network with increased social inequality that is represented by low transitivity and high popularity dyadic effects.
-We include nodal covariate effects for homophily, including age, gender, ritual activity, and the degree of wealth, to test whether nodes with similar attributes tend to form relationships presented by the network ties overall.
+We include burial-specific attributes as covariate effects for homophily, such as age, gender, ritual activity, and the degree of wealth, to test whether burials with similar attributes tend to form relationships.
+For example, age-homophily means people of the same age to have the same burial goods.
 We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
 Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics, with covariates showing the underlying mechanisms.
+This would be indicated by high popularity or degree values in the network statistics, with covariates to control the preferential tendency of formation of relationship.
 According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
@@
 Normal distribution for the priors is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-Thus, we specified the priors as low density for both network models.
-For covariate based on nodal attributes, such as age, gender, and wealth index, we set the priors to follow a normal distribution with mean 0 and standard deviation 5 (i.e.
+Thus, we specified the prior of edge to low density for both network models.
+For covariates based on burial attributes, such as age, gender, and wealth index, we set the priors to follow a normal distribution with mean 0 and standard deviation 5 (i.e.
 N(0, 5)) for both models.
-We set a negative covariate effect based on physical distance between burials (N(-1, 5)).
-To evaluate our anthropological assumption of increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for examining social inequality, especially for gwesp and gwedegree.
-For example, we set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on the ritual (N(2, 5)) for network before European contact to indicate less social inequality.
+For physical distance between burials, we set a negative covariate effect (N(-1, 5)) to infer the kinship-based proximity.
+To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for gwesp and gwedegree representing transitivity and popularity.
+We set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on the ritual (N(2, 5)) for network before European contact to indicate less social inequality.
 On the contrary, we set the prior for the network after the European arrival to lower transitivity (N(4, 8)), and higher popularity (N(8, 8)) to illustrate an increased social inequality.
-This prior information derives from the theory about the social structure of horizontal hierarchy, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate model at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
+This prior information derives from the theory about horizontal hierarchy, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
 The priors for density were also set differently with N(-17, 7) for earlier network and N(-20, 6) for later network.
-We should note that the mean priors and standard deviation priors indicate relative differences for comparison.
+It should be noted the idea of those settings is to reflect the relative differences between two networks, and the standard deviation was adjusted in terms of different networks to indicate uncertainty.
 
-<!-- Goodness-of-fit assessment-->
-The output from our two  models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of the model [@Caimo2011; @Caimo2017].
+<!-- assessment of MCMC and Goodness-of-fit in Bayesian-->
+To assess MCMC analysis, we first evaluate three diagnostics presented as visual summaries, including density plots, trace plots, and autocorrelation plots [@Hamra2013].
+Those plots provide an informal way to diagnose model convergence by identifying any unexpected peaks or shapes in distribution.
+Then we access output from our two  models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
 Our Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
 This provides a statistical approach to check how well the estimated posterior parameter distribution based on our hypotheses can reproduce networks similar to the general structural features of the observed network.
-We then compared the distribution of our observed networks, pre-European and post-European networks, with the distribution of our hypothesized model with higher popularity.
-We expect to see the network after the European presence has a better fit with the hypothesis, indicative of a structure with higher social differentiation.
+We then compared the distribution of our observed networks, networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
+We expect to see the models fit with our hypotheses, indicative of a increased social inequality after the foreign contact.
+In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network.
 
@@
 
-The pre-European network shows three subgroups that possessing the same types of burial goods: medium amount of carnelian bead on the right, medium amount of gold-foil beads on the left, and porcelains in the middle.
-Although burials with different amount of burial goods can be identified, there is no clear clusters demonstrating the sharing of prestige goods only for burials with rich goods.
-The post-European network also presented subgroups with more edges connecting burials.
+We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure.
+For the earlier network, the covariates having significant tendency to form relations between burials includes age-homophily and ritual-homophily.
+The age homophiligy shows negative effect that that indicates burials with the same age does not share similar prestige goods, but different age does instead.
+The positive value of ritual-homophily shows burial with riutal pottery also to form relations.
+Despite other covariates are positive such as sex and burial values, they does not show significant tendency to form relationships.
 

### 2020-11-16 analysis/paper/paper.Rmd
work on result section

@@
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
-Based on our hypothesis, we model a network with increased social inequality that is represented by low transitivity and high popularity dyadic effects.
+Based on our hypothesis, we model a network with increased social inequality that is represented by endogenous network effects, low transitivity and high popularity.
 We include burial-specific attributes as covariate effects for homophily, such as age, gender, ritual activity, and the degree of wealth, to test whether burials with similar attributes tend to form relationships.
@@
 
-We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure.
-For the earlier network, the covariates having significant tendency to form relations between burials includes age-homophily and ritual-homophily.
-The age homophiligy shows negative effect that that indicates burials with the same age does not share similar prestige goods, but different age does instead.
-The positive value of ritual-homophily shows burial with riutal pottery also to form relations.
-Despite other covariates are positive such as sex and burial values, they does not show significant tendency to form relationships.
+We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure (a table here).
+For the former network, the nodal covariates, age-homophily and ritual-homophily, have significant effects on the relations between burials.
+The age homophiligy shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
+Despite other covariates are positive, such as sex-homophily and burial value-homophily, they do not show a significant tendency to forming relations duo to value of zero in a confidence interval.
+Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly.
+For the endogenous network effects, both gwesp and gwdegree present significant positive effects.
+The positive value of gwesp suggests a tendency of burials to be clustered in closed transitive structures, while positive gwdegree indicates a tendency toward centralization that means the presence of a limited number of burials with more relations.
+
+For the network after the arrival of Europeans, there are no significant effects of nodal and dyadic covariates, but we still have a clue of possible tendency.
+The nodal covariates of age-homophily, sex-homophily, ritual-homophily, and dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect.
+The interesting thing is that the burials in the same wealth level may tend not to form relations.
+Similar to the network before the European arrival, the endogenous network variables, gwesp and gwdegree, all demonstrate positive effects significantly.
+However, there are differences in values of positive effects between two networks.
+The later network has a smaller gwesp effect and a larger gwedegree effect that may suggest a less tendency toward clustering but more tendency toward centralization after the European presence.
 

### 2020-11-17 analysis/paper/paper.Rmd
work on network comparison para and adjust the setting for ties

@@
 
-```{r}
+```{r bergm-iteration}
 burn_in <- 200
@@
 aux_iter <- 10000
-...
 ```
@@
 
-We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure (a table here).
+```{r BERGMs}
+```
+
+<!--posterior stats for pre-E-->
+We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure (Table XX).
 For the former network, the nodal covariates, age-homophily and ritual-homophily, have significant effects on the relations between burials.
@@
 
+<!--posterior stats for post-E and comparison-->
 For the network after the arrival of Europeans, there are no significant effects of nodal and dyadic covariates, but we still have a clue of possible tendency.
@@
 The later network has a smaller gwesp effect and a larger gwedegree effect that may suggest a less tendency toward clustering but more tendency toward centralization after the European presence.
+Both posterior estimates presents symmetric distributions because of the posterior means are close to posterior medians (Table XX).
 
-## ERGM parameter estimation
-<!--interpretation of network plots across phases-->
+```{r vertex-bootstrap}
+boostrap <- 1000
+```
+<!--comparison between two networks using bootstrap--> 
+To understand whether the differences in posterior distributions is affected by different network size, we conducted sensitivity test using vertex bootstrap inference on networks that allows comparison for different small networks [@Chen2019].
+Vertex bootstrap is a non-parametric method that conducts resampling of all vertices (i.e.
+node) of a network to quantify standard errors in network statistics estimation of interest.
+This enables the evaluation of uncertainty for networks and the comparison between two networks by examining the confidence interval for network population.
+We set the resampling size to 1,000 for network statistics, including density, clustering, and degree, for two network.
+The 95% confidence interval of each network statistic contains the observed network for both period, which indicate less sample size effects on our models.
 
-```{r network-ERGMs}
-```
-
-<!--interpretation of Bayesian modeling-->
-
-```{r Bayesian-ERGMs}
-```
-
-## Structure Effects
-
-The network statistics presents each θ that corresponds to the parameter specified in ERGM previously.
-In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.
-
-θ1 = number of ties
-θ2 = individuals with the same abundance of burial goods
-θ3 = gwesp, a positive sign suggests a tendency of burials in our sample to group together in closed transitive structures.
-In mortuary context, clustering implies a tendency of burials with the same status indicated by prestige goods to be organized in closed structures.
-Burials with multiple prestige goods in common are more likely to be directed connected.
-θ4 = gwdegree, negative estimates reflect an increased likelihood on ties to higher-degree nodes, or the strong edges are not necessarily centralised or dispersed in the degree distribution.
-
-The plots from left to right show estimated marginal posterior densities, trace plots, and auto correlation plots.
+<!--assessment of MCMC analysis and BGOF-->
 

### 2020-11-17 analysis/paper/001-data-tidy.R

@@ -40,9 +40,9 @@ burial_three_period_age_tidy <-
   mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
          Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
          Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
-         Agate_bead_low = ifelse(Agate_bead == 1, 1, NA),
-         Agate_bead_med = ifelse(Agate_bead > 1 & Agate_bead <10, 1, NA),
-         Agate_bead_high = ifelse(Agate_bead > 10, 1, NA),
+         Agate_bead_low = ifelse(Agate_bead < 3, 1, NA), # set low level to 2
+         Agate_bead_med = ifelse(Agate_bead > 2 & Agate_bead <10, 1, NA),
+         Agate_bead_high = ifelse(Agate_bead > 9, 1, NA), # set high level to 10
          `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` < 100, 1, NA),
          `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
          `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% #based on the result of histogram

### 2020-11-19 analysis/paper/paper.Rmd
adjust network ties and work on result section

@@
 Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
-There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples].
-Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples].
+There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
+Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
 By quantifying those network properties, archaeologists can answer a wide range of research questions.
@@
 Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporates prior information about the network configurations into the model and offers uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
-<!-- add brief sentence to define posterior and prior to help the reader --> The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
+In Bayesian analysis, posterior distribution is an updated probability distribution after combining the prior derived from previous data or assumptions and our data.
+The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
 MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
@@
 Thus, we specified the prior of edge to low density for both network models.
-For covariates based on burial attributes, such as age, gender, and wealth index, we set the priors to follow a normal distribution with mean 0 and standard deviation 5 (i.e.
+For covariates based on burial attributes, such as age, gender, and wealth index, we specified a vague prior information that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
 N(0, 5)) for both models.
@@
 Our Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+We set of 10000 network graphs simulated from the estimated posterior distribution in ERGMs.
 This provides a statistical approach to check how well the estimated posterior parameter distribution based on our hypotheses can reproduce networks similar to the general structural features of the observed network.
@@
 
-<!--posterior stats for post-E and comparison-->
+<!--posterior stats for for post-E and their comparison-->
 For the network after the arrival of Europeans, there are no significant effects of nodal and dyadic covariates, but we still have a clue of possible tendency.
@@
 ```
+
 <!--comparison between two networks using bootstrap--> 
-To understand whether the differences in posterior distributions is affected by different network size, we conducted sensitivity test using vertex bootstrap inference on networks that allows comparison for different small networks [@Chen2019].
-Vertex bootstrap is a non-parametric method that conducts resampling of all vertices (i.e.
-node) of a network to quantify standard errors in network statistics estimation of interest.
-This enables the evaluation of uncertainty for networks and the comparison between two networks by examining the confidence interval for network population.
-We set the resampling size to 1,000 for network statistics, including density, clustering, and degree, for two network.
-The 95% confidence interval of each network statistic contains the observed network for both period, which indicate less sample size effects on our models.
+To understand whether the differences in posterior distributions are affected by different network sizes, we conducted a sensitivity test using vertex bootstrap inference on networks that allows a comparison for multiple networks of different sizes [@Chen2019].
+Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
+node) of a network to quantify standard errors in the network statistics estimation of interest.
+This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population.
+We set the resampling size to 1,000 for network statistics, including density, clustering, and degree, for two networks.
+The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our network is less sensitive to sample size.
+This suggests a limited sample size effect on both models.
+In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
+In general, there are no significant changes in the posterior estimates using different sets of priors.
+But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
+
+```{r Bayesian-gof}
+# the grey lines mark the 95% intervals. X-axis means the proportion of nodes
+```
 
 <!--assessment of MCMC analysis and BGOF--> 
-
-```{r Bayesian-gof}
-```
-
-Figure shows the Bayesian goodness of fit diagnostics plots where the red lines represent the distributions of the observed data and the box plots represent the GOF distributions of 10000 network graphs simulated from ERGMs based on the estimated posterior distribution.
-The grey lines mark the 95% intervals.
-X-axis means the proportion of nodes
+The Bayesian goodness of fit diagnostics plots allows an evaluation of our model fit by comparing our observed data with the simulated distributions.
+We compared two models based on diagnostic plots that a good fit to estimated ERGM is indicated by observed data falls inside the 95% intervals.
 

### 2020-11-19 analysis/paper/003-burials-pre-network-modelling.R

@@ -196,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, 0, 0, 1, 0, 5, 2, -1) # positive prior number for edge means high density
+prior.mean <- c(0, 0, 0, 2, 0, 3, 2, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(1, 5, 5, 3, 5, 3, 3, 3), 8, 8) # covariance matrix structure, uncertainty
+prior.sigma <- diag(c(3, 3, 3, 3, 3, 5, 5, 3), 8, 8) # covariance matrix structure, uncertainty
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) a common prior model
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.
@@ -256,19 +256,19 @@ posterior_distribution_pre <-
              "Observed.degree" = round(c(obs_degree_pre$mean,
                                   obs_degree_pre$sd,
                                   obs_degree_pre$skew), 2),
-             "Model.degree" = round(c(mean(sim_degree_pre$mean),
+             "Modelled.degree" = round(c(mean(sim_degree_pre$mean),
                                mean(sim_degree_pre$sd),
                                mean(sim_degree_pre$skew)),2),
              "Observed.distance" = round(c(obs_dist_pre$mean,
                                            obs_dist_pre$sd,
                                            obs_dist_pre$skew), 2),
-             "Model.distance" = round(c(mean(sim_dist_pre$mean),
+             "Modelled.distance" = round(c(mean(sim_dist_pre$mean),
                                         mean(sim_dist_pre$sd),
                                         mean(sim_dist_pre$skew)),2),
              "Observed.esp" = round(c(obs_esp_pre$mean,
                                       obs_esp_pre$sd,
                                       obs_esp_pre$skew), 2),
-             "Model.esp" = round(c(mean(sim_esp_pre$mean),
+             "Modelled.esp" = round(c(mean(sim_esp_pre$mean),
                                    mean(sim_esp_pre$sd),
                                    mean(sim_esp_pre$skew)),2),
              "Phase" = "pre-European")

### 2020-11-19 analysis/paper/004-burials-post-network-modelling.R

@@ -222,19 +222,19 @@ posterior_distribution_post <-
              "Observed.degree" = round(c(obs_degree_post$mean,
                                   obs_degree_post$sd,
                                   obs_degree_post$skew), 2),
-             "Model.degree" = round(c(mean(sim_degree_post$mean),
+             "Modelled.degree" = round(c(mean(sim_degree_post$mean),
                                mean(sim_degree_post$sd),
                                mean(sim_degree_post$skew)),2),
              "Observed.distance" = round(c(obs_dist_post$mean,
                                            obs_dist_post$sd,
                                            obs_dist_post$skew), 2),
-             "Model.distance" = round(c(mean(sim_dist_post$mean),
+             "Modelled.distance" = round(c(mean(sim_dist_post$mean),
                                         mean(sim_dist_post$sd),
                                         mean(sim_dist_post$skew)),2),
              "Observed.esp" = round(c(obs_esp_post$mean,
                                       obs_esp_post$sd,
                                       obs_esp_post$skew), 2),
-             "Model.esp" = round(c(mean(sim_esp_post$mean),
+             "Modelled.esp" = round(c(mean(sim_esp_post$mean),
                                    mean(sim_esp_post$sd),
                                    mean(sim_esp_post$skew)),2),
              "Phase" = "post-European")

### 2020-11-19 analysis/paper/005-bergm-distribution-stats.R

@@ -4,7 +4,7 @@ colnames(pre_bergm_stats) <- pre_bergm$specs
 
 pre_bergm_stats <-
   cbind(pre_bergm_stats, phase = "pre") %>%
-  rename("gwesp" = "gwesp.fixed.0.8") %>%  #gwnsp" = "gwnsp.fixed.0.8
+  rename("gwesp" = "gwesp.fixed.0.7") %>%  #gwnsp" = "gwnsp.fixed.0.7
   rename("dyadcov.dist" = "dyadcov.pre_distance_n.dist")
 
 # get data from post-European model
@@ -58,26 +58,21 @@ distribution_two_phases <-
 
 distribution_two_phases_longer <-
   distribution_two_phases %>%
-  pivot_longer(cols = starts_with(c("Observed", "Model")),
+  pivot_longer(cols = starts_with(c("Observed", "Modelled")),
                names_to = "parameter",
-               values_to = "number") %>%
-  separate(parameter, c("value", "parameter"))
+               values_to = "value") %>%
+  separate(parameter, c("data", "parameter"))
 
 distribution_two_phases_diff <-
   distribution_two_phases %>%
-  mutate(degree = abs(Observed.degree-Model.degree),
-         distance = abs(Observed.distance-Model.distance),
-         esp = abs(Observed.esp-Model.esp)) %>%
+  mutate(degree = abs(Observed.degree-Modelled.degree),
+         distance = abs(Observed.distance-Modelled.distance),
+         esp = abs(Observed.esp-Modelled.esp)) %>%
   pivot_longer(cols = c("degree", "distance", "esp"),
                names_to = "parameter",
                values_to = "difference")
 
 ggplot(distribution_two_phases_longer,
-       aes(parameter, number)) +
-  ggpointgrid::geom_pointgrid(aes(color = Phase, shape = value), size = 3) +
-  facet_wrap(~moments, scales = "free")
-
-ggplot(distribution_two_phases_diff,
-       aes(parameter, difference)) +
-  geom_point(aes(color = Phase), size = 3) +
+       aes(parameter, value)) +
+  ggpointgrid::geom_pointgrid(aes(color = Phase, shape = data), size = 3) +
   facet_wrap(~moments, scales = "free")

### 2020-11-20 analysis/paper/paper.Rmd
work on BGOF paragraph and add code for classes of goods

@@
 <!--assessment of MCMC analysis and BGOF--> 
-The Bayesian goodness of fit diagnostics plots allows an evaluation of our model fit by comparing our observed data with the simulated distributions.
-We compared two models based on diagnostic plots that a good fit to estimated ERGM is indicated by observed data falls inside the 95% intervals.
+Bayesian goodness of fit (GOF) diagnostics plots show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+the number of edges between node pairs in a shortest path) and the degree distribution.
+For edgewise shared partner (gwesp) distribution, despite some observations falls outside the 95% interval, the fitting is generally good with most observations within it.
+This also suggests that our observed network statistic does not follow a normal distribution, especially for the network after the arrival of the Europeans that demonstrate a bimodal pattern.
+We also compared the first three distribution moments of each observed distributions and their corresponding simulated distributions represented by mean.
+The figure shows that modeled values from pre-Euroepan network are closer to the observed values overall, compared to the values from the post-European network.
+For the variance of degree, the values from modeled post-European distribution are closer to the observed distribution.
+In general, this demonstrates a better fit for the pre-European network.
+However, we should note that the differences between the statistics of the two models are relatively small, which allows direct comparison of two networks.
 

### 2020-11-20 analysis/paper/000-prep2-goods-class-plot.R

@@ -59,6 +59,24 @@ burial_three_period_tidy_explore %>%
   geom_histogram() +
   scale_x_log10()
 
+# long format for bead
+burial_beads <-
+  burial_three_period_tidy_explore %>%
+  select(burial_label, Agate_bead, Golden_bead, Glass_bead, `Indo-Pacific_bead`) %>%
+  rowwise() %>%
+  mutate(combined_glass_bead = sum(Glass_bead, `Indo-Pacific_bead`, na.rm = TRUE)) %>%
+  select(Agate_bead, Golden_bead, combined_glass_bead) %>%
+  pivot_longer(col = ends_with("_bead"),
+               names_to = "type",
+               values_to = "value")
+
+# boxplot for all beads
+burial_beads %>%
+  ggplot(aes(type, value)) +
+  geom_boxplot() +
+  coord_flip() +
+  scale_y_log10()
+
 # orientation
 burial %>%
   mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%

### 2020-11-20 analysis/paper/003-burials-pre-network-modelling.R

@@ -196,9 +196,9 @@ summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(0, 0, 0, 2, 0, 3, 2, -1) # positive prior number for edge means high density
+prior.mean <- c(-2, 0, 0, 2, 0, 3, 2, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(3, 3, 3, 3, 3, 5, 5, 3), 8, 8) # covariance matrix structure, uncertainty
+prior.sigma <- diag(c(7, 3, 3, 3, 3, 5, 5, 3), 8, 8) # covariance matrix structure, uncertainty
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) a common prior model
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2020-11-20 analysis/paper/004-burials-post-network-modelling.R

@@ -191,9 +191,9 @@ plot(post_bergm)
 bgof_post <-
   bgof(post_bergm,
        aux.iters = 10000,
-       n.deg     = 40,
+       n.deg     = 30,
        n.dist    = 15,
-       n.esp     = 35)
+       n.esp     = 30)
 
 summary(bgof_post)
 

### 2020-11-20 analysis/paper/paper.Rmd
edited and fixed typos in the result section

@@
 <!-- assessment of MCMC and Goodness-of-fit in Bayesian-->
-To assess MCMC analysis, we first evaluate three diagnostics presented as visual summaries, including density plots, trace plots, and autocorrelation plots [@Hamra2013].
+To assess MCMC analysis, we first evaluated three diagnostics visual summaries, including density plots, trace plots, and autocorrelation plots [@Hamra2013].
 Those plots provide an informal way to diagnose model convergence by identifying any unexpected peaks or shapes in distribution.
@@
 The age homophiligy shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
-Despite other covariates are positive, such as sex-homophily and burial value-homophily, they do not show a significant tendency to forming relations duo to value of zero in a confidence interval.
+Despite positive value from some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to forming relations duo to value of zero in a confidence interval.
 Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly.
@@
 The nodal covariates of age-homophily, sex-homophily, ritual-homophily, and dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect.
-The interesting thing is that the burials in the same wealth level may tend not to form relations.
-Similar to the network before the European arrival, the endogenous network variables, gwesp and gwdegree, all demonstrate positive effects significantly.
+We also found that the burials in the same wealth level may tend not to form relations.
+Similar to the earlier network, the endogenous network variables, gwesp and gwdegree, all demonstrate positive effects significantly.
 However, there are differences in values of positive effects between two networks.
@@
 Bayesian goodness of fit (GOF) diagnostics plots show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
-the number of edges between node pairs in a shortest path) and the degree distribution.
+the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
 For edgewise shared partner (gwesp) distribution, despite some observations falls outside the 95% interval, the fitting is generally good with most observations within it.
-This also suggests that our observed network statistic does not follow a normal distribution, especially for the network after the arrival of the Europeans that demonstrate a bimodal pattern.
-We also compared the first three distribution moments of each observed distributions and their corresponding simulated distributions represented by mean.
-The figure shows that modeled values from pre-Euroepan network are closer to the observed values overall, compared to the values from the post-European network.
+This also suggests that our observed network statistics do not follow a normal distribution, especially for the network after the arrival of the Europeans that presents a bimodal pattern.
+We also compared the first three distribution moments of each observed distribution and their corresponding simulated distribution represented by mean.
+The figure shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
 For the variance of degree, the values from modeled post-European distribution are closer to the observed distribution.
 In general, this demonstrates a better fit for the pre-European network.
-However, we should note that the differences between the statistics of the two models are relatively small, which allows direct comparison of two networks.
+However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.
 

### 2020-11-20 analysis/paper/paper.Rmd
editing methods and results

@@
 Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods.
-We can identify these high value trade goods had been from historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
+We can identify these high value trade goods from historical documents written by Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
 A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power.
@@
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
-Normal distribution for the priors is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-Thus, we specified the prior of edge to low density for both network models.
-For covariates based on burial attributes, such as age, gender, and wealth index, we specified a vague prior information that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
+Normal distributions for the priors are typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
+Thus, we specified the prior of the edge density parameter to low for both network models.
+For covariates based on burial attributes, such as age, gender, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
 N(0, 5)) for both models.
-For physical distance between burials, we set a negative covariate effect (N(-1, 5)) to infer the kinship-based proximity.
-To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for gwesp and gwedegree representing transitivity and popularity.
-We set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on the ritual (N(2, 5)) for network before European contact to indicate less social inequality.
-On the contrary, we set the prior for the network after the European arrival to lower transitivity (N(4, 8)), and higher popularity (N(8, 8)) to illustrate an increased social inequality.
-This prior information derives from the theory about horizontal hierarchy, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
-The priors for density were also set differently with N(-17, 7) for earlier network and N(-20, 6) for later network.
-It should be noted the idea of those settings is to reflect the relative differences between two networks, and the standard deviation was adjusted in terms of different networks to indicate uncertainty.
+For physical distance between burials, we set a negative covariate effect (N(-1, 5)) to infer kinship-based proximity, e.g.
+stronger correlations for shorter distances.
+To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
+We set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on ritual activity (N(2, 5)) for network before European contact to indicate less social inequality.
+On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(4, 8)), and higher popularity (N(8, 8)) to  model an increased social inequality.
+This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
+The priors for density were also set differently with N(-17, 7) for pre-European network and N(-20, 6) for post-European network.
 
 <!-- assessment of MCMC and Goodness-of-fit in Bayesian-->
-To assess MCMC analysis, we first evaluated three diagnostics visual summaries, including density plots, trace plots, and autocorrelation plots [@Hamra2013].
-Those plots provide an informal way to diagnose model convergence by identifying any unexpected peaks or shapes in distribution.
-Then we access output from our two  models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
-Our Bayesian GOF summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
-We set of 10000 network graphs simulated from the estimated posterior distribution in ERGMs.
-This provides a statistical approach to check how well the estimated posterior parameter distribution based on our hypotheses can reproduce networks similar to the general structural features of the observed network.
-We then compared the distribution of our observed networks, networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
+As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
+Then we summarise output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
+We set 10,000 network graphs simulated from the estimated posterior distribution in ERGMs.
+Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
+We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
 We expect to see the models fit with our hypotheses, indicative of a increased social inequality after the foreign contact.
-In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network.
+In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
 
@@
 
-The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/xxx/xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
@@
 <!--posterior stats for pre-E-->
-We compared estimated statistics from the posterior distributions, networks before and after the European presence, to examine their differences in network structure (Table XX).
-For the former network, the nodal covariates, age-homophily and ritual-homophily, have significant effects on the relations between burials.
-The age homophiligy shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
-Despite positive value from some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to forming relations duo to value of zero in a confidence interval.
-Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly.
-For the endogenous network effects, both gwesp and gwdegree present significant positive effects.
-The positive value of gwesp suggests a tendency of burials to be clustered in closed transitive structures, while positive gwdegree indicates a tendency toward centralization that means the presence of a limited number of burials with more relations.
+We compared estimated statistics from the posterior distributions to examine their differences in  structure of the simulated networks (Table XX).
+For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the relations between burials.
+The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
+Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals.
+Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations.
+For the endogenous network effects, both transitivity (gwesp) and popularity (gwdegree) present significant positive effects.
+The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, while positive popularity indicates a tendency toward centralization that means the presence of a limited number of burials with many more relations than others.
+<!-- what is the social anthropological implication for these values -->
 
 <!--posterior stats for for post-E and their comparison-->
-For the network after the arrival of Europeans, there are no significant effects of nodal and dyadic covariates, but we still have a clue of possible tendency.
-The nodal covariates of age-homophily, sex-homophily, ritual-homophily, and dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect.
-We also found that the burials in the same wealth level may tend not to form relations.
-Similar to the earlier network, the endogenous network variables, gwesp and gwdegree, all demonstrate positive effects significantly.
-However, there are differences in values of positive effects between two networks.
-The later network has a smaller gwesp effect and a larger gwedegree effect that may suggest a less tendency toward clustering but more tendency toward centralization after the European presence.
-Both posterior estimates presents symmetric distributions because of the posterior means are close to posterior medians (Table XX).
+For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect.
+However, there are no significant effects for these nodal and dyadic covariates.
+We also found that the burials in the same wealth level tend not to form relations.
+Similar to the pre-European network, the endogenous network variables, transitivity (gwesp) and popularity (gwdegree), all demonstrate significant positive effects.
+The post-European network has a smaller transitivity effect and a larger popularity effect than the pre-European network.
+This may suggest a reduced tendency toward clustering but higher tendency toward centralization after the European presence.
+Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).
 
@@
 
-<!--comparison between two networks using bootstrap--> 
-To understand whether the differences in posterior distributions are affected by different network sizes, we conducted a sensitivity test using vertex bootstrap inference on networks that allows a comparison for multiple networks of different sizes [@Chen2019].
+<!-- comparison between two networks using patchwork bootstrap --> 
+One key difference between the pre-European and post-European networks is their size, with x burials compared to y burials.
+To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks.
+This allows a comparison for multiple networks of different sizes [@Chen2019].
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) of a network to quantify standard errors in the network statistics estimation of interest.
+node) of a network to quantify standard errors in the network statistics of interest.
 This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population.
-We set the resampling size to 1,000 for network statistics, including density, clustering, and degree, for two networks.
-The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our network is less sensitive to sample size.
-This suggests a limited sample size effect on both models.
+We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks.
+The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our networks are relatively robust to differences in sample size.
+This suggests a very small sample size effect on both models.
 In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
@@
 But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
+<!-- re-write wither to justify why vertex bootstrap is the right method to answer this question, or use patchwork bootstrap and justify that-->
+
+<!-- stopped editing here -->
 

### 2020-11-22 analysis/paper/paper.Rmd
make density plots for goods for class assignment and model it

@@
 For the endogenous network effects, both transitivity (gwesp) and popularity (gwdegree) present significant positive effects.
-The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, while positive popularity indicates a tendency toward centralization that means the presence of a limited number of burials with many more relations than others.
-<!-- what is the social anthropological implication for these values -->
+The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
+The positive popularity shows a tendency toward centralization that reflects a limited number of burials have many more relations than others.
+This implies that a few central individuals developed personal networks through wealth accumulation in terms of the flow of goods.
 

### 2020-11-22 analysis/paper/000-prep2-goods-class-plot.R

@@ -68,7 +68,8 @@ burial_beads <-
   select(Agate_bead, Golden_bead, combined_glass_bead) %>%
   pivot_longer(col = ends_with("_bead"),
                names_to = "type",
-               values_to = "value")
+               values_to = "value") %>%
+  mutate(value = ifelse(is.na(value), 0, value))
 
 # boxplot for all beads
 burial_beads %>%
@@ -76,6 +77,48 @@ burial_beads %>%
   geom_boxplot() +
   coord_flip() +
   scale_y_log10()
+
+# density ridgeline plot 1
+library(ggridges)
+burial_beads %>%
+  ggplot(aes(x = value, y = type, fill = factor(stat(quantile)))) +
+  stat_density_ridges(
+    geom = "density_ridges_gradient", calc_ecdf = TRUE,
+    quantiles = 4, quantile_lines = TRUE) +
+  scale_x_continuous(limits = c(1, 15), expand = c(0.1, 0)) +
+  scale_fill_viridis_d(name = "Quartiles")
+  #stat_density_ridges(quantile_lines = TRUE, quantiles = 3)
+
+# density ridgeline plot 2
+burial_beads %>%
+  ggplot(aes(x = value, y = type, fill = factor(stat(quantile)))) +
+  stat_density_ridges(
+    geom = "density_ridges_gradient", calc_ecdf = TRUE,
+    quantiles = c(0.3, 0.7)) +
+  scale_x_continuous(limits = c(1, 15), expand = c(0.1, 0)) +
+  scale_fill_manual(
+    name = "Probability", values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
+    labels = c("(0, 0.3)", "(0.3, 0.7)", "(0.7, 1)")
+  )
+
+# density ridgeline plot 3
+burial_beads %>%
+  ggplot(aes(x = value, y = type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
+  stat_density_ridges(geom = "density_ridges_gradient",
+                      calc_ecdf = TRUE) +
+  scale_fill_viridis_c(name = "Tail probability",
+                       direction = -1) +
+  scale_x_continuous(limits = c(1, 15), expand = c(0.1, 0))
+
+# density ridgeline plot 4
+burial_beads %>%
+  ggplot(aes(x = value, y = type)) +
+  geom_density_ridges(
[…]

### 2020-11-22 analysis/paper/001-data-tidy.R

@@ -38,14 +38,14 @@ burial_three_period_age_tidy <-
     `Degree_axis` %in% c(0:90) ~ "northeast",
     TRUE ~ "NA")) %>%
   mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
-         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
-         Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
-         Agate_bead_low = ifelse(Agate_bead < 3, 1, NA), # set low level to 2
-         Agate_bead_med = ifelse(Agate_bead > 2 & Agate_bead <10, 1, NA),
-         Agate_bead_high = ifelse(Agate_bead > 9, 1, NA), # set high level to 10
-         `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` < 100, 1, NA),
-         `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
-         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) %>% #based on the result of histogram
+         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <= 3 , 1, NA),
+         Gold_bead_high = ifelse(Golden_bead > 3, 1, NA),
+         Agate_bead_low = ifelse(Agate_bead <= 2, 1, NA), # set low level to 2
+         Agate_bead_med = ifelse(Agate_bead > 2 & Agate_bead <= 6, 1, NA),
+         Agate_bead_high = ifelse(Agate_bead > 6, 1, NA), # set high level to 10
+         `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` == 1, 1, NA),
+         `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 1 & `Indo-Pacific_bead` <= 6, 1, NA),
+         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 6, 1, NA)) %>% #based on the result of histogram
   left_join(burial_with_type_value_class) %>%
   select(burial_label,
          Phase,

### 2020-12-03 analysis/paper/paper.Rmd
work on discussion section and run some more models

@@
 We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
-We ask: (1) does the observed burial data after the start of European presence resemble the of network our hypothesized model?
+We ask: (1) does the observed burial data after the start of European presence resemble the network of our hypothesized model?
 (2) did European colonial activities in 17th century Taiwan result in the emergence of social inequality in an Indigenous society in ways that can be detected by burials networks?
@@
 <!-- comparison between two networks using patchwork bootstrap --> 
-One key difference between the pre-European and post-European networks is their size, with x burials compared to y burials.
+One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
 To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks.
@@
 
-<!-- briefly discuss Chinese burials -->
+<!-- archaeological implication based on the results-->
+The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century.
+This confirms out hypothesized models of chronological differences in the structure of burials based on grave goods and burial attributes associated with demography and mortuary practices.
+The differences are in line with the theoretical assumption derived from the corporate-network continuum for explaining horizontal hierarchy that lacks characteristics of the traditional definition of vertical hierarchy but shows some level of social inequality [@Feinman2000].
+This theoretical framework works better to understand the small scale societies and the strategies for gaining power between ambitious individuals in the pericolonical context.
+
+<!-- methodological implication-->
+The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices.
+The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017].
+In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
+We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation.
+It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects.
+This closely related to the overall structure of burials that can be viewed as a representation of a given society.
+
+<!-- more discussion of nodal attributes -->
 Age and gender attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].

### 2020-12-03 analysis/paper/003-burials-pre-network-modelling.R

@@ -156,7 +156,7 @@ legend("topleft",
 model_pre_1 <- burial_network_pre ~
   edges + # ties, a measure of density, equal to kstar(1) for undirected networks
   density +
-  gwdegree(0.5, fixed = TRUE)  +
+  gwesp(0.5, fixed = TRUE)  +
   triangle # triad relation, a measure of clustering or cohesion, also called transitive triple in undirected network
 summary(model_pre_1)
 
@@ -173,8 +173,8 @@ model_pre_2 <- burial_network_pre ~ edges + # density
   # fixed = TRUE means the scale parameter lambda is fit as a curved exponential-family model
   # not much difference in a range of 0-1.5, the lower the value of the scaling parameter, the less likely the model is to be degenerate
   # ergm can estimate the parameter from the data by using fixed=FALSE
-  gwdegree(0.1, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
-  # tendency of being in contact with multiple partners, measures of centralisation
+  gwdegree(0.3, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
+  # tendency of being in contact with multiple partners, measures of centralization
   # distribution of node-based edge counts, each node counts only once
   # number means weight parameter decay
   # The decay is close to zero, the more gwdegree considers low degree nodes relative to high degree nodes
@@ -191,15 +191,15 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   #absdiff('burial_value') +
   gwesp(0.5, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   #gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
-  gwdegree(0.5, fixed = TRUE) + # prior = 3
+  gwdegree(0.3, fixed = TRUE) + # prior = 3
   dyadcov(pre_distance_n, "dist")
 summary(model_pre_3)
 
 #--------------------Bayesian inference for ERGMs-------------------------
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
-prior.mean <- c(-3, 0, 0, 1, 0, 3, -3, 0) # positive prior number for edge means high density
+prior.mean <- c(-3, 0, 0, 1, 0, 3, -2, 0) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
[…]

### 2020-12-03 analysis/paper/004-burials-post-network-modelling.R

@@ -150,7 +150,7 @@ model.ergm <- burial_network_post ~
 summary(model.ergm)
 
 model.2 <- burial_network_post ~ edges + # density
-  gwesp(1.7, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
+  gwesp(1.4, fixed = TRUE) +  # transitivity(cohesion; triangle), a tendency for those with shared partners to become tied, or tendency of ties to cluster together
   gwdegree(0.8, fixed = TRUE)  # popularity(degree; star), the frequency distribution for nodal degrees
 summary(model.2)
 
@@ -163,23 +163,23 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
   nodematch('value_class') +
   #nodematch('orientation') +
   #absdiff('burial_value') +
-  gwesp(1.7, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
+  gwesp(1.4, fixed = TRUE) + # start close to zero and move up, how well we do in matching the count of triangles
   #gwnsp(1.7, fixed = TRUE) + # original 1.8
-  gwdegree(0.8, fixed = TRUE) +
+  gwdegree(0.5, fixed = TRUE) +
   dyadcov(post_distance_n, "dist")
 
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 0, 0, 0, 0, 4, 7, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(c(3, 5, 5, 5, 5, 3, 3, 3), 8, 8) # covariance matrix structure
+prior.mean <- c(-3, 0, 0, -1, 0, 1, 2, -1) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(c(3, 5, 5, 5, 5, 3, 2, 3), 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,
                  prior.sigma = prior.sigma,
                  burn.in     = 100, # burn-in iterations for every chain of the population, drops the first 200
                  main.iters  = 1000, # iterations for every chain of the population
-                 aux.iters   = 1000, # MCMC steps used for network simulation
+                 aux.iters   = 3000, # MCMC steps used for network simulation
                  nchains     = 16, # number of chains of the population MCMC
[…]

### 2020-12-04 analysis/paper/paper.Rmd
work on discussion section and add references

@@
 
-<!-- archaeological implication based on the results-->
+<!-- theoretical framework for the results-->
 The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century.
-This confirms out hypothesized models of chronological differences in the structure of burials based on grave goods and burial attributes associated with demography and mortuary practices.
-The differences are in line with the theoretical assumption derived from the corporate-network continuum for explaining horizontal hierarchy that lacks characteristics of the traditional definition of vertical hierarchy but shows some level of social inequality [@Feinman2000].
-This theoretical framework works better to understand the small scale societies and the strategies for gaining power between ambitious individuals in the pericolonical context.
+This confirms our hypothesized models of chronological differences in the structure of burials in terms of grave goods and burial attributes associated with demography and mortuary practices.
+The differences are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum, that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000].
+Compared to vertical hierarchy that stresses stratified social characteristics, the corporate-network model is more effective to investigate the way how a small scale society organized and explore changes in social relation in a pericolonial context.
+A corporate-based society stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999].
+In contrast, a network-based society presents wealth accumulation through individual networks, resources control, and trade monopolization.
+
+<!--archaeological implication based on the results-->
+The covariants for both networks before and after the European presence provide further information of which variables relative to the formation of networks with the same prestige goods in common.
+We included age, sex, ritual pottery, and wealth index based on value of grave goods.
+The pre-European network indicates that burials with pottery as ritual vessel tend to be tied, while the burials within the same level of wealth tend to be connected in the post-European network.
+Moreover, the burials with ritual vessel tend not to be tied in the post-European network.
+This illustrates there are different mechanisms to form relations between burials in terms of the access to trade trade goods.
+Before the European arrival, ritual behavior seem to be an emphasized cultural practice relative to the use of prestige goods.
+After the European arrival, the cultural emphasis shifted from ritual to wealth difference.
+The demographic information, such as age and sex, do not show any correlations with the network formation for both periods.
+This may be influenced by some unidentified burials that limits a more insightful result.
+The distance between burials is not a variable that correlates to the network formation either.
+
+<!-- fit into a big anthropological picture-->
 
 <!-- methodological implication-->
+In this paper, we used burial network as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970].
+We also note that the issue of disconnection between the role of the deceased in burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003].
 The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices.

### 2020-12-04 analysis/paper/003-burials-pre-network-modelling.R

@@ -191,7 +191,7 @@ model_pre_3 <- burial_network_pre ~ edges +  # the overall density of the networ
   #absdiff('burial_value') +
   gwesp(0.5, fixed = TRUE) + #start close to zero and move up, how well we do in matching the count of triangles
   #gwnsp(0.8, fixed = TRUE) + #0.75, #prior = -1
-  gwdegree(0.3, fixed = TRUE) + # prior = 3
+  gwdegree(0.5, fixed = TRUE) + # prior = 3
   dyadcov(pre_distance_n, "dist")
 summary(model_pre_3)
 

### 2020-12-04 analysis/paper/004-burials-post-network-modelling.R

@@ -171,8 +171,8 @@ model.post.3 <- burial_network_post ~ edges +  # the overall density of the netw
 summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
-prior.mean <- c(-3, 0, 0, -1, 0, 1, 2, -1) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(c(3, 5, 5, 5, 5, 3, 2, 3), 8, 8) # covariance matrix structure
+prior.mean <- c(-3, 0, 0, 0, 0, 1, 1, 0) # prior mean corresponds to mean for each parameter
+prior.sigma <- diag(c(3, 1, 1, 1, 1, 1, 1, 1), 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,

### 2020-12-04 analysis/paper/paper.Rmd
revise the bootstrap paragraphs

@@
 chi <- burial_phases_sample_size %>% filter(Phase == "chi") %>% pull()
+diff_pre_post <- post_E - pre_E
 ```
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks.
-This allows a comparison for multiple networks of different sizes [@Chen2019].
+To understand if this network size different affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post ` nodes from the post-European network to resemble the size of pre-European network with vertex bootstrap.
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
 node) of a network to quantify standard errors in the network statistics of interest.
-This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population.
+This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population, especially for smaller networks.
 We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks.
-The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our networks are relatively robust to differences in sample size.
-This suggests a very small sample size effect on both models.
+The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but higher extreme network clustering.
+This pattern is consistent before and after node removal for the post-European network, and the difference can be still identified between networks from different phases.
+This indicates our networks are relatively robust to differences in sample size and suggests a very small sample size effect.
+
+Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows a statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999].
+We used t test for simulated networks and the results show that there are significant difference in all network statistics from both periods.
 In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
@@
 But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
-<!-- re-write wither to justify why vertex bootstrap is the right method to answer this question, or use patchwork bootstrap and justify that-->
 
@@
 We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation.
-It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects.
-This closely related to the overall structure of burials that can be viewed as a representation of a given society.
+Although some network statistics of covariants do not showing the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials.
+This sturcutre can be viewed as a representation of a given society.
 

### 2020-12-04 analysis/paper/004-burials-post-network-modelling.R

@@ -172,7 +172,7 @@ summary(model.post.3)
 
 # Specify a prior distribution: normal distribution (low density and high transitivity)
 prior.mean <- c(-3, 0, 0, 0, 0, 1, 1, 0) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(c(3, 1, 1, 1, 1, 1, 1, 1), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(3, 3, 1, 1, 1, 2, 2, 1), 8, 8) # covariance matrix structure
 
 post_bergm <- bergmM(model.post.3,
                  prior.mean  = prior.mean,

### 2020-12-04 analysis/paper/006-bergm-bootstrap.R

@@ -18,7 +18,7 @@ pre_matrix <- as.matrix(as_adjacency_matrix(pre_E_igraph))
 # vertex bootstrap, useful for small network
 library(snowboot)
 set.seed(1)
-number <- 500
+number <- 1000
 pre_Astar <- vertboot(pre_matrix, number)
 
 # calculate 95% bootstrap confidence intervals for the density and bootstrap standard error
@@ -88,7 +88,7 @@ post_all_obs <- rbind(post_density_obs,
 post_matrix <- as.matrix(as_adjacency_matrix(post_E_igraph))
 
 # vertex bootstrap, useful for small network
-post_number <- 500
+post_number <- 1000
 set.seed(1)
 post_Astar <- vertboot(post_matrix, post_number)
 
@@ -97,7 +97,7 @@ post_boot_density <- sapply(1:post_number, function(x)
     graph.density(graph_from_adjacency_matrix(post_Astar[[x]])))
 post_CIvertboot_den <- quantile(post_boot_density, c(0.025, 0.975))
 post_CIvertboot_den
-post_bootstrap_se <- sd(boot_density)
+post_bootstrap_se <- sd(post_boot_density)
 post_bootstrap_se
 
 # confedence interval for mean degree and transitivity
@@ -146,3 +146,52 @@ post_CIpatchwork <- lsmi_cv(post_E_igraph_to_network,
 
 # calculate 95% bootstrap confidence intervals for the density
 post_CIpatchwork$bci/(post_E_igraph_to_network$n - 1)
+
+#----------------------T test for pre and post network statistics---------------------
+# T test for network density, clustering, and degree
+t.test(pre_boot_density, post_boot_density)
+t.test(pre_boot_trans, post_boot_trans)
+t.test(pre_boot_star, post_boot_star)
+
+t.test(post_rm_boot_density, post_boot_density)
+t.test(post_rm_boot_trans, post_boot_trans)
+t.test(post_rm_boot_star, post_boot_star)
+#-----------------------vertex removal for post-E-------------------------------------
+post_E_node_removal <-
+  delete_vertices(post_E_igraph, sample(1:49, 20, replace=F)) # remove 20 nodes
+
+post_rm_density_obs <- graph.density(post_E_node_removal)
+post_rm_mean_star_obs <- mean(degree(post_E_node_removal))
+post_rm_trans_obs <- transitivity(post_E_node_removal)
[…]

### 2020-12-04 analysis/paper/paper.Rmd
fixed typos

@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand if this network size different affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post ` nodes from the post-European network to resemble the size of pre-European network with vertex bootstrap.
+To understand if this network size difference affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post ` nodes from the post-European network to resemble the size of pre-European network with vertex bootstrap.
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
@@
 
-Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows a statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999].
-We used t test for simulated networks and the results show that there are significant difference in all network statistics from both periods.
+Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999].
+We used t-test for simulated networks and the results show that there are significant differences in all network statistics from both periods.
 In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
@@
 The differences are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum, that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000].
-Compared to vertical hierarchy that stresses stratified social characteristics, the corporate-network model is more effective to investigate the way how a small scale society organized and explore changes in social relation in a pericolonial context.
+Compared to vertical hierarchy that stresses stratified social characteristics, the corporate-network model is more effective to investigate the way how a small scale society organized and explore changes in social relations in a pericolonial context.
 A corporate-based society stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999].
-In contrast, a network-based society presents wealth accumulation through individual networks, resources control, and trade monopolization.
+In contrast, a network-based society presents wealth accumulation through individual networks, resource control, and trade monopolization.
 
@@
 The covariants for both networks before and after the European presence provide further information of which variables relative to the formation of networks with the same prestige goods in common.
-We included age, sex, ritual pottery, and wealth index based on value of grave goods.
-The pre-European network indicates that burials with pottery as ritual vessel tend to be tied, while the burials within the same level of wealth tend to be connected in the post-European network.
-Moreover, the burials with ritual vessel tend not to be tied in the post-European network.
-This illustrates there are different mechanisms to form relations between burials in terms of the access to trade trade goods.
+We included age, sex, ritual pottery, and wealth index based on the value of grave goods.
+The pre-European network indicates that burials with pottery as ritual vessels tend to be tied, while the burials within the same level of wealth tend to be connected in the post-European network.
+Moreover, the burials with ritual vessels tend not to be tied in the post-European network.
+This illustrates there are different mechanisms to form relations between burials in terms of access to trade goods.
 Before the European arrival, ritual behavior seem to be an emphasized cultural practice relative to the use of prestige goods.
@@
 <!-- methodological implication-->
-In this paper, we used burial network as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970].
-We also note that the issue of disconnection between the role of the deceased in burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003].
+In this paper, we used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970].
+We also note that the issue of disconnection between the role of the deceased in the burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003].
 The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices.
@@
 We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation.
-Although some network statistics of covariants do not showing the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials.
-This sturcutre can be viewed as a representation of a given society.
+Although some network statistics of covariants do not show the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials.
+This structure can be viewed as a representation of a given society.
 

### 2020-12-04 analysis/paper/paper.Rmd
some edits to the discussion section

@@
 
-<!-- stopped editing here -->
-
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand if this network size difference affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post ` nodes from the post-European network to resemble the size of pre-European network with vertex bootstrap.
+To understand if this network size difference affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post` nodes from the post-European network to resemble the size of pre-European network using a vertex bootstrap procedure.
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
 node) of a network to quantify standard errors in the network statistics of interest.
-This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population, especially for smaller networks.
-We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks.
-The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but higher extreme network clustering.
-This pattern is consistent before and after node removal for the post-European network, and the difference can be still identified between networks from different phases.
-This indicates our networks are relatively robust to differences in sample size and suggests a very small sample size effect.
+This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence intervals for the network population, especially for smaller networks.
+We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for our two networks.
+The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but the observed network clustering values are higher than the 95% CI of the bootstrap distributions.
+This pattern is consistent before and after node removal for the post-European network.
+The bootstrap distributions for the post-European network, when adjusted to be the same size pre-European network, show distinct differences from the pre-European network statistics.
+This indicates our networks are relatively robust to differences in sample size, and suggests a negligible sample size effect.
 
 Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999].
-We used t-test for simulated networks and the results show that there are significant differences in all network statistics from both periods.
-In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
+We used a t-test on the distributions of network statistics from the bootstrapped network samples.
+The results show that there are significant differences in all network statistics from both periods.
+This supports the presence of anthropologically meaningful differences in the pre- and post-European networks.
+
+<!-- move this text to another section that is a more logical fit --> 
+In addition, since the prior settings will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
 In general, there are no significant changes in the posterior estimates using different sets of priors.
 But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
-
-<!-- stopped editing here -->
 
@@
 the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
-For edgewise shared partner (gwesp) distribution, despite some observations falls outside the 95% interval, the fitting is generally good with most observations within it.
+For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
 This also suggests that our observed network statistics do not follow a normal distribution, especially for the network after the arrival of the Europeans that presents a bimodal pattern.
-We also compared the first three distribution moments of each observed distribution and their corresponding simulated distribution represented by mean.
-The figure shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
-For the variance of degree, the values from modeled post-European distribution are closer to the observed distribution.
+We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
+Figure \@ref(fig:xxxx) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
+For the variance of degree, the values from the modeled post-European distribution are closer to the observed distribution.
 In general, this demonstrates a better fit for the pre-European network.
@@
 <!-- theoretical framework for the results-->
-The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century.
-This confirms our hypothesized models of chronological differences in the structure of burials in terms of grave goods and burial attributes associated with demography and mortuary practices.
-The differences are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum, that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000].
-Compared to vertical hierarchy that stresses stratified social characteristics, the corporate-network model is more effective to investigate the way how a small scale society organized and explore changes in social relations in a pericolonial context.
-A corporate-based society stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999].
-In contrast, a network-based society presents wealth accumulation through individual networks, resource control, and trade monopolization.
+A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century.
+This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
+These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000].
+Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan.
+A corporate-based society, such the pre-European situation at Kiwulan, stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999].
+In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.
 
 <!--archaeological implication based on the results-->
-The covariants for both networks before and after the European presence provide further information of which variables relative to the formation of networks with the same prestige goods in common.
-We included age, sex, ritual pottery, and wealth index based on the value of grave goods.
-The pre-European network indicates that burials with pottery as ritual vessels tend to be tied, while the burials within the same level of wealth tend to be connected in the post-European network.
-Moreover, the burials with ritual vessels tend not to be tied in the post-European network.
-This illustrates there are different mechanisms to form relations between burials in terms of access to trade goods.
-Before the European arrival, ritual behavior seem to be an emphasized cultural practice relative to the use of prestige goods.
-After the European arrival, the cultural emphasis shifted from ritual to wealth difference.
-The demographic information, such as age and sex, do not show any correlations with the network formation for both periods.
-This may be influenced by some unidentified burials that limits a more insightful result.
-The distance between burials is not a variable that correlates to the network formation either.
+The covariants for both networks, before and after the European presence, provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:xxx)).
+We included age, sex, ritual pottery, and a wealth index based on the value of grave goods.
+The pre-European network shows that burials with pottery as ritual vessels tend to be tied.
+On the other hand, in the post-European network ties tend to be between burials within the same level of wealth.
+Moreover, in the post-European network, burials with ritual vessels tend not to be tied.
+This illustrates that different social and economic mechanisms determined relations between burials for the two periods.
+Before European arrival, ritual behaviors seem to have been important status indicators, relative to the use of prestige goods.
+After European presence was established, status indicators shifted from ritual to wealth differences.
+
+Distance between burials did not correlate to the network formation... <!-- anthropological concept here -->
+
+Demographic information, such as age and sex, do not show any correlations with network formation for either period.
+This may be due to the unavailability of ge and sex values for many burials in our sample.
+
+<!-- stopped editing here -->
 

### 2020-12-08 analysis/paper/paper.Rmd
work on discussion section and run models

@@
 Based on our hypothesis, we model a network with increased social inequality that is represented by endogenous network effects, low transitivity and high popularity.
-We include burial-specific attributes as covariate effects for homophily, such as age, gender, ritual activity, and the degree of wealth, to test whether burials with similar attributes tend to form relationships.
+We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to test whether burials with similar attributes tend to form relationships.
 For example, age-homophily means people of the same age to have the same burial goods.
@@
 Thus, we specified the prior of the edge density parameter to low for both network models.
-For covariates based on burial attributes, such as age, gender, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
+For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
 N(0, 5)) for both models.
@@
 As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
-Then we summarise output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
+In general, we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
+Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
 We set 10,000 network graphs simulated from the estimated posterior distribution in ERGMs.
-Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
+Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
 We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
@@
 We compared estimated statistics from the posterior distributions to examine their differences in  structure of the simulated networks (Table XX).
-For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the relations between burials.
+For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials.
 The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
@@
 
-<!-- move this text to another section that is a more logical fit --> 
-In addition, since the prior settings will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters.
-In general, there are no significant changes in the posterior estimates using different sets of priors.
-But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
-
 ```{r Bayesian-gof}
@@
 After European presence was established, status indicators shifted from ritual to wealth differences.
-
-Distance between burials did not correlate to the network formation... <!-- anthropological concept here -->
-
-Demographic information, such as age and sex, do not show any correlations with network formation for either period.
-This may be due to the unavailability of ge and sex values for many burials in our sample.
+Distance between burials did not correlate to network formation for both periods that means geographical proximity, implying some kinship association, is not a factor for sharing similar goods between burials.
+This may suggest that the use of burial goods is more individual-oriented instead of kinship-based sharing.
+In addition, demographic information, such as age and sex, do not show any correlations with network formation for either period.
+This may be due to the unavailability of age and sex values for many burials in our sample.
 
@@
 
-<!-- fit into a big anthropological picture-->
+<!-- other evidence and a bigger anthropological picture-->
+We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970].
+We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003].
+Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution since the Spanish and the Dutch were active in northern Taiwan that hints at an increased social differentiation [@LWandBM2020ornament].
+This is consistent with our result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case.
+Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations indicated by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
+After the presence of the Europeans, a more centralized and less cohesive network indicates some individuals were differentiated or privileged to possess more distinctive goods.
+Moreover, the relationship tends to form between individuals with the same wealth level represented by the rarity of goods.
+This uneven prestige accumulation might result from differential access to exotic trade relative to the European colonization.
+The involvement in long-distance exchange often leads to changes in political-economic strategies toward a network mode, especially for small-scale societies, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000].
 
 <!-- methodological implication-->
-In this paper, we used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970].
-We also note that the issue of disconnection between the role of the deceased in the burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003].
 The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices.
@@
 <!-- more discussion of nodal attributes -->
-Age and gender attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].
+Age and sex attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].
 

### 2020-12-09 analysis/paper/paper.Rmd
udpate discussion section

@@
 
-<!-- we should tell the reader what does it mean high-medium-low --> 
+<!-- burial data and the definition of tie formation  --> 
 We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
-Gold-foil and agate beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high (> 10), medium (> 1 and < 10), and low (= 1), of these beads.
-Glass beads are in levels of high (> 100), medium (> 99 and < 900), and low (< 100).
+Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high, medium, and low, according to their distributions across all burials.
+Gold-foil beads are in levels of high (>3), medium (=2-3), and low (=1); glass beads and carnelian beads are in levels of high (>6), medium (=3-6), and low (=1-2).
 That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 <!-- theoretical framework for the results-->
-A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century.
+A striking finding in our results is the change in network properties of Kiwulan burials from a  more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with  (lower transitivity and positive popularity) after the European presence during the 17th century.
 This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
-These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000].
+These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
 Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan.
-A corporate-based society, such the pre-European situation at Kiwulan, stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999].
+A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999].
 In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.
@@
 We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003].
-Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution since the Spanish and the Dutch were active in northern Taiwan that hints at an increased social differentiation [@LWandBM2020ornament].
-This is consistent with our result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case.
-Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations indicated by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
-After the presence of the Europeans, a more centralized and less cohesive network indicates some individuals were differentiated or privileged to possess more distinctive goods.
-Moreover, the relationship tends to form between individuals with the same wealth level represented by the rarity of goods.
-This uneven prestige accumulation might result from differential access to exotic trade relative to the European colonization.
-The involvement in long-distance exchange often leads to changes in political-economic strategies toward a network mode, especially for small-scale societies, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000].
+Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament].
+This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case.
+Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
+After the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
+According to historical records, the Dutch would give local Indigenous leaders symbolic items or trade goods to assert their authority and consolidate the relationship with local people [@Kang2016, pp137-162].
+Uneven distribution of goods in burials we observed could result from differential access to exotic trade relative to the European colonization and gradually led to increased social differentiation.
+The involvement in long-distance exchange often leads to changes in political-economic strategies to a network mode, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000; ].
 
@@
 # Conclusion
+It should be noted that corporate/network spectrum represents a different degree of hierarchical complexity instead of a clear division. 
+This might lead to accumulation of rare prestige goods and competition between individuals that Kiwulan gradually moves to a more network-based society.
 

### 2020-12-10 analysis/paper/paper.Rmd
edit discussion and work on conclusion

@@
 <!-- theoretical framework for the results-->
-A striking finding in our results is the change in network properties of Kiwulan burials from a  more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with  (lower transitivity and positive popularity) after the European presence during the 17th century.
+A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration in a limited burials (lower transitivity and positive popularity) after the European presence during the 17th century.
 This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
@@
 A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999].
-In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.
+In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good system manipulation, and trade monopolization [@Blanton1996].
+It should be noted that corporate/network continuum represents a dynamic process with different degree of hierarchical complexity instead of a static ideal-type stage.
 
@@
 
-<!-- other evidence and a bigger anthropological picture-->
+<!-- other evidence -->
 We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970].
@@
 This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case.
-Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
-After the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
-According to historical records, the Dutch would give local Indigenous leaders symbolic items or trade goods to assert their authority and consolidate the relationship with local people [@Kang2016, pp137-162].
-Uneven distribution of goods in burials we observed could result from differential access to exotic trade relative to the European colonization and gradually led to increased social differentiation.
-The involvement in long-distance exchange often leads to changes in political-economic strategies to a network mode, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000; ].
+Since the burial goods used for making ties were treated as prestige goods or status items in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
 
-<!-- methodological implication-->
-The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices.
-The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017].
-In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation.
-Although some network statistics of covariants do not show the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials.
-This structure can be viewed as a representation of a given society.
-
-<!-- more discussion of nodal attributes -->
-Age and sex attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].
+<!--fit in a big anthropological picture in a periclonial context -->
+The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
+Despite relatively weak control, northeastern Taiwan was incorporated into colonial territories in 1647 after the Dutch conquests [].
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies [@Kang2016].
+At Kiwulan, after the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
+Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation.
+The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
+This may lead to ambitious individuals to compete with each other and accumulate trade goods.
+It demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context, gradually resulting in social changes from a more corporate to a more network strategy.
 
 # Conclusion
-It should be noted that corporate/network spectrum represents a different degree of hierarchical complexity instead of a clear division. 
-This might lead to accumulation of rare prestige goods and competition between individuals that Kiwulan gradually moves to a more network-based society.
+
+<!-- summary of the research, contribution and methodological implication-->
+In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European percicolonial context in northeastern Taiwan.
+We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
+Our results provide a meaningful explanation for the relationship between burials that share similar prestige goods.
+Before European contact in the 17th century, the network properties has a tendency of more closed burial relationships with ritual elements as the key formation mechanism.
+After the European arrival, the network has a tendency of centralization relative to the rarity of goods.
+This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality.
+Based on the burial data, along with historical records and ethnography, we are able to understand the processes of changes in Indigenous societies in the periphery of European colonial bases.
+This study demonstrates not only the pericolonial effect on local Indigenous societies that have not well understood yet, but also the agency of local people to manipulate colonial power instead of passive recipients of trade goods.
+
+This case study highlights the methodological improvement that Bayesian inference on ERGM can inform and enhance studies in rational data in archaeology.
+A Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
+Moreover, the interrelations indicated by the probabilistic inference provide robust results, allowing in-depth discussion for the dynamic processes of relationship formation and their underlying factors to reflect the historical trajectories of sociocultural phenomena.
+This approach can be applied to a wide range of relational data in archaeology to examine the formation of relationships and give an insight into the study of social structure in general.
 

### 2020-12-11 analysis/paper/paper.Rmd
edit conclusion and split code for sensitivity test to two

@@
 This case study highlights the methodological improvement that Bayesian inference on ERGM can inform and enhance studies in rational data in archaeology.
-A Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-Moreover, the interrelations indicated by the probabilistic inference provide robust results, allowing in-depth discussion for the dynamic processes of relationship formation and their underlying factors to reflect the historical trajectories of sociocultural phenomena.
-This approach can be applied to a wide range of relational data in archaeology to examine the formation of relationships and give an insight into the study of social structure in general.
+A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
+Moreover, probabilistic inference provides robust results of interrelations that are not statistically tested by traditional network approaches.
+Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships.
+This allows in-depth discussion of dynamic processes of relationship formation and their underlying factors that are informative of historical trajectories or socio-cultural phenomena.
 

### 2020-12-11 analysis/paper/006-bergm-bootstrap.R

@@ -100,7 +100,7 @@ post_CIvertboot_den
 post_bootstrap_se <- sd(post_boot_density)
 post_bootstrap_se
 
-# confedence interval for mean degree and transitivity
+# confidence interval for mean degree and transitivity
 post_boot_star <- sapply(1:number, function(x)
   mean(degree(graph_from_adjacency_matrix(post_Astar[[x]]), mode = "in")))
 post_CIvertboot_star <- quantile(post_boot_star , c(0.025, 0.975))
@@ -156,42 +156,3 @@ t.test(pre_boot_star, post_boot_star)
 t.test(post_rm_boot_density, post_boot_density)
 t.test(post_rm_boot_trans, post_boot_trans)
 t.test(post_rm_boot_star, post_boot_star)
-#-----------------------vertex removal for post-E-------------------------------------
-post_E_node_removal <-
-  delete_vertices(post_E_igraph, sample(1:49, 20, replace=F)) # remove 20 nodes
-
-post_rm_density_obs <- graph.density(post_E_node_removal)
-post_rm_mean_star_obs <- mean(degree(post_E_node_removal))
-post_rm_trans_obs <- transitivity(post_E_node_removal)
-post_rm_all_obs <- rbind(post_rm_density_obs,
-                         post_rm_mean_star_obs,
-                         post_rm_trans_obs)
-post_rm_matrix <- as.matrix(as_adjacency_matrix(post_E_node_removal))
-
-# vertex bootstrap, useful for small network
-post_number <- 1000
-set.seed(1)
-post_rm_Astar <- vertboot(post_rm_matrix, post_number)
-
-# calculate 95% bootstrap confidence intervals for the density and bootstrap standard error
-post_rm_boot_density <- sapply(1:post_number, function(x)
-  graph.density(graph_from_adjacency_matrix(post_rm_Astar[[x]])))
-post_rm_CIvertboot_den <- quantile(post_rm_boot_density, c(0.025, 0.975))
-post_rm_CIvertboot_den
-post_rm_bootstrap_se <- sd(post_rm_boot_density)
-post_rm_bootstrap_se
-
-# confedence interval for mean degree and transitivity
-post_rm_boot_star <- sapply(1:number, function(x)
-  mean(degree(graph_from_adjacency_matrix(post_rm_Astar[[x]]), mode = "in")))
[…]

### 2020-12-11 analysis/paper/007-sensitivity-test.R

@@ -1,0 +1,40 @@
+# run 006-bergm-bootstrap
+#-----------------------vertex removal for post-E-------------------------------------
+post_E_node_removal <-
+  delete_vertices(post_E_igraph, sample(1:49, 20, replace=F)) # remove 20 nodes
+
+post_rm_density_obs <- graph.density(post_E_node_removal)
+post_rm_mean_star_obs <- mean(degree(post_E_node_removal))
+post_rm_trans_obs <- transitivity(post_E_node_removal)
+post_rm_all_obs <- rbind(post_rm_density_obs,
+                         post_rm_mean_star_obs,
+                         post_rm_trans_obs)
+post_rm_matrix <- as.matrix(as_adjacency_matrix(post_E_node_removal))
+
+# vertex bootstrap, useful for small network
+post_number <- 1000
+set.seed(1)
+post_rm_Astar <- vertboot(post_rm_matrix, post_number)
+
+# calculate 95% bootstrap confidence intervals for the density and bootstrap standard error
+post_rm_boot_density <- sapply(1:post_number, function(x)
+  graph.density(graph_from_adjacency_matrix(post_rm_Astar[[x]])))
+post_rm_CIvertboot_den <- quantile(post_rm_boot_density, c(0.025, 0.975))
+post_rm_CIvertboot_den
+post_rm_bootstrap_se <- sd(post_rm_boot_density)
+post_rm_bootstrap_se
+
+# confedence interval for mean degree and transitivity
+post_rm_boot_star <- sapply(1:number, function(x)
+  mean(degree(graph_from_adjacency_matrix(post_rm_Astar[[x]]), mode = "in")))
+post_rm_CIvertboot_star <- quantile(post_rm_boot_star , c(0.025, 0.975))
+
+post_rm_boot_trans <- sapply(1:number, function(x)
+  transitivity(graph_from_adjacency_matrix(post_rm_Astar[[x]]), type = "undirected"))
+post_rm_CIvertboot_trans <- quantile(post_rm_boot_trans , c(0.025, 0.975))
+
+post_rm_boot_all_stats <-
+  as.data.frame(rbind(post_rm_CIvertboot_den,
+                      post_rm_CIvertboot_star,
+                      post_rm_CIvertboot_trans)) %>%
+  cbind(obs = post_rm_all_obs)

### 2020-12-11 analysis/paper/paper.Rmd
edit discussion and add references

@@
 
-<!--fit in a big anthropological picture in a periclonial context -->
+<!--fit in a bigger historical picture in a periclonial context -->
 The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
 Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
-Despite relatively weak control, northeastern Taiwan was incorporated into colonial territories in 1647 after the Dutch conquests [].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies [@Kang2016].
-At Kiwulan, after the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
+In 1626, the Spanish built their forts in Keelung and Tamsui, northern Taiwan, and later defeated and took over by the Dutch in 1642 [@Andrade2007].
+Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attached by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+
+<!--anthopological implication at Kiwulan and limitation-->
+At Kiwulan, after the European presence, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
 Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation.
 The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
-This may lead to ambitious individuals to compete with each other and accumulate trade goods.
-It demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context, gradually resulting in social changes from a more corporate to a more network strategy.
+This may lead to ambitious individuals to compete with each other for the goods and accumulate wealth.
+It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
+Our interpretation is based on relations indicated by burial data that can be supported by the clustered distribution of trade goods in Kiwulan residential area [eg.
+@LWandBM2020ornament].
+We would still need more other different evidence to understand the process of social change.
+In addition, we acknowledge the fragmentation issue in archaeology data that might limit network interpretation [eg.
+@Erik2015], such as missing age and sex nodal attributes for some burials in our data.
+Despite we lack of informative results about their effect on network formation, the Bayesian approach still provides effective and reliable statistical inferences on the overall network properties by reducing the missing data effect.
 

### 2020-12-13 analysis/paper/paper.Rmd

@@
 
-<!-- comparison between two networks using patchwork bootstrap --> 
+<!-- comparison between two networks using bootstrap --> 
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.

### 2020-12-14 analysis/paper/paper.Rmd
update testing results and edit result section

@@
 Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations.
-For the endogenous network effects, both transitivity (gwesp) and popularity (gwdegree) present significant positive effects.
+For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect.
 The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
-The positive popularity shows a tendency toward centralization that reflects a limited number of burials have many more relations than others.
-This implies that a few central individuals developed personal networks through wealth accumulation in terms of the flow of goods.
+The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones.
+This might imply that individuals have equal access to trade goods in terms of the flow of goods.
 
@@
 We also found that the burials in the same wealth level tend not to form relations.
-Similar to the pre-European network, the endogenous network variables, transitivity (gwesp) and popularity (gwdegree), all demonstrate significant positive effects.
-The post-European network has a smaller transitivity effect and a larger popularity effect than the pre-European network.
-This may suggest a reduced tendency toward clustering but higher tendency toward centralization after the European presence.
+Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect.
+In contrast to the pre-European network, the popularity (gwdegree) shows a significant positive effect.
+This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others.
+This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation.
+In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network.
+This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
 Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand if this network size difference affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post` nodes from the post-European network to resemble the size of pre-European network using a vertex bootstrap procedure.
+To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results based on Bayesian network analysis.
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) of a network to quantify standard errors in the network statistics of interest.
-This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence intervals for the network population, especially for smaller networks.
-We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for our two networks.
-The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but the observed network clustering values are higher than the 95% CI of the bootstrap distributions.
-This pattern is consistent before and after node removal for the post-European network.
-The bootstrap distributions for the post-European network, when adjusted to be the same size pre-European network, show distinct differences from the pre-European network statistics.
-This indicates our networks are relatively robust to differences in sample size, and suggests a negligible sample size effect.
+node) of a network to quantify standard errors in the network statistics of interest [@Chen2019; @Snijders1999].
+This enables the evaluation of uncertainty for networks and tests the difference between networks by examining their confidence intervals for the network population, where networks with unoverlapped intervals represent a significant difference.
+We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks.
+The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
+For density and transitivity, both 95% confidence intervals show some level of overlapping that indicates no significant difference.
+This pattern is also similar to our results of Bayesian modeling, where both networks show similar positive or negative effects with some level of differences.
 
-Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999].
-We used a t-test on the distributions of network statistics from the bootstrapped network samples.
-The results show that there are significant differences in all network statistics from both periods.
-This supports the presence of anthropologically meaningful differences in the pre- and post-European networks.
+Similarly, vertex bootstrap allows statistical testing for multiple networks, such as a t-test method.
+We conducted a t-test on the distributions of network statistics between two networks.
+The results show that there are significant differences in all network statistics.
+This supports the presence of anthropologically meaningful differences in the pre-and post-European networks.
+Moreover, we tested the sample size effect by removing nodes at certain percents (5-30%) and compare their confidence intervals for both networks using the same vertex bootstrap procedure.
+The results show a consistent difference between the two networks that shows they are robust to the different sizes.
+In general, the cross-validation of results from vertex bootstrap and Bayesian modeling illustrates the robust comparison between two networks without sample size effects.
 
@@
 ```
+
+Word count: `r wordcountaddin::word_count()`

### 2020-12-17 analysis/paper/paper.Rmd
edit intro, result, and conclusion section

@@
 Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-This paper will use a novel Bayesian modeling approach to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
+This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
@@
 <!-- archaeological specific case study -->
-In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
+In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
 We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
@@
 We ask: (1) does the observed burial data after the start of European presence resemble the network of our hypothesized model?
-(2) did European colonial activities in 17th century Taiwan result in the emergence of social inequality in an Indigenous society in ways that can be detected by burials networks?
+(2) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by burials networks?
 (3) what are the major variables affecting or forming the higher degree of popularity (e.g.
 a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period?
-By answering these questions, this study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups.
+By answering these questions, this study helps to expand the use of statistical modeling for burials in understanding the indirect effects of a colonial presence on Indigenous groups.
 This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
@@
 and make it clear that we are doing something new -->
-Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+
+Bayesian approach to Exponential Random Graph Models (ERGMs) is an effective way to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Current network analysis used by archaeologist is mostly restricted to a single rational structure without consideration of interaction between network variables.
+Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
+ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
 The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-ERGMs differs from traditional network approaches in that edges are only formed according to their corresponding predictors separately without taking into account the interactions among edges [@Morris2008].
-In ERGMs, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011].
+In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
 For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
 Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007].
-By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks that represent our model.
+By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
 Such distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-This helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
 
@@
 They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
-They also pointed out some limitations of ERGMs, such as static outcomes and sensitivity to missing data.
+However, they also point out some limitations of ERGMs, such as static outcomes and sensitivity to missing data.
 In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
@@
 node) of a network to quantify standard errors in the network statistics of interest [@Chen2019; @Snijders1999].
-This enables the evaluation of uncertainty for networks and tests the difference between networks by examining their confidence intervals for the network population, where networks with unoverlapped intervals represent a significant difference.
+This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population.
+Networks with unoverlapped intervals means there is a significant difference.
 We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks.
 The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
-For density and transitivity, both 95% confidence intervals show some level of overlapping that indicates no significant difference.
-This pattern is also similar to our results of Bayesian modeling, where both networks show similar positive or negative effects with some level of differences.
+For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference.
+This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects some degree of difference.
 
-Similarly, vertex bootstrap allows statistical testing for multiple networks, such as a t-test method.
-We conducted a t-test on the distributions of network statistics between two networks.
+Using the same vertex bootstrap procedure, we can also test the difference of network populations with a t-test on density, transitivity, and popularity.
 The results show that there are significant differences in all network statistics.
-This supports the presence of anthropologically meaningful differences in the pre-and post-European networks.
-Moreover, we tested the sample size effect by removing nodes at certain percents (5-30%) and compare their confidence intervals for both networks using the same vertex bootstrap procedure.
-The results show a consistent difference between the two networks that shows they are robust to the different sizes.
-In general, the cross-validation of results from vertex bootstrap and Bayesian modeling illustrates the robust comparison between two networks without sample size effects.
+Moreover, we explored the sample size effect by removing nodes at certain percents (5-40%) for both networks and compared their confidence intervals.
+The results demonstrate a consistent difference between the two networks that means they are robust to the different sizes.
+In general, those statistical methods support the presence of anthropologically meaningful differences in the pre-and post-European networks.
+The procedure of cross-validation ensures our results can provide robust comparison between two networks without sample size effects.
 
@@
 
-<!-- summary of the research, contribution and methodological implication-->
+<!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
 In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European percicolonial context in northeastern Taiwan.
 We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
-Our results provide a meaningful explanation for the relationship between burials that share similar prestige goods.
-Before European contact in the 17th century, the network properties has a tendency of more closed burial relationships with ritual elements as the key formation mechanism.
+The results support our hypothesize model that the relationship between burials changed after European colonization period in the 17th century.
+Before the arrival of the Europeans, the network properties has a tendency of more closed burial relationships with ritual elements as the key formation mechanism.
 After the European arrival, the network has a tendency of centralization relative to the rarity of goods.
-This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality.
-Based on the burial data, along with historical records and ethnography, we are able to understand the processes of changes in Indigenous societies in the periphery of European colonial bases.
-This study demonstrates not only the pericolonial effect on local Indigenous societies that have not well understood yet, but also the agency of local people to manipulate colonial power instead of passive recipients of trade goods.
+The changes in formation mechanisms for networks from ritual-oriented practices to wealth-oriented possession suggest the behavior of wealth accumulation and an increased differentiation stimulated by the European presence and associated long-distance trade network.
+This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality, indicative of an increased social inequality.
+Using burial data with historical documents, we are able to detect the processes of changes in Indigenous societies in the periphery of European colonial bases to better understand indirect impacts that are underestimated in this region.
+This study not only discuss pericolonal impacts that are not well understood yet, but also points out the agency of local people to manipulate colonial power instead of being passive recipients of trade goods.
 
+<!-- technical and methodological implication-->
 This case study highlights the methodological improvement that Bayesian inference on ERGM can inform and enhance studies in rational data in archaeology.

### 2020-12-18 analysis/paper/paper.Rmd
added a map and references

@@
   comment = "#>",
+  dpi = 300,
   fig.path = "../figures/"
 )
+
+image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
+image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
 ```
@@
 The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
 For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp.
@@
 
+```{r, burial-map, fig.cap = "Map illustrating the location of burials at Kiwulan (suare size is 4 X 4 meters)", fig.width = image_width_2_col_inch}
+knitr::include_graphics(here::here("analysis", "figures", "burial-map.png"))
+```
+
 <!--archaeological setting-->
@@
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with _in-situ_ posts were also excavated.
-The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison <!-- fig reference to a map of burials at the site-->.
+The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:burial-map)).
 Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
@@
 ## Model specification in a Bayesian framework
+
 ```{r ergm-terms-table}
 ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
-kableExtra::kable(ergm_terms,
+knitr::kable(ergm_terms,
              caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
@@
 This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality, indicative of an increased social inequality.
-Using burial data with historical documents, we are able to detect the processes of changes in Indigenous societies in the periphery of European colonial bases to better understand indirect impacts that are underestimated in this region.
+Using burial data with historical documents, we are able to detect the processes of changes in Indigenous societies in the periphery of European colonial bases to better understand indirect impacts that are underestimated in this region [@Acabado2017; @Trabert2017].
 This study not only discuss pericolonal impacts that are not well understood yet, but also points out the agency of local people to manipulate colonial power instead of being passive recipients of trade goods.

### 2020-12-19 analysis/paper/paper.Rmd
Cite a ERGM paper in Intro section

@@
 <!--burial study- network analysis-->
+
 Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
@@
 <!-- ERGM in archaeology, and limitations --> 
-To date, ERGM has only been used once in archaeology, by @Amati2019 who reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms.
+The use of ERGM in archaeology is first introduced by @Brughmans2014, who studied the Iron age settlement pattern based on 159 Iron Age sites in Southern Spain by modeling inter-settlement visibility networks and visual control.
+Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
+Similarly, @Amati2019 reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms.
 These network models evaluated different hypotheses about the importance of three variables, proximity, inter-cultural items, and pottery types, on the presence of hub sites they observed.
 They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
-However, they also point out some limitations of ERGMs, such as static outcomes and sensitivity to missing data.
-In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
+However, those studies also point out some limitations of ERGM, such as sensitivity to missing data and less ability to handle uncertainty.
+In addition, it is difficult in ERGM to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
 A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
@@
 Thus, we specified the prior of the edge density parameter to low for both network models.
-For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
-N(0, 5)) for both models.
-For physical distance between burials, we set a negative covariate effect (N(-1, 5)) to infer kinship-based proximity, e.g.
+For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 3 (i.e.
+N(0, 3)) for both models.
+For physical distance between burials, we set a negative covariate effect (N(-1, 1)) to infer kinship-based proximity, e.g.
 stronger correlations for shorter distances.
 To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(5, 7)), lower popularity (N(4, 7)), and higher covariate effect based on ritual activity (N(2, 5)) for network before European contact to indicate less social inequality.
-On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(4, 8)), and higher popularity (N(8, 8)) to model an increased social inequality.
+We set the priors to higher transitivity (N(2, 4)), lower popularity (N(-1, 1)), and higher covariate effect based on ritual activity (N(1, 5)) for network before European contact to indicate less social inequality.
+On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(1, 2)), and higher popularity (N(1, 2)) to model an increased social inequality.
 This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
-The priors for density were also set differently with N(-17, 7) for pre-European network and N(-20, 6) for post-European network.
+The priors for density were also set the same with N(-3, 3) for both pre- and post-European networks.
 

### 2020-12-19 analysis/paper/003-burials-pre-network-modelling.R

@@ -147,7 +147,6 @@ legend("topleft",
        legend = unique(quantity),# quantity
        title  = 'Burial good counts')
 
-
 # ? can't get the items in legend in order, have tried as.factor for level
 # can't match color with categories, need to adjust manually every time
 
@@ -199,7 +198,7 @@ summary(model_pre_3)
 # prior suggestion: normal distribution (low density and high transitivity), but it also depends on the ERGM netowrk we observed
 prior.mean <- c(-3, 0, 0, 1, 0, 2, -1, -1) # positive prior number for edge means high density
 # follow Alberto Caimo et al. (2015) hospital example
-prior.sigma <- diag(c(4, 5, 3, 5, 5, 5, 2, 3), 8, 8) # covariance matrix structure, uncertainty
+prior.sigma <- diag(c(3, 3, 3, 5, 5, 4, 1, 1), 8, 8) # covariance matrix structure, uncertainty
 
 # normal distribution 𝜃 ∼ Nd (𝜇prior , Σprior ) a common prior model
 # where the dimension d corresponds to the number of parameters, 𝜇 is mean vector and Σprior is a d × d covariance matrix.

### 2021-01-02 analysis/paper/paper.Rmd
added one reference and citation
change title and revise the way for making ties

@@
 ---
-title: "Bayesian exponential random graph modeling of KWL Burial network"
+title: "A Bayesian approach to burial networks to infer social changes in 17th northeastern Taiwan during the European colonization period"
 author:
@@
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  Bayesian modeling; network analysis; Monte-carlo simulation; Burials; European colonialism; social differentiation; Taiwan
 highlights: |
@@
 We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
-Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high, medium, and low, according to their distributions across all burials.
-Gold-foil beads are in levels of high (>3), medium (=2-3), and low (=1); glass beads and carnelian beads are in levels of high (>6), medium (=3-6), and low (=1-2).
+Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials.
+Gold-foil beads are in levels of high (>3), upper-middle (=3), lower-middle (=2), and low (=1); carnelian beads are in levels of high (>6), upper-middle (=4-6), lower-middle (=3), and low (=1-2); glass beads are in levels of high (>6), upper-middle (=3-6), lower-middle (=2), and low (=1).
 That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) of a network to quantify standard errors in the network statistics of interest [@Chen2019; @Snijders1999].
+node) that is useful for quantifying standard errors and estimating sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
 This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population.

### 2021-01-02 analysis/paper/001-data-tidy.R

@@ -46,7 +46,7 @@ burial_three_period_age_tidy <-
          Gold_bead_high = ifelse(Golden_bead > 3, 1, NA),
          Agate_bead_low = ifelse(Agate_bead <= 2, 1, NA), # set low level to 2
          #Agate_bead_med = ifelse(Agate_bead > 2 & Agate_bead <= 6, 1, NA),
-         Agate_bead_med_2 = ifelse(Agate_bead > 2 & Agate_bead <= 3, 1, NA),
+         Agate_bead_med_2 = ifelse(Agate_bead = 3, 1, NA),
          Agate_bead_med_3 = ifelse(Agate_bead > 3 & Agate_bead <= 6, 1, NA),
          Agate_bead_high = ifelse(Agate_bead > 6, 1, NA), # set high level to 10
          all_glass_bead_low = ifelse(all_glass_bead == 1 , 1, NA),

### 2021-01-03 analysis/paper/paper.Rmd
add abstract

@@
 author:
-  - Author One
-  - Author Two
+  - Li-Ying Wang
+  - Ben Marwick
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 abstract: |
-  Text of abstract
+  Burials provide valuable information to study social structures based on the assumption that burials and associated grave goods can represent social roles and relations in a society.
+To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns that can infer interactions and relationship between entities.
+Different from traditional static networks, statistical approaches to network analysis, such as exponential random graph models (ERGMs), provides a way to test hypotheses on dynamic process of network formation.
+However, the computational difficulty and sensitivity to uncertainties limits the application of ERGMs.
+In this paper, we introduced a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification for uncertainty, and robust model evaluation of network properties.
+We tested the hypothesis of social changes relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
+The results indicate a more centralized structure with an emphasis of social differentiation after the European presence in the 17th century.
+Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expand the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
 keywords: |

### 2021-01-04 analysis/paper/paper.Rmd
source R scripts in the Rmd

@@
 <!--burial study- network analysis-->
-
 Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
@@
 
-```{r read-data}
-library(readxl)
-library(tidyverse)
-library(here)
-
-# read data
-burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("000-prep1-value-burials.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "000-prep1-value-burials.R")))
 ```
 
-```{r tidy-data}
-# number of each phase
-burial_phases_sample_size <-
-  burial %>%
-  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
-  filter(!is.na(Phase)) %>% 
-  count(Phase)
-
-pre_E <- burial_phases_sample_size %>% filter(Phase == "pre") %>% pull()
-post_E <- burial_phases_sample_size %>% filter(Phase == "post") %>% pull()
-chi <- burial_phases_sample_size %>% filter(Phase == "chi") %>% pull()
-diff_pre_post <- post_E - pre_E
-```
-
-```{r, burial-map, fig.cap = "Map illustrating the location of burials at Kiwulan (suare size is 4 X 4 meters)", fig.width = image_width_2_col_inch}
-knitr::include_graphics(here::here("analysis", "figures", "burial-map.png"))
+```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("001-data-tidy.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
 ```
@@
 The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.
+
+```{r, burial-map, fig.cap = "Map illustrating the location of burials at Kiwulan (suare size is 4 X 4 meters)", fig.width = image_width_2_col_inch}
+knitr::include_graphics(here::here("analysis", "figures", "burial-map.png"))
+```
+
+```{r sample-size}
+# number of each phase
+library(tidyverse)
+pre_E <- burial_three_period_age_number %>% filter(Phase == "pre") %>% pull()
+post_E <- burial_three_period_age_number %>% filter(Phase == "post") %>% pull()
+chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
+diff_pre_post <- post_E - pre_E
+```
 
@@
 Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials.
-Gold-foil beads are in levels of high (>3), upper-middle (=3), lower-middle (=2), and low (=1); carnelian beads are in levels of high (>6), upper-middle (=4-6), lower-middle (=3), and low (=1-2); glass beads are in levels of high (>6), upper-middle (=3-6), lower-middle (=2), and low (=1).
+Gold-foil beads are in levels of high (> 3), upper-middle (= 3), lower-middle (= 2), and low (= 1); carnelian beads are in levels of high (> 6), upper-middle (= 4-6), lower-middle (= 3), and low (= 1-2); glass beads are in levels of high (> 6), upper-middle (= 3-6), lower-middle (= 2), and low (= 1).
 That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, account for 10%), above average (= 12-28), and below average (< 12), as an index of wealth.
 The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
@@
 ```{r bergm-iteration}
-burn_in <- 200
+burn_in <- 100
 main_iter <- 1000
-aux_iter <- 10000
+aux_iter <- 4000
 ```
@@
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European percicolonial context in northeastern Taiwan.
+In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European pericolonial context in northeastern Taiwan.
 We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
@@
 # Acknowledgments
+
+We would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the data used in this paper.
+We thank the excavation team of Kiwulan led by Yu-pei Chen for their post-excavation analysis and detailed documents about burials that enables this study.
+We also thank Alberto Caimo and Steven M.
+Goodreau for the discussion regarding the methods at the earliest stage fo analysis.
+This work was supported in part by the travel grant from the Department of Anthropology at University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
+We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
+We also acknowledge comments from anonymous reviewers that have greatly improved this manuscript.
 

### 2021-01-04 analysis/paper/001-data-tidy.R

@@ -89,4 +89,3 @@ burial_three_period_age_tidy <-
 burial_three_period_age_number <-
   burial_three_period_age_tidy %>%
   count(Phase)
-

### 2021-01-04 analysis/paper/004-burials-post-network-modelling.R

@@ -144,7 +144,7 @@ summary(model_post_3)
 # Specify a prior distribution
 # normal distribution (low density, low transitivity, high popularity)
 prior.mean <- c(-3, 0, 0, 0, 1, 1, 1, 0) # prior mean corresponds to mean for each parameter
-prior.sigma <- diag(c(2, 3, 3, 2, 3, 3, 4, 2), 8, 8) # covariance matrix structure
+prior.sigma <- diag(c(3, 3, 3, 2, 3, 3, 5, 2), 8, 8) # covariance matrix structure
 
 post_bergm <- bergm(model_post_3,
                  prior.mean  = prior.mean,

### 2021-01-04 analysis/paper/paper.Rmd
editing through to conlcusion

@@
 ---
-title: "A Bayesian approach to burial networks to infer social changes in 17th northeastern Taiwan during the European colonization period"
+title: "A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
 author:
@@
 However, the computational difficulty and sensitivity to uncertainties limits the application of ERGMs.
-In this paper, we introduced a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification for uncertainty, and robust model evaluation of network properties.
+In this paper, we introduce a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification for uncertainty, and robust model evaluation of network properties.
 We tested the hypothesis of social changes relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
-The results indicate a more centralized structure with an emphasis of social differentiation after the European presence in the 17th century.
-Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expand the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
+The results indicate a transition from ties based on ritual objects to wealth objects, and a more centralized structure with social differentiation after European presence in the 17th century.
+Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expands the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
 keywords: |
@@
 
+# A Case Study from Northereastern Taiwan
+
 <!-- archaeological background -->
@@
 
+<!--fit in a bigger historical picture in a periclonial context -->
+The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
+In 1626, the Spanish built their forts in Keelung and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
+Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+
 <!-- archaeological specific case study -->
@@
 We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
-We ask: (1) does the observed burial data after the start of European presence resemble the network of our hypothesized model?
+We ask: (1) does the observed burial data after the start of European presence resemble the network of our model?
 (2) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by burials networks?
@@
 For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`.
-We set the number of iterations used to simulate a network y′ at each iteration to `r aux_iter`.
+We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.
 
@@
 <!-- theoretical framework for the results-->
-A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration in a limited burials (lower transitivity and positive popularity) after the European presence during the 17th century.
-This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
+A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration of connections among fewer burials (lower transitivity and positive popularity) after the European presence during the 17th century.
+This supports our models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
 These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
 Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan.
-A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999].
-In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good system manipulation, and trade monopolization [@Blanton1996].
-It should be noted that corporate/network continuum represents a dynamic process with different degree of hierarchical complexity instead of a static ideal-type stage.
+A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communalrituals, and social inequality, if any, would be associated with groups [@Siegel1999].
+In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good manipulation, and trade monopolization [@Blanton1996].
+It should be noted that the corporate/network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
 
@@
 The covariants for both networks, before and after the European presence, provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:xxx)).
-We included age, sex, ritual pottery, and a wealth index based on the value of grave goods.
+Our covariants were age, sex, ritual pottery, and a wealth index based on the value of grave goods.
 The pre-European network shows that burials with pottery as ritual vessels tend to be tied.
@@
 After European presence was established, status indicators shifted from ritual to wealth differences.
-Distance between burials did not correlate to network formation for both periods that means geographical proximity, implying some kinship association, is not a factor for sharing similar goods between burials.
-This may suggest that the use of burial goods is more individual-oriented instead of kinship-based sharing.
+Physical distance between burials did not correlate to network formation for either period.
+This means geographical proximity, implying some kinship association, is not a factor for sharing similar goods between burials.
+This may suggest that the use of burial goods is more individual-oriented instead of kinship-based.
 In addition, demographic information, such as age and sex, do not show any correlations with network formation for either period.
@@
 
-<!-- stopped editing here -->
-
 <!-- other evidence -->
-We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970].
-We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003].
+We used burial data as a proxy to explore the relations based on the argument that burials  represent social structures [@Binford1971; @Saxe1970].
+We note that there could be an issue of disconnection between the role of the deceased in a burial context and their life, which could be reduced by examining evidence from the residential context [@Chapman2003].
 Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament].
-This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case.
-Since the burial goods used for making ties were treated as prestige goods or status items in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.
-
-<!--fit in a bigger historical picture in a periclonial context -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
-In 1626, the Spanish built their forts in Keelung and Tamsui, northern Taiwan, and later defeated and took over by the Dutch in 1642 [@Andrade2007].
-Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attached by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+This is consistent with the result of a more centralized burial network after the presence of Europeans, which supports a connection between a burial and the social role of the living person.
+Since the burial goods used for making ties were treated as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods observed in the burial data likely reflect social contexts of the living people at Kiwulan in a pericolonial context.
 
 <!--anthopological implication at Kiwulan and limitation-->
-At Kiwulan, after the European presence, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans.
-Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation.
-The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
-This may lead to ambitious individuals to compete with each other for the goods and accumulate wealth.
-It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
-Our interpretation is based on relations indicated by burial data that can be supported by the clustered distribution of trade goods in Kiwulan residential area [eg.
-@LWandBM2020ornament].
-We would still need more other different evidence to understand the process of social change.
-In addition, we acknowledge the fragmentation issue in archaeology data that might limit network interpretation [eg.
-@Erik2015], such as missing age and sex nodal attributes for some burials in our data.
-Despite we lack of informative results about their effect on network formation, the Bayesian approach still provides effective and reliable statistical inferences on the overall network properties by reducing the missing data effect.
+At Kiwulan, changes in interburial networks may result from differential access to exotic trade enabled by European contact, gradually leading to increased social differentiation.
+The image of European power could be embedded into prestige objects pre-existing in local Indigenous culture where their values were amplified with concepts of wealth or power, since their values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
+This may lead to ambitious individuals to competing with each other for traditional goods and accumulating wealth.
+It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
+We recognise that the meaning of the burial goods remains speculative without rich ethnographic data to support our interpretations.
+Furthermore, a detailed material culture context for burial goods in residential settings would also help to strengthen our interpretations, but this is not available from Kiwulan.
+In addition, we acknowledge the fragmentation issue in our archaeological data that limit network interpretations [eg.
+@Erik2015], such as missing age and sex attributes for some burials at Kiwulan.
 
@@
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European pericolonial context in northeastern Taiwan.
+In this paper, we presented a novel approach for studying burial relations using ERGM network analysis within a Bayesian framework.
+We examined social changes in a European pericolonial context in northeastern Taiwan.
 We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
-The results support our hypothesize model that the relationship between burials changed after European colonization period in the 17th century.
-Before the arrival of the Europeans, the network properties has a tendency of more closed burial relationships with ritual elements as the key formation mechanism.
-After the European arrival, the network has a tendency of centralization relative to the rarity of goods.
-The changes in formation mechanisms for networks from ritual-oriented practices to wealth-oriented possession suggest the behavior of wealth accumulation and an increased differentiation stimulated by the European presence and associated long-distance trade network.
-This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality, indicative of an increased social inequality.
-Using burial data with historical documents, we are able to detect the processes of changes in Indigenous societies in the periphery of European colonial bases to better understand indirect impacts that are underestimated in this region [@Acabado2017; @Trabert2017].
-This study not only discuss pericolonal impacts that are not well understood yet, but also points out the agency of local people to manipulate colonial power instead of being passive recipients of trade goods.
+The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
+Before the arrival of Europeans, the burial network has a tendency of more closed relationships with ritual elements as the key formation mechanism.
+After European arrival, the network has a tendency of centralization relative to the rarity of goods.
+The changes in formation mechanisms for networks from ritual-oriented practices to wealth possession suggest increased behavior of wealth accumulation and differentiation stimulated by the European presence and associated long-distance trade network.
+This aligns with changes in social complexity from a more corporate to a more network mode society with different strategies for achieving power and different degrees of social inequality.
+Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
+@Acabado2017; @Trabert2017].
+Furthermore, our study also highlights the agency of local people to manipulate colonial power instead of being passive recipients of trade goods.
 
 <!-- technical and methodological implication-->
-This case study highlights the methodological improvement that Bayesian inference on ERGM can inform and enhance studies in rational data in archaeology.
+This case study demonstrates the methodological improvement that Bayesian inference on ERGM provides to inform and enhance studies of relational data in archaeology.
 A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-Moreover, probabilistic inference provides robust results of interrelations that are not statistically tested by traditional network approaches.
-Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships.
-This allows in-depth discussion of dynamic processes of relationship formation and their underlying factors that are informative of historical trajectories or socio-cultural phenomena.
+Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
+This enables insights into the dynamic processes of relationship formation and the underlying factors of historical trajectories of socio-cultural phenomena.
+
+<!-- stopped editing here -->
 

### 2021-01-05 analysis/paper/paper.Rmd
source 002-005 script files in rmd and cache the data

@@
 image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
+image_width_3_col_inch = 190/20
 ```
@@
 
-# Burial data from Kiwulan, Northeastern Taiwan
+# Materials
 
@@
 <!--archaeological setting-->
-We analyzed burial data collected from the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD) located in northeastern Taiwan, which experienced European colonial impacts in the early 17th century and a large wave of Chinese immigrants in the 19th century [@Chen2007].
+We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD).
+The settlement experienced European colonial impacts in the early 17th century and a large wave of Chinese immigrants in the 19th century [@Chen2007].
 Excavations revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
@@
 
-```{r, burial-map, fig.cap = "Map illustrating the location of burials at Kiwulan (suare size is 4 X 4 meters)", fig.width = image_width_2_col_inch}
-knitr::include_graphics(here::here("analysis", "figures", "burial-map.png"))
+```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("001-data-tidy.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
+```
+
+```{r burial-map, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters).
+The gray dots are post holes.", fig.width = image_width_2_col_inch}
+KWL_burial_map
 ```
@@
 
-## Data and Hypothesis
+```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("003-burials-pre-network-modelling.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
+```
+
+```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("004-burials-post-network-modelling.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
+```
+
+## Hypothesis and construction of networks
+
+<!-- our hypothesis --> 
+Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods.
+We can identify these high value trade goods from historical documents written by Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
+A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power.
+If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
+To test this prediction, we use ERGM in a Bayesian framework to examine the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
+By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
 
 <!-- burial data and the definition of tie formation  --> 
-We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan.
+We used burial goods in common to construct ties for representing a flow of good between burials.
 Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials.
@@
 Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
-
-<!-- our hypothesis --> 
-Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods.
-We can identify these high value trade goods from historical documents written by Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
-A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power.
-If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
-To test this prediction, we use ERGM in a Bayesian framework to examine the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
-By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
 
@@
 
-```{r BERGMs}
+```{r make-network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
 ```
@@
 <!--posterior stats for pre-E-->
+The networks constructed from observed burial data show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:network-diagrams)).
 We compared estimated statistics from the posterior distributions to examine their differences in structure of the simulated networks (Table XX).
@@
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
+
+```{r network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
+The size of the nodes is proportional to the node degrees that means the number of connections to a node.
+The thickness of tie represents the number of goods in common between nodes.", fig.width = image_width_3_col_inch}
+diagrams_two_phases
+```
 

### 2021-01-06 analysis/paper/paper.Rmd
correct chunk header and delete some duplicated figures

@@
 image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
-image_width_3_col_inch = 190/20
+image_width_3_col_inch = 190/19
 ```
@@
 
-```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("001-data-tidy.R"))}
+```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("002-burial-location.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))

### 2021-01-06 analysis/paper/paper.Rmd
change figure names

@@
 image_width_3_col_inch = 190/19
+
 ```
@@
 
-```{r burial-map, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters).
-The gray dots are post holes.", fig.width = image_width_2_col_inch}
-KWL_burial_map
+```{r figure-burial-map, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters). The gray dots are post holes."}
+knitr::include_graphics(here::here("analysis", "figures", "002-burial-map.png"))
 ```
@@
 
-```{r ergm-terms-table}
+```{r table-ergm-terms}
 ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
@@
 
-```{r make-network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
-```
-
 <!--posterior stats for pre-E-->
@@
 
-```{r network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
-The size of the nodes is proportional to the node degrees that means the number of connections to a node.
-The thickness of tie represents the number of goods in common between nodes.", fig.width = image_width_3_col_inch}
-diagrams_two_phases
+```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
+```
+
+```{r figure-network-diagrams, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters). The gray dots are post holes."}
+knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
 ```

### 2021-01-06 analysis/paper/002-burial-location.R

@@ -71,4 +71,5 @@ KWL_burial_map <-
   blank() +
   north(AD_data, scale = 0.2, symbol = 3)
 
-ggsave(here::here("analysis", "figures", "burial-map.png"))
+ggsave(here::here("analysis", "figures", "002-burial-map.png"),
+       w = 8, h = 4)

### 2021-01-06 analysis/paper/005-network-diagrams-two-phases.R

@@ -134,8 +134,7 @@ diagrams_two_phases <-
   plot_grid(pre_diagram,
             post_diagram,
             labels = c('A', 'B'),
-            label_size = 12,
-            w = 10, h = 4)
+            label_size = 12)
 
-ggsave(here::here("analysis", "figures", "networks_two_phases.png"),
-       w = 10, h = 4)
+ggsave(here::here("analysis", "figures", "005-network-diagrams.png"),
+       w = 8, h = 3)

### 2021-01-07 analysis/paper/paper.Rmd
source 006 file in rmd and cross-reference figures

@@
 
+```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
+```
+
 <!-- burial data and the definition of tie formation  --> 
@@
 Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
+The networks constructed based on this principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).
+
+```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
+The size of the nodes is proportional to the node degrees that means the number of connections to a node.
+The thickness of tie represents the number of goods in common between nodes."}
+knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
+```
 
@@
 
+```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("006-bergm-distribution-stats.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "006-bergm-distribution-stats.R")))
+```
+
 <!--posterior stats for pre-E-->
-The networks constructed from observed burial data show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:network-diagrams)).
-We compared estimated statistics from the posterior distributions to examine their differences in structure of the simulated networks (Table XX).
+We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table XX; Figure \@ref(fig:figure-posterior-distribution)).
 For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials.
@@
 
-```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
-```
-
-```{r figure-network-diagrams, fig.cap = "Map illustrating the location of burials by periods at the central excavation area of Kiwulan (each square is 4 X 4 meters). The gray dots are post holes."}
-knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
+```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, GWESP, GWDeg, and distance-dyadcov by phases.
+The pre-European group presents remarkedly larger values for the GWESP, but lower vlaues for the edges and GWDeg parameters comapred to the post-Eruoepan group.
+The distance parameters overlap between two groups."}
+knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
 ```
@@
 This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation.
-In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network.
+In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network .
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
@@
 We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
-Figure \@ref(fig:xxxx) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
+Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
 For the variance of degree, the values from the modeled post-European distribution are closer to the observed distribution.
@@
 However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.
+
+```{r figure-distribution-moments, fig.cap = "Comparison of statistics calculated on the observed and simulated GOF distributions for models between phase groups "}
+knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
+```
 

### 2021-01-07 analysis/paper/006-bergm-distribution-stats.R

@@ -56,8 +56,7 @@ histogram <-
             degree_value,
             dist_value)
 
-# density plots of posterior distribution
-
+# density plots for posterior distribution
 bergm_two_phases_longer <-
   bergm_two_phases %>%
   select(edges, gwesp, gwdeg.fixed.0.5, dyadcov.dist, phase) %>%
@@ -105,3 +104,6 @@ ggplot(distribution_two_phases_longer,
                               size = 3) +
   facet_wrap(~moments, scales = "free") +
   theme_minimal()
+
+ggsave(here::here("analysis", "figures", "006-distribution-moments.png"),
+       w = 8, h = 5)

### 2021-01-08 analysis/paper/paper.Rmd
add table of pre-E posterior stats

@@
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, account for 10%), above average (= 12-28), and below average (< 12), as an index of wealth.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, about 10 percent), above average (= 12-28), and below average (< 12), as an index of wealth.
 The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
@@
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
+
+```{r table-posterior-stats}
+pre_posterior_esti <-
+  data.frame("Parameter" = pre_bergm$specs,
+             round(summary(pre_bergm$Theta)$statistics[,1:2], 2),
+             round(summary(pre_bergm$Theta)$quantiles[,c(1,3,5)], 2),
+             row.names = NULL) %>% 
+  rename("0.025" = X2.5., "Median" = X50., "0.975" = X97.5.)
+
+post_posterior_esti <-
+  data.frame("Parameter" = post_bergm$specs,
+             round(summary(post_bergm$Theta)$statistics[,1:2], 2),
+             round(summary(post_bergm$Theta)$quantiles[,c(1,3,5)], 2),
+             row.names = NULL) %>% 
+  rename("0.025" = X2.5., "Median" = X50., "0.975" = X97.5.)
+
+knitr::kable(pre_posterior_esti,
+             caption = "Estimated posterior means, standard deviation, medians, and 95% credible intervals for the pre-Euroepan model ")
+```
 

### 2021-01-10 analysis/paper/paper.Rmd
update bergm code and cache

--- 2021-01-07 analysis/paper/paper.Rmd
@@
 <!--posterior stats for pre-E-->
-We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table XX; Figure \@ref(fig:figure-posterior-distribution)).
+We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
 For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials.
@@
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
+
+```{r table-posterior-stats}
+# make summary table
+posterior_esti_two_phases<-
+  bergm_two_phases %>%
+  pivot_longer(!phase, names_to = "parameter", values_to = "posterior") %>%
+  mutate(parameter = factor(parameter,
+                            levels=c("edges", "nodematch.age", 
+                                     "nodematch.sex", "nodematch.ritual",
+                                     "nodematch.value", "gwesp",	
+                                     "gwdeg", "dyadcov.distance"))) %>% 
+  group_by(parameter, phase) %>% 
+  summarize(mean = round(mean(posterior), 2),
+            median = round(quantile(posterior, 0.5), 2),
+            `2.5%` = round(quantile(posterior, 0.025), 2),
+            `97.5%` = round(quantile(posterior, 0.075), 2))
+
+knitr::kable(posterior_esti_two_phases,
+             caption = "Estimated posterior means, standard deviation, medians, and 95% credible intervals for the pre-Euroepan model ")
+```
 

### 2021-01-10 analysis/paper/005-network-diagrams-two-phases.R

@@ -49,21 +49,20 @@ relation_tidy_pre %>%
 pre_diagram <-
   ggraph(relation_tidy_pre, layout = "stress") + #graphopt
   geom_edge_link(aes(width = as.factor(common_counts)),
-                 alpha = 0.5) +
+                 alpha = 0.4) +
   geom_node_point(aes(filter = !is.na(connections),
                       size = connections,
-                      color = connections,
-                      alpha = connections)) +
+                      color = connections)) + #alpha = connection
   scale_edge_width_manual(values = c(0.2, 0.5, 1)) +
-  #scale_size_continuous(breaks = c(2, 5, 8, 10, 13)) +
-  scale_color_viridis(direction = -1) +
-  guides(size = guide_legend(), color=guide_legend()) +
-  geom_node_text(aes(filter= !is.na(connections) & connections > 11,
-                     label = burial_label_rm),
-                 repel = TRUE) +
+  scale_size_continuous(breaks = c(2, 5, 8, 11, 13)) +
+  scale_color_viridis(direction = -1, breaks = c(2, 5, 8, 11, 13)) +
+  guides(size = guide_legend(order = 1), color=guide_legend(order = 1)) +
+  #geom_node_text(aes(filter= !is.na(connections) & connections > 11,
+                     #label = burial_label_rm),
+                 #repel = TRUE) + # no.21, 48, 85
   labs(edge_width = "common items") +
   theme_graph() +
-  theme(plot.margin = unit(rep(0.8,4), "cm")) #legend.position="none"
+  theme(plot.margin = unit(rep(0.8, 4), "cm")) #legend.position="none"
 
 #---------------------post-E network diagram using ggraph pkg----------------------------
 # number of connection based on starting nodes
@@ -113,17 +112,16 @@ relation_tidy_post %>%
 post_diagram <-
   ggraph(relation_tidy_post, layout = "stress") + #graphopt
   geom_edge_link(aes(width = as.factor(common_counts)),
-                 alpha = 0.5) +
+                 alpha = 0.4) +
   geom_node_point(aes(filter= !is.na(connections),
                       size = connections,
-                      color = connections,
-                      alpha = connections)) +
+                      color = connections)) +
   scale_edge_width_manual(values=c(0.2, 0.5, 1)) +
   scale_color_viridis_c(direction = -1) +
   guides(size = guide_legend(), color=guide_legend()) +
[…]

### 2021-01-10 analysis/paper/006-bergm-distribution-stats.R

@@ -13,11 +13,16 @@ colnames(post_bergm_stats) <- post_bergm$specs
 
 post_bergm_stats <-
   cbind(post_bergm_stats, phase = "post") %>%
-  rename("gwesp" = "gwesp.fixed.1.2") %>%  #gwnsp" = "gwnsp.fixed.1.2
-  rename("dyadcov.dist" = "dyadcov.post_distance_n.dist")
+  rename(gwesp = gwesp.fixed.1.2,
+         dyadcov.dist = dyadcov.post_distance_n.dist) #gwnsp" = "gwnsp.fixed.1.2
 
 bergm_two_phases <-
-  rbind(pre_bergm_stats, post_bergm_stats)
+  rbind(pre_bergm_stats, post_bergm_stats) %>%
+  rename(gwdeg = gwdeg.fixed.0.5,
+         `dyadcov.distance` = dyadcov.dist,
+         nodematch.sex = nodematch.gender,
+         nodematch.ritual = nodematch.ritual_pottery,
+         nodematch.value = nodematch.value_class)
 
 # histogram of posterior distribution for both phases
 edges_value <-
@@ -36,14 +41,14 @@ trans_value <-
 
 degree_value <-
   ggplot(bergm_two_phases,
-       aes(x = gwdeg.fixed.0.5, fill = phase)) +
+       aes(x = gwdeg, fill = phase)) +
   geom_histogram(alpha = 0.8, position = "identity",
                  binwidth = 0.5) +
   theme_minimal()
 
 dist_value <-
   ggplot(bergm_two_phases,
-         aes(x = dyadcov.dist , fill = phase)) +
+         aes(x = dyadcov.distance, fill = phase)) +
   geom_histogram(alpha = 0.8, position = "identity",
                  binwidth = 0.004) +
   theme_minimal()
@@ -59,17 +64,15 @@ histogram <-
 # density plots for posterior distribution
 bergm_two_phases_longer <-
   bergm_two_phases %>%
-  select(edges, gwesp, gwdeg.fixed.0.5, dyadcov.dist, phase) %>%
-  rename(gwdeg = gwdeg.fixed.0.5,
-         `distance-dyadcov` = dyadcov.dist) %>%
+  select(edges, gwesp, gwdeg, dyadcov.distance, phase) %>%
   pivot_longer(!phase, names_to = "parameter", values_to = "posterior") %>%
   mutate(parameter = factor(parameter,
[…]

### 2021-01-11 analysis/paper/paper.Rmd

@@
 Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
-In 1626, the Spanish built their forts in Keelung and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
+In 1626, the Spanish built their forts in Heping dao and Tamsui (Figure \@ref(fig:figure-KWL-map)), northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
 Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+The Dutch would give l
+ocal Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
 
 <!-- archaeological specific case study -->
-In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
+In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
 We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
@@
 This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
+
+```{r, figure-KWL-map, fig.cap = "The location of Kiwulan and other places with the European forts in northern Taiwan mentioned in the text (map modified from Wang and Marwick, 2020).
+Map data from naturalearthdata.com."}
+
+knitr::include_graphics(here::here("analysis", "figures", "000-KWL-map.png"))
+```
 
@@
 In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with _in-situ_ posts were also excavated.
-The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:burial-map)).
+The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
 Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
@@
 
-```{r table-ergm-terms}
-ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
-knitr::kable(ergm_terms,
-             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
-```
-
 <!-- Network construction and specification -->
 We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
-Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
+Table \@ref(tab:table-ergm-terms) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
@@
 
+```{r table-ergm-terms}
+ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
+knitr::kable(ergm_terms,
+             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
+```
+
 ```{r bergm-iteration}
@@
 aux_iter <- 4000
+bgof_si <-10000
 ```
@@
 Thus, we specified the prior of the edge density parameter to low for both network models.
-For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 3 (i.e.
-N(0, 3)) for both models.
-For physical distance between burials, we set a negative covariate effect (N(-1, 1)) to infer kinship-based proximity, e.g.
+For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
+N(0, 5)) for both models.
+For physical distance between burials, we also set a vague prior effect (N(0, 1)) to explore whether there is kinship-based proximity, e.g.
 stronger correlations for shorter distances.
 To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(2, 4)), lower popularity (N(-1, 1)), and higher covariate effect based on ritual activity (N(1, 5)) for network before European contact to indicate less social inequality.
-On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(1, 2)), and higher popularity (N(1, 2)) to model an increased social inequality.
+We set the priors to higher transitivity (N(2, 2)), lower popularity (N(-2, 3)), and higher covariate effect based on ritual activity (N(1, 5)) for network before European contact to indicate less social inequality and stress ritual element shared in corporate groups.
+On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(1, 3)), higher popularity (N(2, 3)), and higher covariate effect of burial values (N(2, 3)) to model an increased social inequality.
 This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
-The priors for density were also set the same with N(-3, 3) for both pre- and post-European networks.
+The priors for density were also set the same with N(-3, 3) for both pre- and post-European networks based on a common observation of network density.
 
@@
 Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
-We set 10,000 network graphs simulated from the estimated posterior distribution in ERGMs.
+We set `r bgof_si` network graphs simulated from the estimated posterior distribution in ERGMs.
 Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials.
-The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations.
-Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals.
-Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations.
+For the pre-European network simulations, the nodal covariates, ritual-homophily, have significant effects on the formation of relations between burials.
+The positive effect demostrates that burials with ritual pottery tend to form relations.
+Despite positive values for some covariates, such as age-homophily, sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals.
+Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not affect the formation of relations.
 For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect.
-The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
+The high positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
 The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones.
@@
 
-```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, GWESP, GWDeg, and distance-dyadcov by phases.
-The pre-European group presents remarkedly larger values for the GWESP, but lower vlaues for the edges and GWDeg parameters comapred to the post-Eruoepan group.
+```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, gwesp, gwdegree, and distance-dyadcov by phases.
+The pre-European group presents remarkedly larger values for the gwesp, but lower vlaues for the edges and gwdegree parameters comapred to the post-Eruoepan group.
 The distance parameters overlap between two groups."}
@@
 <!--posterior stats for for post-E and their comparison-->
-For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect.
-However, there are no significant effects for these nodal and dyadic covariates.
-We also found that the burials in the same wealth level tend not to form relations.
-Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect.
-In contrast to the pre-European network, the popularity (gwdegree) shows a significant positive effect.
+For the network after the arrival of Europeans, the nodal covariates of burial value-homophily show positive effects, while the ritual-homophily and sex-homophily have a negative effect.
+Despite the effect is not strong, it could indicate the burials in the same wealth level tend to form relations.
+On the contrary, the same sex burials and burial with ritual pottery in common tend not to form relations.
+There are no significant effects for age-homophily and the dyadic covariate, physical distance.
+Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
+In contrast, the popularity (gwdegree) shows a significantly higher positive effect than the effect of pre-European network.
 This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others.
 This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation.
-In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network .
+In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
-Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).
+Both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
 
-```{r vertex-bootstrap}
-boostrap <- 1000
+```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("007-bootstrap-CI.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "007-bootstrap-CI.R")))
 ```
@@
 Networks with unoverlapped intervals means there is a significant difference.
-We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks.
-The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
+We computed and compared endogenous network statistics, including density, popularity, and transitivity for our two networks.
+Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
 For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference.
-This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects some degree of difference.
+This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with smaller degree of difference.
 
-Using the same vertex bootstrap procedure, we can also test the difference of network populations with a t-test on density, transitivity, and popularity.
-The results show that there are significant differences in all network statistics.
+```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, indicated by 0 nodes removal, and networks with different number of node removal.
+The number of bootstrap samples is 1000."}
+knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
+```
+
+```{r bootstrap-t-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("008-bootstrap-t-test.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "008-bootstrap-t-test.R")))
+pval_den <- signif(density_b_test$p.value, 4)
+pval_trans <- signif(trans_b_test$p.value, 4)
+pval_star <- signif(star_b_test$p.value, 4)
+```
+
+The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and popularity (p = $`r pval_star`$).
 Moreover, we explored the sample size effect by removing nodes at certain percents (5-40%) for both networks and compared their confidence intervals.
-The results demonstrate a consistent difference between the two networks that means they are robust to the different sizes.
+The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
 In general, those statistical methods support the presence of anthropologically meaningful differences in the pre-and post-European networks.
@@
 
-```{r Bayesian-gof}
-# the grey lines mark the 95% intervals. X-axis means the proportion of nodes
+```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (left) and the post-European model (right).
+Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
+Red lines represent distributions of observed networks, and gray lines show the 95% intervals."}
+library(magick)
+pre_bgof_img <- image_read(here::here("analysis/figures/003-pre-bgof.png"))
+post_bgof_img <- image_read(here::here("analysis/figures/004-post-bgof.png"))
+image_append(c(pre_bgof_img, post_bgof_img))
 ```
@@
 <!--assessment of MCMC analysis and BGOF--> 
-Bayesian goodness of fit (GOF) diagnostics plots show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
 the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
 For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
-This also suggests that our observed network statistics do not follow a normal distribution, especially for the network after the arrival of the Europeans that presents a bimodal pattern.
+This also suggests that our
+models can reproduce networks that resemble the structural features of our observed networks.
 We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
-Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are closer to the observed values overall, compared to the values from the post-European network.
-For the variance of degree, the values from the modeled post-European distribution are closer to the observed distribution.
+Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network.
 In general, this demonstrates a better fit for the pre-European network.
@@
 
-```{r figure-distribution-moments, fig.cap = "Comparison of statistics calculated on the observed and simulated GOF distributions for models between phase groups "}
+```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, variance) calculated on the observed data and simulated distributions for pre- and post- European models."}
 knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
@@
 On the other hand, in the post-European network ties tend to be between burials within the same level of wealth.
-Moreover, in the post-European network, burials with ritual vessels tend not to be tied.
+Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied.
 This illustrates that different social and economic mechanisms determined relations between burials for the two periods.
 Before European arrival, ritual behaviors seem to have been important status indicators, relative to the use of prestige goods.
-After European presence was established, status indicators shifted from ritual to wealth differences.
+After European presence was established, status indicators shifted from ritual to wealth differences and sex is not associated with status.
 Physical distance between burials did not correlate to network formation for either period.
@@
 This may suggest that the use of burial goods is more individual-oriented instead of kinship-based.
-In addition, demographic information, such as age and sex, do not show any correlations with network formation for either period.
+In addition, demographic information, such as age and sex, do not show any correlations with network formation for the pre-European network and only sex shows small effect for the post European network.
 This may be due to the unavailability of age and sex values for many burials in our sample.

### 2021-01-11 analysis/paper/paper.Rmd
get the caching working as expected with correct file paths for the md5sum fn

@@
 
-```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("000-prep1-value-burials.R"))}
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "000-prep1-value-burials.R")))
@@
 
-```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("001-data-tidy.R"))}
+```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
@@
 
-```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("002-burial-location.R"))}
+```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-burial-location.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
@@
 
-```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("003-burials-pre-network-modelling.R"))}
+```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
@@
 
-```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("004-burials-post-network-modelling.R"))}
+```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-burials-post-network-modelling.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
@@
 
-```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("005-network-diagrams-two-phases.R"))}
+```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/005-network-diagrams-two-phases.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
@@
 
-```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("006-bergm-distribution-stats.R"))}
+```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/006-bergm-distribution-stats.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "006-bergm-distribution-stats.R")))
@@
 For the pre-European network simulations, the nodal covariates, ritual-homophily, have significant effects on the formation of relations between burials.
-The positive effect demostrates that burials with ritual pottery tend to form relations.
+The positive effect demonstrates that burials with ritual pottery tend to form relations.
 Despite positive values for some covariates, such as age-homophily, sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals.
@@
 
-```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("007-bootstrap-CI.R"))}
+```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/007-bootstrap-CI.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "007-bootstrap-CI.R")))
@@
 
-```{r bootstrap-t-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("008-bootstrap-t-test.R"))}
+```{r bootstrap-t-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/008-bootstrap-t-test.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "008-bootstrap-t-test.R")))

### 2021-01-11 analysis/paper/007-bootstrap-CI.R

@@ -272,7 +272,7 @@ CIs_two_nets_tidy %>%
                   position = position_dodge(0.3)) +
   #scale_color_viridis(discrete= T) +
   theme_minimal() +
-  labs(x = "confidensce intervals (95%)",
+  labs(x = "confidence intervals (95%)",
        y = "node removal (%)") +
   facet_wrap(~variable,
              scales = "free_x")

### 2021-01-12 analysis/paper/paper.Rmd
include cache directory in gitignore and delete '=' in the rmd

--- 2021-01-11 analysis/paper/paper.Rmd
@@
 Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials.
-Gold-foil beads are in levels of high (> 3), upper-middle (= 3), lower-middle (= 2), and low (= 1); carnelian beads are in levels of high (> 6), upper-middle (= 4-6), lower-middle (= 3), and low (= 1-2); glass beads are in levels of high (> 6), upper-middle (= 3-6), lower-middle (= 2), and low (= 1).
+Gold-foil beads are in levels of high (>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (>6), upper-middle (3-6), lower-middle (2), and low (1).
 That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>= 29, about 10 percent), above average (= 12-28), and below average (< 12), as an index of wealth.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>28, about 10 percent), above average (12-28), and below average (<12), as an index of wealth.
 The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.

### 2021-01-12 analysis/paper/paper.Rmd
working on fixing typos

--- 2021-01-11 analysis/paper/paper.Rmd
@@
 <!--burial study- network analysis-->
-Burials analysis by archaeologists is an approach to understand past social structure through the study of the physical traces of mortuary practices.
-The material cultures and biological records of burial behaviors can represent the social ranking or identity of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
+Burial analysis by archaeologists is an approach to understand past social structures through the study of the physical traces of mortuary practices.
+The material cultures and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
 Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
-For example, mortuary practices, including burial forms, grave goods, and ritual behaviors, that structure the material configuration of burials can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
-Based on network science and graph theories, the recent development of social network analysis provides many new tools to visualize and analyze relationships for archaeological data [@Borgatti2009; @Brughmans2013].
+For example, mortuary practices, including burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
+Based on network science and graph theory, the recent development of social network analysis provides many new tools to visualize and analyze relationships in archaeological data [@Borgatti2009; @Brughmans2013].
 Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013; @Collar2015].
-That is, a network can be viewed as a patterned aggregation that includes individual elements (i.e.
+A network can be viewed as a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
@@
 In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
-For example, past trade can be conceptualized as a network of individual entities connected by shared similarity, the flow of goods, to represent their interactions [@Collar2015, pp.
+For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
 4].
@@
 In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
-We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
+We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
 We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
-We ask: (1) does the observed burial data after the start of European presence resemble the network of our model?
-(2) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by burials networks?
+We ask: (1) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by analysis of burial networks?
 (3) what are the major variables affecting or forming the higher degree of popularity (e.g.
@@
 By answering these questions, this study helps to expand the use of statistical modeling for burials in understanding the indirect effects of a colonial presence on Indigenous groups.
-This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.
+This is important to understand reactions to indirect impacts of European colonization.
 
@@
 
-<!-- current approaches to network analysis: important to show how what we are doing is amazing and new and not done before by other archaeologists.
-We need a sentence or two to say what do archaeologists usually do for network statistics...
-and make it clear that we are doing something new -->
-
-Bayesian approach to Exponential Random Graph Models (ERGMs) is an effective way to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
-Current network analysis used by archaeologist is mostly restricted to a single rational structure without consideration of interaction between network variables.
-Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
-ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
-The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
-For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
-Such distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
-
 <!-- ERGM in archaeology, and limitations --> 
-The use of ERGM in archaeology is first introduced by @Brughmans2014, who studied the Iron age settlement pattern based on 159 Iron Age sites in Southern Spain by modeling inter-settlement visibility networks and visual control.
+ERGMs first appeared in archaeology with @Brughmans2014, who studied  Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
 Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
-Similarly, @Amati2019 reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms.
+Similarly, @Amati2019 reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms.
 These network models evaluated different hypotheses about the importance of three variables, proximity, inter-cultural items, and pottery types, on the presence of hub sites they observed.
@@
 
+<!-- current approaches to network analysis: important to show how what we are doing is amazing and new and not done before by other archaeologists.
+We need a sentence or two to say what do archaeologists usually do for network statistics...
+and make it clear that we are doing something new -->
+
+Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
+Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
+ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
+For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
+By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
+Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
+ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+
 <!-- Bayesian inference for Exponential Random Graph Models --> 
-Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporates prior information about the network configurations into the model and offers uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
+Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporate prior information about the network configurations into the model and offer uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
 In Bayesian analysis, posterior distribution is an updated probability distribution after combining the prior derived from previous data or assumptions and our data.

### 2021-01-13 analysis/paper/paper.Rmd
make a table with multiple column headers and knit

@@
 ```{r table-posterior-stats}
-# make summary table
+# tidy data for table 
 posterior_esti_two_phases<-
   bergm_two_phases %>%
-  pivot_longer(!phase, names_to = "parameter", values_to = "posterior") %>%
+  pivot_longer(!phase, 
+               names_to = "parameter", 
+               values_to = "posterior") %>%
   mutate(parameter = factor(parameter,
@@
             `2.5%` = round(quantile(posterior, 0.025), 2),
-            `97.5%` = round(quantile(posterior, 0.075), 2))
+            `97.5%` = round(quantile(posterior, 0.075), 2)) %>% 
+  pivot_wider(names_from = phase, 
+              values_from = c(mean, median, `2.5%`, `97.5%`)) %>% 
+  select(parameter, ends_with(c("pre", "post")))
 
-knitr::kable(posterior_esti_two_phases,
-             caption = "Estimated posterior means, standard deviation, medians, and 95% credible intervals for the pre-Euroepan model ")
+library(flextable)
+# add column labels
+ft <- flextable(posterior_esti_two_phases) %>% 
+  set_header_labels(mean_pre = "pre-Euorpean", median_pre = "pre-Euorpean", 
+                    `2.5%_pre` = "pre-Euorpean", `97.5%_pre` = "pre-Euorpean",
+                    mean_post = "post-Euorpean", median_post = "post-Euorpean", 
+                    `2.5%_post` = "post-Euorpean", `97.5%_post` = "post-Euorpean") %>%
+  merge_at(i = 1, j = 2:5, part = "header") %>% 
+  merge_at(i = 1, j = 6:9, part = "header") %>% 
+  add_header_row(values = c("", "mean", "median", "2.5%", "97.5%",
+                            "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
+  set_caption("Estimated posterior means, standard deviation, medians, and 95% credible intervals for the pre-Euroepan model.") 
+ft
 ```
@@
 
-```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, variance) calculated on the observed data and simulated distributions for pre- and post- European models."}
+```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, variance) calculated on the observed data and simulated distributions for the pre- and post- European models."}
 knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))

### 2021-01-13 analysis/paper/paper.Rmd
edit captions in the rmd and the ergm table

--- 2021-01-12 analysis/paper/paper.Rmd
@@
 knitr::kable(ergm_terms,
-             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
+             caption = "Network variables used for model specifications with assocaited interpeation for burial relations.")
 ```
@@
 ```{r table-posterior-stats}
-# make summary table
+# tidy data for the table of posterior stats 
 posterior_esti_two_phases<-
   bergm_two_phases %>%
-  pivot_longer(!phase, names_to = "parameter", values_to = "posterior") %>%
+  pivot_longer(!phase, 
+               names_to = "parameter", 
+               values_to = "posterior") %>%
   mutate(parameter = factor(parameter,
@@
             `2.5%` = round(quantile(posterior, 0.025), 2),
-            `97.5%` = round(quantile(posterior, 0.075), 2))
+            `97.5%` = round(quantile(posterior, 0.075), 2)) %>% 
+  pivot_wider(names_from = phase, 
+              values_from = c(mean, median, `2.5%`, `97.5%`)) %>% 
+  select(parameter, ends_with(c("pre", "post")))
 
-knitr::kable(posterior_esti_two_phases,
-             caption = "Estimated posterior means, standard deviation, medians, and 95% credible intervals for the pre-Euroepan model ")
+library(flextable)
+# add column labels
+ft <- flextable(posterior_esti_two_phases) %>% 
+  set_header_labels(mean_pre = "pre-Euorpean", median_pre = "pre-Euorpean", 
+                    `2.5%_pre` = "pre-Euorpean", `97.5%_pre` = "pre-Euorpean",
+                    mean_post = "post-Euorpean", median_post = "post-Euorpean", 
+                    `2.5%_post` = "post-Euorpean", `97.5%_post` = "post-Euorpean") %>%
+  merge_at(i = 1, j = 2:5, part = "header") %>% 
+  merge_at(i = 1, j = 6:9, part = "header") %>% 
+  add_header_row(values = c("", "mean", "median", "2.5%", "97.5%",
+                            "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
+  set_caption("Estimated posterior means, medians, and 95% credible intervals for each network parameter of two models.") 
+ft
 ```
@@
 
-```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, indicated by 0 nodes removal, and networks with different number of node removal.
+```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, the top one with 0 nodes removal, and networks with different number of node removal.
 The number of bootstrap samples is 1000."}
@@
 
-```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, variance) calculated on the observed data and simulated distributions for pre- and post- European models."}
+```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post- European models."}
 knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))

### 2021-01-13 analysis/paper/paper.Rmd
align up the text in Table 2 and adjust cell widths

@@
                             "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
+  align(j = 2:9, align = "right", part = "header") %>% 
+  align(i = 1, align = "center", part = "header") %>% 
+  width(j = ~ parameter, width = 1.3) %>% 
   set_caption("Estimated posterior means, medians, and 95% credible intervals for each network parameter of two models.")

### 2021-01-16 analysis/paper/paper.Rmd
update cross referencing and adjust table cell size

--- 2021-01-13 analysis/paper/paper.Rmd
@@
                             "mean", "median", "2.5%", "97.5%"), top = FALSE ) %>% 
+  align(j = 2:9, align = "right", part = "header") %>% 
+  align(i = 1, align = "center", part = "header") %>% 
+  width(j = ~ parameter, width = 1.3) %>% 
+  width(j = 2:9, width = 0.65) %>% 
   set_caption("Estimated posterior means, medians, and 95% credible intervals for each network parameter of two models.")
@@
 <!--archaeological implication based on the results-->
-The covariants for both networks, before and after the European presence, provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:xxx)).
+The covariants for both networks, before and after the European presence, provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
 Our covariants were age, sex, ritual pottery, and a wealth index based on the value of grave goods.

### 2021-01-18 analysis/paper/paper.Rmd
identify some paras to move to SI to keep in word count

@@
 
-<!-- intro to networks -->
+<!-- intro to networks: move to SI -->
 A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
@@
 
-<!-- two types of network analysis -->
+<!-- two types of network analysis: move to SI -->
 Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
@@
 Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
+<!-- keep --> This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
-# A Case Study from Northereastern Taiwan
+# A Case Study from Northeastern Taiwan
 
@@
 
-<!--fit in a bigger historical picture in a periclonial context -->
+<!--fit in a bigger historical picture in a periclonial context: move to SI -->
 The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
@@
 Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give l
-ocal Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
 
@@
 
-<!-- current approaches to network analysis: important to show how what we are doing is amazing and new and not done before by other archaeologists.
-We need a sentence or two to say what do archaeologists usually do for network statistics...
-and make it clear that we are doing something new -->
-
+<!-- current approaches to network analysis: move to SI -->
 Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
@@
 
-<!--archaeological setting-->
+<!-- archaeological setting -->
 We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD).
@@
 
-<!--chronology-->
+<!-- chronology -->
 To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
@@
 
-<!--Model fitting--> 
+<!--Model fitting: move to SI --> 
 After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
@@
 
-<!-- assessment of MCMC and Goodness-of-fit in Bayesian-->
+<!-- assessment of MCMC and Goodness-of-fit in Bayesian: move to SI -->
 As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
@@
 For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
-This also suggests that our
-models can reproduce networks that resemble the structural features of our observed networks.
+This also suggests that our models can reproduce networks that resemble the structural features of our observed networks.
 We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.

### 2021-01-19 analysis/paper/supplementary-material.Rmd
create supplementary rmd and delete test files

---
title: "Supplementary material- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
author:
  - Li-Ying Wang
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
---

### 2021-01-19 analysis/paper/paper.Rmd
commit before pulling

@@
 
-```{r, figure-KWL-map, fig.cap = "The location of Kiwulan and other places with the European forts in northern Taiwan mentioned in the text (map modified from Wang and Marwick, 2020).
+```{r site-map, figure-KWL-map, fig.cap = "The location of Kiwulan and other places with the European forts in northern Taiwan mentioned in the text (map modified from Wang and Marwick, 2020).
 Map data from naturalearthdata.com."}

### 2021-01-19 analysis/paper/paper.Rmd
move paragraphs to SI rmd

@@
 image_width_3_col_inch = 190/19
-
 ```
@@
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
+The details about archaeological networks provided in Supplementary Materials.
+This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
 <!-- intro to networks: move to SI -->
-A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
-The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
-For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
-4].
-Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
-Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
-Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
-@Coward2013, pp.
-252].
-This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
 
 <!-- two types of network analysis: move to SI -->
-Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
-There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
-Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
-By quantifying those network properties, archaeologists can answer a wide range of research questions.
-Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
-Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-<!-- keep --> This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
-# A Case Study from Northeastern Taiwan
+## A Case Study from Northeastern Taiwan
 
@@
 <!--fit in a bigger historical picture in a periclonial context: move to SI -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
-In 1626, the Spanish built their forts in Heping dao and Tamsui (Figure \@ref(fig:figure-KWL-map)), northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
-Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
 
@@
 <!-- ERGM in archaeology, and limitations --> 
-ERGMs first appeared in archaeology with @Brughmans2014, who studied  Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
+Exponential random graph models first appeared in archaeology with @Brughmans2014, who studied  Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
 Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
@@
 <!-- current approaches to network analysis: move to SI -->
-Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
-Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
-Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
-ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
-The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
-For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
-Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
 
@@
 <!--Model fitting: move to SI -->
-After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
-MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
-We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm.
-This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.
-More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
-In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
-In our case, we set the number of chains to six.
-For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`.
-We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.
 
@@
 <!-- assessment of MCMC and Goodness-of-fit in Bayesian: move to SI -->
-As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
-In general, we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
-Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
-We set `r bgof_si` network graphs simulated from the estimated posterior distribution in ERGMs.
-Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
-This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
-We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
-We expect to see the models fit with our hypotheses, indicative of a increased social inequality after the foreign contact.
-In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
 

### 2021-01-19 analysis/paper/supplementary-material.Rmd
move paragraphs to SI rmd

@@
 ---
+
+```{r setup, echo = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  warning = FALSE,
+  message = FALSE,
+  echo = FALSE,
+  comment = "#>",
+  dpi = 300,
+  fig.path = "../figures/"
+)
+
+image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
+image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
+```
+
+<!-- intro to networks: move to SI -->
+A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
+The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
+For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
+4].
+Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
+Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
+Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
+@Coward2013, pp.
+252].
+This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
+
+<!-- two types of network analysis: move to SI -->
+Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
+There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
+Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
+By quantifying those network properties, archaeologists can answer a wide range of research questions.
+Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
+Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
+Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+
+<!--fit in a bigger historical picture in a periclonial context: move to SI -->
+The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
+In 1626, the Spanish built their forts in Heping dao and Tamsui (Figure \@ref(fig:figure-KWL-map)), northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
+Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+
+<!-- current approaches to network analysis: move to SI -->
+Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
+Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
+ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
+For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
+By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
+Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
+ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+
+```{r bergm-iteration}
+burn_in <- 100
+main_iter <- 1000
+aux_iter <- 4000
+bgof_si <-10000
+```
+
+<!--Model fitting: move to SI --> 
+After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
+MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
+We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm.
+This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.
+More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
+In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
+In our case, we set the number of chains to six.
+For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`.
+We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.
+
+<!-- assessment of MCMC and Goodness-of-fit in Bayesian: move to SI -->
+As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
+In general, we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
+Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
+We set `r bgof_si` network graphs simulated from the estimated posterior distribution in ERGMs.
+Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
+We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
+We expect to see the models fit with our hypotheses, indicative of a increased social inequality after the foreign contact.
+In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.

### 2021-01-19 analysis/paper/paper.Rmd
add inline code for priors

@@
 
-```{r bergm-iteration}
-burn_in <- 100
-main_iter <- 1000
-aux_iter <- 4000
+```{r bergm-prior-and-iteration}
+prior_mean <- as_tibble(cbind(pre_prior_mean, 
+                              post_prior_mean)) %>% 
+  mutate(mean_same = ifelse(pre_prior_mean == post_prior_mean, pre_prior_mean, NA))
+
+prior_sigma <- as_tibble(cbind(diag(pre_prior_sigma), 
+                               diag(post_prior_sigma))) %>% 
+  rename(pre_prior_sigma = V1, post_prior_sigma = V2) %>% 
+  mutate(sigma_same = ifelse(pre_prior_sigma == post_prior_sigma, pre_prior_sigma, NA))
+
+priors <- cbind(prior_mean, prior_sigma)
 bgof_si <-10000
@@
 Thus, we specified the prior of the edge density parameter to low for both network models.
-For covariates based on burial attributes, such as age, sex, and our wealth index, we specified a vague prior that follows a normal distribution with mean at 0 and standard deviation at 5 (i.e.
-N(0, 5)) for both models.
-For physical distance between burials, we also set a vague prior effect (N(0, 1)) to explore whether there is kinship-based proximity, e.g.
+For covariates about biological features, such as age, sex, we specified a vague prior that follows a normal distribution with mean at `r priors$mean_same[2]` and standard deviations at `r priors$sigma_same[2]`for both models to explore their effects.
+For physical distance between burials, we also set a vague prior effect (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) to explore whether there is kinship-based proximity, e.g.
 stronger correlations for shorter distances.
 To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(2, 2)), lower popularity (N(-2, 3)), and higher covariate effect based on ritual activity (N(1, 5)) for network before European contact to indicate less social inequality and stress ritual element shared in corporate groups.
-On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(1, 3)), higher popularity (N(2, 3)), and higher covariate effect of burial values (N(2, 3)) to model an increased social inequality.
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect based on ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for network before European contact to indicate less social inequality and stress ritual element shared in corporate groups.
+On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social inequality.
 This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
-The priors for density were also set the same with N(-3, 3) for both pre- and post-European networks based on a common observation of network density.
+The priors for density were also set the same with N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`) for both pre- and post-European networks based on a common observation of network density.
 

### 2021-01-20 analysis/paper/supplementary-material.Rmd
move the description of bead level to SI and add raincloud plot

@@
 ---
-title: "Supplementary material- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
+title: "Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
 author:
@@
       reference_docx: "../templates/template.docx" # Insert path for the DOCX file
+bibliography: references.bib
 ---
@@
 ```
+
+# Introduction to network analysis and archaeological applications
 
@@
 
-<!--fit in a bigger historical picture in a periclonial context: move to SI -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
-In 1626, the Spanish built their forts in Heping dao and Tamsui (Figure \@ref(fig:figure-KWL-map)), northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
-Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
-
 <!-- current approaches to network analysis: move to SI -->
@@
 ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+
+# Kiwulan in the 17th northern Taiwan
+
+<!--fit in a bigger historical picture in a periclonial context: move to SI -->
+The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
+In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
+Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
+The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+
+# Method
+
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "000-prep1-value-burials.R")))
+```
+
+```{r explore-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "000-prep2-goods-class-plot.R")))
+```
+
+<!-- burial data and the definition of tie formation: partially move to SI  -->
+
+For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)).
+Gold-foil beads are in levels of high (>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (>6), upper-middle (3-6), lower-middle (2), and low (1).
+If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+
+```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
+# combine ridge plots 1 & 4 in the 000-prep2 file
+library(tidyverse)
+library(ggridges)
+ridge_1_and_4 <-
+  burial_beads %>%
+  mutate(type = case_when(
+    type == "Golden_bead" ~"gold-foil bead",
+    type == "all_glass_bead" ~"glass bead",
+    type == "Agate_bead" ~"carnelian bead")) %>% 
+  ggplot(aes(x = value,
+             y = type,
+             fill = factor(stat(quantile)))) +
+  stat_density_ridges(
+    geom = "density_ridges_gradient",
+    calc_ecdf = TRUE,
+    quantiles = 4,
+    quantile_lines = TRUE,
+    jittered_points = TRUE,
+    alpha = 0.7,
+    vline_size = 0.5,
+    vline_color = "grey10",
+    point_size = 2.5,
+    point_alpha = 0.4,
+    position = position_raincloud(adjust_vlines = TRUE),
+    rel_min_height = 0.01,
+    scale = 0.5, # so the filled regions don't overlap on the points
+    size = 0.2) +
+  scale_x_continuous(limits = c(1, 15),
+                     expand = c(0.1, 0)) +
+  scale_fill_viridis_d(name = "Quartiles") +
+  theme_minimal()
+
+#ggsave(here::here("analysis", "figures", "000-raincloud-beads.png"),
+       #w = 6, h = 4)
+```
+
+```{r figure-raincloud-beads, fig.cap = "The distributions of different types of trade beads across burials analyzed in this paper."}
+
+knitr::include_graphics(here::here("analysis", "figures", "000-raincloud-beads.png"))
+```
 
@@
 Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
-We set `r bgof_si` network graphs simulated from the estimated posterior distribution in ERGMs.
+We set $`r bgof_si`$ network graphs simulated from the estimated posterior distribution in ERGMs.
 Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
@@
 In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
+
+# References
+
+<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
+<div id="refs"></div>
+
+##### pagebreak
+
+### Colophon
+
+This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+
+```{r colophon, cache = FALSE}
+# which R packages and versions?
+if ("devtools" %in% installed.packages()) devtools::session_info()
+```
+
+The current Git commit details are:
+
+```{r}
+# what commit is this file at? 
+if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
+```
+
+Word count: `r wordcountaddin::word_count()`

### 2021-01-20 analysis/paper/000-prep2-goods-class-plot.R

@@ -104,32 +104,6 @@ library(cowplot)
 plot_grid(ridge_1, ridge_2, ridge_3, ridge_4,
           ncol = 2)
 
-# combine ridge plots 1 & 4 for supplementary materials
-ridge_1_and_4 <-
-  burial_beads %>%
-  ggplot(aes(x = value,
-             y = type,
-             fill = factor(stat(quantile)))) +
-  stat_density_ridges(
-    geom = "density_ridges_gradient",
-    calc_ecdf = TRUE,
-    quantiles = 4,
-    quantile_lines = TRUE,
-    jittered_points = TRUE,
-    alpha = 0.7,
-    vline_size = 0.5,
-    vline_color = "grey10",
-    point_size = 2.5,
-    point_alpha = 0.4,
-    position = position_raincloud(adjust_vlines = TRUE),
-    rel_min_height = 0.01,
-    scale = 0.5, # so the filled regions don't overlap on the points
-    size = 0.2) +
-  scale_x_continuous(limits = c(1, 15),
-                     expand = c(0.1, 0)) +
-  scale_fill_viridis_d(name = "Quartiles") +
-  theme_minimal()
-
 # orientation
 burial %>%
   mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%

### 2021-01-21 analysis/paper/paper.Rmd

@@
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
-The details about archaeological networks provided in Supplementary Materials.
-This paper will use a novel Bayesian approach on exponential random graph model to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
+The details about archaeological networks provided in Supplementary Online Materials.
+This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
 
@@
 
-```{r site-map, figure-KWL-map, fig.cap = "The location of Kiwulan and other places with the European forts in northern Taiwan mentioned in the text (map modified from Wang and Marwick, 2020).
+```{r figure-KWL-map, fig.cap = "The location of Kiwulan and the European forts at Heping dao and Tamsui in northern Taiwan (modified from Wang and Marwick, 2020).
 Map data from naturalearthdata.com."}
@@
 The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and `in-situ` wooden posts, indicating a well organized spatial arrangement.
+Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, indicating a well organized spatial arrangement.
 Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
@@
 @Coward2013, pp.
-252].Therefore, we built networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
+252].
+Therefore, we built networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
 The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments.
@@
 We used burial goods in common to construct ties for representing a flow of good between burials.
-Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials.
-Gold-foil beads are in levels of high (>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (>6), upper-middle (3-6), lower-middle (2), and low (1).
-That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
 For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e.

### 2021-01-21 analysis/paper/supplementary-material.Rmd
stash

@@
 
-```{r explore-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
+```{r explore-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep2-goods-class-plot.R"))}
 suppressWarnings(source(here::here("analysis", "paper", "000-prep2-goods-class-plot.R")))
+```
+
+```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
+```
+
+```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-burial-location.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
 ```
@@
 <!-- assessment of MCMC and Goodness-of-fit in Bayesian: move to SI -->
+
+```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
+```
+
+```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
+```
+
+```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-burials-post-network-modelling.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
+```
+
+```{r diag-plots}
+png(filename = here::here("analysis", "figures", "008-pre-diag.png"),
+    width = 10, height = 8, units = "in", res = 360)
+plot(pre_bergm)
+dev.off()
+```
+
 As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.

### 2021-01-21 analysis/paper/supplementary-material.Rmd
delete replicated code chunk
get mcmc diagnostic plots all on one plot

@@
 
-```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
-```
-
 ```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
@@
 ```{r diag-plots}
+# get the first set of diagnostic plots
+x <- pre_bergm
+
 png(filename = here::here("analysis", "figures", "008-pre-diag.png"),
-    width = 10, height = 8, units = "in", res = 360)
-plot(pre_bergm)
+    width = 10, height = 12, units = "in", res = 360)
+
+seqq <- 4
+par(mfrow = c(x$dim, 3), 
+    oma   = c(0, 0, 3, 0), 
+    mar   = c(4, 3, 1.5, 1))
+
+for (i in 1:x$dim) {
+
+  plot(density(x$Theta[, i]), 
+       main = "", 
+       axes = FALSE, 
+       xlab = bquote(paste(theta[.(i)], " (", .(x$specs[i]), ")")),
+       ylab = "", lwd = 2)
+  axis(1)
+  axis(2)
+  coda::traceplot(x$Theta[, i], type = "l", xlab = "Iterations", ylab = "")
+  coda::autocorr.plot(x$Theta[, i], auto.layout = FALSE)
+  if (x$dim > 4) seqq <- seq(4, x$dim, 4)
+  if (i %in% union(x$dim, seqq)) {
+    title(paste("MCMC output for Model: y ~", x$formula[3]), outer = TRUE)
+  }
+}
+
 dev.off()
@@
 In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
+
+```{r xx1, fig.cap = "xx1"}
+knitr::include_graphics(here::here("analysis", "figures", "008-pre-diag.png"))
+```
 

### 2021-01-22 analysis/paper/supplementary-material.Rmd
try code to save another set of plot

@@
 bibliography: references.bib
+csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
 ---
@@
 
-```{r diag-plots}
+```{r pre-diag-plots, include= FALSE}
 # get the first set of diagnostic plots
@@
 
-As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots [@Hamra2013], which we include in our Supplementary Online Materials.
-In general, we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.
+```{r post-diag-plots, include= FALSE}
+# get the first set of diagnostic plots
+y <- post_bergm
+
+png(filename = here::here("analysis", "figures", "009-post-diag.png"),
+    width = 10, height = 12, units = "in", res = 360)
+
+seqq <- 4
+par(mfrow = c(y$dim, 3), 
+    oma   = c(0, 0, 3, 0), 
+    mar   = c(4, 3, 1.5, 1))
+
+for (i in 1:y$dim) {
+
+  plot(density(y$Theta[, i]), 
+       main = "", 
+       axes = FALSE, 
+       xlab = bquote(paste(theta[.(i)], " (", .(y$specs[i]), ")")),
+       ylab = "", lwd = 2)
+  axis(1)
+  axis(2)
+  coda::traceplot(y$Theta[, i], type = "l", xlab = "Iterations", ylab = "")
+  coda::autocorr.plot(y$Theta[, i], auto.layout = FALSE)
+  if (y$dim > 4) seqq <- seq(4, y$dim, 4)
+  if (i %in% union(y$dim, seqq)) {
+    title(paste("MCMC output for Model: y ~", y$formula[3]), outer = TRUE)
+  }
+}
+
+dev.off()
+```
+
+As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots for both models (Figure \@ref(fig:figure-pre-diag) and Figure \@ref(fig:figure-post-diag)) [@Hamra2013].
+In general, models with informative priors have a better convergence of the MCMC.
+We can see that the diagnostic plots show stationary distributions and a trend of decreased autocorrelation.
 Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
@@
 
-```{r xx1, fig.cap = "xx1"}
+```{r figure-pre-diag, fig.cap = "MCMC output of the pre-European model. In the order of edges, nodematch-age, nodematch-sex, nodematch-ritual, nodematch-value, gwesp, gwdeg, and dyadcov.distance."}
 knitr::include_graphics(here::here("analysis", "figures", "008-pre-diag.png"))
+```
+
+```{r figure-post-diag, fig.cap = "MCMC output of the post-European model. In the order of edges, nodematch-age, nodematch-sex, nodematch-ritual, nodematch-value, gwesp, gwdeg, and dyadcov.distance."}
+knitr::include_graphics(here::here("analysis", "figures", "009-post-diag.png"))
 ```

### 2021-01-23 analysis/paper/supplementary-material.Rmd
removed characters on the top of MCMC diagnostic plots

@@
   if (x$dim > 4) seqq <- seq(4, x$dim, 4)
-  if (i %in% union(x$dim, seqq)) {
-    title(paste("MCMC output for Model: y ~", x$formula[3]), outer = TRUE)
-  }
 }
@@
   if (y$dim > 4) seqq <- seq(4, y$dim, 4)
-  if (i %in% union(y$dim, seqq)) {
-    title(paste("MCMC output for Model: y ~", y$formula[3]), outer = TRUE)
-  }
 }

### 2021-01-24 analysis/paper/paper.Rmd
correct variable name and knit

@@
 
-# Introduction
+# Introduction {-}
 
@@
 
-## A Case Study from Northeastern Taiwan
+## A Case Study from Northeastern Taiwan  {-}
 
@@
 
-# Exponential random graph models in a Baysesian framework
+# Exponential random graph models in a Baysesian framework  {-}
 
@@
 
-# Materials
+# Materials  {-}
 
@@
 
-# Methods
+# Methods {-}
 
@@
 
-## Hypothesis and construction of networks
+## Hypothesis and construction of networks {-}
 
@@
 
-## Model specification in a Bayesian framework
+## Model specification in a Bayesian framework {-}
 
@@
 
-# Reproducibility and open source materials
+# Reproducibility and open source materials {-}
 
@@
 
-# Results
+# Results {-}
 
@@
 
-# Discussion
+# Discussion {-}
 
@@
 
-# Conclusion
+# Conclusion {-}
 
@@
 
-# Acknowledgments
+# Acknowledgments {-}
 
@@
 
-# References 
+# References {-} 
 <!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

### 2021-01-24 analysis/paper/006-bergm-distribution-stats.R

@@ -20,7 +20,6 @@ bergm_two_phases <-
   rbind(pre_bergm_stats, post_bergm_stats) %>%
   rename(gwdeg = gwdeg.fixed.0.5,
          `dyadcov.distance` = dyadcov.dist,
-         nodematch.sex = nodematch.gender,
          nodematch.ritual = nodematch.ritual_pottery,
          nodematch.value = nodematch.value_class)
 

### 2021-01-25 analysis/paper/paper.Rmd
combine prep1 and prep2 in one script file, edit code, and knit

@@
 
-```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "000-prep1-value-burials.R")))
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep-data.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "000-prep-data.R")))
 ```

### 2021-01-25 analysis/paper/supplementary-material.Rmd
correct variable name and knit
combine prep1 and prep2 in one script file, edit code, and knit

@@
 
-# Introduction to network analysis and archaeological applications
+# Introduction to network analysis and archaeological applications {-}
 
@@
 
-# Kiwulan in the 17th northern Taiwan
+# Kiwulan in the 17th northern Taiwan {-}
 
@@
 
-# Method
+# Method {-}
 
-```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep1-value-burials.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "000-prep1-value-burials.R")))
-```
-
-```{r explore-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep2-goods-class-plot.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "000-prep2-goods-class-plot.R")))
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep-data.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "000-prep-data.R")))
 ```
@@
 ```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
-# combine ridge plots 1 & 4 in the 000-prep2 file
+# combine ridge plots 1 & 4 in the 000-prep-data file
 library(tidyverse)
@@
 
-# References
+# References {-}
 

### 2021-01-25 analysis/paper/000-prep-data.R

@@ -1,0 +1,170 @@
+# Explore and prepare burial data for network analysis:
+# 1. create burial value as a new variable
+# 2. examine bead distribution to guide group assignment for making ties in 001 file
+library(readxl)
+library(tidyverse)
+library(here)
+
+burial <- read_excel(here("analysis", "data", "raw_data", "Kiwulan_Burials.xlsx"))
+
+# tidy up data, select and combine parameters
+burial_goods_selected <-
+  burial %>%
+  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
+  filter(!is.na(Phase)) %>%
+  mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
+         Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
+         Glass_bead = ifelse(Glass_bead == "shatter", "1", Glass_bead),
+         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
+  mutate_at(21:ncol(.), as.numeric) %>%
+  janitor::remove_empty(which = "cols") %>%
+  mutate(total = rowSums(.[c(21:48, 50, 55, 56)], na.rm = TRUE)) %>% # prestige goods
+  mutate(Porcelain = rowSums(.[c(40:48, 56)], na.rm = TRUE)) %>% # B&W, porcelain, kendi
+  mutate(Porcelain = ifelse(Porcelain == 0, NA, Porcelain)) %>%
+  mutate(Stonewares = rowSums(.[c(50, 55)], na.rm = TRUE)) %>% # stoneware, Anping jars
+  mutate(Stonewares = ifelse(Stonewares == 0, NA, Stonewares)) %>%
+  mutate(Metal_bangles = rowSums(.[c(26, 28, 29)], na.rm = TRUE)) %>% # different bangles
+  mutate(Metal_bangles = ifelse(Metal_bangles == 0, NA, Metal_bangles)) %>%
+  mutate(Metal_pendant = rowSums(.[c(34, 36)], na.rm = TRUE)) %>% # different bangles
+  mutate(Metal_pendant= ifelse(Metal_pendant == 0, NA, Metal_pendant)) %>%
+  mutate(all_glass_bead = rowSums(.[c(23, 24)], na.rm = TRUE)) %>%
+  select(ID,
+         Phase,
+         Degree_axis,
+         Golden_bead,
+         Agate_bead,
[…]

### 2021-01-25 analysis/paper/000-prep2-goods-class-plot.R

@@ -1,114 +1,0 @@
-# this file explores burial data that can be ran individually and separately
-# examines distribution of beads to guide group assignment for making ties in 001 file
-
-# tidy up
-burial_three_period_tidy_explore <-
-  burial %>%
-  rename(burial_label = ID) %>%
-  mutate(Phase = ifelse(Phase == 'euro', 'post', Phase)) %>%
-  filter(!is.na(Phase)) %>%
-  mutate(Gold_leaf = ifelse(Gold_leaf == "shatter", "1", Gold_leaf),
-         Stoneware = ifelse(Stoneware == "base", "1", Stoneware),
-         Stamped_ceramic = ifelse(Stamped_ceramic == "cluster", "1", Stamped_ceramic)) %>%
-  mutate_at(21:ncol(.), as.numeric) %>%
-  janitor::remove_empty(which = "cols") %>%
-  mutate(total = rowSums(.[c(21:48, 50, 55, 56)], na.rm = TRUE)) %>%
-  mutate(quantity = case_when(
-    total == 0 ~ "none",
-    total > 0 & total <= 7 ~ "low",
-    total > 7 & total < 100 ~ "medium",
-    total >= 100 ~ "high",
-    TRUE ~ "other")) %>% # the classification is based on the result of histogram
-  rowwise() %>%
-  mutate(all_glass_bead = sum(Glass_bead, `Indo-Pacific_bead`, na.rm = TRUE)) %>%
-  mutate(Gold_bead_low = ifelse(Golden_bead == 1, 1, NA),
-         Gold_bead_med = ifelse(Golden_bead > 1 & Golden_bead <10, 1, NA),
-         Gold_bead_high = ifelse(Golden_bead > 10, 1, NA),
-         Agate_bead_low = ifelse(Agate_bead == 1, 1, NA),
-         Agate_bead_med = ifelse(Agate_bead > 1 & Agate_bead <10, 1, NA),
-         Agate_bead_high = ifelse(Agate_bead > 10, 1, NA),
-         `Indo-Pacific_bead_low` = ifelse(`Indo-Pacific_bead` < 100, 1, NA),
-         `Indo-Pacific_bead_med` = ifelse(`Indo-Pacific_bead` > 100 & `Indo-Pacific_bead` < 900, 1, NA),
-         `Indo-Pacific_bead_high` = ifelse(`Indo-Pacific_bead` > 900, 1, NA)) #based on the result of histogram
-
-# long format for bead
[…]

### 2021-01-25 analysis/paper/004-burials-post-network-modelling.R

@@ -43,7 +43,7 @@ common_counts_lst_post <-
          rowwise() %>%
          mutate(common_counts = sum(count...3, count...6)))
 
-# count of ornament types in common between each pair of burials
+  # count of ornament types in common between each pair of burials
 common_counts_vct_post <- map_int(common_counts_lst_post, ~sum(!is.na(.x$common_counts)))
 
 burial_comb_with_common_counts_post <-

### 2021-01-25 analysis/paper/paper.Rmd
add a reference and highlights

@@
 highlights: |
-  These are the highlights. 
+  1. We present Bayesian inference on Exponential random graph models (ERGMs) to model archaeological networks.
+  2. We test social processes by simulating two sets of burial data from different time.
+  3. Differences in network statistics implies changes in social structures influenced by foreign presence. 
+  4. Bayesian ERGMs is efficient and useful for evaluating and interpreting networks. 
 ---
@@
 Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
-The introduction of prestige goods to local Indigenous communities might occur earlier before a European presence became established, but was amplified during the European colonial period [@Borao2009; @LiandWu2006; @LWandBM2020ornament].
+The use of trade items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
 A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks [@Brumfiel1994; @Clark1994].

### 2021-01-26 analysis/paper/paper.Rmd
edit the paper rmd from intro to method

@@
 Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
-For example, mortuary practices, including burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, can represent social relations between members of a community, allowing the inference of social structures [@Byrd1995; @Seikel2011].
+By studying burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, we can infer social status and relations between members of a community that give an insight to social structures [@Byrd1995; @Seikel2011].
 Based on network science and graph theory, the recent development of social network analysis provides many new tools to visualize and analyze relationships in archaeological data [@Borgatti2009; @Brughmans2013].
-Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation, based on our specification in terms of network theories [@Brandes2013; @Collar2015].
+Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation [@Brandes2013; @Collar2015].
 A network can be viewed as a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
-The details about archaeological networks provided in Supplementary Online Materials.
-This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organization.
+Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
+@Coward2013, pp.
+252].
+This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign presence on social organization.
 
@@
 <!-- archaeological background -->
-Archaeological sites in northeastern Taiwan show evidence of imported prestige goods, such as ornaments and porcelains, in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007; @LWandBM2020ornament].
-The use of imported goods is also mentioned in historical records from the Spanish, indicative of the pursuit of prestige or wealth in those Indigenous societies [@Borao2009; @LiandWu2006].
+Archaeological sites in northeastern Taiwan show evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and the use as burial goods around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament].
+Imported goods were treated as prestige or wealth in those Indigenous societies that were commonly mentioned in Spanish historical records [@Borao2009; @LiandWu2006].
 For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006].
 In addition, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
-Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear.
-The use of trade items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
-A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks [@Brumfiel1994; @Clark1994].
-The observed uneven distribution may be explained as a result of social change from less social inequality to more social inequality when Indigenous societies in Northeastern Taiwan were involved in the complex trade network stimulated by the Europeans.
+Despite we observed the uneven distribution of imported goods in burials, the degree of differentiation across burials over time remain unclear.
+The use of imported items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
+A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade networks when Indigenous societies were involved in the complex trade network stimulated by the Europeans.
 
@@
 <!-- archaeological specific case study -->
-In this study, we explore the changes in social inequality in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
+In this study, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
 We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
@@
 We ask: (1) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by analysis of burial networks?
-(3) what are the major variables affecting or forming the higher degree of popularity (e.g.
+(2) what are the major variables affecting or forming the higher degree of popularity (e.g.
 a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period?
@@
 Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
-Similarly, @Amati2019 reconstructed three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms.
-These network models evaluated different hypotheses about the importance of three variables, proximity, inter-cultural items, and pottery types, on the presence of hub sites they observed.
-They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
+Similarly, @Amati2019 modeled three networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms, including proximity, inter-cultural items, and pottery types.
+By comparing hypothesized networks with the observed sites, they found that the presence of hub sites can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
 However, those studies also point out some limitations of ERGM, such as sensitivity to missing data and less ability to handle uncertainty.
@@
 <!-- Bayesian inference for Exponential Random Graph Models --> 
-Bayesian approaches to ERGMs are efficient computational tools for social network analysis because they incorporate prior information about the network configurations into the model and offer uncertainty quantification by evaluating the posterior distributions of the parameters associated with network configurations [@Caimo2015; @Nemmers2019].
-In Bayesian analysis, posterior distribution is an updated probability distribution after combining the prior derived from previous data or assumptions and our data.
+Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
 The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
@@
 This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-By fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters.
+After fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information derived from previous data or our assumptions through summary statistics from estimates of the ERGM parameters [@Caimo2015; @Nemmers2019].
+This also allows uncertainty quantification.
 Posterior probability estimates the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
@@
 <!-- archaeological setting -->
-We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD).
-The settlement experienced European colonial impacts in the early 17th century and a large wave of Chinese immigrants in the 19th century [@Chen2007].
-Excavations revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007].
-In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with _in-situ_ posts were also excavated.
-The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, indicating a well organized spatial arrangement.
+We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD) [@Chen2007].
+Excavations revealed abundant trade goods, including imported ceramics and stonewares, metal artifacts, glass beads, and stone beads, indicative of frequent trade activities [@Chen2007].
+A total of 90 burials were unearthed that are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
+Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well organized spatial arrangement.
 Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
@@
 However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods.
-She found that the burials with high value burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
+She found that the burials with high value goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
 One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization.
@@
 chi <- burial_three_period_age_number %>% filter(Phase == "chi") %>% pull()
+disturb <- burial_three_period_age_number %>% filter(Phase == "disturbed") %>% pull()
 diff_pre_post <- post_E - pre_E
@@
 We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
-There are eight burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology.
-We reconstructed social relations between individuals by linking them according to similar prestige goods [cf.
+There are `r disturb` burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology.
+Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf.
 @Coward2013, pp.
-252].
-Therefore, we built networks where burials represent actors (nodes in the network) that are linked when they have the same prestige goods in common.
-The prestige goods we identified include agate beads, gold-foil beads, imported Chinese porcelains, gold foils, and fish-shaped ornaments.
+252], we built networks where burials (nodes in the network) that are linked when they have the same prestige goods in common.
+The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments.
 Previous studies indicate that these items are considered as high-value across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
-Historical accounts also support these items as prestige goods [@Borao2009; @LiandWu2006].
+Historical accounts written by Europeans and the Han Chinese also support these items as prestige goods [@Borao2009; @LiandWu2006].
 
@@
 <!-- our hypothesis --> 
-Our pericolonial impact hypothesis is that rare, high quality grave goods accumulated in only a few individuals' burials after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access to these goods.
-We can identify these high value trade goods from historical documents written by Europeans and the Han Chinese [@Borao2009; @LiandWu2006].
-A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power.
-If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status).
-To test this prediction, we use ERGM in a Bayesian framework to examine the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
+Our pericolonial impact hypothesis is that rare, high quality burial goods accumulated in only a few individuals' burials after the European presence because of increased social inequality at Kiwulan due to differential access to these goods.
+A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power, recognized as network strategies to achieve power [@Drennan2010; @Feinman2000].
+If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity) and less transitivity (cohesion), along with greater differences in wealth accumulation according to burial goods.
+To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
 By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
@@
 
+```{r burial-value}
+ave_value <- round(mean_burial_value,1) 
+high_value <- round(high_burial_value,1) 
+```
+
 <!-- burial data and the definition of tie formation  --> 
 We used burial goods in common to construct ties for representing a flow of good between burials.
-Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
-For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e.
+Since trade beads are commonly found across burials with substantial differences in quantities, we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
+For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods (i.e.
 presence or absence).
 Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index.
-Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>28, about 10 percent), above average (12-28), and below average (<12), as an index of wealth.
+Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting based on historical records.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
 The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
-Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
+Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edit the paper rmd from intro to method

@@
 Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
-Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
-@Coward2013, pp.
-252].
 This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
@@
 <!-- current approaches to network analysis: move to SI -->
-Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective ways to examine relationships by incorporating prior information into networks to better understand dependencies of variables in complex networks and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
 Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.

### 2021-01-26 analysis/paper/paper.Rmd
edit model specification in paper rmd and SI

@@
 <!-- Network construction and specification -->
-We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
-Table \@ref(tab:table-ergm-terms) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence.
+We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
+Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification with the corresponding archaeological evidence.
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
-Based on our hypothesis, we model a network with increased social inequality that is represented by endogenous network effects, low transitivity and high popularity.
-We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to test whether burials with similar attributes tend to form relationships.
-For example, age-homophily means people of the same age to have the same burial goods.
-We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
+Based on our hypothesis, we model a network with increased social inequality that is represented by endogenous network effects: low transitivity and high popularity.
+We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore and test the tendency of tie forming.
+For example, age-homophily means people of the same age tend to have the same burial goods.
+We also include the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
 Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
@@
 knitr::kable(ergm_terms,
-             caption = "Network variables used for model specifications with assocaited interpeation for burial relations.")
+             caption = "Network parameters in ERGMs used for model specifications with assocaited interpeation for burial relations.")
 ```

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edit model specification in paper rmd and SI

@@
 ```
+
+This supplementary document contains information on the additional review of network analysis in archaeology, bigger historical background of northeastern Taiwan in a pericolonial context, definition of tie making, methods about modeling fitting, and assessment of MCMC and Goodness-of-fit in a Bayesian framework.
 
@@
 
-# Method {-}
+# Materials and Method {-}
 
@@
 <!-- burial data and the definition of tie formation: partially move to SI  --> 
-
 For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)).

### 2021-01-26 analysis/paper/paper.Rmd
looking at highlights
edit highlights

@@
 highlights: |
-  1. We present Bayesian inference on Exponential random graph models (ERGMs) to model archaeological networks.
-  2. We test social processes by simulating two sets of burial data from different time.
-  3. Differences in network statistics implies changes in social structures influenced by foreign presence. 
-  4. Bayesian ERGMs is efficient and useful for evaluating and interpreting networks. 
+  1. We present Bayesian inference on exponential random graph models (ERGMs) to model archaeological networks.
+  2. We test social processes by simulating two sets of burial data from different time periods.
+  3. Differences in network statistics imply change over time in social structures influenced by foreign presence. 
+  4. Bayesian ERGMs are efficient and useful for evaluating and interpreting networks. 
 ---

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edits to SI

@@
 
-This supplementary document contains information on the additional review of network analysis in archaeology, bigger historical background of northeastern Taiwan in a pericolonial context, definition of tie making, methods about modeling fitting, and assessment of MCMC and Goodness-of-fit in a Bayesian framework.
+This supplementary document contains:
 
-# Introduction to network analysis and archaeological applications {-}
+- further historical background of northeastern Taiwan in a pericolonial context
+- further definition of our approach to tie formation with the beads from Kiwulan burials
+- further review of network analysis in archaeology
+- further details on methods of modeling fitting, and
+- assessment of MCMC output and Goodness-of-fit diagnostics for our models
 
-<!-- intro to networks: move to SI -->
-A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
-The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
-For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
-4].
-Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
-Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
-This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
-
-<!-- two types of network analysis: move to SI -->
-Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
-There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
-Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
-By quantifying those network properties, archaeologists can answer a wide range of research questions.
-Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
-Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-
-<!-- current approaches to network analysis: move to SI -->
-Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
-Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
-ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
-The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
-For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
-Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
-
-# Kiwulan in the 17th northern Taiwan {-}
+## Kiwulan in the 17th century northern Taiwan {-}
 
@@
 The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
-
-# Materials and Method {-}
 
@@
 ```
+
+## Tie formation using bead data from Kiwulan
 
@@
 
+## Review of network analysis and archaeological applications {-}
+
+<!-- intro to networks: move to SI -->
+A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
+The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
+For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
+4].
+Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
+Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
+This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
+
+<!-- two types of network analysis: move to SI -->
+Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
+There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
+Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
+By quantifying those network properties, archaeologists can answer a wide range of research questions.
+
+Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
+Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
+Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+
+<!-- current approaches to network analysis: move to SI -->
+Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
+Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
+ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+
+The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
+For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
+
+By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
+Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
+ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+
+## Further details on methods of modeling fitting {-}
+
 ```{r bergm-iteration}
@@
 More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
+
 In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
@@
 
+## Assessment of MCMC output and Goodness-of-fit diagnostics for our models
+
 As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots for both models (Figure \@ref(fig:figure-pre-diag) and Figure \@ref(fig:figure-post-diag)) [@Hamra2013].
@@
 We set $`r bgof_si`$ network graphs simulated from the estimated posterior distribution in ERGMs.
-Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+
+Our Bayesian GOF diagnostics, shown in the main text, summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
 This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.

### 2021-01-26 analysis/paper/supplementary-material.Rmd
edit SI
showing visual md editing

@@
 
-## Kiwulan in the 17th century northern Taiwan {-}
+## Kiwulan in the 17th century northern Taiwan {.unnumbered}
 
 <!--fit in a bigger historical picture in a periclonial context: move to SI -->
+
 The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
@@
 <!-- burial data and the definition of tie formation: partially move to SI  -->
+
 For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)).
-Gold-foil beads are in levels of high (>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (>6), upper-middle (3-6), lower-middle (2), and low (1).
+Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1).
 If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
@@
 
-## Review of network analysis and archaeological applications {-}
+## Review of network analysis and archaeological applications {.unnumbered}
 
 <!-- intro to networks: move to SI -->
+
 A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
@@
 <!-- two types of network analysis: move to SI -->
+
 Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
@@
 <!-- current approaches to network analysis: move to SI -->
+
 Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
@@
 
-## Further details on methods of modeling fitting {-}
+## Further details on methods of modeling fitting {.unnumbered}
 
@@
 <!--Model fitting: move to SI -->
+
 After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.
@@
 
-# References {-}
+# References {.unnumbered}
 
 <!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
+
+::: {#refs}
+:::
 

### 2021-01-26 analysis/paper/supplementary-material.Rmd
resolve merge conflict

@@
 
-## Kiwulan in the 17th century northern Taiwan {.unnumbered}
+# Kiwulan in the 17th northern Taiwan {-}
 
 <!--fit in a bigger historical picture in a periclonial context: move to SI -->
-
 The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].

### 2021-01-26 analysis/paper/supplementary-material.Rmd
merge conflicts and edit SI

@@
 
--   further historical background of northeastern Taiwan in a pericolonial context
+-  further historical background of 17th century northeastern Taiwan in a pericolonial context
 -  further definition of our approach to tie formation with the beads from Kiwulan burials
--   further review of network analysis in archaeology
+-  further review of network analysis and archaeological applications
 -  further details on methods of modeling fitting, and
@@
 
-# Kiwulan in the 17th northern Taiwan {-}
+## Kiwulan in the 17th century northeastern Taiwan {.unnumbered}
 
@@
 
-## Tie formation using bead data from Kiwulan
+## Tie formation with the beads from Kiwulan burials {.unnumbered}
 
@@
 ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
-
 The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
@@
 More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
-
 In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
@@
 
-## Assessment of MCMC output and Goodness-of-fit diagnostics for our models
+## Assessment of MCMC output and Goodness-of-fit diagnostics for our models {.unnumbered}
 
@@
 
-### Colophon
+### Colophon {.unnumbered}
 

### 2021-01-27 analysis/paper/paper.Rmd
merge changes in highlights
edit ERGMs background and result sections in paper rmd, fix a typo in SI

@@
   1. We present Bayesian inference on exponential random graph models (ERGMs) to model archaeological networks.
-  2. We test social processes by simulating two sets of burial data from different time periods.
+  2. We explore social processes and network structures in archaeological data from from two time periods.
   3. Differences in network statistics imply change over time in social structures influenced by foreign presence. 
@@
 
-<!--fit in a bigger historical picture in a periclonial context: move to SI -->
-
 <!-- archaeological specific case study -->
@@
 Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Prior information is derived from previous data or assumptions of our data.
 The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
-MCMC avoids doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
+Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
 This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-After fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporates our sample data and prior information derived from previous data or our assumptions through summary statistics from estimates of the ERGM parameters [@Caimo2015; @Nemmers2019].
-This also allows uncertainty quantification.
-Posterior probability estimates the effect of ERGM parameters by looking at the posterior mean and 95% credible intervals, which can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
-The typical criteria for interpreting the posteriors is that an odds ratio greater than one means a positive effect of a parameter, while odds ratios less than one represents negative effect.
+After fitting an ERMG with an MCMC algorithm, the Bayesian approach generates posterior probabilities incorporating our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
+This allows uncertainty quantification by looking at the posterior mean and 95% credible intervals, or the odds ratios (conditional probability of observing an edge) that can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
+The typical criteria for interpreting an significant effect of a parameter is that its credible interval does not contain zero, or an odds ratio is greater/less than one, meaning a significant positive/negative effect.
 In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
@@
 
-<!--Model fitting: move to SI -->
-
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For the pre-European network simulations, the nodal covariates, ritual-homophily, have significant effects on the formation of relations between burials.
-The positive effect demonstrates that burials with ritual pottery tend to form relations.
-Despite positive values for some covariates, such as age-homophily, sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals.
+For nodal covariates in the pre-European model, ritual-homophily has a significant effect on the formation of relations between burials, while burial value-homophily has a negative effect.
+This demonstrates that burials with ritual pottery tend to form relations, but burials in the same value level tend not to.
+Despite positive mean value for some covariates, such as age-homophily, sex-homophily, they do not show a significant tendency duo to value of zero in the confidence intervals.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not affect the formation of relations.
@@
 The high positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
-The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones.
+In contrast, the strong negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones.
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
@@
 <!--posterior stats for for post-E and their comparison-->
-For the network after the arrival of Europeans, the nodal covariates of burial value-homophily show positive effects, while the ritual-homophily and sex-homophily have a negative effect.
-Despite the effect is not strong, it could indicate the burials in the same wealth level tend to form relations.
+For the network model after the arrival of Europeans, the nodal covariates of burial value-homophily show significant positive effects, while the ritual-homophily and sex-homophily have a negative effect.
+This indicates the burials in the same wealth level tend to form relations.
 On the contrary, the same sex burials and burial with ritual pottery in common tend not to form relations.
@@
 This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others.
-This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation.
-In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network.
+This implies that a few central individuals might have better access to trade goods and developed personal networks through wealth accumulation.
+In general, the post-European network model has a smaller transitivity effect and a positive popularity effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
-Both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
+In addtion, both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
 

### 2021-01-27 analysis/paper/supplementary-material.Rmd
edit ERGMs background and result sections in paper rmd, fix a typo in SI

@@
 
-## Kiwulan in the 17th century northeastern Taiwan {.unnumbered}
+# Kiwulan in the 17th northstern Taiwan in a pericolonial context {-}
 
-<!--fit in a bigger historical picture in a periclonial context: move to SI -->
+<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
 The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].

### 2021-01-29 analysis/paper/paper.Rmd
edit the method, results, discussion, and conclusion

@@
 Normal distributions for the priors are typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-Thus, we specified the prior of the edge density parameter to low for both network models.
-For covariates about biological features, such as age, sex, we specified a vague prior that follows a normal distribution with mean at `r priors$mean_same[2]` and standard deviations at `r priors$sigma_same[2]`for both models to explore their effects.
+Thus, we specified the prior of the edge density parameter to low for both network models, which follows a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
+For covariates about biological features, such as age, sex, we specified a vague prior (N(`r priors$mean_same[2]`, `r priors$sigma_same[2]`)) for both models to explore their effects.
 For physical distance between burials, we also set a vague prior effect (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) to explore whether there is kinship-based proximity, e.g.
@@
 To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect based on ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for network before European contact to indicate less social inequality and stress ritual element shared in corporate groups.
-On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social inequality.
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
+On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
 This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
-The priors for density were also set the same with N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`) for both pre- and post-European networks based on a common observation of network density.
 
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results based on Bayesian network analysis.
+To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results of Bayesian ERGMs.
 Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
@@
 For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference.
-This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with smaller degree of difference.
+This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with some degree of difference.
 
@@
 <!--assessment of MCMC analysis and BGOF--> 
-Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) show that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
 the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
 For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
-This also suggests that our models can reproduce networks that resemble the structural features of our observed networks.
-We compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
+This suggests that our models can reproduce networks that resemble the structural features of our observed networks.
+We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
 Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network.
@@
 A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration of connections among fewer burials (lower transitivity and positive popularity) after the European presence during the 17th century.
-This supports our models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes.
+This supports our models of chronological change from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
 These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
 Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan.
-A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communalrituals, and social inequality, if any, would be associated with groups [@Siegel1999].
-In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good manipulation, and trade monopolization [@Blanton1996].
+A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
+In contrast, a network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good manipulation, and trade monopolization [@Blanton1996].
 It should be noted that the corporate/network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
@@
 <!--archaeological implication based on the results-->
-The covariants for both networks, before and after the European presence, provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
+The covariants for networks provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
 Our covariants were age, sex, ritual pottery, and a wealth index based on the value of grave goods.
@@
 This may suggest that the use of burial goods is more individual-oriented instead of kinship-based.
-In addition, demographic information, such as age and sex, do not show any correlations with network formation for the pre-European network and only sex shows small effect for the post European network.
+In addition, demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows small effect for the post European network.
 This may be due to the unavailability of age and sex values for many burials in our sample.
@@
 We note that there could be an issue of disconnection between the role of the deceased in a burial context and their life, which could be reduced by examining evidence from the residential context [@Chapman2003].
-Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament].
+Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial situation [@LWandBM2020ornament].
 This is consistent with the result of a more centralized burial network after the presence of Europeans, which supports a connection between a burial and the social role of the living person.
@@
 It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
-We recognise that the meaning of the burial goods remains speculative without rich ethnographic data to support our interpretations.
-Furthermore, a detailed material culture context for burial goods in residential settings would also help to strengthen our interpretations, but this is not available from Kiwulan.
+We recognize that the meaning of the burial goods remains speculative without rich ethnographic data to support our interpretations.
 In addition, we acknowledge the fragmentation issue in our archaeological data that limit network interpretations [eg.
@@
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-In this paper, we presented a novel approach for studying burial relations using ERGM network analysis within a Bayesian framework.
+In this paper, we presented a novel approach for studying burial relations using ERGMs within a Bayesian framework.
 We examined social changes in a European pericolonial context in northeastern Taiwan.
@@
 The changes in formation mechanisms for networks from ritual-oriented practices to wealth possession suggest increased behavior of wealth accumulation and differentiation stimulated by the European presence and associated long-distance trade network.
-This aligns with changes in social complexity from a more corporate to a more network mode society with different strategies for achieving power and different degrees of social inequality.
+This aligns with social changes from a more corporate to a more network mode with different strategies for achieving power and different degrees of social inequality.
 Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
@@
 <!-- technical and methodological implication-->
-This case study demonstrates the methodological improvement that Bayesian inference on ERGM provides to inform and enhance studies of relational data in archaeology.
+This case study demonstrates the methodological improvement that Bayesian inference on ERGMs provides to inform and enhance studies of relational data in archaeology.
 A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.

### 2021-01-31 analysis/paper/paper.Rmd
fix typo and add a citation

@@
   1. We present Bayesian inference on exponential random graph models (ERGMs) to model archaeological networks.
-  2. We explore social processes and network structures in archaeological data from from two time periods.
+  2. We explore social processes and network structures in archaeological data from two time periods.
   3. Differences in network statistics imply change over time in social structures influenced by foreign presence. 
@@
 This allows uncertainty quantification by looking at the posterior mean and 95% credible intervals, or the odds ratios (conditional probability of observing an edge) that can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
-The typical criteria for interpreting an significant effect of a parameter is that its credible interval does not contain zero, or an odds ratio is greater/less than one, meaning a significant positive/negative effect.
+The typical criteria for interpreting an significant effect of a parameter is that its credible interval does not contain zero, or an odds ratio is greater/less than one, meaning a significant positive/negative effect [@Harris2013, pp.
+51].
 In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.

### 2021-02-03 analysis/paper/paper.Rmd
edit based on PL's comments

@@
 Burial analysis by archaeologists is an approach to understand past social structures through the study of the physical traces of mortuary practices.
-The material cultures and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
+The material culture and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
 Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
@@
 252].
-This paper will use a novel Bayesian approach on exponential random graph model for networks to investigate the formation of relationship between burials in northeastern Taiwan to explore the indirect impact of foreign presence on social organization.
+This paper will use a novel Bayesian approach on exponential random graph models for networks to investigate the formation of relationships between burials in northeastern Taiwan to explore the indirect impact of foreign presence on social organization.
 
@@
 <!-- archaeological background -->
-Archaeological sites in northeastern Taiwan show evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and the use as burial goods around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament].
-Imported goods were treated as prestige or wealth in those Indigenous societies that were commonly mentioned in Spanish historical records [@Borao2009; @LiandWu2006].
-For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006].
-In addition, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
-Despite we observed the uneven distribution of imported goods in burials, the degree of differentiation across burials over time remain unclear.
-The use of imported items as prestige goods in local Indigenous communities might occur earlier before a European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
+Archaeological sites in northeastern Taiwan show the evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and burial contexts around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament].
+Imported goods were treated as prestige or wealth in those Indigenous societies that were commonly mentioned in Spanish historical records in the 17th century (1624-1642) [@Borao2009; @LiandWu2006].
+For example, a document written by a Spanish priest described a Spanish soldier using carnelian beads to exchange for natural resources with local Indigenous people, because of their high value in Indigenous culture [@LiandWu2006].
+In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
+Despite the fact that we recorded an uneven distribution of imported goods in burials, the degree of differentiation across burials over time remains unclear.
+The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
 A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade networks when Indigenous societies were involved in the complex trade network stimulated by the Europeans.
+The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade networks when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
 
 <!-- archaeological specific case study -->
-In this study, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
+In this study, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which was occupied from 14th to 19th centuries.
+This occupation period includes the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
 We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
@@
 <!-- ERGM in archaeology, and limitations --> 
-Exponential random graph models first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
+Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
 Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
@@
 <!-- Bayesian inference for Exponential Random Graph Models --> 
-Bayesian approaches to Exponential Random Graph Models (ERGMs) are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
 Prior information is derived from previous data or assumptions of our data.
 The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
-Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives a better convergence results.
-This enable us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
+Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, where the acceptance probability is tractable that give a better convergence result.
+This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
 After fitting an ERMG with an MCMC algorithm, the Bayesian approach generates posterior probabilities incorporating our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
@@
 In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
-@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data missing.
+@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data is missing.
 
@@
 Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period.
-The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well studied.
+The discord over the discussion of Kiwulan social organization could be associated with chronological differences that have not yet been well studied.
 
@@
 <!-- burial data and the definition of tie formation  --> 
-We used burial goods in common to construct ties for representing a flow of good between burials.
+We used burial goods in common to construct ties for representing a flow of goods between burials.
 Since trade beads are commonly found across burials with substantial differences in quantities, we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
@@
 The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
-The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial.
+The number is the sum of values of each type of prestige good, which is calculated by the total number of a prestige item from burial context over the number of prestige items in a burial.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
-The networks constructed based on this principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).
+The networks constructed based on these principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).
 
@@
 Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics, with covariates to control the preferential tendency of formation of relationship.
+This would be indicated by high popularity or degree values in the network statistics, with covariates to control the preferential tendency of formation of relationships.
 According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
@@
 knitr::kable(ergm_terms,
-             caption = "Network parameters in ERGMs used for model specifications with assocaited interpeation for burial relations.")
+             caption = "Network parameters in ERGMs used for model specifications with associated interpretation for burial relations.")
 ```
@@
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not affect the formation of relations.
-For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect.
+For the endogenous network effects, transitivity (gwesp) presents significant positive effect, while popularity (gwdegree) demonstrates negative effect.
 The high positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
@@
 ft <- flextable(posterior_esti_two_phases) %>% 
-  set_header_labels(mean_pre = "pre-Euorpean", median_pre = "pre-Euorpean", 
-                    `2.5%_pre` = "pre-Euorpean", `97.5%_pre` = "pre-Euorpean",
-                    mean_post = "post-Euorpean", median_post = "post-Euorpean", 
-                    `2.5%_post` = "post-Euorpean", `97.5%_post` = "post-Euorpean") %>%
+  set_header_labels(mean_pre = "pre-European", median_pre = "pre-European", 
+                    `2.5%_pre` = "pre-European", `97.5%_pre` = "pre-European",
+                    mean_post = "post-European", median_post = "post-European", 
+                    `2.5%_post` = "post-European", `97.5%_post` = "post-European") %>%
   merge_at(i = 1, j = 2:5, part = "header") %>% 
@@
 ```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, gwesp, gwdegree, and distance-dyadcov by phases.
-The pre-European group presents remarkedly larger values for the gwesp, but lower vlaues for the edges and gwdegree parameters comapred to the post-Eruoepan group.
+The pre-European group presents remarkable larger values for the gwesp, but lower values for the edges and gwdegree parameters compared to the post-European group.
 The distance parameters overlap between two groups."}
@@
 Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
-In contrast, the popularity (gwdegree) shows a significantly higher positive effect than the effect of pre-European network.
-This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others.
-This implies that a few central individuals might have better access to trade goods and developed personal networks through wealth accumulation.
+In contrast, the popularity (gwdegree) shows a significantly higher positive effect than the effect of the pre-European network.
+This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
+This implies that the presence of a better access to trade goods and the behavior of wealth accumulation or display in burial events.
 In general, the post-European network model has a smaller transitivity effect and a positive popularity effect than the pre-European network model.
@@
 
-```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, the top one with 0 nodes removal, and networks with different number of node removal.
+```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, the top one with 0 nodes removal, and networks with different numbers of node removal.
 The number of bootstrap samples is 1000."}
@@
 The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and popularity (p = $`r pval_star`$).
-Moreover, we explored the sample size effect by removing nodes at certain percents (5-40%) for both networks and compared their confidence intervals.
+Moreover, we explored the sample size effect by removing nodes at certain percent (5-40%) for both networks and compared their confidence intervals.
 The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
@@
 
-```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (left) and the post-European model (right).
+```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
 Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
@@
 post_bgof_img <- image_read(here::here("analysis/figures/004-post-bgof.png"))
-image_append(c(pre_bgof_img, post_bgof_img))
+image_append(c(pre_bgof_img, post_bgof_img), stack = T)
 ```
@@
 A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
-In contrast, a network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good manipulation, and trade monopolization [@Blanton1996].
+In contrast, a network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
 It should be noted that the corporate/network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
@@
 This may lead to ambitious individuals to competing with each other for traditional goods and accumulating wealth.
-It also demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
+It also demonstrates that local Indigenous societies could actively incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
 We recognize that the meaning of the burial goods remains speculative without rich ethnographic data to support our interpretations.
-In addition, we acknowledge the fragmentation issue in our archaeological data that limit network interpretations [eg.
+Also, the fragmentation issue in our archaeological data could limit network interpretations [eg.
 @Erik2015], such as missing age and sex attributes for some burials at Kiwulan.
+Human skeletal remains can provide information about health for interpreting social status; however, the preservation of Kiwulan human remains was generally not good to yield useful data.
 
@@
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-The changes in formation mechanisms for networks from ritual-oriented practices to wealth possession suggest increased behavior of wealth accumulation and differentiation stimulated by the European presence and associated long-distance trade network.
+The changes in formation mechanisms for networks from ritual practice connections to wealth level connections suggest increased behavior of wealth accumulation and differentiation stimulated by the European presence and associated long-distance trade network.
 This aligns with social changes from a more corporate to a more network mode with different strategies for achieving power and different degrees of social inequality.
@@
 @Acabado2017; @Trabert2017].
-Furthermore, our study also highlights the agency of local people to manipulate colonial power instead of being passive recipients of trade goods.
+Furthermore, our study also highlights the possibility of uneven access to foreign trade goods that allows active accumulation or even manipulation their values by local people instead of being a passive recipients of trade goods.
 
@@
 We would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the data used in this paper.
-We thank the excavation team of Kiwulan led by Yu-pei Chen for their post-excavation analysis and detailed documents about burials that enables this study.
+We thank the excavation team of Kiwulan led by Yu-pei Chen for their post-excavation analysis and detailed documentation about burials that enable this study.
 We also thank Alberto Caimo and Steven M.
-Goodreau for the discussion regarding the methods at the earliest stage fo analysis.
+Goodreau for the discussion regarding the methods at the earliest stage of analysis.
 This work was supported in part by the travel grant from the Department of Anthropology at University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).

### 2021-02-03 analysis/paper/supplementary-material.Rmd
edit based on PL's comments

@@
 <!--fit in a bigger historical picture in a pericolonial context: move to SI -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007].
+The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007].
 In 1626, the Spanish built their forts in Heping dao and Tamsui, northern Taiwan, and later were defeated and taken over by the Dutch in 1642 [@Andrade2007].
-Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch, who also asked them to pay annual tributes in a flexible way [@Borao2009; @Kang2016].
-The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies.
+Indigenous people in northeastern Taiwan first encountered colonial power in 1632 when they were attacked by the Spanish and later in 1647 by the Dutch.
+The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016].
+In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.
 
@@
 In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
-For example, past trade can be conceptualized as a network of individual entities connected by shared similarities, the flow of goods, to represent their interactions [@Collar2015, pp.
+For example, past trade can be conceptualized as a network of individual entities connected by shared similar objects or elements, such as the flow of goods, to represent their interactions [@Collar2015, pp.
 4].
@@
 
-Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Regarding burials, @Sosna2013 examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling evaluates networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], or exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks.
-Other examples show the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
+Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
+Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
 
@@
 More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
-In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
+In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pairs of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
 In our case, we set the number of chains to six.
@@
 We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
-We expect to see the models fit with our hypotheses, indicative of a increased social inequality after the foreign contact.
+We expect to see the models fit with our hypotheses, indicative of an increased social inequality after the foreign contact.
 In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.

### 2021-02-04 analysis/paper/paper.Rmd
edit based on BF's comments, stop at model specification

@@
   Burials provide valuable information to study social structures based on the assumption that burials and associated grave goods can represent social roles and relations in a society.
-To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns that can infer interactions and relationship between entities.
-Different from traditional static networks, statistical approaches to network analysis, such as exponential random graph models (ERGMs), provides a way to test hypotheses on dynamic process of network formation.
+To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns that can infer interactions and relationships between entities.
+Statistical approaches to network analysis, such as exponential random graph models (ERGMs), provide a way to test hypotheses on dynamic process of network formation.
 However, the computational difficulty and sensitivity to uncertainties limits the application of ERGMs.
@@
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
-Burials with the same prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
+Burials with the same kinds of prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
 @Coward2013, pp.
@@
 For example, a document written by a Spanish priest described a Spanish soldier using carnelian beads to exchange for natural resources with local Indigenous people, because of their high value in Indigenous culture [@LiandWu2006].
-In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006].
-Despite the fact that we recorded an uneven distribution of imported goods in burials, the degree of differentiation across burials over time remains unclear.
+In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status in their community [@LiandWu2006].
+Despite the fact that we recorded an uneven distribution of imported goods in burials, the degree of differentiation across burials over time remained unclear.
 The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
-A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade networks when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
+A general model to summarize this situation in northeastern Taiwan might assume that the influence of a colonial power, combined with high local values attached to imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
 
 <!-- archaeological specific case study -->
-In this study, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which was occupied from 14th to 19th centuries.
-This occupation period includes the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007].
-We assume that social changes at Kiwulan would be supported by chronological differences in the structure of burial networks.
-We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival.
-We ask: (1) did European colonial activities in 17th century Taiwan result in an increased social inequality in an Indigenous society in ways that can be detected by analysis of burial networks?
-(2) what are the major variables affecting or forming the higher degree of popularity (e.g.
-a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period?
-By answering these questions, this study helps to expand the use of statistical modeling for burials in understanding the indirect effects of a colonial presence on Indigenous groups.
-This is important to understand reactions to indirect impacts of European colonization.
+In what follows, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which was occupied from 14th to 19th centuries.
+This occupation period includes the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
+We argue that the involvement in the long-distance trade stimulated by the Europeans might lead to increased social inequality at Kiwulan.
+This argument would be supported by changes in the structure of networks built on burial data based on the assumption that burials can reflect social structure.
+We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival in 17th century Taiwan.
+We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
+(2) what are the major variables affecting or forming the unequal social positions represented by higher degree of network popularity (e.g., a few nodes have more relationships with other nodes) that might hint at social heterogeneity influence by the European arrival?
+By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in an indirect way by examining burial data.
+We also introduces a novel Bayesian approach on statistical modeling for networks that is suitable for networks with smaller sizes.
 
@@
 In addition, it is difficult in ERGM to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
-A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian].
+A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with a larger set of networks [@Caimo2017bayesian].
 This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
-Model degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-One solution to these limitations is available by implementing ERGMs in a Bayesian framework.
+Model degeneracy is another issue where probability models tend to overestimate a small number of extreme graphs by assigning too much weight, e.g., in extreme cases such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
+One solution to these limitations is available by implementing ERGMs by means of a Bayesian framework.
 
@@
 The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
-Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, where the acceptance probability is tractable that give a better convergence result.
+With the exchange algorithm, Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives better convergence results.
 This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-After fitting an ERMG with an MCMC algorithm, the Bayesian approach generates posterior probabilities incorporating our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
-This allows uncertainty quantification by looking at the posterior mean and 95% credible intervals, or the odds ratios (conditional probability of observing an edge) that can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
+By fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
+This allows uncertainty quantification by examining the posterior mean and 95% credible intervals, or the odds ratios (conditional probability of observing an edge) that can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
 The typical criteria for interpreting an significant effect of a parameter is that its credible interval does not contain zero, or an odds ratio is greater/less than one, meaning a significant positive/negative effect [@Harris2013, pp.
@@
 A total of 90 burials were unearthed that are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well organized spatial arrangement.
+Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement.
 Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
@@
 However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods.
-She found that the burials with high value goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
+She found that the burials with rare goods, also prestige goods in local culture, were usually associated with elders, which might indicate achieved, rather than inherited, status.
 One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization.
@@
 The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments.
-Previous studies indicate that these items are considered as high-value across different archaeological contexts [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
-Historical accounts written by Europeans and the Han Chinese also support these items as prestige goods [@Borao2009; @LiandWu2006].
+Previous studies indicate that these items are considered as high-value across different archaeological contexts based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+Historical accounts written by Europeans mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
 
@@
 <!-- our hypothesis --> 
-Our pericolonial impact hypothesis is that rare, high quality burial goods accumulated in only a few individuals' burials after the European presence because of increased social inequality at Kiwulan due to differential access to these goods.
+Our "pericolonial impact hypothesis" is that rare, high quality burial goods accumulated in only a few individuals' burials after the European presence because of increased social inequality at Kiwulan due to differential access to these goods.
 A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power, recognized as network strategies to achieve power [@Drennan2010; @Feinman2000].
@@
 Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting based on historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three levels, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
-The number is the sum of values of each type of prestige good, which is calculated by the total number of a prestige item from burial context over the number of prestige items in a burial.
+The burial value index is an important attribute for economical inequality that we assigned the burials into three ranks, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
+The number is the sum of values of each type of prestige good, which is calculated by the total number of prestige item from burial context over the number of prestige items in a burial.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
-Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.
-The networks constructed based on these principles show that the network after the European presence has higher node degrees in general (Figure \@ref(fig:figure-network-diagrams)).
+Our networks are non-directed, which means a mutual relationship where the tie between any two actors is bidirectional.
+The networks constructed based on these principles show that the network after the European presence has more node connectivity in general with some nodes having a larger number of connections (Figure \@ref(fig:figure-network-diagrams)).
 
 ```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
-The size of the nodes is proportional to the node degrees that means the number of connections to a node.
-The thickness of tie represents the number of goods in common between nodes."}
+The size of each node is proportional to "node degrees" or the number of connections to a node.
+The thickness of tie represents the number of goods in common between two nodes."}
 knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))

### 2021-02-04 analysis/paper/supplementary-material.Rmd
edit based on BF's comments, stop at model specification

@@
 <!--fit in a bigger historical picture in a pericolonial context: move to SI -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
 Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007].
@@
 4].
-Similar concepts can be applied to a wide range of archaeological data with relational assumptions, such as burial contexts.
+This concept can be applied to burial contexts where each burial is an actor linked by similar elements, such as burial goods.
 Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
@@
 There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
-Node level focuses on the role of node in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
+Node level focuses on node properties in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
 By quantifying those network properties, archaeologists can answer a wide range of research questions.
 
-Examples includes exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
 Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically testing the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-Such application includes simulations and testing food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest simulation [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
+Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
 Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
@@
 In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
-For example, nodes with similar attributes are more likely to form a relationship, such as friendship between people with the same hobby.
+For example, nodes with similar attributes are more likely to have a relationship, such as friendship between people with the same hobby.
 Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].

### 2021-02-05 analysis/paper/paper.Rmd
edit the result section based on BF's comments

@@
 She found that the burials with rare goods, also prestige goods in local culture, were usually associated with elders, which might indicate achieved, rather than inherited, status.
-One important limitation of these previous studies is that they did not use analytical units suitable for comparing pre-European social organization with post-European social organization.
-Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period to the post-European period.
+One important limitation of these previous studies is that they did not use analytical units suitable for comparing the phases before and after European arrival.
+Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until Han Chinese immigrants arrival).
 The discord over the discussion of Kiwulan social organization could be associated with chronological differences that have not yet been well studied.
@@
 <!-- Network construction and specification -->
-We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008].
+We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], where model specification and simulation procedures is based on statnet package [@Hunter2008; @Morris2008; @Handcock2008].
 Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification with the corresponding archaeological evidence.
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
-Based on our hypothesis, we model a network with increased social inequality that is represented by endogenous network effects: low transitivity and high popularity.
-We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore and test the tendency of tie forming.
+Based on our hypothesis, we model a network with increased social inequality to be represented by endogenous network effects: low transitivity and high popularity.
+We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore the tendency of tie forming.
 For example, age-homophily means people of the same age tend to have the same burial goods.
@@
 Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics, with covariates to control the preferential tendency of formation of relationships.
+This would be indicated by high popularity or degree values in the network statistics, with covariates to understand the preferential tendency of formation of relationships.
 According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, ritual-homophily has a significant effect on the formation of relations between burials, while burial value-homophily has a negative effect.
-This demonstrates that burials with ritual pottery tend to form relations, but burials in the same value level tend not to.
-Despite positive mean value for some covariates, such as age-homophily, sex-homophily, they do not show a significant tendency duo to value of zero in the confidence intervals.
-Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not affect the formation of relations.
-For the endogenous network effects, transitivity (gwesp) presents significant positive effect, while popularity (gwdegree) demonstrates negative effect.
-The high positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common.
-In contrast, the strong negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones.
+A significant effect is determined by consistent positive/negative estimates in a 95% confidence interval.
+For nodal covariates in the pre-European model, ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while wealth rank represented by burial values (value-homophily) has a negative effect.
+This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
+Despite positive mean for some covariates, such as age (age-homophily) and sex (sex-homophily), they do not show a significant tendency due to value of zero in the confidence intervals.
+Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
+For the endogenous network effects, transitivity (gwesp) presents a significant positive effect, while popularity (gwdegree) demonstrates a negative effect.
+The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common.
+In contrast, the strong negative popularity shows there is a tendency toward decentralization that reflects most burials have a similar number of ties without any prominent ones.
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
@@
 <!--posterior stats for for post-E and their comparison-->
-For the network model after the arrival of Europeans, the nodal covariates of burial value-homophily show significant positive effects, while the ritual-homophily and sex-homophily have a negative effect.
-This indicates the burials in the same wealth level tend to form relations.
-On the contrary, the same sex burials and burial with ritual pottery in common tend not to form relations.
-There are no significant effects for age-homophily and the dyadic covariate, physical distance.
+For the network model after the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects, while ritual and sex variables have negative effects.
+This indicates the burials in the same wealth level tend to have relations.
+On the contrary, the same sex burials and burial with ritual pottery in common tend not to have relations.
+There are no significant effects for variables of age and physical distance.
 Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
@@
 This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population.
-Networks with unoverlapped intervals means there is a significant difference.
+Networks with nonoverlapped intervals means there is a significant difference.
 We computed and compared endogenous network statistics, including density, popularity, and transitivity for our two networks.
@@
 The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
-In general, those statistical methods support the presence of anthropologically meaningful differences in the pre-and post-European networks.
+In general, those statistical methods support the presence of anthropologically meaningful differences for density, transitivity, and popularity (i.e.
+centrality) in the pre-and post-European networks.
 The procedure of cross-validation ensures our results can provide robust comparison between two networks without sample size effects.

### 2021-02-05 analysis/paper/supplementary-material.Rmd
edit the result section based on BF's comments

@@
 MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
-We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm.
+We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on he approximate exchange algorithm [@Caimo2011].
 This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.

### 2021-02-06 analysis/paper/paper.Rmd
edit discussion and conclusion based on BF's comments

@@
 ```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
-The size of each node is proportional to "node degrees" or the number of connections to a node.
+The size of each node is proportional to 'node degrees' or the number of connections to a node.
 The thickness of tie represents the number of goods in common between two nodes."}
@@
 <!-- theoretical framework for the results-->
-A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration of connections among fewer burials (lower transitivity and positive popularity) after the European presence during the 17th century.
+A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration of connections among fewer burials (lower transitivity and positive popularity) after the European arrival in the 17th century.
 This supports our models of chronological change from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
@@
 A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
-In contrast, a network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
+In contrast, a network-based society, as suggested the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
 It should be noted that the corporate/network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
@@
 The covariants for networks provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
-Our covariants were age, sex, ritual pottery, and a wealth index based on the value of grave goods.
-The pre-European network shows that burials with pottery as ritual vessels tend to be tied.
-On the other hand, in the post-European network ties tend to be between burials within the same level of wealth.
-Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied.
-This illustrates that different social and economic mechanisms determined relations between burials for the two periods.
-Before European arrival, ritual behaviors seem to have been important status indicators, relative to the use of prestige goods.
-After European presence was established, status indicators shifted from ritual to wealth differences and sex is not associated with status.
-Physical distance between burials did not correlate to network formation for either period.
-This means geographical proximity, implying some kinship association, is not a factor for sharing similar goods between burials.
-This may suggest that the use of burial goods is more individual-oriented instead of kinship-based.
-In addition, demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows small effect for the post European network.
-This may be due to the unavailability of age and sex values for many burials in our sample.
+In other words, we can know what kinds of socio-cultural factors associated with the acquisition and distribution of foreign prestige goods reflected by burial networks.
+Our covariants were age, sex, ritual pottery, and wealth ranks calculated based on the value of grave goods.
+The results indicate that the major factor contributing to ties between burials is ritual for the pre-European network, and wealth for the post-European network.
+Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied that suggest they are not associated with the network pattern we observed.
+Those results illustrate that different social and economic mechanisms determined relations between burials, or the interconnections of prestige goods, for the two periods.
+Before European arrival, ritual behaviors seem to have been important factors that structure the interconnections of prestige goods in the burial context.
+This might imply that ritual could be a status indicator when the foreign goods were first introduced to the local society.
+After European presence was established, status indicators shifted from ritual to wealth differences with the least sex status discrimination.
+Moreover, geographical proximity, implying some kinship association, is not a factor for sharing similar prestige items between burials for both phases.
+This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
+Demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows a negative effect for the post-European network.
+This may also support a shift to more wealth and individual based status competition without sex differences in the later phase.
+But we also note that no effects may be due to the unavailability of age and sex values for many burials in our sample.
 
 <!-- other evidence -->
-We used burial data as a proxy to explore the relations based on the argument that burials represent social structures [@Binford1971; @Saxe1970].
-We note that there could be an issue of disconnection between the role of the deceased in a burial context and their life, which could be reduced by examining evidence from the residential context [@Chapman2003].
+We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on the assumption that burials can represent social structures [@Binford1971; @Saxe1970].
+We argue that the social relation at Kiwulan changed because of the contact with European colonial power and associated foreign trade.
+This argument is supported by the changes in burial network structures from before and after European arrival and their associated underlying mechanisms.
+It should be noted that there could be an issue of disconnection between the role of the deceased in a burial context and their life.
+However, this issue can reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
 Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial situation [@LWandBM2020ornament].
@@
 The image of European power could be embedded into prestige objects pre-existing in local Indigenous culture where their values were amplified with concepts of wealth or power, since their values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
-This may lead to ambitious individuals to competing with each other for traditional goods and accumulating wealth.
+This may lead to ambitious individuals competing with each other for traditional goods and accumulating wealth.
 It also demonstrates that local Indigenous societies could actively incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
-We recognize that the meaning of the burial goods remains speculative without rich ethnographic data to support our interpretations.
-Also, the fragmentation issue in our archaeological data could limit network interpretations [eg.
-@Erik2015], such as missing age and sex attributes for some burials at Kiwulan.
-Human skeletal remains can provide information about health for interpreting social status; however, the preservation of Kiwulan human remains was generally not good to yield useful data.
+We recognize that the meaning of the burial goods could be refined with more ethnographic data to support our interpretations.
+Despite fragmentation in our archaeological data, such as missing age and sex attributes for some burials at Kiwulan, we can still capture the general patterns of network structures related to social implication [eg.
+@Erik2015].
+Human skeletal remains can provide information about health for interpreting social status; however, the preservation of human remains at Kiwulan was generally not good to yield useful data.
+By focusing on burial goods, we seek to understand the social changes at Kiwulan in relation to the contact of Europeans.
 
@@
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-In this paper, we presented a novel approach for studying burial relations using ERGMs within a Bayesian framework.
+In this paper, we presented a novel approach for studying burials to interpret social structures using ERGMs within a Bayesian framework.
 We examined social changes in a European pericolonial context in northeastern Taiwan.
@@
 The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
-Before the arrival of Europeans, the burial network has a tendency of more closed relationships with ritual elements as the key formation mechanism.
+Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism.
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-The changes in formation mechanisms for networks from ritual practice connections to wealth level connections suggest increased behavior of wealth accumulation and differentiation stimulated by the European presence and associated long-distance trade network.
-This aligns with social changes from a more corporate to a more network mode with different strategies for achieving power and different degrees of social inequality.
+The changes in formation mechanisms for networks from ritual practice connections to wealth level connections suggest increased behavior of wealth accumulation stimulated by the European presence and associated long-distance trade network.
+This aligns with changes in horizontal hierarchy from a more corporate to a more network mode that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
 Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
@@
 <!-- technical and methodological implication-->
-This case study demonstrates the methodological improvement that Bayesian inference on ERGMs provides to inform and enhance studies of relational data in archaeology.
+Furthermore, this case study demonstrates the methodological improvement that Bayesian inference on ERGMs provides to inform and enhance studies of relational data in archaeology.
 A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.

### 2021-02-08 analysis/paper/paper.Rmd
reorganize the Intro and the background sections

@@
 
-<!--burial study- network analysis-->
-Burial analysis by archaeologists is an approach to understand past social structures through the study of the physical traces of mortuary practices.
-The material culture and biological records of burial behaviors can represent the social ranking or identities of the deceased [@Binford1971; @Drennan2010; @Saxe1970].
-Despite criticism that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982], burials can still provide valuable information to infer past societies [@Chapman2003].
-By studying burial forms, grave goods, and ritual behaviors that structure the material configuration of burials, we can infer social status and relations between members of a community that give an insight to social structures [@Byrd1995; @Seikel2011].
-Based on network science and graph theory, the recent development of social network analysis provides many new tools to visualize and analyze relationships in archaeological data [@Borgatti2009; @Brughmans2013].
-Network science is the study of relational data where a phenomenon is conceptualized as a network through two steps, abstraction and representation [@Brandes2013; @Collar2015].
-A network can be viewed as a patterned aggregation that includes individual elements (i.e.
+<!--study social changes in pericolonial context using burial data as a proxy-->
+Social changes in Indigenous societies when faced with colonial powers are commonly observed in many parts of the world, especially for European colonies where Indigenous economic, culture, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
+Recent studies demonstrate that the indirect effect of colonialism or involvement in long-distance trade could also have impacts on Indigenous societies in a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
+To explore social changes, burials are important evidence to understand past societies because material culture and biological records of burial behaviors can represent the social ranking or identities of the deceased, and social relations between them [@Binford1971; @Drennan2010; @Saxe1970].
+Burial treatments can further imply social complexity or social inequality where foreign goods from long-distance trade were often used by Indigenous people to express status [@Carter2015; @Dolfini2019].
+Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
+Recent development of network modeling provides new tools to explore the structure of the material configuration in archaeological data and dependencies between variables to understand past relationships, interactions, and underlying mechanisms [@Brughmans2013; @Brughmans2018; @Crabtree2015; @Erik2015; @Freeman2004; @Salvini2010; @Ostborn2015].
+
+<!-- intro to networks: move to SI -->
+<!-- two types of network analysis: move to SI -->
+
+<!---network analysis and the basic concepts-->
+Different from a static characterization of burial variables, network modeling approaches have the potential to study social changes in a dynamic way by examining the intercorrelations between burials.
+Drawing largely on network science and graph theory, a key assumption in the interpretation of archaeological networks is that a phenomenon can be conceptualized as a network through abstraction and representation [@Brandes2013; @Collar2015; @Mills2017; @Peeples2019].
+A network is viewed as a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
@@
 252].
-This paper will use a novel Bayesian approach on exponential random graph models for networks to investigate the formation of relationships between burials in northeastern Taiwan to explore the indirect impact of foreign presence on social organization.
+Among the network modeling methods, exponential random graph models (ERGMs) are one important family of statistical graphs to explore complex networks [@Ghafouri2020].
+However, the computational difficulties and the sensitivity to uncertainties to some extent limits its applications [@Caimo2014].
+In this paper, we try to go beyond ERGMs with a novel Bayesian approach to investigate the formation of material connections between burials and changes in network structures by examining evidence from a case study in northeastern Taiwan.
 
-<!-- intro to networks: move to SI -->
+<!--research aim and questions with archaeological specific case study-->
+We study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan, which was occupied from the 14th to 19th centuries.
+This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
+We argue that the involvement in the long-distance trade stimulated by the Europeans might lead to increased social inequality at Kiwulan based on the theoretical framework of corporate-network strategies [@Blanton1996; @Feinman2000].
+We test our hypothesis that changes in power structure from a corporate mode to a network mode can be observed in the burial network after the European arrival in 17th century Taiwan.
+We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
+(2) if so, what are the major variables affecting or forming the unequal social positions represented that might hint at social heterogeneity influence by the European arrival?
+By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
+In addition, we introduce a novel Bayesian inference on statistical modeling for networks that is suitable for networks with smaller sample sizes.
 
-<!-- two types of network analysis: move to SI -->
+# Background {-}
 
@@
 
-<!-- archaeological background -->
+<!-- archaeological background in NE Taiwan in general -->
 Archaeological sites in northeastern Taiwan show the evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and burial contexts around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament].
@@
 Despite the fact that we recorded an uneven distribution of imported goods in burials, the degree of differentiation across burials over time remained unclear.
+
+<!--specific case from Kiwulan-->
+To explore indirect colonial impacts on Indigenous societies in northwestern Taiwan, our case study focuses on Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age settlement (1350-1850 AD) [@Chen2007].
+Excavations revealed abundant trade goods, including imported ceramics and stonewares, metal artifacts, glass beads, and stone beads, indicative of frequent trade activities.
 The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
@@
 The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
-
-<!-- archaeological specific case study -->
-In what follows, we explore social changes in a pericolonial context at Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age site in northeastern Taiwan, which was occupied from 14th to 19th centuries.
-This occupation period includes the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
-We argue that the involvement in the long-distance trade stimulated by the Europeans might lead to increased social inequality at Kiwulan.
-This argument would be supported by changes in the structure of networks built on burial data based on the assumption that burials can reflect social structure.
-We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival in 17th century Taiwan.
-We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
-(2) what are the major variables affecting or forming the unequal social positions represented by higher degree of network popularity (e.g., a few nodes have more relationships with other nodes) that might hint at social heterogeneity influence by the European arrival?
-By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in an indirect way by examining burial data.
-We also introduces a novel Bayesian approach on statistical modeling for networks that is suitable for networks with smaller sizes.
 
@@
 
-# Exponential random graph models in a Baysesian framework  {-}
+## Exponential random graph models in a Bayesian framework  {-}
 
@@
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
-Although their models do not fit perfectly with the observed network of archaeological data, they proposed that ERGM is a promising method for exploration and hypothesis testing for social processes.
+By fitting models with the observed network of archaeological data, they proposed that ERGMs is a promising method for exploration and hypothesis testing for social processes.
 Similarly, @Amati2019 modeled three networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms, including proximity, inter-cultural items, and pottery types.
 By comparing hypothesized networks with the observed sites, they found that the presence of hub sites can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
-However, those studies also point out some limitations of ERGM, such as sensitivity to missing data and less ability to handle uncertainty.
-In addition, it is difficult in ERGM to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
+However, those studies also point out some limitations of ERGMs, such as sensitivity to missing data and less able to handle uncertainty.
+Also, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
 A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with a larger set of networks [@Caimo2017bayesian].
@@
 This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-By fitting an ERMG with an MCMC algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
-This allows uncertainty quantification by examining the posterior mean and 95% credible intervals, or the odds ratios (conditional probability of observing an edge) that can replace p-values for assessing a null hypothesis [@Caimo2017bayesian].
-The typical criteria for interpreting an significant effect of a parameter is that its credible interval does not contain zero, or an odds ratio is greater/less than one, meaning a significant positive/negative effect [@Harris2013, pp.
+By fitting an ERMG with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
+This allows uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].
+The typical criteria for interpreting a significant effect of a parameter is that its credible interval does not contain zero [@Harris2013, pp.
 51].
-In addition, Bayesian approaches are also useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
+In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
 @Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data is missing.
@@
 <!-- archaeological setting -->
-We analyzed burial data collected from the published excavation reports, and the original fieldwork notes for the upper component of Kiwulan, an Iron Age settlement (1350-1850 AD) [@Chen2007].
-Excavations revealed abundant trade goods, including imported ceramics and stonewares, metal artifacts, glass beads, and stone beads, indicative of frequent trade activities [@Chen2007].
-A total of 90 burials were unearthed that are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
+We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) (Figure \@ref(fig:figure-KWL-map)) [@Chen2007].
+A total of 90 burials were unearthed from adjacent squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
 Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement.
@@
 For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society.
-However, @Hsieh2012 suggested a relatively egalitarian structure based on comparative analysis of the frequencies of all burial goods.
-She found that the burials with rare goods, also prestige goods in local culture, were usually associated with elders, which might indicate achieved, rather than inherited, status.
+However, @Hsieh2012 suggested a relatively egalitarian structure based on the comparative analysis of the frequencies of all burial goods.
+She found that the burials with rare prestige goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
 One important limitation of these previous studies is that they did not use analytical units suitable for comparing the phases before and after European arrival.
@@
 The burial value index is an important attribute for economical inequality that we assigned the burials into three ranks, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
-The number is the sum of values of each type of prestige good, which is calculated by the total number of prestige item from burial context over the number of prestige items in a burial.
+The number is the sum of values of each type of prestige good, which is calculated by the total number of prestige items from burial context over the number of prestige items in a burial.
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], where model specification and simulation procedures is based on statnet package [@Hunter2008; @Morris2008; @Handcock2008].
-Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification with the corresponding archaeological evidence.
+Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification.
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
@@
 This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
-This implies that the presence of a better access to trade goods and the behavior of wealth accumulation or display in burial events.
+This implies that the presence of better access to trade goods and the behavior of wealth accumulation or display in burial events.
 In general, the post-European network model has a smaller transitivity effect and a positive popularity effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
-In addtion, both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
+In addition, both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
 
@@
 Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
-For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference.
+For density and transitivity, the 95% confidence intervals from the two networks overlapped with each other that indicates no significant difference.
 This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with some degree of difference.
@@
 Demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows a negative effect for the post-European network.
-This may also support a shift to more wealth and individual based status competition without sex differences in the later phase.
+This may also support a shift to more wealth and individual-based status competition without sex differences in the later phase.
 But we also note that no effects may be due to the unavailability of age and sex values for many burials in our sample.
@@
 This argument is supported by the changes in burial network structures from before and after European arrival and their associated underlying mechanisms.
-It should be noted that there could be an issue of disconnection between the role of the deceased in a burial context and their life.
-However, this issue can reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
+It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
+However, this issue can be reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
 Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial situation [@LWandBM2020ornament].
@@
 @Acabado2017; @Trabert2017].
-Furthermore, our study also highlights the possibility of uneven access to foreign trade goods that allows active accumulation or even manipulation their values by local people instead of being a passive recipients of trade goods.
+Furthermore, our study also highlights the possibility of uneven access to foreign trade goods that allows active accumulation or even manipulation their values by local people instead of being passive recipients of trade goods.
 
@@
 Goodreau for the discussion regarding the methods at the earliest stage of analysis.
-This work was supported in part by the travel grant from the Department of Anthropology at University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
+This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
 We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.

### 2021-02-08 analysis/paper/paper.Rmd
edit to reduce word counts

@@
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
-Burials with the same kinds of prestige goods could indicate some underlying social relations where individuals share similar access to trade, exchange, and gifting networks according to their status [eg.
-@Coward2013, pp.
-252].
 Among the network modeling methods, exponential random graph models (ERGMs) are one important family of statistical graphs to explore complex networks [@Ghafouri2020].
@@
 By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
-In addition, we introduce a novel Bayesian inference on statistical modeling for networks that is suitable for networks with smaller sample sizes.
 
@@
 <!-- theoretical framework for the results-->
-A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration of connections among fewer burials (lower transitivity and positive popularity) after the European arrival in the 17th century.
-This supports our models of chronological change from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
+A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups toward a more centralized network with concentration of connections among fewer burials after the European arrival in the 17th century.
+This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
 These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
-Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan.
+Compared to vertical hierarchies that stress stratified social characteristics, the corporate-network strategies may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan [cf.
+@Ueda2016].
 A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
 In contrast, a network-based society, as suggested the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
-It should be noted that the corporate/network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
+It should be noted that the corporate-network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
 
@@
 The covariants for networks provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
-In other words, we can know what kinds of socio-cultural factors associated with the acquisition and distribution of foreign prestige goods reflected by burial networks.
+In other words, we can know what kinds of socio-cultural factors associated with the acquisition and distribution of foreign prestige goods.
 Our covariants were age, sex, ritual pottery, and wealth ranks calculated based on the value of grave goods.
-The results indicate that the major factor contributing to ties between burials is ritual for the pre-European network, and wealth for the post-European network.
+The results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
 Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied that suggest they are not associated with the network pattern we observed.
@@
 Before European arrival, ritual behaviors seem to have been important factors that structure the interconnections of prestige goods in the burial context.
-This might imply that ritual could be a status indicator when the foreign goods were first introduced to the local society.
+This might imply that ritual could be a status indicator when the foreign goods were first introduced to Indigenous societies.
 After European presence was established, status indicators shifted from ritual to wealth differences with the least sex status discrimination.
@@
 This may also support a shift to more wealth and individual-based status competition without sex differences in the later phase.
-But we also note that no effects may be due to the unavailability of age and sex values for many burials in our sample.
+But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.
 
 <!-- other evidence -->
-We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on the assumption that burials can represent social structures [@Binford1971; @Saxe1970].
+We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks according to their status [eg.
+@Coward2013, pp.
+252].
 We argue that the social relation at Kiwulan changed because of the contact with European colonial power and associated foreign trade.
-This argument is supported by the changes in burial network structures from before and after European arrival and their associated underlying mechanisms.
+This argument is supported by the changes in burial network structures through time and their associated underlying mechanisms.
 It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
 However, this issue can be reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
-Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial situation [@LWandBM2020ornament].
-This is consistent with the result of a more centralized burial network after the presence of Europeans, which supports a connection between a burial and the social role of the living person.
-Since the burial goods used for making ties were treated as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods observed in the burial data likely reflect social contexts of the living people at Kiwulan in a pericolonial context.
+Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation [@LWandBM2020ornament].
+This is consistent with the result of a more centralized burial network after the presence of Europeans, that supports a connection between a burial and the social role of the living person.
+Also, the burial goods used for making ties were treated as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
 
@@
 We recognize that the meaning of the burial goods could be refined with more ethnographic data to support our interpretations.
-Despite fragmentation in our archaeological data, such as missing age and sex attributes for some burials at Kiwulan, we can still capture the general patterns of network structures related to social implication [eg.
-@Erik2015].
 Human skeletal remains can provide information about health for interpreting social status; however, the preservation of human remains at Kiwulan was generally not good to yield useful data.
 By focusing on burial goods, we seek to understand the social changes at Kiwulan in relation to the contact of Europeans.
+Despite fragmentation in our archaeological data, such as missing age and sex attributes for some burials at Kiwulan, we can still capture the general patterns of network structures related to social implication.
 
@@
 We examined social changes in a European pericolonial context in northeastern Taiwan.
-We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
+We tested the hypotheses of changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
 The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
@@
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-The changes in formation mechanisms for networks from ritual practice connections to wealth level connections suggest increased behavior of wealth accumulation stimulated by the European presence and associated long-distance trade network.
-This aligns with changes in horizontal hierarchy from a more corporate to a more network mode that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
+The changes in formation mechanisms for networks from ritual practice connections to wealth level connections could suggest increased behavior of wealth accumulation stimulated by the European presence and associated long-distance trade network.
+This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
 Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
 @Acabado2017; @Trabert2017].
-Furthermore, our study also highlights the possibility of uneven access to foreign trade goods that allows active accumulation or even manipulation their values by local people instead of being passive recipients of trade goods.
 

### 2021-02-09 analysis/paper/paper.Rmd
edit based on BM's comments

@@
   Burials provide valuable information to study social structures based on the assumption that burials and associated grave goods can represent social roles and relations in a society.
-To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns that can infer interactions and relationships between entities.
-Statistical approaches to network analysis, such as exponential random graph models (ERGMs), provide a way to test hypotheses on dynamic process of network formation.
+To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns to infer interactions and relationships between entities.
+Statistical approaches to network analysis, such as exponential random graph models (ERGMs), provide a way to test hypotheses on dynamic processes of network formation.
 However, the computational difficulty and sensitivity to uncertainties limits the application of ERGMs.
-In this paper, we introduce a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification for uncertainty, and robust model evaluation of network properties.
-We tested the hypothesis of social changes relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
-The results indicate a transition from ties based on ritual objects to wealth objects, and a more centralized structure with social differentiation after European presence in the 17th century.
+In this paper, we introduce a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification of uncertainty, and robust model evaluation of network properties.
+We tested a hypothesis of social change relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
+The results indicate a transition from ties based on ritual objects to wealth objects, and a more centralized structure with increased social differentiation after European presence in the 17th century.
 Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expands the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
@@
 
+<!---network analysis and the basic concepts-->
+Network analysis has been increasingly used by archaeologists to explore rational data to understand past relationships, interactions, or overall structures of observed phenomena [@Brandes2013; @Mills2017; @Peeples2019].
+A network can be viewed as a patterned aggregation that includes individual elements, pair-wise relationships, and an overall structure showing global patterns represented in the data.
+Recently, network modeling provides many new tools to examine statistically the intercorrelations between individual elements in a dynamic way.
+Among network modeling methods, exponential random graph models (ERGMs) are one important family of statistical graphs to explore complex networks [@Harris2013; @Ghafouri2020].
+However, the computational difficulties and the sensitivity to uncertainties to some extent limits its practical applications [@Caimo2014].
+In this paper, we go beyond ERGMs with a novel Bayesian approach to investigate the formation of material connections using burial data from a case study in Iron Age northeastern Taiwan to understand social changes during foreign contact.
+
 <!--study social changes in pericolonial context using burial data as a proxy-->
-Social changes in Indigenous societies when faced with colonial powers are commonly observed in many parts of the world, especially for European colonies where Indigenous economic, culture, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
-Recent studies demonstrate that the indirect effect of colonialism or involvement in long-distance trade could also have impacts on Indigenous societies in a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
+Social changes in Indigenous societies, when faced with colonial powers, are commonly observed in many parts of the world, especially European colonies where Indigenous economic, cultural, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
+Recent studies demonstrate that the indirect effect of colonialism, or involvement in long-distance trade, could also have impacts on Indigenous societies, this is known as a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
 To explore social changes, burials are important evidence to understand past societies because material culture and biological records of burial behaviors can represent the social ranking or identities of the deceased, and social relations between them [@Binford1971; @Drennan2010; @Saxe1970].
-Burial treatments can further imply social complexity or social inequality where foreign goods from long-distance trade were often used by Indigenous people to express status [@Carter2015; @Dolfini2019].
+Burial treatments can further indicate social complexity or inequality where foreign goods from long-distance trade were used by Indigenous people to express status [@Carter2015; @Dolfini2019].
 Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
-Recent development of network modeling provides new tools to explore the structure of the material configuration in archaeological data and dependencies between variables to understand past relationships, interactions, and underlying mechanisms [@Brughmans2013; @Brughmans2018; @Crabtree2015; @Erik2015; @Freeman2004; @Salvini2010; @Ostborn2015].
+Recent development of network modeling provides new tools to explore the structure of the material configurations and dependencies between variables to understand past relationships, interactions, and underlying mechanisms [@Brughmans2013; @Brughmans2018; @Crabtree2015; @Erik2015; @Freeman2004; @Salvini2010; @Ostborn2015].
 
@@
 
-<!---network analysis and the basic concepts-->
-Different from a static characterization of burial variables, network modeling approaches have the potential to study social changes in a dynamic way by examining the intercorrelations between burials.
-Drawing largely on network science and graph theory, a key assumption in the interpretation of archaeological networks is that a phenomenon can be conceptualized as a network through abstraction and representation [@Brandes2013; @Collar2015; @Mills2017; @Peeples2019].
-A network is viewed as a patterned aggregation that includes individual elements (i.e.
-individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data.
-Among the network modeling methods, exponential random graph models (ERGMs) are one important family of statistical graphs to explore complex networks [@Ghafouri2020].
-However, the computational difficulties and the sensitivity to uncertainties to some extent limits its applications [@Caimo2014].
-In this paper, we try to go beyond ERGMs with a novel Bayesian approach to investigate the formation of material connections between burials and changes in network structures by examining evidence from a case study in northeastern Taiwan.
+# Exponential random graph models in a Bayesian framework  {-}
 
-<!--research aim and questions with archaeological specific case study-->
-We study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan, which was occupied from the 14th to 19th centuries.
-This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
-We argue that the involvement in the long-distance trade stimulated by the Europeans might lead to increased social inequality at Kiwulan based on the theoretical framework of corporate-network strategies [@Blanton1996; @Feinman2000].
-We test our hypothesis that changes in power structure from a corporate mode to a network mode can be observed in the burial network after the European arrival in 17th century Taiwan.
-We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
-(2) if so, what are the major variables affecting or forming the unequal social positions represented that might hint at social heterogeneity influence by the European arrival?
-By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
-
-# Background {-}
-
-## A Case Study from Northeastern Taiwan  {-}
-
-<!-- archaeological background in NE Taiwan in general -->
-Archaeological sites in northeastern Taiwan show the evidence of accumulation patterns of imported goods, such as ornaments and porcelains, in residential areas and burial contexts around the same time as European contacts in the 17th century [@Chen2007; @LWandBM2020ornament].
-Imported goods were treated as prestige or wealth in those Indigenous societies that were commonly mentioned in Spanish historical records in the 17th century (1624-1642) [@Borao2009; @LiandWu2006].
-For example, a document written by a Spanish priest described a Spanish soldier using carnelian beads to exchange for natural resources with local Indigenous people, because of their high value in Indigenous culture [@LiandWu2006].
-In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status in their community [@LiandWu2006].
-Despite the fact that we recorded an uneven distribution of imported goods in burials, the degree of differentiation across burials over time remained unclear.
-
-<!--specific case from Kiwulan-->
-To explore indirect colonial impacts on Indigenous societies in northwestern Taiwan, our case study focuses on Kiwulan (Figure \@ref(fig:figure-KWL-map)), an Iron Age settlement (1350-1850 AD) [@Chen2007].
-Excavations revealed abundant trade goods, including imported ceramics and stonewares, metal artifacts, glass beads, and stone beads, indicative of frequent trade activities.
-The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
-A general model to summarize this situation in northeastern Taiwan might assume that the influence of a colonial power, combined with high local values attached to imported goods, might lead to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-The observed uneven distribution in Northeastern Taiwan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
-
-```{r figure-KWL-map, fig.cap = "The location of Kiwulan and the European forts at Heping dao and Tamsui in northern Taiwan (modified from Wang and Marwick, 2020).
-Map data from naturalearthdata.com."}
-
-knitr::include_graphics(here::here("analysis", "figures", "000-KWL-map.png"))
-```
-
-## Exponential random graph models in a Bayesian framework  {-}
-
-<!-- ERGM in archaeology, and limitations --> 
+<!-- ERGMs in archaeology--> 
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
@@
 By comparing hypothesized networks with the observed sites, they found that the presence of hub sites can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
+
+<!--limitations of ERGMs-->
 However, those studies also point out some limitations of ERGMs, such as sensitivity to missing data and less able to handle uncertainty.
@@
 A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with a larger set of networks [@Caimo2017bayesian].
-This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
+This is also termed “doubly intractable” since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
 Model degeneracy is another issue where probability models tend to overestimate a small number of extreme graphs by assigning too much weight, e.g., in extreme cases such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
@@
 This allows uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].
-The typical criteria for interpreting a significant effect of a parameter is that its credible interval does not contain zero [@Harris2013, pp.
-51].
 In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
-@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on the average, 80% of the ties when a third of data is missing.
+@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on average, 80% of the ties when a third of data is missing.
 
-# Materials {-}
+# Background {-}
+
+## A Case Study from Northeastern Taiwan  {-}
+
+<!--archaeological specific case study and our model-->
+We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries.
+This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
+Our general model proposes that the influence of a colonial power, combined with high local values attached to imported goods, leads to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
+
+<!--theoretical framework and research questions-->
+Using the theoretical framework of corporate-network strategies [@Blanton1996; @Feinman2000], we test a hypothesis that changes from a corporate mode (…) to a network mode (…) can be observed in the Kiwulan burial network after European arrival.
+We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
+(2) if so, what are the major variables affecting or forming unequal social positions that might hint at social heterogeneity?
+By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
+
+```{r figure-KWL-map, fig.cap = "The location of Kiwulan and the European forts at Heping dao and Tamsui in northern Taiwan (modified from Wang and Marwick, 2020).
+Map data from naturalearthdata.com."}
+
+knitr::include_graphics(here::here("analysis", "figures", "000-KWL-map.png"))
+```
+
+## Materials {-}
 
@@
 
-<!-- archaeological setting -->
-We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) (Figure \@ref(fig:figure-KWL-map)) [@Chen2007].
+<!-- Kiwulan burials and a review of current studies of social organization -->
+We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) [@Chen2007].
 A total of 90 burials were unearthed from adjacent squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-Burials are oriented in an east-west direction on the north side of the residential area, indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement.
+Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses.
 Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
@@
 One important limitation of these previous studies is that they did not use analytical units suitable for comparing the phases before and after European arrival.
-Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until Han Chinese immigrants arrival).
-The discord over the discussion of Kiwulan social organization could be associated with chronological differences that have not yet been well studied.
+Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until arrival of Han Chinese immigrants).
 
@@
 
-<!-- chronology -->
+<!-- Our chronology and principles of network building -->
 To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
-Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [@LWandBM2020ornament; @Chen2007].
+Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament].
 We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
-There are `r disturb` burials heavily disturbed by modern construction that are also excluded because we cannot determine their chronology.
+We also excluded `r disturb` burials that were heavily disturbed by modern construction because we cannot determine their chronology.
 Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf.
@@
 The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments.
-Previous studies indicate that these items are considered as high-value across different archaeological contexts based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
-Historical accounts written by Europeans mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
+These items are considered as high-value across different archaeological contexts based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
+For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads’ high value in Indigenous culture [@LiandWu2006].
+In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as having higher status in their community [@LiandWu2006].
 
@@
 ## Hypothesis and construction of networks {-}
-
-<!-- our hypothesis --> 
-Our "pericolonial impact hypothesis" is that rare, high quality burial goods accumulated in only a few individuals' burials after the European presence because of increased social inequality at Kiwulan due to differential access to these goods.
-A related effect of the European presence in northeastern Taiwan may have been manipulation of the European colonial image by ambitious local Indigenous individuals for building their personal status and power, recognized as network strategies to achieve power [@Drennan2010; @Feinman2000].
-If social inequality gradually increased in Kiwulan as we hypothesize, then we expect to observe a network with higher degree (popularity) and less transitivity (cohesion), along with greater differences in wealth accumulation according to burial goods.
+<!-- our expectation based on hypothesis --> 
+Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to an increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Feinman2000].
+This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
+If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher prominence (popularity) and less transitivity (cohesion), along with associations with wealth accumulation according to burial goods.
 To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
@@
 <!-- burial data and the definition of tie formation  --> 
-We used burial goods in common to construct ties for representing a flow of goods between burials.
-Since trade beads are commonly found across burials with substantial differences in quantities, we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
-For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods (i.e.
+Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
+For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they both possess each type of goods (i.e.
 presence or absence).
-Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index.
-Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting based on historical records.
-The burial value index is an important attribute for economical inequality that we assigned the burials into three ranks, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
-The number is the sum of values of each type of prestige good, which is calculated by the total number of prestige items from burial context over the number of prestige items in a burial.
-Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
+Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index.
+Ritual pottery was identified as locally made ceramics placed above an individual.
+Historical records link these ceramics to funeral feasting [@citation].
+We calculated our burial value by summing values of all types of prestige goods in each burial context.
+The value of each type of prestige good is calculated by taking the number of prestige items in all burials, and then dividing this by total number of prestige items in one burial context.
+We then assigned the burials into three ranks according to their burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
+Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
@@
 ```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
-The size of each node is proportional to 'node degrees' or the number of connections to a node.
+The size of each node is proportional to node degrees or the number of connections to a node.
 The thickness of tie represents the number of goods in common between two nodes."}
@@
 <!-- Network construction and specification -->
-We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], where model specification and simulation procedures is based on statnet package [@Hunter2008; @Morris2008; @Handcock2008].
-Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification.
+We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014].
+Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification [@Morris2008].
 Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
 Based on our hypothesis, we model a network with increased social inequality to be represented by endogenous network effects: low transitivity and high popularity.
-We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore the tendency of tie forming.
-For example, age-homophily means people of the same age tend to have the same burial goods.
-We also include the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006].
-Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics, with covariates to understand the preferential tendency of formation of relationships.
+We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore the importance of these variables in tie forming.
+For example, if age-homophily is important here, then people of the same age should have the same burial goods.
+We also include the physical distance between burials as an indicator of a kinship-based relations since the deceased from the same family were buried nearby [@LiandWu2006].
+Our model may reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
+This would be indicated by high popularity or degree values in the network statistics.
 According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
@@
 knitr::kable(ergm_terms,
-             caption = "Network parameters in ERGMs used for model specifications with associated interpretation for burial relations.")
+             caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.")
 ```
@@
 
+## Choice of prior values to evaluate our anthropological model
+
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
@@
 stronger correlations for shorter distances.
-To evaluate our anthropological assumption about increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
-On the contrary, we set the priors for the network after the European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
+To evaluate our anthropological model of increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
+Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
 This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
@@
             `2.5%` = round(quantile(posterior, 0.025), 2),
-            `97.5%` = round(quantile(posterior, 0.075), 2)) %>% 
+            `97.5%` = round(quantile(posterior, 0.975), 2)) %>% 
   pivot_wider(names_from = phase, 
@@
   width(j = 2:9, width = 0.65) %>% 
-  set_caption("Estimated posterior means, medians, and 95% credible intervals for each network parameter of two models.")
+  set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models.
+Confidence intervals that do not include zero idicate significant effects of parameters (Caimo, 2017).")
 ft

### 2021-02-09 analysis/paper/supplementary-material.Rmd
edit based on BM's comments

@@
 -  further historical background of 17th century northeastern Taiwan in a pericolonial context
+-  further review of network analysis and archaeological applications
 -  further definition of our approach to tie formation with the beads from Kiwulan burials
--  further review of network analysis and archaeological applications
 -  further details on methods of modeling fitting, and
@@
 <!--fit in a bigger historical picture in a pericolonial context: move to SI -->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017].
+The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
 Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007].
@@
 In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.
+The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
 
@@
 
+## Review of network analysis and archaeological applications {.unnumbered}
+
+<!-- intro to networks: move to SI -->
+Drawing largely on network science and graph theory, a key assumption in the construction of archaeological networks is that a phenomenon can be conceptualized as a network through abstraction and representation [@Brandes2013; @Collar2015; @Mills2017; @Peeples2019].
+A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
+The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
+In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
+For example, past trade can be conceptualized as a network of individual entities connected by shared similar objects or elements, such as the flow of goods, to represent their interactions [@Collar2015, pp.
+4].
+This concept can be applied to burial contexts where each burial is an actor linked by similar elements, such as burial goods.
+Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
+This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
+
+<!-- two types of network analysis: move to SI -->
+Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
+There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
+Node level focuses on node properties in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
+By quantifying those network properties, archaeologists can answer a wide range of research questions.
+
+Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
+Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
+Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
+Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+
+<!-- current approaches to network analysis: move to SI -->
+Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
+Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
+ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
+In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
+For example, nodes with similar attributes are more likely to have a relationship, such as friendship between people with the same hobby.
+Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
+Those configurations represent the structure or the property of a network and can be expressed by network statistics.
+
+By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
+Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
+ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+
 ## Tie formation with the beads from Kiwulan burials {.unnumbered}
@@
 <!-- burial data and the definition of tie formation: partially move to SI  -->
-
 For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)).
@@
 
-## Review of network analysis and archaeological applications {.unnumbered}
-
-<!-- intro to networks: move to SI -->
-
-A social network is generally visualized as a graph consisting of a set of socially-relevant nodes/actors, connected by edges/ties representing one or more relations, such as friendship, collaborations, information flow, trade ties, or any other forms of connection of interest [@Marin2011; @Wasserman1994].
-The ties can be classified into four major types, including similarities, social relations, interactions, and flows [@Borgatti2009].
-In archaeology, actors can be people, groups, objects, places, or events, with ties built on similarity, proximity, or co-presence of material culture to create patterns reflecting influence, geographical distance, or affiliations in social groups [@Mills2017; @Brughmans2018; @Peeples2019].
-For example, past trade can be conceptualized as a network of individual entities connected by shared similar objects or elements, such as the flow of goods, to represent their interactions [@Collar2015, pp.
-4].
-This concept can be applied to burial contexts where each burial is an actor linked by similar elements, such as burial goods.
-Burial goods, especially high value goods, can reflect social practices in broader cultural contexts to represent personal wealth or social status from which we can infer social differentiation or complexity [@Gamble2002; @Janes2013].
-This enables the exploration of the structure of the past social organization through the identification of the relationships among burials.
-
-<!-- two types of network analysis: move to SI -->
-
-Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
-There are two common approaches to characterize network properties at two distinct scales: node/edge level and graph level [@Peeples2019].
-Node level focuses on node properties in a network, such as centrality, representing the individual influence or social prominence in a group, while graph level assesses the whole network attributes, such as density, clustering in a network, to generalize relationship patterns [@Mills2017; @Peeples2019].
-By quantifying those network properties, archaeologists can answer a wide range of research questions.
-
-Examples include exploring the political centralization in the Kofun period in Japan through the hierarchical communication network constructed by prestige goods [@Mizoguchi2013], and the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013].
-Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
-Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
-
-<!-- current approaches to network analysis: move to SI -->
-
-Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
-Our use of Bayesian inference on ERGMs is the first application to archaeological data that can bring new insights to understand past social structures by characterizing network properties as a whole.
-ERGMs are an important family of statistical models for networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
-The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007].
-In other words, networks in ERGMs are viewed as dependent variables, where network dependencies and the attributes of nodes/edges can influence the formation of a tie [@Snijders2011].
-For example, nodes with similar attributes are more likely to have a relationship, such as friendship between people with the same hobby.
-Ties form a small structure in a network called a graph configuration, that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity or clustering (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008].
-Those configurations represent the structure or the property of a network and can be expressed by network statistics.
-
-By modeling those network statistics as direct functions of ties by specifying the forms of configurations, we can generate a distribution of random networks that represent our hypothesis-based model [@Morris2008; @Robins2007].
-Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
-
 ## Further details on methods of modeling fitting {.unnumbered}
@@
 <!--Model fitting: move to SI -->
-
 After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm.

### 2021-02-10 analysis/paper/paper.Rmd
revise Intro section to stress the method

@@
 We tested a hypothesis of social change relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
-The results indicate a transition from ties based on ritual objects to wealth objects, and a more centralized structure with increased social differentiation after European presence in the 17th century.
+The results indicate a transition in the process that led to the formation of the burial network, from ties based on ritual objects to wealth objects, and a more centralized structure with increased social differentiation after European presence in the 17th century.
 Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expands the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
@@
 <!---network analysis and the basic concepts-->
-Network analysis has been increasingly used by archaeologists to explore rational data to understand past relationships, interactions, or overall structures of observed phenomena [@Brandes2013; @Mills2017; @Peeples2019].
-A network can be viewed as a patterned aggregation that includes individual elements, pair-wise relationships, and an overall structure showing global patterns represented in the data.
-Recently, network modeling provides many new tools to examine statistically the intercorrelations between individual elements in a dynamic way.
-Among network modeling methods, exponential random graph models (ERGMs) are one important family of statistical graphs to explore complex networks [@Harris2013; @Ghafouri2020].
-However, the computational difficulties and the sensitivity to uncertainties to some extent limits its practical applications [@Caimo2014].
-In this paper, we go beyond ERGMs with a novel Bayesian approach to investigate the formation of material connections using burial data from a case study in Iron Age northeastern Taiwan to understand social changes during foreign contact.
+Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing rational data [@Brandes2013; @Mills2017; @Peeples2019].
+The recent development of modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures inherit in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+By comparing hypothesized networks with the observed network, network modeling can answer a wide range of anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
+Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulating network patterns [@Harris2013; @Ghafouri2020].
+Since relationships emerging relative to their position in a network can be hypothesized, ERGMs allows evaluation of dynamic social processes behind observed archaeological networks [@Brughmans2014].
+Although ERGMs has potential to study social processes, computational difficulties and the sensitivity to uncertainties to some extent limits the practical applications of ERGMs, where results could be challenging for interpretation [@Caimo2014].
 
 <!--study social changes in pericolonial context using burial data as a proxy-->
+In this paper, we go beyond ERGMs with a novel Bayesian approach that can alleviate the computational issues in ERGMs to better interpret the process of network formation for past phenomena.
+We use Baysien ERGMs to explore social changes by investigating the formation of material connections between burials from a case study in Iron Age northeastern Taiwan.
 Social changes in Indigenous societies, when faced with colonial powers, are commonly observed in many parts of the world, especially European colonies where Indigenous economic, cultural, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
@@
 Burial treatments can further indicate social complexity or inequality where foreign goods from long-distance trade were used by Indigenous people to express status [@Carter2015; @Dolfini2019].
+
 Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
-Recent development of network modeling provides new tools to explore the structure of the material configurations and dependencies between variables to understand past relationships, interactions, and underlying mechanisms [@Brughmans2013; @Brughmans2018; @Crabtree2015; @Erik2015; @Freeman2004; @Salvini2010; @Ostborn2015].
+We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
+individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data [@Brandes2013].
 
@@
 <!-- ERGMs in archaeology-->
+
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.

### 2021-02-11 analysis/paper/paper.Rmd
edit the intro and background sections

@@
 
-<!---network analysis and the basic concepts-->
+<!---network analysis approaches and basic concepts-->
 Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing rational data [@Brandes2013; @Mills2017; @Peeples2019].
-The recent development of modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures inherit in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-By comparing hypothesized networks with the observed network, network modeling can answer a wide range of anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
-Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulating network patterns [@Harris2013; @Ghafouri2020].
-Since relationships emerging relative to their position in a network can be hypothesized, ERGMs allows evaluation of dynamic social processes behind observed archaeological networks [@Brughmans2014].
-Although ERGMs has potential to study social processes, computational difficulties and the sensitivity to uncertainties to some extent limits the practical applications of ERGMs, where results could be challenging for interpretation [@Caimo2014].
+The recent development of modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures inherent in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+By comparing hypothesized networks with the observed network, network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
+Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Ghafouri2020].
+Since relationships in a network emerging relative to their position in a network can be hypothesized, ERGMs are promising to evaluate dynamic social processes behind observed archaeological networks [@Brughmans2014].
+However, computational difficulties and the sensitivity to uncertainties to some extent limits the practical applications of ERGMs, where results could be challenging for interpretation [@Caimo2014].
 
@@
 In this paper, we go beyond ERGMs with a novel Bayesian approach that can alleviate the computational issues in ERGMs to better interpret the process of network formation for past phenomena.
-We use Baysien ERGMs to explore social changes by investigating the formation of material connections between burials from a case study in Iron Age northeastern Taiwan.
+Using burial data from an Iron Age site in northeastern Taiwan as a case study, we explore social changes by investigating the formation of material connections between burials.
 Social changes in Indigenous societies, when faced with colonial powers, are commonly observed in many parts of the world, especially European colonies where Indigenous economic, cultural, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
@@
 
+<!--Our burial network and research implication-->
 Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
-We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
+Here we treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data [@Brandes2013].
+Our case study illustrates how a Bayesian approach to network modeling provides new insights for studying the formation of archaeological networks and allows testing on theoretical assumptions to better understand dynamic social processes.
 
@@
 <!-- ERGMs in archaeology--> 
-
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
@@
 Our general model proposes that the influence of a colonial power, combined with high local values attached to imported goods, leads to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+Increased social inequality associated with the use of foreign prestige goods is recognized as network strategies for gaining power [@Blanton1996; @Feinman2000].
+Network-based societies are identified in some places in Asia with involvment in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
@@
 <!--theoretical framework and research questions-->
-Using the theoretical framework of corporate-network strategies [@Blanton1996; @Feinman2000], we test a hypothesis that changes from a corporate mode (…) to a network mode (…) can be observed in the Kiwulan burial network after European arrival.
+The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000], while corporate-based societies stress communal ritual elements, shared power, and wealth differentiation, if any, would be associated with corporate groups, such as age groups [@Siegel1999].
+Using the theoretical framework of corporate-network strategies, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
 We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
@@
 
-## Choice of prior values to evaluate our anthropological model
+## Choice of prior values to evaluate our anthropological model {-}
 

### 2021-02-12 analysis/paper/paper.Rmd
edit the whole paper and change variable names

@@
 <!-- Bayesian inference for Exponential Random Graph Models --> 
-Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
+Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
 Prior information is derived from previous data or assumptions of our data.
@@
 <!-- our expectation based on hypothesis --> 
-Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to an increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Feinman2000].
+Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to an increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
 This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
-If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher prominence (popularity) and less transitivity (cohesion), along with associations with wealth accumulation according to burial goods.
+If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with associations with wealth accumulation, indicative of a network-based society.
+In contrast, the network before the European arrival will show low centralization and high transitivity, associated with ritual element, or age/sex groups to indicate a corporate-based society.
 To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
@@
 Ritual pottery was identified as locally made ceramics placed above an individual.
-Historical records link these ceramics to funeral feasting [@citation].
+Those pots mostly have soot on vessel exteriors and charred residues on interiors that are interpreted as vessels for funeral feasting [@Hsieh2009].
 We calculated our burial value by summing values of all types of prestige goods in each burial context.
-The value of each type of prestige good is calculated by taking the number of prestige items in all burials, and then dividing this by total number of prestige items in one burial context.
-We then assigned the burials into three ranks according to their burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
+The value of prestige good is calculated by taking the total number of burials in Kiwulan, and then dividing this by the number of burials with each type of prestige good [@Jorgensen1991].
+We then assigned the burials into three ranks according to their burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (Details about other node attributes in Supplementary Online Materials).
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014].
-Table \@ref(tab:table-ergm-terms) lists the network parameters we used for the model specification [@Morris2008].
-Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of grave goods between two burials.
-Based on our hypothesis, we model a network with increased social inequality to be represented by endogenous network effects: low transitivity and high popularity.
+Table \@ref(tab:table-ergm-terms) lists the network parameters we used for dependence assumptions that define our models [@Morris2008].
+Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of prestige goods between two burials.
+Based on our hypothesis, we model a network with increased social inequality to be represented by endogenous network effects: low transitivity and high centralization.
 We include burial-specific attributes as covariate effects for homophily, such as age, sex, ritual activity, and the degree of wealth, to explore the importance of these variables in tie forming.
@@
 We also include the physical distance between burials as an indicator of a kinship-based relations since the deceased from the same family were buried nearby [@LiandWu2006].
-Our model may reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others.
-This would be indicated by high popularity or degree values in the network statistics.
-According to our hypothesis, the burial evidence from after the European arrival will show higher popularity compared with the burial evidence from before European contact.
+Our model may reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others, with more wealth differentiation after the European arrival.
 
@@
 ergm_terms <- readxl::read_excel(here::here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
-knitr::kable(ergm_terms,
+ergm_terms_selected <-
+  ergm_terms %>% 
+  select(-"Description", -"note")
+knitr::kable(ergm_terms_selected,
              caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.")
@@
 stronger correlations for shorter distances.
-To evaluate our anthropological model of increased social inequality over time, we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and popularity (gwedegree).
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower popularity (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
-Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher popularity (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
-This prior information derives from theory about horizontal hierarchies, which can be viewed as a spectrum that illustrates an increasing social inequality from a corporate mode at one end toward a network mode at the other end [@Drennan2010; @Feinman2000].
+To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and centralization (gwedegree).
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
+Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
 
@@
 A significant effect is determined by consistent positive/negative estimates in a 95% confidence interval.
-For nodal covariates in the pre-European model, ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while wealth rank represented by burial values (value-homophily) has a negative effect.
+For nodal covariates in the pre-European model, ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while wealth rank represented by burial values (wealth-homophily) has a negative effect.
 This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
@@
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
-For the endogenous network effects, transitivity (gwesp) presents a significant positive effect, while popularity (gwdegree) demonstrates a negative effect.
+For the endogenous network effects, transitivity (gwesp) presents a significant positive effect, while centralization (gwdegree) demonstrates a negative effect.
 The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common.
-In contrast, the strong negative popularity shows there is a tendency toward decentralization that reflects most burials have a similar number of ties without any prominent ones.
+In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials have a similar number of ties without any prominent ones.
 This might imply that individuals have equal access to trade goods in terms of the flow of goods.
@@
                values_to = "posterior") %>%
+  mutate(parameter= case_when(parameter == "edges" ~ "Density",
+                              parameter == "nodematch.age" ~ "Age-homophily",
+                              parameter == "nodematch.sex" ~ "Sex-homophily",
+                              parameter == "nodematch.ritual" ~ "Ritual-homophily",
+                              parameter == "nodematch.value" ~ "Wealth-homophily",
+                              parameter == "gwesp" ~ "tansitivity",
+                              parameter == "gwdeg" ~ "centralization",
+                              parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
   mutate(parameter = factor(parameter,
-                            levels=c("edges", "nodematch.age", 
-                                     "nodematch.sex", "nodematch.ritual",
-                                     "nodematch.value", "gwesp",	
-                                     "gwdeg", "dyadcov.distance"))) %>% 
+                            levels=c("Density", "Age-homophily", 
+                                     "Sex-homophily", "Ritual-homophily",
+                                     "Wealth-homophily", "tansitivity",	
+                                     "centralization", "Physical distance"))) %>% 
   group_by(parameter, phase) %>% 
@@
 Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
-In contrast, the popularity (gwdegree) shows a significantly higher positive effect than the effect of the pre-European network.
+In contrast, the centralization (gwdegree) shows a significantly higher positive effect than the effect of the pre-European network.
 This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
 This implies that the presence of better access to trade goods and the behavior of wealth accumulation or display in burial events.
-In general, the post-European network model has a smaller transitivity effect and a positive popularity effect than the pre-European network model.
+In general, the post-European network model has a smaller transitivity effect and a positive centralization effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
@@
 Networks with nonoverlapped intervals means there is a significant difference.
-We computed and compared endogenous network statistics, including density, popularity, and transitivity for our two networks.
-Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach.
+We computed and compared endogenous network statistics, including density, centralization, and transitivity for our two networks.
+Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network centralization between the two networks, which is consistent with our finding of negative centralization in the pre-European period and positive centralization in the post-European period using the Bayesian approach.
 For density and transitivity, the 95% confidence intervals from the two networks overlapped with each other that indicates no significant difference.
@@
 
-The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and popularity (p = $`r pval_star`$).
+The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and centralization (p = $`r pval_star`$).
 Moreover, we explored the sample size effect by removing nodes at certain percent (5-40%) for both networks and compared their confidence intervals.
 The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
-In general, those statistical methods support the presence of anthropologically meaningful differences for density, transitivity, and popularity (i.e.
+In general, those statistical methods support the presence of anthropologically meaningful differences for density, transitivity, and centralization (i.e.
 centrality) in the pre-and post-European networks.

### 2021-02-12 analysis/paper/supplementary-material.Rmd
edit the whole paper and change variable names

@@
 -  further review of network analysis and archaeological applications
--  further definition of our approach to tie formation with the beads from Kiwulan burials
+-  further definition of our approach to tie formation and node attributes from Kiwulan burials
 -  further details on methods of modeling fitting, and
@@
 
-## Tie formation with the beads from Kiwulan burials {.unnumbered}
+## Tie formation and node attributes from Kiwulan burials {.unnumbered}
 
@@
 If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+For burial attributes, we assigned burials to different groups according to their age, sex, burial value and the presence/absence of ritual pottery.
+We have three groups for the age attribute, 0-12, 12-20, and above 20; two groups for the sex attribute, male and female, and four groups for the burial value that are specified in the main paper document.
 

### 2021-02-18 analysis/paper/paper.Rmd
edit the main text in paper rmd

@@
 By comparing hypothesized networks with the observed network, network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
-Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Ghafouri2020].
+Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Handcock2008; @Ghafouri2020].
 Since relationships in a network emerging relative to their position in a network can be hypothesized, ERGMs are promising to evaluate dynamic social processes behind observed archaeological networks [@Brughmans2014].
@@
 The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000], while corporate-based societies stress communal ritual elements, shared power, and wealth differentiation, if any, would be associated with corporate groups, such as age groups [@Siegel1999].
-Using the theoretical framework of corporate-network strategies, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
+Using the theoretical framework of corporate-network strategies, or horizontal hierarchy, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
 We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
@@
 A total of 90 burials were unearthed from adjacent squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
+The preservation of human remains at Kiwulan was generally poor, so we have incomplete age and sex data (details in Supplementary Online Materials) and lack health status data.
 Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses.
@@
 We also excluded `r disturb` burials that were heavily disturbed by modern construction because we cannot determine their chronology.
+
 Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf.
@@
 <!-- our expectation based on hypothesis --> 
-Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to an increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
+Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
 This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
@@
 presence or absence).
-Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index.
+Node attributes here include osteologically determined age and sex, cultural data, such as ritual pottery, and a burial value index.
 Ritual pottery was identified as locally made ceramics placed above an individual.
@@
 We calculated our burial value by summing values of all types of prestige goods in each burial context.
-The value of prestige good is calculated by taking the total number of burials in Kiwulan, and then dividing this by the number of burials with each type of prestige good [@Jorgensen1991].
-We then assigned the burials into three ranks according to their burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (Details about other node attributes in Supplementary Online Materials).
+For example, for gold-foil beads, we take the total number of burials (90) and divide by number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96.
+We repeat this for each type of prestige item.
+For each burial we sum these values of each type of prestige item to get a burial value [@Jorgensen1991].
+
+We then assigned the burials into three ranks according to breaks in the distribution of burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (details in Supplementary Online Materials).
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 knitr::kable(ergm_terms_selected,
-             caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.")
+             caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.
+ERGM terminology is from the R package statnet (Handcock, 2008; Morris, 2008)")
 ```
@@
 stronger correlations for shorter distances.
-To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and centralization (gwedegree).
+To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity and centralization.
 We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-A significant effect is determined by consistent positive/negative estimates in a 95% confidence interval.
-For nodal covariates in the pre-European model, ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while wealth rank represented by burial values (wealth-homophily) has a negative effect.
-This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
-Despite positive mean for some covariates, such as age (age-homophily) and sex (sex-homophily), they do not show a significant tendency due to value of zero in the confidence intervals.
+For nodal covariates in the pre-European model, the ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
+On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect.
+This demonstrates that burials in the pre-European period with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
+Despite positive mean values for some covariates, such as age and sex, their confidence intervals include zero, so they do not show a significant influence.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
-For the endogenous network effects, transitivity (gwesp) presents a significant positive effect, while centralization (gwdegree) demonstrates a negative effect.
+For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect.
 The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common.
-In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials have a similar number of ties without any prominent ones.
-This might imply that individuals have equal access to trade goods in terms of the flow of goods.
+In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties, without any prominent burials.
+This might imply that individuals have equal access to the flow of trade goods.
 
@@
                               parameter == "nodematch.value" ~ "Wealth-homophily",
-                              parameter == "gwesp" ~ "tansitivity",
+                              parameter == "gwesp" ~ "transitivity",
                               parameter == "gwdeg" ~ "centralization",
@@
                                      "Sex-homophily", "Ritual-homophily",
-                                     "Wealth-homophily", "tansitivity",	
+                                     "Wealth-homophily", "transitivity",	
                                      "centralization", "Physical distance"))) %>% 
@@
   set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models.
-Confidence intervals that do not include zero idicate significant effects of parameters (Caimo, 2017).")
+Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017).
+Posterior means are generally close to posterior medians.")
 ft
@@
 The pre-European group presents remarkable larger values for the gwesp, but lower values for the edges and gwdegree parameters compared to the post-European group.
-The distance parameters overlap between two groups."}
+The distance parameters overlap between two groups.
+Both posterior estimates present generally symmetric distributions."}
 knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
@@
 <!--posterior stats for for post-E and their comparison-->
-For the network model after the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects, while ritual and sex variables have negative effects.
-This indicates the burials in the same wealth level tend to have relations.
-On the contrary, the same sex burials and burial with ritual pottery in common tend not to have relations.
-There are no significant effects for variables of age and physical distance.
-Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
-In contrast, the centralization (gwdegree) shows a significantly higher positive effect than the effect of the pre-European network.
+After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects, while the ritual and sex variables have negative effects.
+This indicates that burials in the same wealth level tend to be more connected with each other.
+On the contrary, the same sex burials, and burials with ritual pottery in common tend, not to have relations.
+In the post-European burials there are no significant effects for variables of age and physical distance.
+Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network.
+In contrast, centralization has a significantly higher positive effect than the pre-European network.
 This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
-This implies that the presence of better access to trade goods and the behavior of wealth accumulation or display in burial events.
-In general, the post-European network model has a smaller transitivity effect and a positive centralization effect than the pre-European network model.
+This implies more access to trade goods and the increased wealth accumulation and display behaviors in burial events.
+In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
-In addition, both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
 
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results of Bayesian ERGMs.
-Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) that is useful for quantifying standard errors and estimating sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
-This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population.
-Networks with nonoverlapped intervals means there is a significant difference.
-We computed and compared endogenous network statistics, including density, centralization, and transitivity for our two networks.
-Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network centralization between the two networks, which is consistent with our finding of negative centralization in the pre-European period and positive centralization in the post-European period using the Bayesian approach.
-For density and transitivity, the 95% confidence intervals from the two networks overlapped with each other that indicates no significant difference.
-This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with some degree of difference.
+To understand the robustness of comparison between two networks, we used a vertex bootstrap technique to cross-validate the results of our Bayesian ERGMs.
+The vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
+node) to quantify standard errors and estimate sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
+This enables the evaluation of uncertainty for networks and tests the difference between multiple networks of different sizes by examining their confidence intervals for the network population.
+We used the vertex bootstrap to compute endogenous network statistics, including density, centralization, and transitivity for our two networks.
+We explored the sample size effect by removing nodes at certain percentages (5-40%) for both networks and comparing their confidence intervals.
+The results demonstrate a consistent difference between the two networks up to 30% node removals (Figure \@ref(fig:figure-bootstrap-CI)).
+This indicates that the network variables under investigation are robust to the different network size of our sample.
+Figure \@ref(fig:figure-bootstrap-CI also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
+Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
 
-```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, the top one with 0 nodes removal, and networks with different numbers of node removal.
-The number of bootstrap samples is 1000."}
+```{r figure-bootstrap-CI, fig.cap = "Results of the vertex bootstrap analysis by three network variables, showing the 95% bootstrap confidence intervals for the bootstrapped (n = 1000) networks, for each category of node removal.
+The 0% node removal category is the original data for the two phases.
+Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variables."}
 knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
 ```
-
-```{r bootstrap-t-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/008-bootstrap-t-test.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "008-bootstrap-t-test.R")))
-pval_den <- signif(density_b_test$p.value, 4)
-pval_trans <- signif(trans_b_test$p.value, 4)
-pval_star <- signif(star_b_test$p.value, 4)
-```
-
-The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and centralization (p = $`r pval_star`$).
-Moreover, we explored the sample size effect by removing nodes at certain percent (5-40%) for both networks and compared their confidence intervals.
-The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
-In general, those statistical methods support the presence of anthropologically meaningful differences for density, transitivity, and centralization (i.e.
-centrality) in the pre-and post-European networks.
-The procedure of cross-validation ensures our results can provide robust comparison between two networks without sample size effects.
 
@@
 <!--assessment of MCMC analysis and BGOF--> 
+Our computational models reproduce networks that resemble the structural features of our observed networks.
 Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
-the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
-For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
-This suggests that our models can reproduce networks that resemble the structural features of our observed networks.
+the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
+For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
 We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
 Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network.
-In general, this demonstrates a better fit for the pre-European network.
-However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.
+In general, this demonstrates a slightly better fit between model and data for the pre-European network.
 
-```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post- European models."}
+```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post-European models."}
 knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
@@
 This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
-These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
-Compared to vertical hierarchies that stress stratified social characteristics, the corporate-network strategies may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan [cf.
-@Ueda2016].
-A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
-In contrast, a network-based society, as suggested the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
-It should be noted that the corporate-network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
+A corporate-based society, as we interpret the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Drennan2010; @Feinman2000; @Feinman2000political].
+In contrast, a network-based society, as we interpret the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
+It should be noted that the corporate-network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages [@Feinman2000].
 
 <!--archaeological implication based on the results-->
-The covariants for networks provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
-In other words, we can know what kinds of socio-cultural factors associated with the acquisition and distribution of foreign prestige goods.
-Our covariants were age, sex, ritual pottery, and wealth ranks calculated based on the value of grave goods.
-The results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
-Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied that suggest they are not associated with the network pattern we observed.
-Those results illustrate that different social and economic mechanisms determined relations between burials, or the interconnections of prestige goods, for the two periods.
-Before European arrival, ritual behaviors seem to have been important factors that structure the interconnections of prestige goods in the burial context.
-This might imply that ritual could be a status indicator when the foreign goods were first introduced to Indigenous societies.
-After European presence was established, status indicators shifted from ritual to wealth differences with the least sex status discrimination.
-Moreover, geographical proximity, implying some kinship association, is not a factor for sharing similar prestige items between burials for both phases.
+Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
+Our results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
+We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
+Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts.
+This might imply that ritual could be a status indicator at the time before foreign goods were first introduced to Indigenous societies.
+After the European presence was established, status indicators shifted from ritual to wealth differences.
+Moreover, geographical proximity, implying some kinship association, has not effect on sharing similar prestige items between burials for both phases.
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
 Demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows a negative effect for the post-European network.
-This may also support a shift to more wealth and individual-based status competition without sex differences in the later phase.
+This may also support a shift to more wealth and individual-based status competition without sex differences, in the later phase.
 But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.
@@
 <!-- other evidence -->
-We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks according to their status [eg.
+We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
 @Coward2013, pp.
 252].
-We argue that the social relation at Kiwulan changed because of the contact with European colonial power and associated foreign trade.
-This argument is supported by the changes in burial network structures through time and their associated underlying mechanisms.
-It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
+We argue that the social relations at Kiwulan changed because of contact with European colonial powers and associated foreign trade.
+This argument is supported by changes in burial network structures through time.
+It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person’s status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
 However, this issue can be reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
@@
 This is consistent with the result of a more centralized burial network after the presence of Europeans, that supports a connection between a burial and the social role of the living person.
-Also, the burial goods used for making ties were treated as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
-
-<!--anthopological implication at Kiwulan and limitation-->
-At Kiwulan, changes in interburial networks may result from differential access to exotic trade enabled by European contact, gradually leading to increased social differentiation.
-The image of European power could be embedded into prestige objects pre-existing in local Indigenous culture where their values were amplified with concepts of wealth or power, since their values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
-This may lead to ambitious individuals competing with each other for traditional goods and accumulating wealth.
-It also demonstrates that local Indigenous societies could actively incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
-We recognize that the meaning of the burial goods could be refined with more ethnographic data to support our interpretations.
-Human skeletal remains can provide information about health for interpreting social status; however, the preservation of human remains at Kiwulan was generally not good to yield useful data.
-By focusing on burial goods, we seek to understand the social changes at Kiwulan in relation to the contact of Europeans.
-Despite fragmentation in our archaeological data, such as missing age and sex attributes for some burials at Kiwulan, we can still capture the general patterns of network structures related to social implication.
+Also, ethnographic records support our interpretation of burial goods as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], so the interrelations represented by the flow of these goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
 
@@
 
+<!-- technical and methodological implication-->
+We presented a novel approach for studying burials to interpret social structures using ERGMs within a Bayesian framework.
+This case study demonstrates the methodological benefits of Bayesian inference on ERGMs to inform and enhance studies of relational data in archaeology.
+A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
+Another valuable feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological models, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
+Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
+This enables insights into the dynamic processes of relationship formation and the underlying factors of historical trajectories of socio-cultural phenomena.
+
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-In this paper, we presented a novel approach for studying burials to interpret social structures using ERGMs within a Bayesian framework.
 We examined social changes in a European pericolonial context in northeastern Taiwan.
-We tested the hypotheses of changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
+We tested the hypotheses of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
 The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
@@
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-The changes in formation mechanisms for networks from ritual practice connections to wealth level connections could suggest increased behavior of wealth accumulation stimulated by the European presence and associated long-distance trade network.
+We interpret the changes in formation mechanisms of networks, from ritual practice to wealth levels, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
 This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
-Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
+Using burial data with historical documents, we are to better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
 @Acabado2017; @Trabert2017].
-
-<!-- technical and methodological implication-->
-Furthermore, this case study demonstrates the methodological improvement that Bayesian inference on ERGMs provides to inform and enhance studies of relational data in archaeology.
-A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
-This enables insights into the dynamic processes of relationship formation and the underlying factors of historical trajectories of socio-cultural phenomena.
-
-<!-- stopped editing here -->
 

### 2021-02-18 analysis/paper/paper.Rmd
update figure labels and result section after rerunning models

@@
 Increased social inequality associated with the use of foreign prestige goods is recognized as network strategies for gaining power [@Blanton1996; @Feinman2000].
-Network-based societies are identified in some places in Asia with involvment in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+Network-based societies are identified in some places in Asia with involvement in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
@@
 For nodal covariates in the pre-European model, the ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
-On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect.
-This demonstrates that burials in the pre-European period with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
-Despite positive mean values for some covariates, such as age and sex, their confidence intervals include zero, so they do not show a significant influence.
+On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero.
+This demonstrates that burials in the pre-European period with ritual pottery tend to form relations.
+Despite positive mean values for some covariates, such as age and sex, they do not show a significant influence.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
@@
                               parameter == "nodematch.value" ~ "Wealth-homophily",
-                              parameter == "gwesp" ~ "transitivity",
-                              parameter == "gwdeg" ~ "centralization",
+                              parameter == "gwesp" ~ "Transitivity",
+                              parameter == "gwdeg" ~ "Centralization",
                               parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
@@
                                      "Sex-homophily", "Ritual-homophily",
-                                     "Wealth-homophily", "transitivity",	
-                                     "centralization", "Physical distance"))) %>% 
+                                     "Wealth-homophily", "Transitivity",	
+                                     "Centralization", "Physical distance"))) %>% 
   group_by(parameter, phase) %>% 
@@
 
-```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, gwesp, gwdegree, and distance-dyadcov by phases.
-The pre-European group presents remarkable larger values for the gwesp, but lower values for the edges and gwdegree parameters compared to the post-European group.
+```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases.
+The pre-European group presents remarkable larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group.
 The distance parameters overlap between two groups.
@@
 <!--posterior stats for for post-E and their comparison-->
-After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects, while the ritual and sex variables have negative effects.
-This indicates that burials in the same wealth level tend to be more connected with each other.
-On the contrary, the same sex burials, and burials with ritual pottery in common tend, not to have relations.
-In the post-European burials there are no significant effects for variables of age and physical distance.
+After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects.
+This indicates burials in the same wealth level tend to be more connected with each other.
+Despite negative effects for ritual and sex variables, they are not significant.
+Also, there are no significant effects for variables of age and physical distance.
 Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network.
@@
 This indicates that the network variables under investigation are robust to the different network size of our sample.
-Figure \@ref(fig:figure-bootstrap-CI also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
+Figure \@ref(fig:figure-bootstrap-CI) also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
 Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
@@
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
-Demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows a negative effect for the post-European network.
-This may also support a shift to more wealth and individual-based status competition without sex differences, in the later phase.
+Demographic information, such as age and sex, do not have any effects on network formation for both networks.
 But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.

### 2021-02-18 analysis/paper/004-burials-post-network-modelling.R

@@ -194,24 +194,24 @@ sim_esp_post <-
 
 # make dataframes for distribution stats
 posterior_distribution_post <-
-  data.frame("moments" = c("Mean", "Variance", "Skewness"),
-             "Observed.degree" = round(c(obs_degree_post$mean,
+  data.frame("moments" = c("mean", "variance", "skewness"),
+             "observed.degree" = round(c(obs_degree_post$mean,
                                   obs_degree_post$sd,
                                   obs_degree_post$skew), 2),
-             "Modelled.degree" = round(c(mean(sim_degree_post$mean),
+             "modelled.degree" = round(c(mean(sim_degree_post$mean),
                                mean(sim_degree_post$sd),
                                mean(sim_degree_post$skew)),2),
-             "Observed.distance" = round(c(obs_dist_post$mean,
+             "observed.distance" = round(c(obs_dist_post$mean,
                                            obs_dist_post$sd,
                                            obs_dist_post$skew), 2),
-             "Modelled.distance" = round(c(mean(sim_dist_post$mean),
+             "modelled.distance" = round(c(mean(sim_dist_post$mean),
                                         mean(sim_dist_post$sd),
                                         mean(sim_dist_post$skew)),2),
-             "Observed.esp" = round(c(obs_esp_post$mean,
+             "observed.esp" = round(c(obs_esp_post$mean,
                                       obs_esp_post$sd,
                                       obs_esp_post$skew), 2),
-             "Modelled.esp" = round(c(mean(sim_esp_post$mean),
+             "modelled.esp" = round(c(mean(sim_esp_post$mean),
                                    mean(sim_esp_post$sd),
                                    mean(sim_esp_post$skew)),2),
-             "Phase" = "post-European")
+             "phase" = "post-European")
 

### 2021-02-18 analysis/paper/006-bergm-distribution-stats.R

@@ -63,9 +63,19 @@ histogram <-
 bergm_two_phases_longer <-
   bergm_two_phases %>%
   select(edges, gwesp, gwdeg, dyadcov.distance, phase) %>%
+  mutate(phase = case_when(
+    phase == "pre" ~ "pre-European",
+    phase == "post" ~ "post-European",
+    TRUE ~ "")) %>%
+  rename(transitivity = gwesp,
+         centralization = gwdeg,
+         "physical distance" = dyadcov.distance) %>%
   pivot_longer(!phase, names_to = "parameter", values_to = "posterior") %>%
   mutate(parameter = factor(parameter,
-                            levels=c("edges","gwesp","gwdeg","dyadcov.distance")))
+                            levels=c("edges",
+                                     "transitivity",
+                                     "centralization",
+                                     "physical distance")))
 
 density_posterior_facet <-
   ggplot(bergm_two_phases_longer,
@@ -74,6 +84,7 @@ density_posterior_facet <-
   facet_wrap(~parameter, scales = "free") +
   theme_minimal()
 
+# Figure 4
 ggsave(here::here("analysis", "figures", "006-posterior-distribution.png"),
        w = 8, h = 5)
 
@@ -84,16 +95,16 @@ distribution_two_phases <-
 
 distribution_two_phases_longer <-
   distribution_two_phases %>%
-  pivot_longer(cols = starts_with(c("Observed", "Modelled")),
+  pivot_longer(cols = starts_with(c("observed", "modelled")),
                names_to = "parameter",
                values_to = "value") %>%
   separate(parameter, c("data", "parameter"))
 
 distribution_two_phases_diff <-
   distribution_two_phases %>%
-  mutate(degree = abs(Observed.degree-Modelled.degree),
-         distance = abs(Observed.distance-Modelled.distance),
-         esp = abs(Observed.esp-Modelled.esp)) %>%
+  mutate(degree = abs(observed.degree-modelled.degree),
+         distance = abs(observed.distance-modelled.distance),
+         esp = abs(observed.esp-modelled.esp)) %>%
   pivot_longer(cols = c("degree", "distance", "esp"),
                names_to = "parameter",
                values_to = "difference")
@@ -101,11 +112,12 @@ distribution_two_phases_diff <-
 # https://github.com/nevrome/ggpointgrid
 ggplot(distribution_two_phases_longer,
        aes(parameter, value)) +
-  ggpointgrid::geom_pointgrid(aes(color = Phase,
[…]

### 2021-02-18 analysis/paper/007-bootstrap-CI.R

@@ -252,7 +252,15 @@ CIs_two_nets_tidy <-
   mutate(variable = case_when(
     variable == "1" ~ "density",
     variable == "2" ~ "transitivity",
-    variable == "3" ~ "popularity",
+    variable == "3" ~ "centralization",
+    TRUE ~ "")) %>%
+  mutate(variable = factor(variable,
+                            levels=c("density",
+                                     "transitivity",
+                                     "centralization"))) %>%
+  mutate(phase = case_when(
+    phase == "pre" ~ "pre-European",
+    phase == "post" ~ "post-European",
     TRUE ~ "")) %>%
   mutate(removal = as.factor(removal))
 
@@ -277,5 +285,6 @@ CIs_two_nets_tidy %>%
   facet_wrap(~variable,
              scales = "free_x")
 
+# Figure 5
 ggsave(here::here("analysis", "figures", "007-bootstrap-CI.png"),
        w = 8, h = 5)

### 2021-02-19 analysis/paper/paper.Rmd
fix label names in plots

@@
 diff_pre_post <- post_E - pre_E
+
+# missing values
+age_miss <- sum(NAs_age$n)
+sex_miss <- sum(NAs_sex$n)
 ```
@@
 ft <- flextable(posterior_esti_two_phases) %>% 
+  colformat_double(digits = 2) %>% 
   set_header_labels(mean_pre = "pre-European", median_pre = "pre-European", 
@@
   align(i = 1, align = "center", part = "header") %>% 
-  width(j = ~ parameter, width = 1.3) %>% 
+  width(j = ~ parameter, width = 1.35) %>% 
   width(j = 2:9, width = 0.65) %>% 
@@
 Demographic information, such as age and sex, do not have any effects on network formation for both networks.
-But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.
+But we also note that no effects could be due to the unavailability of age and sex values for many burials (missing age = `r age_miss` and missing sex = `r sex_miss`) in our sample.
 

### 2021-02-19 analysis/paper/001-data-tidy.R

@@ -90,3 +90,15 @@ burial_three_period_age_number <-
   burial_three_period_age_tidy %>%
   count(Phase)
 
+# missing values
+NAs_age <-
+  burial_three_period_age_tidy%>%
+  filter(Age_scale == "NA") %>%
+  filter(Phase != "disturbed") %>%
+  count(Phase)
+
+NAs_sex <-
+  burial_three_period_age_tidy%>%
+  filter(gender == "NA") %>%
+  filter(Phase != "chi") %>%
+  count(Phase)

### 2021-02-19 analysis/paper/006-bergm-distribution-stats.R

@@ -115,6 +115,9 @@ ggplot(distribution_two_phases_longer,
   ggpointgrid::geom_pointgrid(aes(color = phase,
                                   shape = data),
                               size = 3) +
+  scale_x_discrete(labels=c("degree",
+                            "minimum\ngeodetic\ndistance",
+                            "edge-wise\nshared\npartners")) +
   facet_wrap(~moments, scales = "free") +
   theme_minimal()
 

### 2021-02-19 analysis/paper/supplementary-material.Rmd
edit the main text in paper rmd
add a table to SI

@@
 For burial attributes, we assigned burials to different groups according to their age, sex, burial value and the presence/absence of ritual pottery.
-We have three groups for the age attribute, 0-12, 12-20, and above 20; two groups for the sex attribute, male and female, and four groups for the burial value that are specified in the main paper document.
+We have three groups for the age attribute, 0-12, 12-20, and above 20; two groups for the sex attribute, male and female; and four groups for the burial value that are specified in the main paper document.
+Table \@ref(tab:age-sex-table) shows the number of burials, of which sex and age can be determined.
 
@@
 # combine ridge plots 1 & 4 in the 000-prep-data file
-library(tidyverse)
 library(ggridges)
@@
        #w = 6, h = 4)
+```
+
+```{r age-sex-table}
+library(tidyverse)
+library(flextable)
+age_sex_by_phase <-
+  burial_three_period_age_tidy %>% 
+  select(Age_scale, gender, Phase) %>% 
+  group_by(Phase, gender, Age_scale) %>% 
+  filter(!Phase %in% c("disturbed", "chi")) %>% 
+  count() %>% 
+  pivot_wider(names_from = c(Phase, gender), 
+              values_from = n) %>% 
+  mutate(Age_scale = factor(Age_scale,
+                            levels = c("0-12", "12-20", "+20", "NA"))) %>% 
+  arrange(Age_scale) %>% 
+  rename(age = Age_scale) %>% 
+  select(age, pre_male, pre_female, pre_NA, post_male, post_female, post_NA)
+
+ft_age_sex <- flextable(age_sex_by_phase) %>% 
+  set_header_labels(pre_male = "pre-European", pre_female = "pre-European",
+                    pre_NA = "pre-European", post_male = "post-European", 
+                    post_female ="post-European", post_NA = "post-European") %>%
+  merge_at(i = 1, j = 2:4, part = "header") %>% 
+  merge_at(i = 1, j = 5:7, part = "header") %>% 
+  add_header_row(values = c("","male", "female", "NA", 
+                            "male", "female", "NA"), top = FALSE) %>% 
+  align(j = 2:7, align = "right", part = "header") %>% 
+  align(i = 1, align = "center", part = "header") %>% 
+  set_caption("The number of burials with identification of age and sex at Kiwulan. The NA means no sufficient skeletal remains for determination of age or sex")
+ft_age_sex
 ```

### 2021-02-19 analysis/paper/001-data-tidy.R

@@ -24,7 +24,7 @@ burial_three_period_age_tidy <-
     TRUE ~ "")) %>% # the classification is based on the result of histogram
   mutate(Age_scale = case_when(
     `Age` %in% c("1","2") ~ "0-12",
-    `Age` == "3" ~ "12~20",
+    `Age` == "3" ~ "12-20",
     `Age` %in% c("4","5","6","7","8") ~ "+20",
     TRUE ~ "NA")) %>%
   mutate(gender = case_when(

### 2021-02-19 analysis/paper/paper.Rmd
fix all typos

--- 2021-02-18 analysis/paper/paper.Rmd
@@
 diff_pre_post <- post_E - pre_E
+pre_and_post <- pre_E + post_E
+
+# missing values
+age_miss <- sum(NAs_age$n)
+sex_miss <- sum(NAs_sex$n)
 ```
@@
 ## Hypothesis and construction of networks {-}
+
 <!-- our expectation based on hypothesis --> 
-Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
+Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
 This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, the ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
+For nodal covariates in the pre-European model, the ritual element represented by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
 On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero.
@@
                values_to = "posterior") %>%
-  mutate(parameter= case_when(parameter == "edges" ~ "Density",
-                              parameter == "nodematch.age" ~ "Age-homophily",
-                              parameter == "nodematch.sex" ~ "Sex-homophily",
-                              parameter == "nodematch.ritual" ~ "Ritual-homophily",
-                              parameter == "nodematch.value" ~ "Wealth-homophily",
-                              parameter == "gwesp" ~ "Transitivity",
-                              parameter == "gwdeg" ~ "Centralization",
-                              parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
+  mutate(parameter= case_when(parameter == "edges" ~ "density",
+                              parameter == "nodematch.age" ~ "age-homophily",
+                              parameter == "nodematch.sex" ~ "sex-homophily",
+                              parameter == "nodematch.ritual" ~ "ritual-homophily",
+                              parameter == "nodematch.value" ~ "wealth-homophily",
+                              parameter == "gwesp" ~ "transitivity",
+                              parameter == "gwdeg" ~ "centralization",
+                              parameter == "dyadcov.distance" ~ "physical distance")) %>% 
   mutate(parameter = factor(parameter,
-                            levels=c("Density", "Age-homophily", 
-                                     "Sex-homophily", "Ritual-homophily",
-                                     "Wealth-homophily", "Transitivity",	
-                                     "Centralization", "Physical distance"))) %>% 
+                            levels=c("density", "age-homophily", 
+                                     "sex-homophily", "ritual-homophily",
+                                     "wealth-homophily", "transitivity",	
+                                     "centralization", "physical distance"))) %>% 
   group_by(parameter, phase) %>% 
@@
   align(i = 1, align = "center", part = "header") %>% 
-  width(j = ~ parameter, width = 1.3) %>% 
-  width(j = 2:9, width = 0.65) %>% 
+  width(j = ~ parameter, width = 1.36) %>% 
+  width(j = 2:9, width = 0.67) %>% 
   set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models.
@@
 ```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases.
-The pre-European group presents remarkable larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group.
+The pre-European group presents remarkably larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group.
 The distance parameters overlap between two groups.
@@
 This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
-This implies more access to trade goods and the increased wealth accumulation and display behaviors in burial events.
+This implies more access to trade goods, increased wealth accumulation and display behaviors in burial events.
 In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model.
@@
 The 0% node removal category is the original data for the two phases.
-Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variables."}
+Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variable."}
 knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
@@
 Our computational models reproduce networks that resemble the structural features of our observed networks.
-Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+Bayesian goodness of fit diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases); details in Supplementary Online Materials) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
 the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
@@
 Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts.
-This might imply that ritual could be a status indicator at the time before foreign goods were first introduced to Indigenous societies.
+This might imply that ritual could be a status indicator at the time before many foreign goods were introduced to Indigenous societies.
 After the European presence was established, status indicators shifted from ritual to wealth differences.
-Moreover, geographical proximity, implying some kinship association, has not effect on sharing similar prestige items between burials for both phases.
+Moreover, geographical proximity, implying some kinship association, has no effect on sharing similar prestige items between burials for both phases.
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
 Demographic information, such as age and sex, do not have any effects on network formation for both networks.
-But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.
+But we also note that no effects could be due to the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
 
@@
 This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
-Using burial data with historical documents, we are to better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
+Using burial data with historical documents, we can better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
 @Acabado2017; @Trabert2017].

### 2021-02-19 analysis/paper/supplementary-material.Rmd
fix all typos

--- 2021-02-12 analysis/paper/supplementary-material.Rmd
@@
 
-# Kiwulan in the 17th northstern Taiwan in a pericolonial context {-}
+# Kiwulan in the 17th northeastern Taiwan in a pericolonial context {-}
 
@@
 The Dutch asked local settlements to pay deer hides or rice as annual tributes but negotiating a delayed payment is allowed if it is needed [@Borao2009; @Kang2016].
-In the annual event hold by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan stick or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.
+In the annual event held by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan sticks or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.
 The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
@@
 Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
-Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015].
+Another example is the assessment of hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Gjesfjeld2015].
 
@@
 
+```{r age-sex-table}
+library(flextable)
+age_sex_by_phase <-
+  burial_three_period_age_tidy %>% 
+  select(Age_scale, gender, Phase) %>% 
+  group_by(Phase, gender, Age_scale) %>% 
+  filter(!Phase %in% c("disturbed", "chi")) %>% 
+  count() %>% 
+  pivot_wider(names_from = c(Phase, gender), 
+              values_from = n) %>% 
+  mutate(Age_scale = factor(Age_scale,
+                            levels = c("0-12", "12-20", "+20", "NA"))) %>% 
+  arrange(Age_scale) %>% 
+  rename(age = Age_scale) %>% 
+  select(age, pre_male, pre_female, pre_NA, post_male, post_female, post_NA)
+
+ft_age_sex <- flextable(age_sex_by_phase) %>% 
+  set_header_labels(pre_male = "pre-European", pre_female = "pre-European",
+                    pre_NA = "pre-European", post_male = "post-European", 
+                    post_female ="post-European", post_NA = "post-European") %>%
+  merge_at(i = 1, j = 2:4, part = "header") %>% 
+  merge_at(i = 1, j = 5:7, part = "header") %>% 
+  add_header_row(values = c("","male", "female", "NA", 
+                            "male", "female", "NA"), top = FALSE) %>% 
+  align(j = 2:7, align = "right", part = "header") %>% 
+  align(i = 1, align = "center", part = "header") %>% 
+  set_caption("The number of burials with identification of age and sex at Kiwulan. The NA means no sufficient skeletal remains for determination of age or sex")
+ft_age_sex
+```
+
 ```{r figure-raincloud-beads, fig.cap = "The distributions of different types of trade beads across burials analyzed in this paper."}
@@
 MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
-We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on he approximate exchange algorithm [@Caimo2011].
+We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on the approximate exchange algorithm [@Caimo2011].
 This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.

### 2021-02-25 analysis/paper/paper.Rmd
why do I still see t-test in paper.docx?

--- 2021-02-18 analysis/paper/paper.Rmd
@@
 
-# Introduction {-}
+# Introduction {.unnumbered}
 
 <!---network analysis approaches and basic concepts-->
+
 Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing rational data [@Brandes2013; @Mills2017; @Peeples2019].
@@
 By comparing hypothesized networks with the observed network, network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
-Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Handcock2008; @Ghafouri2020].
+Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Ghafouri2020].
 Since relationships in a network emerging relative to their position in a network can be hypothesized, ERGMs are promising to evaluate dynamic social processes behind observed archaeological networks [@Brughmans2014].
@@
 <!--study social changes in pericolonial context using burial data as a proxy-->
+
 In this paper, we go beyond ERGMs with a novel Bayesian approach that can alleviate the computational issues in ERGMs to better interpret the process of network formation for past phenomena.
@@
 <!--Our burial network and research implication-->
+
 Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
@@
 <!-- intro to networks: move to SI -->
+
 <!-- two types of network analysis: move to SI -->
 
-# Exponential random graph models in a Bayesian framework {-}
+# Exponential random graph models in a Bayesian framework {.unnumbered}
 
 <!-- ERGMs in archaeology-->
+
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
@@
 <!--limitations of ERGMs-->
+
 However, those studies also point out some limitations of ERGMs, such as sensitivity to missing data and less able to handle uncertainty.
@@
 A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with a larger set of networks [@Caimo2017bayesian].
-This is also termed “doubly intractable” since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
+This is also termed "doubly intractable" since both the likelihood normalizing constant and the marginal likelihood (the evidence of the posterior) are hard to derive [@Caimo2013; @Lyne2015].
 Model degeneracy is another issue where probability models tend to overestimate a small number of extreme graphs by assigning too much weight, e.g., in extreme cases such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
@@
 <!-- Bayesian inference for Exponential Random Graph Models -->
+
 Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
@@
 
-# Background {-}
+# Background {.unnumbered}
 
-## A Case Study from Northeastern Taiwan {-}
+## A Case Study from Northeastern Taiwan {.unnumbered}
 
 <!--archaeological specific case study and our model-->
+
 We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries.
@@
 Increased social inequality associated with the use of foreign prestige goods is recognized as network strategies for gaining power [@Blanton1996; @Feinman2000].
-Network-based societies are identified in some places in Asia with involvement in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+Network-based societies are identified in some places in Asia with involvment in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
@@
 <!--theoretical framework and research questions-->
+
 The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000], while corporate-based societies stress communal ritual elements, shared power, and wealth differentiation, if any, would be associated with corporate groups, such as age groups [@Siegel1999].
-Using the theoretical framework of corporate-network strategies, or horizontal hierarchy, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
+Using the theoretical framework of corporate-network strategies, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
 We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
@@
 
-## Materials {-}
+## Materials {.unnumbered}
 
@@
 <!-- Kiwulan burials and a review of current studies of social organization -->
+
 We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) [@Chen2007].
 A total of 90 burials were unearthed from adjacent squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-The preservation of human remains at Kiwulan was generally poor, so we have incomplete age and sex data (details in Supplementary Online Materials) and lack health status data.
-Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses.
+Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and *in-situ* wooden posts, suggesting a well-organized spatial arrangement of houses.
 Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences.
@@
 <!-- Our chronology and principles of network building -->
+
 To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
@@
 We also excluded `r disturb` burials that were heavily disturbed by modern construction because we cannot determine their chronology.
-
 Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf.
@@
 European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
-For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads’ high value in Indigenous culture [@LiandWu2006].
+For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads' high value in Indigenous culture [@LiandWu2006].
 In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as having higher status in their community [@LiandWu2006].
 
-# Methods {-}
+# Methods {.unnumbered}
 
@@
 
-## Hypothesis and construction of networks {-}
+## Hypothesis and construction of networks {.unnumbered}
+
 <!-- our expectation based on hypothesis -->
-Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
+
+Our pericolonial impact hypothesis is that differential access to foreign prestige goods after the European presence led to an increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
 This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
@@
 <!-- burial data and the definition of tie formation  -->
+
 Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
@@
 presence or absence).
-Node attributes here include osteologically determined age and sex, cultural data, such as ritual pottery, and a burial value index.
+Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index.
 Ritual pottery was identified as locally made ceramics placed above an individual.
@@
 We calculated our burial value by summing values of all types of prestige goods in each burial context.
-For example, for gold-foil beads, we take the total number of burials (90) and divide by number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96.
-We repeat this for each type of prestige item.
-For each burial we sum these values of each type of prestige item to get a burial value [@Jorgensen1991].
-
-We then assigned the burials into three ranks according to breaks in the distribution of burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (details in Supplementary Online Materials).
+The value of prestige good is calculated by taking the total number of burials in Kiwulan, and then dividing this by the number of burials with each type of prestige good [@Jorgensen1991].
+We then assigned the burials into three ranks according to their burial values, high (\>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (\<`r ave_value`), as an index of wealth (Details about other node attributes in Supplementary Online Materials).
 Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
@@
 
-## Model specification in a Bayesian framework {-}
+## Model specification in a Bayesian framework {.unnumbered}
 
 <!-- Network construction and specification -->
+
 We quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014].
@@
 knitr::kable(ergm_terms_selected,
-             caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.
-ERGM terminology is from the R package statnet (Handcock, 2008; Morris, 2008)")
+             caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.")
 ```
@@
 
-## Choice of prior values to evaluate our anthropological model {-}
+## Choice of prior values to evaluate our anthropological model {.unnumbered}
 
 <!-- add text about prior setting and anthropological assumptions of our prior values -->
+
 Normal distributions for the priors are typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
@@
 stronger correlations for shorter distances.
-To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity and centralization.
+To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity (gwesp) and centralization (gwedegree).
 We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
@@
 
-# Reproducibility and open source materials {-}
+# Reproducibility and open source materials {.unnumbered}
 
-The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/xxx/xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at <https://doi.org/xxx/xxx> to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
@@
 
-# Results {-}
+# Results {.unnumbered}
 
@@
 <!--posterior stats for pre-E-->
+
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, the ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
-On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero.
-This demonstrates that burials in the pre-European period with ritual pottery tend to form relations.
-Despite positive mean values for some covariates, such as age and sex, they do not show a significant influence.
+A significant effect is determined by consistent positive/negative estimates in a 95% confidence interval.
+For nodal covariates in the pre-European model, ritual element represent by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while wealth rank represented by burial values (wealth-homophily) has a negative effect.
+This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
+Despite positive mean for some covariates, such as age (age-homophily) and sex (sex-homophily), they do not show a significant tendency due to value of zero in the confidence intervals.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
-For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect.
+For the endogenous network effects, transitivity (gwesp) presents a significant positive effect, while centralization (gwdegree) demonstrates a negative effect.
 The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common.
-In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties, without any prominent burials.
-This might imply that individuals have equal access to the flow of trade goods.
+In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials have a similar number of ties without any prominent ones.
+This might imply that individuals have equal access to trade goods in terms of the flow of goods.
 
@@
                               parameter == "nodematch.value" ~ "Wealth-homophily",
-                              parameter == "gwesp" ~ "Transitivity",
-                              parameter == "gwdeg" ~ "Centralization",
+                              parameter == "gwesp" ~ "tansitivity",
+                              parameter == "gwdeg" ~ "centralization",
                               parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
@@
                                      "Sex-homophily", "Ritual-homophily",
-                                     "Wealth-homophily", "Transitivity",	
-                                     "Centralization", "Physical distance"))) %>% 
+                                     "Wealth-homophily", "tansitivity",	
+                                     "centralization", "Physical distance"))) %>% 
   group_by(parameter, phase) %>% 
@@
   set_caption("Estimated posterior means, medians, and 95% confidence intervals for each network parameter of two models.
-Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017).
-Posterior means are generally close to posterior medians.")
+Confidence intervals that do not include zero idicate significant effects of parameters (Caimo, 2017).")
 ft
@@
 
-```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases.
-The pre-European group presents remarkable larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group.
-The distance parameters overlap between two groups.
-Both posterior estimates present generally symmetric distributions."}
+```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, gwesp, gwdegree, and distance-dyadcov by phases.
+The pre-European group presents remarkable larger values for the gwesp, but lower values for the edges and gwdegree parameters compared to the post-European group.
+The distance parameters overlap between two groups."}
 knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
@@
 <!--posterior stats for for post-E and their comparison-->
-After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects.
-This indicates burials in the same wealth level tend to be more connected with each other.
-Despite negative effects for ritual and sex variables, they are not significant.
-Also, there are no significant effects for variables of age and physical distance.
-Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network.
-In contrast, centralization has a significantly higher positive effect than the pre-European network.
+
+For the network model after the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects, while ritual and sex variables have negative effects.
+This indicates the burials in the same wealth level tend to have relations.
+On the contrary, the same sex burials and burial with ritual pottery in common tend not to have relations.
+There are no significant effects for variables of age and physical distance.
+Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect but much lower than the effect of pre-European network.
+In contrast, the centralization (gwdegree) shows a significantly higher positive effect than the effect of the pre-European network.
 This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
-This implies more access to trade goods and the increased wealth accumulation and display behaviors in burial events.
-In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model.
+This implies that the presence of better access to trade goods and the behavior of wealth accumulation or display in burial events.
+In general, the post-European network model has a smaller transitivity effect and a positive centralization effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
+In addition, both posterior estimates present symmetric distributions, with the posterior means close to posterior medians (Table \@ref(tab:table-posterior-stats)).
 
@@
 <!-- comparison between two networks using bootstrap -->
+
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand the robustness of comparison between two networks, we used a vertex bootstrap technique to cross-validate the results of our Bayesian ERGMs.
-The vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) to quantify standard errors and estimate sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
-This enables the evaluation of uncertainty for networks and tests the difference between multiple networks of different sizes by examining their confidence intervals for the network population.
-We used the vertex bootstrap to compute endogenous network statistics, including density, centralization, and transitivity for our two networks.
-We explored the sample size effect by removing nodes at certain percentages (5-40%) for both networks and comparing their confidence intervals.
-The results demonstrate a consistent difference between the two networks up to 30% node removals (Figure \@ref(fig:figure-bootstrap-CI)).
-This indicates that the network variables under investigation are robust to the different network size of our sample.
-Figure \@ref(fig:figure-bootstrap-CI) also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
-Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
+To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results of Bayesian ERGMs.
+Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
+node) that is useful for quantifying standard errors and estimating sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
+This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population.
+Networks with nonoverlapped intervals means there is a significant difference.
+We computed and compared endogenous network statistics, including density, centralization, and transitivity for our two networks.
+Figure \@ref(fig:figure-bootstrap-CI) shows a significant difference in observed network centralization between the two networks, which is consistent with our finding of negative centralization in the pre-European period and positive centralization in the post-European period using the Bayesian approach.
+For density and transitivity, the 95% confidence intervals from the two networks overlapped with each other that indicates no significant difference.
+This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects with some degree of difference.
 
-```{r figure-bootstrap-CI, fig.cap = "Results of the vertex bootstrap analysis by three network variables, showing the 95% bootstrap confidence intervals for the bootstrapped (n = 1000) networks, for each category of node removal.
-The 0% node removal category is the original data for the two phases.
-Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variables."}
+```{r figure-bootstrap-CI, fig.cap = "The 95% bootstrap confidence intervals for the observed networks, the top one with 0 nodes removal, and networks with different numbers of node removal.
+The number of bootstrap samples is 1000."}
 knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
 ```
+
+```{r bootstrap-t-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/008-bootstrap-t-test.R"))}
+suppressWarnings(source(here::here("analysis", "paper", "008-bootstrap-t-test.R")))
+pval_den <- signif(density_b_test$p.value, 4)
+pval_trans <- signif(trans_b_test$p.value, 4)
+pval_star <- signif(star_b_test$p.value, 4)
+```
+
+The vertex bootstrap procedure with a t-test on the difference of network populations also shows that there are significant differences in network density(p = $`r pval_den`$), transitivity (p = $`r pval_trans`$), and centralization (p = $`r pval_star`$).
+Moreover, we explored the sample size effect by removing nodes at certain percent (5-40%) for both networks and compared their confidence intervals.
+The results (Figure \@ref(fig:figure-bootstrap-CI)) demonstrate a consistent difference between the two networks within 30 % node removals that suggest they are robust to the different sizes.
+In general, those statistical methods support the presence of anthropologically meaningful differences for density, transitivity, and centralization (i.e.
+centrality) in the pre-and post-European networks.
+The procedure of cross-validation ensures our results can provide robust comparison between two networks without sample size effects.
 
@@
 <!--assessment of MCMC analysis and BGOF-->
-Our computational models reproduce networks that resemble the structural features of our observed networks.
+
 Bayesian goodness of fit (GOF) diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
-the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
-For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
+the number of edges between node pairs in a shortest path [@Hunter2008]) and the degree distribution.
+For edgewise shared partner (gwesp) distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
+This suggests that our models can reproduce networks that resemble the structural features of our observed networks.
 We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.
 Figure \@ref(fig:figure-distribution-moments) shows that modeled values from the pre-European network are slightly closer to the observed values for the mean and the variance of distributions, compared to the values from the post-European network.
-In general, this demonstrates a slightly better fit between model and data for the pre-European network.
+In general, this demonstrates a better fit for the pre-European network.
+However, it should be noted that the differences between the statistics of two models are relatively small, which still allows direct comparison of two networks.
 
-```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post-European models."}
+```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post- European models."}
 knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
@@
 
-# Discussion {-}
+# Discussion {.unnumbered}
 
 <!-- theoretical framework for the results-->
+
 A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups toward a more centralized network with concentration of connections among fewer burials after the European arrival in the 17th century.
 This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
-A corporate-based society, as we interpret the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Drennan2010; @Feinman2000; @Feinman2000political].
-In contrast, a network-based society, as we interpret the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
-It should be noted that the corporate-network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages [@Feinman2000].
+These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political].
+Compared to vertical hierarchies that stress stratified social characteristics, the corporate-network strategies may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan [cf.
+@Ueda2016].
+A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Siegel1999].
+In contrast, a network-based society, as suggested the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige goods manipulation, and trade monopolization [@Blanton1996].
+It should be noted that the corporate-network continuum represents a dynamic process with different degrees of hierarchical complexity instead of static ideal-type stages.
 
 <!--archaeological implication based on the results-->
-Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
-Our results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
-We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
-Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts.
-This might imply that ritual could be a status indicator at the time before foreign goods were first introduced to Indigenous societies.
-After the European presence was established, status indicators shifted from ritual to wealth differences.
-Moreover, geographical proximity, implying some kinship association, has not effect on sharing similar prestige items between burials for both phases.
+
+The covariants for networks provide further information to give insights into how specific elements of the archaeological record relate to the formation of networks (Table \@ref(tab:table-posterior-stats)).
+In other words, we can know what kinds of socio-cultural factors associated with the acquisition and distribution of foreign prestige goods.
+Our covariants were age, sex, ritual pottery, and wealth ranks calculated based on the value of grave goods.
+The results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
+Moreover, in the post-European network, burials with ritual vessels or same sex tend not to be tied that suggest they are not associated with the network pattern we observed.
+Those results illustrate that different social and economic mechanisms determined relations between burials, or the interconnections of prestige goods, for the two periods.
+Before European arrival, ritual behaviors seem to have been important factors that structure the interconnections of prestige goods in the burial context.
+This might imply that ritual could be a status indicator when the foreign goods were first introduced to Indigenous societies.
+After European presence was established, status indicators shifted from ritual to wealth differences with the least sex status discrimination.
+Moreover, geographical proximity, implying some kinship association, is not a factor for sharing similar prestige items between burials for both phases.
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
-Demographic information, such as age and sex, do not have any effects on network formation for both networks.
+Demographic information, such as age and sex, do not have any effects on network formation for the pre-European network and only sex shows a negative effect for the post-European network.
+This may also support a shift to more wealth and individual-based status competition without sex differences in the later phase.
 But we also note that no effects could be due to the unavailability of age and sex values for many burials in our sample.
@@
 <!-- other evidence -->
-We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
+
+We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks according to their status [eg.
 @Coward2013, pp.
 252].
-We argue that the social relations at Kiwulan changed because of contact with European colonial powers and associated foreign trade.
-This argument is supported by changes in burial network structures through time.
-It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person’s status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
+We argue that the social relation at Kiwulan changed because of the contact with European colonial power and associated foreign trade.
+This argument is supported by the changes in burial network structures through time and their associated underlying mechanisms.
+It should be noted that manipulation of burial rituals by the living can cause a disconnect between a person's status in life and their status represented by burial contexts [@Hodder1980; @Pearson1982].
 However, this issue can be reduced by comparing the results with evidence from other archaeological contexts, such as residential areas [@Chapman2003].
@@
 This is consistent with the result of a more centralized burial network after the presence of Europeans, that supports a connection between a burial and the social role of the living person.
-Also, ethnographic records support our interpretation of burial goods as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], so the interrelations represented by the flow of these goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
+Also, the burial goods used for making ties were treated as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
 
-# Conclusion {-}
+<!--anthopological implication at Kiwulan and limitation-->
+
+At Kiwulan, changes in interburial networks may result from differential access to exotic trade enabled by European contact, gradually leading to increased social differentiation.
+The image of European power could be embedded into prestige objects pre-existing in local Indigenous culture where their values were amplified with concepts of wealth or power, since their values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009].
+This may lead to ambitious individuals competing with each other for traditional goods and accumulating wealth.
+It also demonstrates that local Indigenous societies could actively incorporate and manipulate trade items in a pericolonial context with weak colonial control, gradually resulting in social changes from a more corporate to a more network strategy.
+We recognize that the meaning of the burial goods could be refined with more ethnographic data to support our interpretations.
+Human skeletal remains can provide information about health for interpreting social status; however, the preservation of human remains at Kiwulan was generally not good to yield useful data.
+By focusing on burial goods, we seek to understand the social changes at Kiwulan in relation to the contact of Europeans.
+Despite fragmentation in our archaeological data, such as missing age and sex attributes for some burials at Kiwulan, we can still capture the general patterns of network structures related to social implication.
+
+# Conclusion {.unnumbered}
+
+<!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
+
+In this paper, we presented a novel approach for studying burials to interpret social structures using ERGMs within a Bayesian framework.
+We examined social changes in a European pericolonial context in northeastern Taiwan.
+We tested the hypotheses of changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
+The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
+Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism.
+After European arrival, the network has a tendency of centralization relative to the rarity of goods.
+The changes in formation mechanisms for networks from ritual practice connections to wealth level connections could suggest increased behavior of wealth accumulation stimulated by the European presence and associated long-distance trade network.
+This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
+Using burial data with historical documents, we are able to detect changes in Indigenous societies indicative of increased social inequality to better understand indirect European colonial impacts that have been underestimated in this region [e.g.
+@Acabado2017; @Trabert2017].
 
 <!-- technical and methodological implication-->
-We presented a novel approach for studying burials to interpret social structures using ERGMs within a Bayesian framework.
-This case study demonstrates the methodological benefits of Bayesian inference on ERGMs to inform and enhance studies of relational data in archaeology.
+
+Furthermore, this case study demonstrates the methodological improvement that Bayesian inference on ERGMs provides to inform and enhance studies of relational data in archaeology.
 A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-Another valuable feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological models, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
 Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
@@
 
-<!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-We examined social changes in a European pericolonial context in northeastern Taiwan.
-We tested the hypotheses of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
-The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
-Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism.
-After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-We interpret the changes in formation mechanisms of networks, from ritual practice to wealth levels, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
-This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
-Using burial data with historical documents, we are to better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
-@Acabado2017; @Trabert2017].
+<!-- stopped editing here -->
 
-# Acknowledgments {-}
+# Acknowledgments {.unnumbered}
 
@@
 Goodreau for the discussion regarding the methods at the earliest stage of analysis.
-This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
+This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project \#DF009-A-18).
 We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
@@
 <!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
+
 ##### pagebreak
 
-# References {-} 
+# References {.unnumbered}
+
 <!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
+
+::: {#refs}
+:::
 

### 2021-02-25 analysis/paper/paper.Rmd
edit title, abstract, and intro based on BM's comment

--- 2021-02-19 analysis/paper/paper.Rmd
@@
 ---
-title: "A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
+title: "A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period"
 author:
@@
   Burials provide valuable information to study social structures based on the assumption that burials and associated grave goods can represent social roles and relations in a society.
-To study social relationships, network analysis has been increasingly applied to archaeological data to explore network structures and patterns to infer interactions and relationships between entities.
-Statistical approaches to network analysis, such as exponential random graph models (ERGMs), provide a way to test hypotheses on dynamic processes of network formation.
-However, the computational difficulty and sensitivity to uncertainties limits the application of ERGMs.
+To study social relationships, network analysis has been increasingly applied to archaeological data to infer interactions and relationships between entities.
+Statistical approaches to network analysis, such as exponential random graph models (ERGMs), provide a way to test hypotheses about dynamic processes of network formation.
+However, computational difficulties and sensitivity to uncertainties limit the application of ERGMs.
 In this paper, we introduce a Bayesian framework on ERGMs that enables an efficient computational process, effective quantification of uncertainty, and robust model evaluation of network properties.
 We tested a hypothesis of social change relative to the arrival of Europeans by studying burial data from Kiwulan, an Iron Age site in northeastern Taiwan.
-The results indicate a transition in the process that led to the formation of the burial network, from ties based on ritual objects to wealth objects, and a more centralized structure with increased social differentiation after European presence in the 17th century.
+The results indicate a transition among the burials from network ties based on ritual objects to wealth objects, and a more centralized structure with increased social differentiation after the European presence was established in the 17th century.
 Our case study demonstrates the effectiveness of Bayesian network analysis for archaeological data, and expands the use of burials in understanding the impacts of colonial presence on Indigenous groups in a pericolonial context.
@@
 <!---network analysis approaches and basic concepts-->
-Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing rational data [@Brandes2013; @Mills2017; @Peeples2019].
-The recent development of modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures inherent in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
-By comparing hypothesized networks with the observed network, network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
+Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing relational data [@Brandes2013; @Mills2017; @Peeples2019].
+Recent developments in modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+By comparing hypothesized networks with the observed network, statistical network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way.
 Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Handcock2008; @Ghafouri2020].
-Since relationships in a network emerging relative to their position in a network can be hypothesized, ERGMs are promising to evaluate dynamic social processes behind observed archaeological networks [@Brughmans2014].
-However, computational difficulties and the sensitivity to uncertainties to some extent limits the practical applications of ERGMs, where results could be challenging for interpretation [@Caimo2014].
+ERGMs are promising for evaluating the dynamic social processes behind observed archaeological networks [@Brughmans2014].
+However, computational difficulties, sensitivities to uncertainties, and ambiguities in interpretation limit the practical applications of ERGMs [@Caimo2014].
 
 <!--study social changes in pericolonial context using burial data as a proxy-->
-In this paper, we go beyond ERGMs with a novel Bayesian approach that can alleviate the computational issues in ERGMs to better interpret the process of network formation for past phenomena.
+In this paper, we go beyond ERGMs with a novel Bayesian approach that alleviates the computational issues and other limitations of ERGMs to enable clearer and more robust interpretations of the processes of network formation.
 Using burial data from an Iron Age site in northeastern Taiwan as a case study, we explore social changes by investigating the formation of material connections between burials.
 Social changes in Indigenous societies, when faced with colonial powers, are commonly observed in many parts of the world, especially European colonies where Indigenous economic, cultural, or socio-political aspects were substantially impacted [@Dietler2005; @Silliman2005; @Voss2005].
-Recent studies demonstrate that the indirect effect of colonialism, or involvement in long-distance trade, could also have impacts on Indigenous societies, this is known as a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
-To explore social changes, burials are important evidence to understand past societies because material culture and biological records of burial behaviors can represent the social ranking or identities of the deceased, and social relations between them [@Binford1971; @Drennan2010; @Saxe1970].
+Recent studies demonstrate that the indirect effects of colonialism, or involvement in long-distance trade, may also have impacted Indigenous societies, this is known as a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
+Burials are important for understanding past societies because material culture and biological records of burial behaviors can represent the social ranking of the deceased, and social relations between them [@Binford1971; @Drennan2010; @Saxe1970].
 Burial treatments can further indicate social complexity or inequality where foreign goods from long-distance trade were used by Indigenous people to express status [@Carter2015; @Dolfini2019].
@@
 <!--Our burial network and research implication-->
-Usually, burial studies are based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining those individual observations to infer past societies [@Byrd1995; @Seikel2011].
+Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011].
 Here we treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
-individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing global patterns represented in the data [@Brandes2013].
-Our case study illustrates how a Bayesian approach to network modeling provides new insights for studying the formation of archaeological networks and allows testing on theoretical assumptions to better understand dynamic social processes.
+individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013].
+Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.
 

### 2021-02-25 analysis/paper/paper.Rmd
edit ERGMs and background based on BM's comments

@@
 Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
-By fitting models with the observed network of archaeological data, they proposed that ERGMs is a promising method for exploration and hypothesis testing for social processes.
+By fitting models with the observed network of archaeological data, they proposed that ERGMs are a promising approach for exploration and hypothesis testing of social processes.
 Similarly, @Amati2019 modeled three networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore interaction mechanisms, including proximity, inter-cultural items, and pottery types.
@@
 <!--limitations of ERGMs-->
-However, those studies also point out some limitations of ERGMs, such as sensitivity to missing data and less able to handle uncertainty.
-Also, it is difficult in ERGMs to estimate model parameters and interpret the result due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
+However, those studies also illustrate some limitations of ERGMs, such as sensitivity to missing data and limited ability to incorporate and represent uncertainty.
+Also, it is difficult in ERGMs to estimate model parameters and interpret the results due to intractable likelihood normalizing constants and model degeneracy [@Caimo2014; @Jin2013].
 A normalizing constant is a function of the model parameter for making probability distributions integrate to one, which becomes harder to compute with a larger set of networks  [@Caimo2017bayesian].
@@
 Model degeneracy is another issue where probability models tend to overestimate a small number of extreme graphs by assigning too much weight, e.g., in extreme cases such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014].
-One solution to these limitations is available by implementing ERGMs by means of a Bayesian framework.
+One solution to these limitations is to implement ERGMs in a Bayesian framework.
 
@@
 <!-- Bayesian inference for Exponential Random Graph Models --> 
-Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and improve computational issues in ERGMs [@Caimo2017; @Lehmann2020].
-Prior information is derived from previous data or assumptions of our data.
-The advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
+Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and overcome computational issues of ERGMs [@Caimo2017; @Lehmann2020].
+Prior information is derived from previous data or assumptions derived from the context of our data.
+An important advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
 With the exchange algorithm, Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives better convergence results.
 This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-By fitting an ERMG with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information that estimates the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
-This allows uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].
+By fitting an ERMG with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
+This allows robust and interpretable uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].
 In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
-@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can predict, on average, 80% of the ties when a third of data is missing.
+@Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can accurately predict, on average, 80% of the ties when a third of data is missing.
 
@@
 This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
-Our general model proposes that the influence of a colonial power, combined with high local values attached to imported goods, leads to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-Increased social inequality associated with the use of foreign prestige goods is recognized as network strategies for gaining power [@Blanton1996; @Feinman2000].
+Our anthropological model proposes that the influence of a colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+Increased social inequality associated with the use of foreign prestige goods is recognized as a ‘network’ strategy for gaining power [@Blanton1996; @Feinman2000].
 Network-based societies are identified in some places in Asia with involvement in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
@@
 <!--theoretical framework and research questions-->
-The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000], while corporate-based societies stress communal ritual elements, shared power, and wealth differentiation, if any, would be associated with corporate groups, such as age groups [@Siegel1999].
-Using the theoretical framework of corporate-network strategies, or horizontal hierarchy, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation ) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
+The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000].
+These may be contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age groups, rather than individuals [@Siegel1999].
+Using the theoretical framework of corporate-network strategies, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
 We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
-(2) if so, what are the major variables affecting or forming unequal social positions that might hint at social heterogeneity?
+and if so, (2) what are the major variables affecting or forming unequal social positions that might hint at social heterogeneity?
 By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
@@
 We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) [@Chen2007].
-A total of 90 burials were unearthed from adjacent squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
+A total of 90 burials were unearthed from adjacent excavation squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
 The preservation of human remains at Kiwulan was generally poor, so we have incomplete age and sex data (details in Supplementary Online Materials) and lack health status data.
@@
 For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society.
-However, @Hsieh2012 suggested a relatively egalitarian structure based on the comparative analysis of the frequencies of all burial goods.
+However, @Hsieh2012 suggested a relatively egalitarian structure based on a comparative analysis of the frequencies of all burial goods.
 She found that the burials with rare prestige goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
-One important limitation of these previous studies is that they did not use analytical units suitable for comparing the phases before and after European arrival.
+One important limitation of these previous studies is that they did not use analytical units suitable for comparing behaviors before and after European arrival.
 Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until arrival of Han Chinese immigrants).
@@
 We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
-We also excluded `r disturb` burials that were heavily disturbed by modern construction because we cannot determine their chronology.
+We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology.
 
@@
 The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments.
-These items are considered as high-value across different archaeological contexts based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+These items are considered as high-value across different archaeological contexts throughout Taiwan based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
 European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
 For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads’ high value in Indigenous culture [@LiandWu2006].
-In addition, Spanish visitors observed that an Indigenous person possessing more imported goods may have been recognized as having higher status in their community [@LiandWu2006].
+In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006].
 

### 2021-03-01 analysis/paper/paper.Rmd
edit the whole rmd based on BM's comments

@@
 <!-- our expectation based on hypothesis --> 
-Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
+Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where the social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
 This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
-If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with associations with wealth accumulation, indicative of a network-based society.
-In contrast, the network before the European arrival will show low centralization and high transitivity, associated with ritual element, or age/sex groups to indicate a corporate-based society.
-To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shape relationships between people at Kiwulan.
+If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with network ties determined by wealth, indicative of a network-based society.
+In contrast, we predict the network before European arrival will show low centralization and high transitivity, and network ties associated with ritual elements, indicative of a corporate-based society.
+To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shaped relationships between people at Kiwulan.
 By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
@@
 presence or absence).
-Node attributes here include osteologically determined age and sex, cultural data, such as ritual pottery, and a burial value index.
-Ritual pottery was identified as locally made ceramics placed above an individual.
-Those pots mostly have soot on vessel exteriors and charred residues on interiors that are interpreted as vessels for funeral feasting [@Hsieh2009].
+Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index.
+Ritual pottery was identified as locally-made ceramics placed above an individual.
+These pots mostly have sooty vessel exteriors and charred residues on interiors that are interpreted as vessels used for funeral feasting [@Hsieh2009].
 We calculated our burial value by summing values of all types of prestige goods in each burial context.
@@
 The size of each node is proportional to node degrees or the number of connections to a node.
-The thickness of tie represents the number of goods in common between two nodes."}
+The thickness of ties represents the number of goods in common between two nodes."}
 knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
@@
 Table \@ref(tab:table-ergm-terms) lists the network parameters we used for dependence assumptions that define our models [@Morris2008].
-Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations in terms of prestige goods between two burials.
+Every parameter in an ERGM has an associated algorithm for computing the probability of observing relations between two burials.
 Based on our hypothesis, we model a network with increased social inequality to be represented by endogenous network effects: low transitivity and high centralization.
@@
 Normal distributions for the priors are typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-Thus, we specified the prior of the edge density parameter to low for both network models, which follows a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
+Thus, we specified the prior of the edge density parameter to low for both network models, following a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
 For covariates about biological features, such as age, sex, we specified a vague prior (N(`r priors$mean_same[2]`, `r priors$sigma_same[2]`)) for both models to explore their effects.
-For physical distance between burials, we also set a vague prior effect (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) to explore whether there is kinship-based proximity, e.g.
+For physical distance between burials, we also set a vague prior (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) to explore whether there is kinship-based proximity, e.g.
 stronger correlations for shorter distances.
 To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity and centralization.
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and stress ritual element shared in corporate groups.
-Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and higher covariate effect of burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation.
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and emphasize the ritual element shared in corporate groups.
+Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and a higher covariate effect for burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation after European contact.
 
@@
 The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/xxx/xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
-Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here.
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here and additional details of definition of our approach to tie formation, model fitting, assessment of MCMC output and goodness-of-fit diagnostics for our models.
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, the ritual element represented by pots (ritual-homophily) has a significant effect on the formation of relations between burials.
-On the other hand, the wealth rank, represented by burial values (wealth-homophily), has a negative effect but not significant since its confidence intervals include zero.
-This demonstrates that burials in the pre-European period with ritual pottery tend to form relations.
-Despite positive mean values for some covariates, such as age and sex, they do not show a significant influence.
+For nodal covariates in the pre-European model, the ritual element represented by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while the wealth rank represented by burial values (wealth-homophily) has a negative effect but not significant since its confidence intervals include zero.
+This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
+Despite positive mean for some covariates, such as age (age-homophily) and sex (sex-homophily), they do not show a significant tendency.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
@@
 The high positive value for transitivity suggests a tendency of burials with similar burial goods to be clustered as connected communities, indicative of the presence of multiple corporate groups sharing burial goods in common.
-In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties, without any prominent burials.
-This might imply that individuals have equal access to the flow of trade goods.
+In contrast, the strong negative centralization shows there is a tendency toward decentralization that reflects most burials having a similar number of ties without any prominent burials.
+This might imply that individuals have equal access to trade goods in terms of the flow of goods.
 
@@
 <!--posterior stats for for post-E and their comparison-->
-After the arrival of Europeans, the nodal covariates of wealth rank show significant positive effects.
-This indicates burials in the same wealth level tend to be more connected with each other.
-Despite negative effects for ritual and sex variables, they are not significant.
-Also, there are no significant effects for variables of age and physical distance.
-Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than in the pre-European network.
-In contrast, centralization has a significantly higher positive effect than the pre-European network.
-This means there is a tendency toward centralization that reflects a limited number of burials having many more ties than others.
-This implies more access to trade goods, increased wealth accumulation and display behaviors in burial events.
-In general, the post-European network model has a smaller transitivity effect and a positive centralization effect compared to the pre-European network model.
+For the post-European network model, the nodal covariate of wealth rank shows significant positive effects.
+This indicates the burials in the same wealth level tend to have relations.
+Despite negative effects for the ritual and sex variables, they are not significant.
+Also, there are no significant effects for the variables of age and physical distance.
+Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than the effect of the pre-European network.
+In contrast, centralization has a significantly higher positive effect than the effect of the pre-European network.
+This means there is a tendency toward centralization, reflecting a limited number of burials having many more ties than others.
+This implies that the presence of better access to trade goods and the behavior of wealth accumulation and display in burial events.
+In general, the post-European network model has a smaller transitivity effect and a positive centralization effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
@@
 Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
-Our results indicate that the major factor contributing to burial networks is ritual for the pre-European network, and wealth for the post-European network.
+Our results indicate that the major factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth.
 We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
@@
 After the European presence was established, status indicators shifted from ritual to wealth differences.
-Moreover, geographical proximity, implying some kinship association, has no effect on sharing similar prestige items between burials for both phases.
+Geographical proximity of burials, implying kinship association, has no effect on sharing similar prestige items between burials for both phases.
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
 Demographic information, such as age and sex, do not have any effects on network formation for both networks.
-But we also note that no effects could be due to the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
+We also note that no effects could be due to the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
 
@@
 This case study demonstrates the methodological benefits of Bayesian inference on ERGMs to inform and enhance studies of relational data in archaeology.
-A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations.
-Another valuable feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological models, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
+A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data and incorporating prior information and MCMC estimations to quantify uncertainty and provide interpretable output.
+An important feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological models, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
 Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
@@
 <!-- summary of the research by answering questions raised in Intro, and anthropological contribution -->
-We examined social changes in a European pericolonial context in northeastern Taiwan.
-We tested the hypotheses of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
+Our case study examined social changes in a European pericolonial context in northeastern Taiwan.
+We tested a hypothesis of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
 The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
@@
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-We interpret the changes in formation mechanisms of networks, from ritual practice to wealth levels, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
+We interpret the changes in formation mechanisms of networks, observed as a change from ties formed by ritual to wealth, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
 This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].

### 2021-03-02 analysis/paper/supplementary-material.Rmd
revised the SI and the table in it

@@
 -  further historical background of 17th century northeastern Taiwan in a pericolonial context
--  further review of network analysis and archaeological applications
--  further definition of our approach to tie formation and node attributes from Kiwulan burials
--  further details on methods of modeling fitting, and
--  assessment of MCMC output and Goodness-of-fit diagnostics for our models
+-  further review of network analysis and archaeological applications relevant to this study
+-  further definitions of our approach to tie formation and node attributes from Kiwulan burials
+-  further details on our methods of modeling fitting, and
+-  assessment of our MCMC output and Goodness-of-fit diagnostics for our models
 
@@
               values_from = n) %>% 
-  mutate(Age_scale = factor(Age_scale,
-                            levels = c("0-12", "12-20", "+20", "NA"))) %>% 
-  arrange(Age_scale) %>% 
   rename(age = Age_scale) %>% 
-  select(age, pre_male, pre_female, pre_NA, post_male, post_female, post_NA)
+  mutate(age = ifelse(age == "NA", "unknown", age)) %>% 
+  mutate(age = factor(age,
+                      levels = c("0-12", "12-20", "+20", "unknown"))) %>% 
+  arrange(age) %>% 
+  select(age, pre_male, pre_female, pre_NA, post_male, post_female, post_NA) %>% 
+  mutate_all(., funs(replace_na(., 0)))
 
 ft_age_sex <- flextable(age_sex_by_phase) %>% 
+  colformat_double(digits = 0) %>%
   set_header_labels(pre_male = "pre-European", pre_female = "pre-European",
@@
   merge_at(i = 1, j = 5:7, part = "header") %>% 
-  add_header_row(values = c("","male", "female", "NA", 
-                            "male", "female", "NA"), top = FALSE) %>% 
+  add_header_row(values = c("","male", "female", "unknown", 
+                            "male", "female", "unknown"), top = FALSE) %>% 
   align(j = 2:7, align = "right", part = "header") %>% 

### 2021-03-02 analysis/paper/supplementary-material.Rmd
adjust table and knit the rmd and SI

@@
   align(i = 1, align = "center", part = "header") %>% 
-  set_caption("The number of burials with identification of age and sex at Kiwulan. The NA means no sufficient skeletal remains for determination of age or sex")
+  width(j = ~ age, width = 1) %>% 
+  width(j = 2:7, width = 0.8) %>% 
+  set_caption("The number of burials with identification of age and sex at Kiwulan. The unknown means no sufficient skeletal remains for determination of age or sex")
 ft_age_sex

### 2021-03-02 analysis/paper/supplementary-material.Rmd
change title for SI

@@
 ---
-title: "Supplementary Materials- A Bayesian approach to burial networks to infer social changes in northeastern Taiwan during the European colonization period"
+title: "Supplementary Materials- A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period"
 author:

### 2021-03-02 analysis/paper/paper.Rmd
add an OSF link

@@
 
-The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/xxx/xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at http://doi.org/10.17605/OSF.IO/XGA6N to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here and additional details of definition of our approach to tie formation, model fitting, assessment of MCMC output and goodness-of-fit diagnostics for our models.

### 2021-03-03 analysis/paper/paper.Rmd
submit to JAS

@@
 Goodreau for the discussion regarding the methods at the earliest stage of analysis.
-This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
+This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington, the predissertation summer travel grant from the Henry Luce Foundation/ACLS Program in China Studies, and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
 We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.

### 2021-04-24 analysis/paper/paper.Rmd
fixed the typo of ERGM

@@
 This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
-By fitting an ERMG with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
+By fitting an ERGM with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
 This allows robust and interpretable uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].

### 2021-05-01 analysis/paper/paper.Rmd
edit the introduction based on the R1's comments

@@
 This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
-Our anthropological model proposes that the influence of a colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-Increased social inequality associated with the use of foreign prestige goods is recognized as a ‘network’ strategy for gaining power [@Blanton1996; @Feinman2000].
-Network-based societies are identified in some places in Asia with involvement in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004].
+We assume that the use of imported goods and their values were increased and stressed after the European arrival.
+Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities stimulated by the Europeans.
+This paper focuses on the effect of European colonization in a pericolonial context that is not well understood compared to places that were colonial centers [@Acabado2017; @Trabert2017].
+
+The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Network-based societies are identified in some places in Asia that were invovled in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
+Here we define trade goods as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with the symbol of personal influence described in the Spanish records [@LiandWu2006].
+In addition, the spatial pattern of trade goods in residential presents a more concentration pattern statistically after the European presence.
+The evidence from different archaeological contexts indicates the possible use as prestige goods.
 

### 2021-05-04 analysis/paper/paper.Rmd
edit the caption of figure 3 and color the ties

@@
 The size of each node is proportional to node degrees or the number of connections to a node.
-The thickness of ties represents the number of goods in common between two nodes."}
+The thickness of ties represents the number of goods in common between two nodes.
+Isolated nodes are not displayed in the plots"}
 knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))

### 2021-05-05 analysis/paper/paper.Rmd
edit discussion to reply to R1's comments

@@
 Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004].
+Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004; @Cheng2017].
 We assume that the use of imported goods and their values were increased and stressed after the European arrival.
-Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities stimulated by the Europeans.
-This paper focuses on the effect of European colonization in a pericolonial context that is not well understood compared to places that were colonial centers [@Acabado2017; @Trabert2017].
+Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2004; @Cheng2017].
+This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that is still not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].
 
@@
 The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
-Here we define trade goods as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with the symbol of personal influence described in the Spanish records [@LiandWu2006].
-In addition, the spatial pattern of trade goods in residential presents a more concentration pattern statistically after the European presence.
-The evidence from different archaeological contexts indicates the possible use as prestige goods.
+Here we assume that trade goods could be used as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with the symbol as personal influence described in the Spanish records [@LiandWu2006].
+In addition, the study of spatial distribution of trade ornaments in residential at Kiwulan presents a more clustered pattern after the European presence [@LWandBM2020ornament].
 
@@
 <!--archaeological implication based on the results-->
-Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
+Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestigegoods.
 Our results indicate that the major factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth.
@@
 
-<!-- other evidence -->
+<!-- other evidence supporting the change in burial networks -->
 We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
@@
 Also, ethnographic records support our interpretation of burial goods as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], so the interrelations represented by the flow of these goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
+
+<!-- trade dynamic during this period and future work for comparison -->
+We also recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have had contributed to social changes at Kiwulan.
+However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival.
+Historical records also support the interaction and trade activities between Europeans and local Indigenous people.
+The trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at Kiwulan were frequently found from archaeological sites with Dutch occupation, such as Fort Zeelandia.
+As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides an ideal example to explore the reaction of Indigenous people to European colonial presence at a settlement scale.
+We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Helping Dao and Tamsui.
+However, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence are rare to offer good comparison currently and future work is still needed.
 

### 2021-05-06 analysis/paper/paper.Rmd
edit intro section based on R2's comments

@@
 Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011].
-Here we treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
+Here we use network analysis to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations.
+We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013].
+Instead of exchange or transmission, our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg.
+@Collar2015, pp.
+4].
+This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013].
 Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.
@@
 As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides an ideal example to explore the reaction of Indigenous people to European colonial presence at a settlement scale.
-We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Helping Dao and Tamsui.
+We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Heping Dao and Tamsui.
 However, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence are rare to offer good comparison currently and future work is still needed.

### 2021-05-07 analysis/paper/paper.Rmd
continue revising based on R2's comments

@@
 We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology.
+Differences in network size were evaluated using vertex bootstrap in a later section.
 
@@
 In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006].
+Besides, those goods are made of durable materials that can reduce the issue of absence due to taphonomic processes.
 
@@
 We then assigned the burials into three ranks according to breaks in the distribution of burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (details in Supplementary Online Materials).
-Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
+Since burials tend to have multiple goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
@@
 
-```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival, B: Burial network after the European arrival.
+```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival (14th-16th centuries), B: Burial network after the European arrival (17th-18th centuries).
 The size of each node is proportional to node degrees or the number of connections to a node.

### 2021-05-07 analysis/paper/supplementary-material.Rmd
edit the introduction based on the R1's comments
edit intro section based on R2's comments
continue revising based on R2's comments

@@
 The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
+Increased social inequality associated with the use of exotic trade goods is recognized as a ‘network’ strategy for gaining power [@Blanton1996; @Feinman2000].
 Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007].
@@
 If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments.
+Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can robustly represent network structure.
+Also, each burial has a clear boundary as our analytical unit that reduces sampling errors.
+
 For burial attributes, we assigned burials to different groups according to their age, sex, burial value and the presence/absence of ritual pottery.

### 2021-05-08 analysis/paper/paper.Rmd
Clarify the use of priors

@@
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
-Normal distributions for the priors are typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].
-Thus, we specified the prior of the edge density parameter to low for both network models, following a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
+Priors are the distinctive aspect of Bayesian methods to estimate population by combining one's beliefs with data [@Gill2014].
+Normal distribution N($\mu$, $\sigma^2$), where $\mu$ is mean and $\sigma^2$ is standard deviation, is typical for prior specification to model network parameters because it is sensible and convenient, allowing adjustments for a wide range of prior information [@Caimo2017; @Normand1992].
+The larger the mean of the prior distribution, the more positive effect of that parameter on network formation.
+For standard deviation, if it is larger, it presents a relatively vague prior enabling comparison between variables that we do not have specific assumptions.
+Since low density is commonly found in the real world [@Caimo2017], we specified the prior of the edge density parameter to low for both network models, following a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
+
+To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we specify priors that are meaningful for social inequality, especially for the variables of transitivity and centralization.
+We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and emphasize the ritual element shared in corporate groups.
+Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and a higher covariate effect for burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation after European contact.
 For covariates about biological features, such as age, sex, we specified a vague prior (N(`r priors$mean_same[2]`, `r priors$sigma_same[2]`)) for both models to explore their effects.
@@
 stronger correlations for shorter distances.
-To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we incorporated different prior information for the network variables that are meaningful for social inequality, especially for transitivity and centralization.
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and emphasize the ritual element shared in corporate groups.
-Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and a higher covariate effect for burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation after European contact.
 

### 2021-05-12 analysis/paper/paper.Rmd
moved figure 6 from the main text to the SI

@@
 
-```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
-Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
-Red lines represent distributions of observed networks, and gray lines show the 95% intervals."}
-library(magick)
-pre_bgof_img <- image_read(here::here("analysis/figures/003-pre-bgof.png"))
-post_bgof_img <- image_read(here::here("analysis/figures/004-post-bgof.png"))
-image_append(c(pre_bgof_img, post_bgof_img), stack = T)
-```
-
 <!--assessment of MCMC analysis and BGOF--> 
-Our computational models reproduce networks that resemble the structural features of our observed networks.
-Bayesian goodness of fit diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases); details in Supplementary Online Materials) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
-the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
-For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
+Our computational models reproduce networks that resemble the structural features of our observed networks according to Bayesian goodness of fit diagnostics (details in Supplementary Online Materials), showing both models fit the observed networks very well.
 We also compared the first three distribution moments of each observed distribution and their corresponding simulated distributions, represented by means.

### 2021-05-12 analysis/paper/supplementary-material.Rmd
moved figure 6 from the main text to the SI

@@
 
-Our Bayesian GOF diagnostics, shown in the main text, summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
-This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
-We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
-We expect to see the models fit with our hypotheses, indicative of an increased social inequality after the foreign contact.
-In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
-
 ```{r figure-pre-diag, fig.cap = "MCMC output of the pre-European model. In the order of edges, nodematch-age, nodematch-sex, nodematch-ritual, nodematch-value, gwesp, gwdeg, and dyadcov.distance."}
@@
 knitr::include_graphics(here::here("analysis", "figures", "009-post-diag.png"))
+```
+
+Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
+The diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
+For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
+This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
+We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
+We expect to see the models fit with our hypotheses, indicative of an increased social inequality after the foreign contact.
+In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
+
+```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
+Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
+Red lines represent distributions of observed networks, and gray lines show the 95% intervals."}
+library(magick)
+pre_bgof_img <- image_read(here::here("analysis/figures/003-pre-bgof.png"))
+post_bgof_img <- image_read(here::here("analysis/figures/004-post-bgof.png"))
+image_append(c(pre_bgof_img, post_bgof_img), stack = T)
 ```

### 2021-05-12 analysis/paper/paper.Rmd
add the raincloud plot to the main text and edit the SI

@@
 <!-- burial data and the definition of tie formation  --> 
-Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials).
+Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (Figure \@ref(fig:raincloud-beads); details in Supplementary Online Materials).
 For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they both possess each type of goods (i.e.
@@
 For each burial we sum these values of each type of prestige item to get a burial value [@Jorgensen1991].
+
+```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
+knitr::include_graphics(here::here("analysis", "figures", "000-raincloud-beads.png"))
+```
 

### 2021-05-12 analysis/paper/supplementary-material.Rmd
add the raincloud plot to the main text and edit the SI

@@
 <!-- burial data and the definition of tie formation: partially move to SI  -->
-For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)).
+For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads), also in the main text).
 Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1).
@@
 The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments.
-Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can robustly represent network structure.
-Also, each burial has a clear boundary as our analytical unit that reduces sampling errors.
+In addition, each burial has a clear boundary as our analytical unit that reduces sampling errors.
+Thus, we are confident that the quantities and distributions of beads in burial contexts can reflect the original differences between burials.
+Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can effectively represent structures of burial networks.
 

### 2021-05-12 analysis/paper/paper.Rmd
finished editing for R2's comments

@@
 <!-- ERGMs in archaeology--> 
-Exponential random graph models (ERGMs) first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
+Exponential random graph models (ERGMs) are statistical models for social networks that first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
 By fitting models with the observed network of archaeological data, they proposed that ERGMs are a promising approach for exploration and hypothesis testing of social processes.
@@
 <!-- Bayesian inference for Exponential Random Graph Models -->
-Bayesian approaches to ERGMs are effective tools for network modeling by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and overcome computational issues of ERGMs [@Caimo2017; @Lehmann2020].
-Prior information is derived from previous data or assumptions derived from the context of our data.
+
+In this paper, we used Bayesian inference on ERGMs to test our hypotheses.
+ERGMs are the tool that provides the conceptual background for network formation and structure, allowing testing social processes through covariates.
+The benefit of the Bayesian part is to deal with computational issues of ERGMs mentioned above.
+Bayesian approaches to ERGMs are effective because it incorporates prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and uncertainty [@Caimo2017; @Lehmann2020].
+Prior information can be derived from previous data or assumptions derived from the context of our data.
 An important advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
@@
 This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].
+
 By fitting an ERGM with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].

### 2021-05-14 analysis/paper/paper.Rmd
start editing for R3's comments

@@
 <!-- ERGMs in archaeology--> 
-Exponential random graph models (ERGMs) are statistical models for social networks that first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites.
+Exponential random graph models (ERGMs) are statistical models for social networks that first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites and at 190 sites for a later study [@Brughmans2015].
 By fitting models with the observed network of archaeological data, they proposed that ERGMs are a promising approach for exploration and hypothesis testing of social processes.
@@
 We explored the sample size effect by removing nodes at certain percentages (5-40%) for both networks and comparing their confidence intervals.
-The results demonstrate a consistent difference between the two networks up to 30% node removals (Figure \@ref(fig:figure-bootstrap-CI)).
-This indicates that the network variables under investigation are robust to the different network size of our sample.
+The results demonstrate a consistent difference between the two networks up to 30% node removals as shown in Figure \@ref(fig:figure-bootstrap-CI).
+This indicates that the network variables, including density, centralization, and transitivity, are robust to the different network size of our sample.
 Figure \@ref(fig:figure-bootstrap-CI) also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.

### 2021-05-15 analysis/paper/paper.Rmd
justify the connection between theory and network expectation

@@
 Recent studies demonstrate that the indirect effects of colonialism, or involvement in long-distance trade, may also have impacted Indigenous societies, this is known as a pericolonial context [@Acabado2017; @Trabert2017; @LWandBM2020ornament].
-Burials are important for understanding past societies because material culture and biological records of burial behaviors can represent the social ranking of the deceased, and social relations between them [@Binford1971; @Drennan2010; @Saxe1970].
+Burials are important for understanding past societies because material culture and biological records of burial behaviors can represent the social ranking of the deceased, and social relations between them [@Binford1971; @Curet1998; @Drennan2010; @Saxe1970].
 Burial treatments can further indicate social complexity or inequality where foreign goods from long-distance trade were used by Indigenous people to express status [@Carter2015; @Dolfini2019].
@@
 The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Network-based societies are identified in some places in Asia that were invovled in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
-The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.
-Here we assume that trade goods could be used as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with the symbol as personal influence described in the Spanish records [@LiandWu2006].
-In addition, the study of spatial distribution of trade ornaments in residential at Kiwulan presents a more clustered pattern after the European presence [@LWandBM2020ornament].
+Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan.
+Foreign trade goods could be used as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006].
+In addition, the study of spatial distribution of trade ornaments in residential area at Kiwulan demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament].
 
@@
 The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000].
-These may be contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age groups, rather than individuals [@Siegel1999].
-Using the theoretical framework of corporate-network strategies, we test a hypothesis that changes from a corporate mode (a structure showing more subgroups with less wealth differentiation) to a network mode (a centralized structure with more wealth differentiation) can be observed in the Kiwulan burial network after European arrival.
-We ask: (1) did European colonial activities result in increased social inequality in Indigenous society in ways that can be detected by analysis of burial networks?
-and if so, (2) what are the major variables affecting or forming unequal social positions that might hint at social heterogeneity?
+These are contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999].
+The European contact and associated trade activities could be the stimulus of competition for foreign resources among aggrandizers that led to social changes.
+Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network.
+We ask: (1) Do the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact?
+and if so, (2) what are the major variables affecting or forming these two networks?
+Can we identify the shift in variables from ritual to wealth that associated with two different strategies?
 By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
@@
 <!-- our expectation based on hypothesis --> 
-Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where the social structure changed from a more corporate strategy to a more network strategy [@Blanton1996; @Drennan2010; @Feinman2000].
-This would be reflected by the pattern of intercorrelations of foreign prestige goods in burial networks.
-If social inequality gradually increased as we hypothesize, then we expect to observe a network structure with higher centralization (or popularity) and less transitivity (or clustering), along with network ties determined by wealth, indicative of a network-based society.
+Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where the social structure changed from a more corporate strategy to a more network strategy, presenting different organizational modes of societies [@Blanton1996; @Drennan2010; @Feinman2000].
+Corporate-based societies stress shared power, multiple segments, and social cohesion through communal rituals, while network-based societies emphasize the accumulation of individual prestige, wealth, and resources by building personal networks.
+We used burial data as a proxy to explore social changes from a corporate strategy to a network strategy.
+This change can be identified though the pattern of intercorrelations of foreign prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
+@Coward2013, pp.
+252].
+
+If social inequality gradually increased after European arrival as we hypothesize, then we expect to observe a network structure with higher centralization and less transitivity, along with network ties determined by wealth, indicative of a network-based society.
 In contrast, we predict the network before European arrival will show low centralization and high transitivity, and network ties associated with ritual elements, indicative of a corporate-based society.
+High transitivity (or clustering) indicates multiple subgroups with shared burial goods, representing social cohesion instead of individual accumulation in an earlier time.
 To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shaped relationships between people at Kiwulan.
@@
 This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
+Changes in mortuary practices could reflect the development of social or political complexity [eg.
+@Curet1998].
 A corporate-based society, as we interpret the pre-European situation at Kiwulan, stresses shared power between individuals, communal rituals, and social inequality, if any, would be associated with groups [@Drennan2010; @Feinman2000; @Feinman2000political].
@@
 <!--archaeological implication based on the results-->
-Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestigegoods.
+Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
 Our results indicate that the major factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth.
@@
 This might imply that ritual could be a status indicator at the time before many foreign goods were introduced to Indigenous societies.
-After the European presence was established, status indicators shifted from ritual to wealth differences.
+After the European presence was established, status indicators shifted from ritual to wealth differences, corresponding with the characteristics of corporate/network strategies.
 Geographical proximity of burials, implying kinship association, has no effect on sharing similar prestige items between burials for both phases.
@@
 <!-- other evidence supporting the change in burial networks -->
-We used burial data as a proxy to explore social relations reflected by interconnections of prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
-@Coward2013, pp.
-252].
 We argue that the social relations at Kiwulan changed because of contact with European colonial powers and associated foreign trade.
@@
 We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Heping Dao and Tamsui.
-However, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence are rare to offer good comparison currently and future work is still needed.
+Unfortunately, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence as Kiwulan are rare to offer good comparison and future work is still needed.
 

### 2021-05-19 analysis/paper/paper.Rmd
cite two papers

@@
 By comparing hypothesized networks with the observed sites, they found that the presence of hub sites can be efficiently explained by multiple interdependent mechanisms instead of only one variable exclusively.
+Other examples, such as exploring drivers of social stability for the Valencian Bronze Age settlements in eastern Spain [@Cegielski2020], or reconstructing structurally efficient networks of Middle Bronze Age settlements in the Aegean [@Amati2018], demonstrate the usefulness of ERGMs for hypothesis testing for social processes.
 

### 2021-05-19 analysis/paper/paper.Rmd
add citations to the main text

@@
 Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011].
-Here we use network analysis to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations.
+Recent studies on burials using network analysis demonstrates it is a promising tool to test expectations derived from hypotheses [@Sosna2013].
+Here we use network modeling to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations.
 We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
 individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013].
-Instead of exchange or transmission, our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg.
+Our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg.
 @Collar2015, pp.
-4].
+4; eg.
+@Sosna2013].
 This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013].
@@
 In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006].
-Besides, those goods are made of durable materials that can reduce the issue of absence due to taphonomic processes.
+Besides, those goods are made of durable materials that is reasonable to assume that they were less affected by taphonomic processes.
 

### 2021-05-19 analysis/paper/supplementary-material.Rmd
edit the SI

@@
 
-## Assessment of MCMC output and Goodness-of-fit diagnostics for our models {.unnumbered}
+## Assessment of priors, MCMC outputs and Goodness-of-fit diagnostics for our models {.unnumbered}
 
-As an informal way to diagnose model convergence, we first evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots for both models (Figure \@ref(fig:figure-pre-diag) and Figure \@ref(fig:figure-post-diag)) [@Hamra2013].
-In general, models with informative priors have a better convergence of the MCMC.
-We can see that the diagnostic plots show stationary distributions and a trend of decreased autocorrelation.
+We conducted a sensitivity analysis to understand the effect of the prior specification on our two models.
+We set all the priors to be vague normal distribution (mean is 0 with standard deviation is 10) and a range of plausible priors (mean from 0 to 5 and standard deviation from 0 to 10).
+The comparison demonstrates that our results are robust to different priors that preserve the differences between our two models, especially for density, transitivity, and centralization.
+In general, models with informative priors have a better convergence of the MCMC presented by a trend of decreased autocorrelation for some of the parameters in both models.
+However, the posterior estimates in 95% confidence intervals could vary for some covariates with different priors.
+
+As an informal way to diagnose model convergence, we evaluated three diagnostic visual summaries of our MCMC output, including density plots, trace plots, and autocorrelation plots for both models (Figure \@ref(fig:figure-pre-diag) and Figure \@ref(fig:figure-post-diag)) [@Hamra2013].
+We can see that the diagnostic plots show a trend of stationary distributions for most of the variables.
+However, the degree of autocorrelation is high at 40 lag for all variable but the dyadcov variable that indicates slow mixing and strong correlations between our samples [@Lynch2007; @Roy2020].
 Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].

### 2021-05-27 analysis/paper/paper.Rmd
edit the main text based on results of 30000 iterations with more chains

@@
 We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, the ritual element represented by pots (ritual-homophily) has a significant effect on the formation of relations between burials, while the wealth rank represented by burial values (wealth-homophily) has a negative effect but not significant since its confidence intervals include zero.
+For nodal covariates in the pre-European model, the values of ritual element represented by pots (ritual-homophily) are mostly greater than zero that suggests a positive effect on the formation of relations between burials, while the values of wealth rank represented by burial values (wealth-homophily) are mostly negative that indicates it has less influence on the tie formation.
 This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
-Despite positive mean for some covariates, such as age (age-homophily) and sex (sex-homophily), they do not show a significant tendency.
+For other covariates with values mostly around zero, such as age (age-homophily) and sex (sex-homophily), they do not show any clear tendency.
 Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
@@
 <!--posterior stats for for post-E and their comparison-->
-For the post-European network model, the nodal covariate of wealth rank shows significant positive effects.
-This indicates the burials in the same wealth level tend to have relations.
-Despite negative effects for the ritual and sex variables, they are not significant.
-Also, there are no significant effects for the variables of age and physical distance.
-Similar to the pre-European network, transitivity demonstrates a significant positive effect, but much weaker than the effect of the pre-European network.
+For the post-European network model, two nodal covariates, wealth rank and age, both show tendencies towards positive values.
+This indicates the burials in the same wealth level and age groups tend to form relations.
+However, there is no clear pattern for ritual and sex variables since most values around zero.
+Also, there are no significant effects for the variables of physical distance.
+Similar to the pre-European network, transitivity demonstrates a significant positive effect, but weaker than the effect of the pre-European network.
 In contrast, centralization has a significantly higher positive effect than the effect of the pre-European network.
@@
 Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
-Our results indicate that the major factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth.
+Our results indicate that the potential factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth and age groups.
 We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
-Before European arrival, ritual behaviors were important factors that structured the interconnections of prestige goods in the burial contexts.
-This might imply that ritual could be a status indicator at the time before many foreign goods were introduced to Indigenous societies.
+Before European arrival, ritual behaviors could be important factors that structured the interconnections of prestige goods in the burial contexts.
+This might imply that ritual as a status indicator at the time before many foreign goods were introduced to Indigenous societies.
 After the European presence was established, status indicators shifted from ritual to wealth differences, corresponding with the characteristics of corporate/network strategies.
@@
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
-Demographic information, such as age and sex, do not have any effects on network formation for both networks.
-We also note that no effects could be due to the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
+Another demographic information, sex, do not have any effects on network formation for both networks.
+We also note that their weak effects may be impacted by the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
 
@@
 After European arrival, the network has a tendency of centralization relative to the rarity of goods.
-We interpret the changes in formation mechanisms of networks, observed as a change from ties formed by ritual to wealth, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
+We interpret the changes in formation mechanisms of networks, observed as a change from tie formation dominated by ritual to wealth, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
 This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
@@
 This work was supported in part by the travel grant from the Department of Anthropology at the University of Washington, the predissertation summer travel grant from the Henry Luce Foundation/ACLS Program in China Studies, and the Doctoral Dissertation Fellowship from the Chiang Ching-kuo Foundation for International Scholarly Exchange (Project #DF009-A-18).
-We thank Ben Fitzhugh and Peter Lape for their insightful comments on an early draft.
+We thank Ben Fitzhugh, Peter Lape, Erik Gjesfjeld, Matthew Peeples for their insightful comments on an early draft.
 We also acknowledge comments from anonymous reviewers that have greatly improved this manuscript.

### 2021-06-04 analysis/paper/paper.Rmd
edit the main text

@@
 )
-
-image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
-image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
-image_width_3_col_inch = 190/19
 ```
@@
 
-<!---network analysis approaches and basic concepts-->
+<!--network analysis approaches and basic concepts-->
 Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing relational data [@Brandes2013; @Mills2017; @Peeples2019].
@@
 Recent studies on burials using network analysis demonstrates it is a promising tool to test expectations derived from hypotheses [@Sosna2013].
-Here we use network modeling to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations.
+Here we use network modeling to explore similarities and differences between burials by identifying underlying variable structuring burial relations.
 We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e.
-individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013].
+individual burials), pairwise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013].
 Our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg.
@@
 
-<!-- intro to networks: move to SI -->
-<!-- two types of network analysis: move to SI -->
+<!--intro to networks: moved to the SI-->
+<!--two types of network analysis: moved to the SI-->
 
@@
 
-<!-- current approaches to network analysis: move to SI -->
+<!--current approaches to network analysis: move to the SI-->
 
-<!-- Bayesian inference for Exponential Random Graph Models -->
-
+<!--Bayesian inference on Exponential Random Graph Models--> 
 In this paper, we used Bayesian inference on ERGMs to test our hypotheses.
-ERGMs are the tool that provides the conceptual background for network formation and structure, allowing testing social processes through covariates.
-The benefit of the Bayesian part is to deal with computational issues of ERGMs mentioned above.
-Bayesian approaches to ERGMs are effective because it incorporates prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and uncertainty [@Caimo2017; @Lehmann2020].
-Prior information can be derived from previous data or assumptions derived from the context of our data.
+ERGMs are used as the tool to provide the conceptual background for network formation and structure, allowing testing social processes through covariates.
+The benefit of the Bayesian part is to deal with computational issues of ERGMs by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and uncertainty [@Caimo2017; @Lehmann2021].
+Prior information can be derived from previous data or assumptions based on the context of data.
 An important advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011].
@@
 By fitting an ERGM with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019].
-This allows robust and interpretable uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian].
+This allows robust and interpretable uncertainty quantification by examining the posterior distribution [@Caimo2017bayesian].
 In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies.
@@
 
-<!--archaeological specific case study and our model-->
+<!--Intro to our case study and model-->
 We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries.
-This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007].
+This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally an influx of Han Chinese in the 19th century [@Chen2007].
 Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
-Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004; @Cheng2017].
-We assume that the use of imported goods and their values were increased and stressed after the European arrival.
-Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2004; @Cheng2017].
+Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2005; @Cheng2017].
+We assume that the use of imported goods and their values were amplified after the European arrival.
+Different from Heping dao and Tamsui in northern Taiwan where European forts are located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2005; @Cheng2017].
 This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that is still not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].
 
-The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
-The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan.
-Foreign trade goods could be used as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006].
-In addition, the study of spatial distribution of trade ornaments in residential area at Kiwulan demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament].
-
-<!--theoretical framework and research questions-->
+<!--theoretical framework: corporate/network strategies-->
+The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign goods [@Blanton1996; @Feinman2000; @Klehm2017].
 The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000].
 These are contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999].
+Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+
+<!--archaeological context of Kiwulan-->
+The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan.
+Foreign trade goods were believed to be used as prestige goods based on archaeological contexts at Kiwulan [@Hsieh2009; @Wang2011].
+There are very few trade goods found in middens but more in residential area and burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006].
+In addition, the study of spatial distribution of trade ornaments in Kiwulan residential area demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament].
 The European contact and associated trade activities could be the stimulus of competition for foreign resources among aggrandizers that led to social changes.
+
+<!--hypothesis and research questions-->
 Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network.
@@
 and if so, (2) what are the major variables affecting or forming these two networks?
-Can we identify the shift in variables from ritual to wealth that associated with two different strategies?
+Can we identify the shift in variables from ritual to wealth associated with two different strategies?
 By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.
@@
 
-```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep-data.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "000-prep-data.R")))
+```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/000-prep-data.R"))}
+suppressWarnings(source(here::here("analysis", "code", "000-prep-data.R")))
 ```
 
-```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
+```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/001-data-tidy.R"))}
+suppressWarnings(source(here::here("analysis", "code", "001-data-tidy.R")))
 ```
@@
 A total of 90 burials were unearthed from adjacent excavation squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-The preservation of human remains at Kiwulan was generally poor, so we have incomplete age and sex data (details in Supplementary Online Materials) and lack health status data.
+The preservation of human remains at Kiwulan was generally poor, so we have incomplete data of age and sex (details in Supplementary Online Materials) and lack health status data.
 Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses.
@@
 She found that the burials with rare prestige goods were usually associated with elders, which might indicate achieved, rather than inherited, status.
-One important limitation of these previous studies is that they did not use analytical units suitable for comparing behaviors before and after European arrival.
-Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until arrival of Han Chinese immigrants).
+One important limitation of these previous studies is that they did not use analytical units suitable for comparing behaviors from different phases.
+Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until the arrival of Han Chinese immigrants).
 
-```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-burial-location.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
+```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/002-burial-location.R"))}
+suppressWarnings(source(here::here("analysis", "code", "002-burial-location.R")))
 ```
@@
 To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
-Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament].
+Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament; @LWandBM2020pot].
 We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
@@
 @Coward2013, pp.
-252], we built networks where burials (nodes in the network) that are linked when they have the same prestige goods in common.
-The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments.
-These items are considered as high-value across different archaeological contexts throughout Taiwan based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+252], we built networks where burials (nodes in the network) are linked when they have the same trade goods in common.
+The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also less affected by taphonomic processes because of durable materials.
+These items are considered as prestige goods due to their rarity in burial contexts and descriptions in historical records [@Chen2007; @Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
 European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
@@
 In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006].
-Besides, those goods are made of durable materials that is reasonable to assume that they were less affected by taphonomic processes.
 
@@
 
-```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
+```{r custom-function, include=FALSE}
+suppressWarnings(source(here::here("analysis", "code", "999-bgof-custom-function.R")))
 ```
 
-```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-burials-post-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
+```{r bergm-pre, include=FALSE}
+suppressWarnings(source(here::here("analysis", "code", "003-burials-pre-network-modelling.R")))
+```
+
+```{r bergm-post, include=FALSE}
+suppressWarnings(source(here::here("analysis", "code", "004-burials-post-network-modelling.R")))
 ```
@@
 
-```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/005-network-diagrams-two-phases.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
+```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/005-network-diagrams-two-phases.R"))}
+suppressWarnings(source(here::here("analysis", "code", "005-network-diagrams-two-phases.R")))
 ```
@@
 ```{r bergm-prior-and-iteration}
-prior_mean <- as_tibble(cbind(pre_prior_mean, 
+prior_mean<- data.frame(cbind(pre_prior_mean, 
                               post_prior_mean)) %>% 
@@
 
-prior_sigma <- as_tibble(cbind(diag(pre_prior_sigma), 
+prior_sigma <- data.frame(cbind(diag(pre_prior_sigma), 
                                 diag(post_prior_sigma))) %>% 
-  rename(pre_prior_sigma = V1, post_prior_sigma = V2) %>% 
+  rename(pre_prior_sigma = X1, post_prior_sigma = X2) %>% 
   mutate(sigma_same = ifelse(pre_prior_sigma == post_prior_sigma, pre_prior_sigma, NA))
@@
 
-```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/006-bergm-distribution-stats.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "006-bergm-distribution-stats.R")))
+```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/006-bergm-distribution-stats.R"))}
+suppressWarnings(source(here::here("analysis", "code", "006-bergm-distribution-stats.R")))
 ```
@@
 
-```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/007-bootstrap-CI.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "007-bootstrap-CI.R")))
+```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/007-bootstrap-CI.R"))}
+suppressWarnings(source(here::here("analysis", "code", "007-bootstrap-CI.R")))
 ```

### 2021-06-05 analysis/paper/supplementary-material.Rmd
edit the bib file and the SI rmd
edit the SI

@@
 
-image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
-image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
 ```
@@
 
-<!--fit in a bigger historical picture in a pericolonial context: move to SI -->
+<!--fit in a bigger historical picture in a pericolonial context: move to SI-->
 The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
@@
 
-```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep-data.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "000-prep-data.R")))
-```
-
-```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
-```
-
-```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-burial-location.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
+```{r read-models-data, include=FALSE}
+load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
+readRDS(here::here("analysis", "data", "derived_data", "pre_bergm.rds"))
+readRDS(here::here("analysis", "data", "derived_data", "post_bergm.rds"))
 ```
@@
 
-<!-- intro to networks: move to SI -->
+<!--intro to networks: moved from the main text-->
 Drawing largely on network science and graph theory, a key assumption in the construction of archaeological networks is that a phenomenon can be conceptualized as a network through abstraction and representation [@Brandes2013; @Collar2015; @Mills2017; @Peeples2019].
@@
 
-<!-- two types of network analysis: move to SI -->
+<!--two types of network analysis: moved from the main text-->
 Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms.
@@
 
-<!-- current approaches to network analysis: move to SI -->
+<!--current approaches to network analysis: move from the main text-->
 Current approaches to network analysis used by archaeologists are mostly restricted to a single rational structure without consideration of interaction between network variables.
@@
 Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1).
-If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them.
+If burial 1 and burial 2 both have high quantities of carnelian beads, then there will be a tie connecting them.
 The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments.
@@
 Thus, we are confident that the quantities and distributions of beads in burial contexts can reflect the original differences between burials.
-Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can effectively represent structures of burial networks.
 
@@
 Table \@ref(tab:age-sex-table) shows the number of burials, of which sex and age can be determined.
-
-```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
-# combine ridge plots 1 & 4 in the 000-prep-data file
-library(tidyverse)
-library(ggridges)
-ridge_1_and_4 <-
-  burial_beads %>%
-  mutate(type = case_when(
-    type == "Golden_bead" ~"gold-foil bead",
-    type == "all_glass_bead" ~"glass bead",
-    type == "Agate_bead" ~"carnelian bead")) %>% 
-  ggplot(aes(x = value,
-             y = type,
-             fill = factor(stat(quantile)))) +
-  stat_density_ridges(
-    geom = "density_ridges_gradient",
-    calc_ecdf = TRUE,
-    quantiles = 4,
-    quantile_lines = TRUE,
-    jittered_points = TRUE,
-    alpha = 0.7,
-    vline_size = 0.5,
-    vline_color = "grey10",
-    point_size = 2.5,
-    point_alpha = 0.4,
-    position = position_raincloud(adjust_vlines = TRUE),
-    rel_min_height = 0.01,
-    scale = 0.5, # so the filled regions don't overlap on the points
-    size = 0.2) +
-  scale_x_continuous(limits = c(1, 15),
-                     expand = c(0.1, 0)) +
-  scale_fill_viridis_d(name = "Quartiles") +
-  theme_minimal()
-
-#ggsave(here::here("analysis", "figures", "000-raincloud-beads.png"),
-       #w = 6, h = 4)
-```
+Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can effectively represent structures of burial networks.
 
@@
 
-```{r figure-raincloud-beads, fig.cap = "The distributions of different types of trade beads across burials analyzed in this paper."}
-
-knitr::include_graphics(here::here("analysis", "figures", "000-raincloud-beads.png"))
-```
-
 ## Further details on methods of modeling fitting {.unnumbered}
@@
 ```{r bergm-iteration}
-burn_in <- 100
-main_iter <- 1000
-aux_iter <- 4000
+burn_in <- 1000
+main_iter <- 40000
+aux_iter <- 5000
 bgof_si <-10000
@@
 <!-- assessment of MCMC and Goodness-of-fit in Bayesian: move to SI -->
-
-```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
-```
-
-```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-burials-post-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
-```
 

### 2021-06-09 analysis/paper/paper.Rmd
finished editing for reviews

@@
 
-# Exponential random graph models in a Bayesian framework {-}
+## Exponential random graph models in a Bayesian framework {-}
 
@@
 
-# Background {-}
-
-## A Case Study from Northeastern Taiwan {-}
+# Case Study: Exploring Social Changes in Northeastern Taiwan {-}
 
@@
 This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally an influx of Han Chinese in the 19th century [@Chen2007].
-Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994].
+Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals (details in Supplementary Online Materials) [@Brumfiel1994; @Clark1994].
 Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2005; @Cheng2017].
@@
 <!--theoretical framework: corporate/network strategies-->
-The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign goods [@Blanton1996; @Feinman2000; @Klehm2017].
-The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000].
-These are contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999].
-Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+The involvement in long-distance exchange is often associated with an increased social inequality where ambitious individuals could gain power by building personal influence through the acquisition and distribution of foreign goods, leading stoical change from corporate based to network based societies [@Blanton1996; @Feinman2000; @Klehm2017].
+Corporate-network societies represent two distinct political-economic strategies for obtaining power, viewed as a horizontal axis crosscutting traditional vertical hierarchy [@Drennan2010; @Feinman2000].
+The features of network strategies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks.
+These are contrasted with corporate strategies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999].
+In Asia, network-based societies are identified, especially for places involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 
@@
 <!--hypothesis and research questions-->
-Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network.
+Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network [@Blanton1996; @Drennan2010; @Feinman2000].
+We explore whether this change can be identified through the pattern of intercorrelations of foreign prestige goods among burials.
 We ask: (1) Do the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact?
@@
 
-## Materials {-}
+# Materials {-}
 
@@
 
-<!-- Our chronology and principles of network building -->
+<!--assigning burials based on our chronology-->
 To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`).
@@
 Differences in network size were evaluated using vertex bootstrap in a later section.
+By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
 
-Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf.
+## Contextualizing Burial Networks {-}
+
+<!--the context of burial networks-->
+We built networks where burials (nodes in the network) are linked when they have the same trade goods in common.
+This is based on the assumptions that social status and associated relations can be represented by sharing similar goods and burials can reflect social structures [@Binford1971; @Saxe1970; cf.
 @Coward2013, pp.
-252], we built networks where burials (nodes in the network) are linked when they have the same trade goods in common.
-The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also less affected by taphonomic processes because of durable materials.
-These items are considered as prestige goods due to their rarity in burial contexts and descriptions in historical records [@Chen2007; @Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
+252].
+Thus, individuals with the same prestige goods reflect similar access to trade, exchange, and gifting networks, indicative of the similarity among Kiwulan residents and social inequality.
+The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also items less affected by taphonomic processes because of durable materials.
+Those goods are considered as prestige goods due to their rarity in burial contexts and descriptions in historical records [@Chen2007; @Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
 European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006].
@@
 
-```{r bergm-pre, include=FALSE}
+```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/003-burials-post-network-modelling.R"))}
 suppressWarnings(source(here::here("analysis", "code", "003-burials-pre-network-modelling.R")))
@@
 
-```{r bergm-post, include=FALSE}
+```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/004-burials-post-network-modelling.R"))}
 suppressWarnings(source(here::here("analysis", "code", "004-burials-post-network-modelling.R")))
@@
 
-## Hypothesis and construction of networks {-}
+## Hypothesis and prediction {-}
 
-<!-- our expectation based on hypothesis --> 
+<!-- archaeological predictions --> 
 Our social change hypothesis is that differential access to foreign prestige goods after the European presence led to increased social inequality at Kiwulan, where the social structure changed from a more corporate strategy to a more network strategy, presenting different organizational modes of societies [@Blanton1996; @Drennan2010; @Feinman2000].
-Corporate-based societies stress shared power, multiple segments, and social cohesion through communal rituals, while network-based societies emphasize the accumulation of individual prestige, wealth, and resources by building personal networks.
-We used burial data as a proxy to explore social changes from a corporate strategy to a network strategy.
-This change can be identified though the pattern of intercorrelations of foreign prestige goods based on assumptions that burials can represent social structures [@Binford1971; @Saxe1970], and individuals with similar prestige goods reflect similar access to trade, exchange, and gifting networks [eg.
-@Coward2013, pp.
-252].
-
 If social inequality gradually increased after European arrival as we hypothesize, then we expect to observe a network structure with higher centralization and less transitivity, along with network ties determined by wealth, indicative of a network-based society.
@@
 To test this hypothesis, we use Bayesian ERGMs to model the formation of network ties and the underlying mechanisms that shaped relationships between people at Kiwulan.
-By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
 
@@
 
+## Construction of Network Variables {-}
+
 <!-- burial data and the definition of tie formation  --> 
-Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (Figure \@ref(fig:raincloud-beads); details in Supplementary Online Materials).
-For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they both possess each type of goods (i.e.
+Trade beads are found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (Figure \@ref(fig:raincloud-beads); details in Supplementary Online Materials).
+For less frequent prestige goods, including imported ceramics, gold foils, fish-shaped ornaments, and coins, we linked two burials when they both possess each type of goods (i.e.
 presence or absence).
 Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index.
-Ritual pottery was identified as locally-made ceramics placed above an individual.
-These pots mostly have sooty vessel exteriors and charred residues on interiors that are interpreted as vessels used for funeral feasting [@Hsieh2009].
+Ritual pottery was identified as locally-made ceramics placed above an individual, which is interpreted as vessels used for funeral feasting or offering [@Hsieh2009].
 We calculated our burial value by summing values of all types of prestige goods in each burial context.
-For example, for gold-foil beads, we take the total number of burials (90) and divide by number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96.
+For example, for gold-foil beads, we take the total number of burials (90) and divide by the number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96.
 We repeat this for each type of prestige item.
@@
 
-We then assigned the burials into three ranks according to breaks in the distribution of burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth (details in Supplementary Online Materials).
+We then assigned the burials into three ranks according to breaks in the distribution of burial values, high (>`r high_value`, top 10 percent), above average (`r ave_value`-`r high_value`), and below average (<`r ave_value`), as an index of wealth.
 Since burials tend to have multiple goods in common, the network ties are weighted instead of binary (the value 1 represents a tie and the value 0 otherwise) [@Snijders2011].
 For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2.
+We are able to weigh ties due to durable materials of trade goods that are preserved well in each burial.
 Our networks are non-directed, which means a mutual relationship where the tie between any two actors is bidirectional.
@@
 ```{r figure-network-diagrams, fig.cap = "A: Burial network before the European arrival (14th-16th centuries), B: Burial network after the European arrival (17th-18th centuries).
+Nodes represent burials, tied when they share the same trade goods.
 The size of each node is proportional to node degrees or the number of connections to a node.
 The thickness of ties represents the number of goods in common between two nodes.
-Isolated nodes are not displayed in the plots"}
+Isolated nodes are not displayed in the plots."}
 knitr::include_graphics(here::here("analysis", "figures", "005-network-diagrams.png"))
@@
 For example, if age-homophily is important here, then people of the same age should have the same burial goods.
-We also include the physical distance between burials as an indicator of a kinship-based relations since the deceased from the same family were buried nearby [@LiandWu2006].
+We also include the physical distance between burials as an indicator of a kinship-based relations that we assume the deceased from the same family were buried nearby [@Chen2007].
 Our model may reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others, with more wealth differentiation after the European arrival.
@@
 
-To evaluate our anthropological model of increased social inequality over time based on assumptions of corporate-network strategies [@Drennan2010; @Feinman2000], we specify priors that are meaningful for social inequality, especially for the variables of transitivity and centralization.
+To evaluate our anthropological model of increased social inequality over time to indicate a shift from a corporate to a network society [@Drennan2010; @Feinman2000], we specify priors that are meaningful for social change, especially for the variables of transitivity and centralization.
 We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and emphasize the ritual element shared in corporate groups.
@@
 For covariates about biological features, such as age, sex, we specified a vague prior (N(`r priors$mean_same[2]`, `r priors$sigma_same[2]`)) for both models to explore their effects.
-For physical distance between burials, we also set a vague prior (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) to explore whether there is kinship-based proximity, e.g.
+Similarly, we used the same prior (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) for physical distance between burials, to explore whether there is kinship-based proximity, e.g.
 stronger correlations for shorter distances.
@@
 
+```{r bgof-assessment, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/007-bgof-assessment-vis.R"))}
+suppressWarnings(source(here::here("analysis", "code", "007-bgof-assessment-vis.R")))
+```
+
 <!--posterior stats for pre-E-->
@@
 For other covariates with values mostly around zero, such as age (age-homophily) and sex (sex-homophily), they do not show any clear tendency.
-Similarly, the dyadic covariate, physical distance, shows no significant effect, indicating that physical proximity between burials does not reflect similarity in burial treatment.
+For the physical distance variable (dyadic covariate), it shows that most values are positive, indicating that physical proximity between burials could reflect similarity in burial treatment.
 For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect.
@@
 Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017).
-Posterior means are generally close to posterior medians.")
+Some posterior means close to posterior medians that indicate symmetric distributions.")
 ft
@@
 ```{r figure-posterior-distribution, fig.cap = "Posterior density estimates for the parameters associated with edges, transitivity, centralization, and physical distance by phases.
-The pre-European group presents remarkably larger values for the transitivity, but lower values for the edges and centralization parameters compared to the post-European group.
-The distance parameters overlap between two groups.
-Both posterior estimates present generally symmetric distributions."}
+The pre-European group presents a long tailed distribution with larger values for the transitivity, but negative values for the centralization parameter compared to the post-European group.
+The edges and distance parameters for the pre-European group have larger values generally compared to pos-European one."}
 knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
@@
 However, there is no clear pattern for ritual and sex variables since most values around zero.
-Also, there are no significant effects for the variables of physical distance.
+Also, there are no effects for the variable of physical distance.
 Similar to the pre-European network, transitivity demonstrates a significant positive effect, but weaker than the effect of the pre-European network.
@@
 This implies that the presence of better access to trade goods and the behavior of wealth accumulation and display in burial events.
-In general, the post-European network model has a smaller transitivity effect and a positive centralization effect than the pre-European network model.
+In general, the post-European network model has a smaller transitivity effect and higher centralization effect than the pre-European network model.
 This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence.
 
-```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/007-bootstrap-CI.R"))}
-suppressWarnings(source(here::here("analysis", "code", "007-bootstrap-CI.R")))
+```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/008-bootstrap-CI.R"))}
+suppressWarnings(source(here::here("analysis", "code", "008-bootstrap-CI.R")))
 ```
@@
 We explored the sample size effect by removing nodes at certain percentages (5-40%) for both networks and comparing their confidence intervals.
-The results demonstrate a consistent difference between the two networks up to 30% node removals as shown in Figure \@ref(fig:figure-bootstrap-CI).
+The results demonstrate a consistent difference between the two networks up to 35% node removals for the centralization variable as shown in Figure \@ref(fig:figure-bootstrap-CI).
+This shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
+Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
 This indicates that the network variables, including density, centralization, and transitivity, are robust to the different network size of our sample.
-Figure \@ref(fig:figure-bootstrap-CI) also shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
-Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
 
@@
 Networks with nonoverlapping confidence intervals indicate a significant difference between the two phases for that network variable."}
-knitr::include_graphics(here::here("analysis", "figures", "007-bootstrap-CI.png"))
+knitr::include_graphics(here::here("analysis", "figures", "008-bootstrap-CI.png"))
 ```
@@
 ```{r figure-distribution-moments, fig.cap = "Distribution moments (mean, skewness, and variance) calculated on the observed data and simulated distributions for the pre- and post-European models."}
-knitr::include_graphics(here::here("analysis", "figures", "006-distribution-moments.png"))
+knitr::include_graphics(here::here("analysis", "figures", "007-distribution-moments.png"))
 ```
@@
 A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups toward a more centralized network with concentration of connections among fewer burials after the European arrival in the 17th century.
-This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the grave goods and burial attributes.
+This supports our pericolonial impact model that Kiwulan changed from a more corporate-based to a more networked-based society, as indicated in the burial goods and burial attributes.
 Changes in mortuary practices could reflect the development of social or political complexity [eg.
@@
 Network covariates give insights into the socio-cultural factors that were associated with the acquisition and distribution of foreign prestige goods.
-Our results indicate that the potential factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth and age groups.
+The general pattern of posterior distributions indicates that the main factor contributing to burial networks for the pre-European network was ritual, and for the post-European network was wealth.
 We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
@@
 After the European presence was established, status indicators shifted from ritual to wealth differences, corresponding with the characteristics of corporate/network strategies.
-Geographical proximity of burials, implying kinship association, has no effect on sharing similar prestige items between burials for both phases.
-This may suggest that the use of prestige items is more individual-oriented instead of kinship-based.
-Another demographic information, sex, do not have any effects on network formation for both networks.
-We also note that their weak effects may be impacted by the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
+In addition, geographical proximity of burials, implying kinship association, has some effect on sharing similar prestige items between burials in the pre-European period, but not in the post-European period.
+This may suggest that the use of prestige items is more individual-oriented instead of kinship-based after European arrival.
+Demographic variable, age, also indicates that burials in the same age may tend to have similar goods after European arrival.
+However, another demographic varialbe, sex, do not have any effects on network formation for both networks.
+We should also note that their effects may be impacted by the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
 
@@
 This is consistent with the result of a more centralized burial network after the presence of Europeans, that supports a connection between a burial and the social role of the living person.
-Also, ethnographic records support our interpretation of burial goods as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], so the interrelations represented by the flow of these goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
+Also, ethnographic records correspond to our interpretation of burial goods as prestige goods or status items in the local Indigenous culture [@Borao2009; @LiandWu2006], so the interrelations represented by the flow of these goods observed in the burial data likely reflect social contexts of the living people at Kiwulan.
 
 <!-- trade dynamic during this period and future work for comparison -->
-We also recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have had contributed to social changes at Kiwulan.
+We recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have contributed to social changes at Kiwulan as well.
 However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival.
-Historical records also support the interaction and trade activities between Europeans and local Indigenous people.
-The trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at Kiwulan were frequently found from archaeological sites with Dutch occupation, such as Fort Zeelandia.
-As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides an ideal example to explore the reaction of Indigenous people to European colonial presence at a settlement scale.
-We still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence from the periphery with colonial bases, such as Heping Dao and Tamsui.
-Unfortunately, Indigenous sites related to the 17th century with intensive excavation and sufficient evidence as Kiwulan are rare to offer good comparison and future work is still needed.
+The frequent trade activities stimulated by the presence of Europeans led to many trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at archaeological sites with Dutch occupation at different parts of Taiwan, such as Fort Zeelandia.
+As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale.
+However, we still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui.
+Unfortunately, Indigenous sites related to European contact with intensive excavation and sufficient evidence are rare to offer suitable comparison with Kiwulan.
+Future work is still needed.
 
@@
 This case study demonstrates the methodological benefits of Bayesian inference on ERGMs to inform and enhance studies of relational data in archaeology.
-A Bayesian framework can reduce the effects of small sample size or missing data commonly present in archaeological data and incorporating prior information and MCMC estimations to quantify uncertainty and provide interpretable output.
-An important feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological models, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
+A Bayesian framework can better estimate archaeological data by incorporating prior information with MCMC estimations to quantify uncertainty and provide interpretable output.
+An important feature of Bayesian network modeling is the ability to set prior values to evaluate competing anthropological hypotheses, enabling a comparison of anthropologically-relevant models, rather than testing a hypothesis of no difference or randomness.
 Bayesian network modeling can be applied to a wide range of archaeological data to examine the formation of relationships using robust probabilistic inference.
@@
 We tested a hypothesis of changes in burial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects.
-The results support our model that the relationship between burials changed after the European colonization period in the 17th century.
+The results support our model that the intercorrelation between burials changed after the European colonization period in the 17th century.
 Before the arrival of Europeans, the burial network has a tendency of more clustered subgroups with pottery as ritual practice as the key formation mechanism.
@@
 We interpret the changes in formation mechanisms of networks, observed as a change from tie formation dominated by ritual to wealth, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
-This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
+This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power (Feinman, 2000).
 Using burial data with historical documents, we can better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.

### 2021-06-10 analysis/paper/paper.Rmd
minor edits to the main text

@@
 <!-- ERGMs in archaeology--> 
-Exponential random graph models (ERGMs) are statistical models for social networks that first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites and at 190 sites for a later study [@Brughmans2015].
+Exponential random graph models (ERGMs) are statistical models for social networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+The application of ERGMs first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites and at 190 sites for a later study [@Brughmans2015].
 By fitting models with the observed network of archaeological data, they proposed that ERGMs are a promising approach for exploration and hypothesis testing of social processes.
@@
 stronger correlations for shorter distances.
+We then examine the fit of the observed data to the two models we specified to evaluate our hypotheses.
 

### 2021-06-10 analysis/paper/supplementary-material.Rmd
finished edits to the SI

@@
 -  further details on our methods of modeling fitting, and
--  assessment of our MCMC output and Goodness-of-fit diagnostics for our models
+-  assessment of priors, MCMC outputs and Goodness-of-fit diagnostics for our models
 
@@
 The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017].
-Increased social inequality associated with the use of exotic trade goods is recognized as a ‘network’ strategy for gaining power [@Blanton1996; @Feinman2000].
+Increased social inequality associated with the use of exotic trade goods is recognized as a "network" strategy for gaining power [@Feinman2000].
 Northeastern Taiwan became more involved in regional exchange with East Asia since the 14th century and long-distance trade with Southeast Asia brought by Europeans in the 17th century [@Chen2005; @Wang2007].
@@
 In the annual event held by the Dutch with participants of local Indigenous representatives [@Andrade1997], the Dutch would give those representatives symbolic items such as rattan sticks or foreign goods to assert their political authority and consolidate their relationship with Indigenous settlements.
-The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for power [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
+The introduction of foreign prestige goods to local Indigenous communities might have occurred before European presence became established, but was maintained and amplified after European arrival due to local competition for prestige accumulation [@Borao2009; @LiandWu2006; @LWandBM2020ornament; @Ueda2016].
 
@@
 Such a distribution consists of a large number of possible networks that enables statistical inference and comparison with an observed network [@Robins2007].
-ERGMs helps us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
+ERGMs help us understand whether an observed network shows significantly more or less of a property of interest than the random networks generated from our model assumptions.
 
@@
 <!-- burial data and the definition of tie formation: partially move to SI  -->
-For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads), also in the main text).
+For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure is in the main text).
 Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1).
@@
 The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments.
-In addition, each burial has a clear boundary as our analytical unit that reduces sampling errors.
+In addition, each burial has a clear boundary as a single analytical unit that reduces sampling errors.
 Thus, we are confident that the quantities and distributions of beads in burial contexts can reflect the original differences between burials.
@@
 library(flextable)
+library(tidyverse)
 age_sex_by_phase <-
@@
 aux_iter <- 5000
-bgof_si <-10000
+bgof_net <-100
+bgof_si <- 5000
+chains <- 32
 ```
@@
 MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013].
-We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on the approximate exchange algorithm [@Caimo2011].
+We obtained a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on the approximate exchange algorithm [@Caimo2011].
 This enables a random or stochastic simulation in a long run where each move does not depend on the previous move.
-More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence.
+More chains can ensure a more desirable posterior distribution that is close to the target distribution under study, or convergence.
 In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pairs of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011].
-In our case, we set the number of chains to six.
-For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`.
-We set the number of iterations used to simulate a network y' at each iteration to `r aux_iter`.
+In our case, we set the number of chains to `r chains`.
+For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was $`r main_iter`$.
+We set the number of iterations used to simulate a network at each iteration to `r aux_iter`.
 
@@
 The comparison demonstrates that our results are robust to different priors that preserve the differences between our two models, especially for density, transitivity, and centralization.
-In general, models with informative priors have a better convergence of the MCMC presented by a trend of decreased autocorrelation for some of the parameters in both models.
 However, the posterior estimates in 95% confidence intervals could vary for some covariates with different priors.
+In general, models with our informative priors have a better convergence of the MCMC presented by the diagnostic plots.
 
@@
 We can see that the diagnostic plots show a trend of stationary distributions for most of the variables.
-However, the degree of autocorrelation is high at 40 lag for all variable but the dyadcov variable that indicates slow mixing and strong correlations between our samples [@Lynch2007; @Roy2020].
+However, the degree of autocorrelation is high at 40 lag for all variables but the dyadcov variable, which indicates slow mixing and strong correlations between our variables [@Lynch2007; @Roy2020].
 Then we summarize output from our two models by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distributions of the parameters of each model [@Caimo2011; @Caimo2017].
-We set $`r bgof_si`$ network graphs simulated from the estimated posterior distribution in ERGMs.
+We set `r bgof_net` network graphs simulated from the estimated posterior distribution in ERGMs, and `r aux_iter` iterations used for network simulation.
 
@@
 Our Bayesian GOF diagnostics summarized three distributions, including degree, minimum geodesic distance, and edgewise shared partner distributions.
-The diagnostics plots (Figure \@ref(fig:figure-bgof-two-phases)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
-the number of edges between node pairs in a shortest path, @Hunter2008) and the degree distribution.
-For edgewise shared partner distribution, despite some observations falling outside the 95% interval, the fit is generally good with most observations within it.
+The diagnostics plots (Figure \@ref(fig:figure-bgof-pre-E) and Figure \@ref(fig:figure-bgof-post-E)) demonstrate that both models fit the observed networks very well for the minimum geodesic distance distribution (i.e.
+the number of edges between node pairs in a shortest path, @Hunter2008).
+The pre-European model has a better fit for the degree distribution and edgewise shared partner distribution, compared to the post-European model.
+Despite some observations falling outside the 95% interval for the post-European model, the fit is generally good with most observations within it.
 This provides a statistical approach to check how well the estimated posterior parameter distribution, based on our hypotheses, can reproduce networks with similar general structural features of the observed networks.
-We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models.
-We expect to see the models fit with our hypotheses, indicative of an increased social inequality after the foreign contact.
-In addition, the covariates can give some more clues for the underlying mechanisms for the formations of each network, such as the relative importance of age and sex.
+We then compared the distribution of our observed networks, the networks before and after the arrival of Europeans, with the distribution of our hypothesized models to test an increased social inequality after the foreign contact.
 
-```{r figure-bgof-two-phases, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
+```{r figure-bgof-pre-E, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
 Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
-Red lines represent distributions of observed networks, and gray lines show the 95% intervals."}
-library(magick)
-pre_bgof_img <- image_read(here::here("analysis/figures/003-pre-bgof.png"))
-post_bgof_img <- image_read(here::here("analysis/figures/004-post-bgof.png"))
-image_append(c(pre_bgof_img, post_bgof_img), stack = T)
+Red lines represent distributions of observed networks, and black lines show the 95% intervals."}
+knitr::include_graphics(here::here("analysis/figures/003-pre-bgof.png"))
+```
+
+```{r figure-bgof-post-E, fig.cap = "Goodness-of-fit diagnostics for the post-European model (below).
+Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
+Red lines represent distributions of observed networks, and black lines show the 95% intervals."}
+knitr::include_graphics(here::here("analysis/figures/004-post-bgof.png"))
 ```

### 2021-06-11 analysis/paper/supplementary-material.Rmd
correct the caption of bgof figures

@@
 
-```{r figure-bgof-pre-E, fig.cap = "Goodness-of-fit diagnostics for the pre-European model (top) and the post-European model (below).
+```{r figure-bgof-pre-E, fig.cap = "Goodness-of-fit diagnostics for the pre-European model and the post-European model (below).
 Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.
@@
 
-```{r figure-bgof-post-E, fig.cap = "Goodness-of-fit diagnostics for the post-European model (below).
+```{r figure-bgof-post-E, fig.cap = "Goodness-of-fit diagnostics for the post-European model.
 Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.

### 2021-06-21 analysis/paper/paper.Rmd
edits to the text after peer review and revisions

@@
 @Sosna2013].
-This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013].
+In this case, network ties do not indicate exchange or transmission but rather similarities.
+This is because burial goods, especially high value ones, are normally cultural symbols of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013].
 Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.
@@
 <!--theoretical framework: corporate/network strategies-->
-The involvement in long-distance exchange is often associated with an increased social inequality where ambitious individuals could gain power by building personal influence through the acquisition and distribution of foreign goods, leading stoical change from corporate based to network based societies [@Blanton1996; @Feinman2000; @Klehm2017].
+The involvement in long-distance exchange is often associated with an increased social inequality where ambitious individuals could gain power by building personal influence through the acquisition and distribution of foreign goods, leading to social change from corporate to network societies [@Blanton1996; @Feinman2000; @Klehm2017].
 Corporate-network societies represent two distinct political-economic strategies for obtaining power, viewed as a horizontal axis crosscutting traditional vertical hierarchy [@Drennan2010; @Feinman2000].
@@
 These are contrasted with corporate strategies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999].
-In Asia, network-based societies are identified, especially for places involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
+In Asia, network-based societies are frequently identified, especially for places involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].
 
 <!--archaeological context of Kiwulan-->
-The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan.
+The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade networks stimulated by European colonial activities in northern Taiwan.
 Foreign trade goods were believed to be used as prestige goods based on archaeological contexts at Kiwulan [@Hsieh2009; @Wang2011].
-There are very few trade goods found in middens but more in residential area and burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006].
+There are very few trade goods found in middens but more in the residential area and burials.
+According to Spanish historical records they are a symbol of personal influence [@LiandWu2006].
 In addition, the study of spatial distribution of trade ornaments in Kiwulan residential area demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament].
@@
 We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
+For the two periods studied here, we evaluate differences in network size were using a vertex bootstrap method.
 We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology.
-Differences in network size were evaluated using vertex bootstrap in a later section.
+Most burials have clear stratigraphic contexts so modern or ancient disturbance is readily visible (i.e.
+if a burial has been cut into by later burials, very few cases of this have been documented at Kiwulan).
 By comparing networks from the pre-European period and the post-European period we can examine the effects of foreign contact on community relationships at Kiwulan.
@@
 Thus, individuals with the same prestige goods reflect similar access to trade, exchange, and gifting networks, indicative of the similarity among Kiwulan residents and social inequality.
-The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also items less affected by taphonomic processes because of durable materials.
+The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also items less affected by taphonomic processes because of their durable materials.
 Those goods are considered as prestige goods due to their rarity in burial contexts and descriptions in historical records [@Chen2007; @Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011].
@@
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
-Priors are the distinctive aspect of Bayesian methods to estimate population by combining one's beliefs with data [@Gill2014].
-Normal distribution N($\mu$, $\sigma^2$), where $\mu$ is mean and $\sigma^2$ is standard deviation, is typical for prior specification to model network parameters because it is sensible and convenient, allowing adjustments for a wide range of prior information [@Caimo2017; @Normand1992].
+Priors are a distinctive aspect of Bayesian methods that enable us combining our beliefs based on prior research with the newly observed data [@Gill2014].
+The normal distribution N($\mu$, $\sigma^2$), where $\mu$ is the mean and $\sigma^2$ is the standard deviation, is typically used for priors when modelling network parameters because it is sensible and convenient, allowing adjustments for a wide range of prior information [@Caimo2017; @Normand1992].
 The larger the mean of the prior distribution, the more positive effect of that parameter on network formation.
-For standard deviation, if it is larger, it presents a relatively vague prior enabling comparison between variables that we do not have specific assumptions.
-Since low density is commonly found in the real world [@Caimo2017], we specified the prior of the edge density parameter to low for both network models, following a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
+A larger standard deviation makes the prior more vague, enabling comparison between variables for which we do not have specific assumptions.
+Since low network densities are commonly found in the real world [@Caimo2017], we specified the prior of the edge density parameter to low for both network models, following a normal distribution with mean at `r priors$mean_same[1]`, and standard deviations at `r priors$sigma_same[1]` (i.e (N(`r priors$mean_same[1]`, `r priors$sigma_same[1]`))).
 
-To evaluate our anthropological model of increased social inequality over time to indicate a shift from a corporate to a network society [@Drennan2010; @Feinman2000], we specify priors that are meaningful for social change, especially for the variables of transitivity and centralization.
-We set the priors to higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)) for the network before European contact to indicate less social differentiation and emphasize the ritual element shared in corporate groups.
-Conversely, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and a higher covariate effect for burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)) to model an increased social differentiation after European contact.
+To evaluate our anthropological model of increased social inequality, specifically a shift from a corporate to network society [@Drennan2010; @Feinman2000], we specified priors that are meaningful for social inequality, especially for the variables of transitivity and centralization.
+For the network before European contact, we set the priors to model less social differentiation and emphasize the ritual element shared in corporate groups: higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and a higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)).
+Conversely, to model an increased social differentiation after European contact, we set the priors for the network after European arrival to lower transitivity (N(`r priors$post_prior_mean[6]`, `r priors$post_prior_sigma[6]`)), higher centralization (N(`r priors$post_prior_mean[7]`, `r priors$post_prior_sigma[7]`)), and a higher covariate effect for burial values (N(`r priors$post_prior_mean[5]`, `r priors$post_prior_sigma[5]`)).
 For covariates about biological features, such as age, sex, we specified a vague prior (N(`r priors$mean_same[2]`, `r priors$sigma_same[2]`)) for both models to explore their effects.
-Similarly, we used the same prior (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) for physical distance between burials, to explore whether there is kinship-based proximity, e.g.
+Similarly, we used the same vague prior (N(`r priors$mean_same[8]`, `r priors$sigma_same[8]`)) for physical distance between burials, to explore whether there is kinship-based proximity, e.g.
 stronger correlations for shorter distances.
@@
 The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at http://doi.org/10.17605/OSF.IO/XGA6N to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
-Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here and additional details of definition of our approach to tie formation, model fitting, assessment of MCMC output and goodness-of-fit diagnostics for our models.
+Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here and additional details of definition of our approach to tie formation, model fitting, MCMC run details, assessment of MCMC convergence and output, and goodness-of-fit diagnostics for our models.
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 <!--posterior stats for pre-E-->
-We examine the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
-For nodal covariates in the pre-European model, the values of ritual element represented by pots (ritual-homophily) are mostly greater than zero that suggests a positive effect on the formation of relations between burials, while the values of wealth rank represented by burial values (wealth-homophily) are mostly negative that indicates it has less influence on the tie formation.
+We examined the estimates from the posterior distributions to compare their differences in structure of the simulated networks (Table \@ref(tab:table-posterior-stats); Figure \@ref(fig:figure-posterior-distribution)).
+For nodal covariates in the pre-European model, the values of ritual element represented by pots (ritual-homophily) are mostly greater than zero.
+This suggests a positive effect on the formation of relations between burials, while the values of wealth rank represented by burial values (wealth-homophily) are mostly negative, indicating it has less influence on the tie formation.
 This demonstrates that burials with ritual pottery tend to form relations, but burials in the same wealth level tend not to.
-For other covariates with values mostly around zero, such as age (age-homophily) and sex (sex-homophily), they do not show any clear tendency.
-For the physical distance variable (dyadic covariate), it shows that most values are positive, indicating that physical proximity between burials could reflect similarity in burial treatment.
+For other covariates with values mostly around zero, such as age (age-homophily) and sex (sex-homophily), we do not see any clear tendency.
+Most values are positive for the physical distance variable (dyadic covariate), indicating that physical proximity between burials influences similarity in burial treatment.
 For the endogenous network effects, transitivity presents a significant positive effect, while centralization demonstrates a negative effect.
@@
 Confidence intervals that do not include zero indicate significant effects of parameters (Caimo, 2017).
-Some posterior means close to posterior medians that indicate symmetric distributions.")
+Posterior means close to posterior medians that indicate symmetric distributions.")
 ft
@@
 <!--posterior stats for for post-E and their comparison-->
-For the post-European network model, two nodal covariates, wealth rank and age, both show tendencies towards positive values.
+For the post-European network model, two nodal covariates, wealth rank and age, both show tendencies towards positive values in the posterior distributions.
 This indicates the burials in the same wealth level and age groups tend to form relations.
-However, there is no clear pattern for ritual and sex variables since most values around zero.
-Also, there are no effects for the variable of physical distance.
+However, there is no clear pattern for ritual and sex variables since most values are around zero.
+There are also no effects for the variable of physical distance.
 Similar to the pre-European network, transitivity demonstrates a significant positive effect, but weaker than the effect of the pre-European network.
@@
 One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials.
-To understand the robustness of comparison between two networks, we used a vertex bootstrap technique to cross-validate the results of our Bayesian ERGMs.
+To confirm the robustness of our comparison between these two different-sized networks, we used a vertex bootstrap technique to cross-validate the results of our Bayesian ERGMs.
 The vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e.
-node) to quantify standard errors and estimate sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
+nodes) to quantify standard errors and estimate sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021].
 This enables the evaluation of uncertainty for networks and tests the difference between multiple networks of different sizes by examining their confidence intervals for the network population.
@@
 Similarly consistent with our earlier result, the vertex bootstrap shows no significant difference on density and transitivity.
-This indicates that the network variables, including density, centralization, and transitivity, are robust to the different network size of our sample.
+This indicates that the network variables, including density, centralization, and transitivity, are robust to the different network sizes of our sample.
 
@@
 We interpret these results as illustrating different social and economic mechanisms influencing relations between burials for the two periods.
-Before European arrival, ritual behaviors could be important factors that structured the interconnections of prestige goods in the burial contexts.
-This might imply that ritual as a status indicator at the time before many foreign goods were introduced to Indigenous societies.
+Before European arrival, ritual behaviors could have been important factors that structured the interconnections of prestige goods in the burial contexts.
+This might imply that ritual was status indicator at the time before many foreign goods were introduced to Indigenous societies.
 After the European presence was established, status indicators shifted from ritual to wealth differences, corresponding with the characteristics of corporate/network strategies.
@@
 This may suggest that the use of prestige items is more individual-oriented instead of kinship-based after European arrival.
-Demographic variable, age, also indicates that burials in the same age may tend to have similar goods after European arrival.
-However, another demographic varialbe, sex, do not have any effects on network formation for both networks.
+Burials in the same age group tend to have similar goods after European arrival.
+However the sex of the buried individual did not have any effects on network formation for both networks.
 We should also note that their effects may be impacted by the unavailability of age (`r age_miss` out of `r pre_and_post` burials) and sex (`r sex_miss` out of `r pre_and_post` burials) values in our sample.
@@
 <!-- trade dynamic during this period and future work for comparison -->
-We recognize that trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have contributed to social changes at Kiwulan as well.
+We recognize that complex local and international trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have contributed to social changes at Kiwulan as well.
 However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival.
@@
 As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale.
-However, we still need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui.
-Unfortunately, Indigenous sites related to European contact with intensive excavation and sufficient evidence are rare to offer suitable comparison with Kiwulan.
-Future work is still needed.
+Was its experience with the European presence unique or typical of the region?
+We need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui.
+Future work should focus on collecting data from Indigenous sites related to European contact for comparison with Kiwulan.
 

### 2021-06-22 analysis/paper/paper.Rmd
replace code for caching the models with code that reads their outputs

@@
 
-```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/003-burials-post-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "code", "003-burials-pre-network-modelling.R")))
-```
-
-```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/004-burials-post-network-modelling.R"))}
-suppressWarnings(source(here::here("analysis", "code", "004-burials-post-network-modelling.R")))
+```{r read-models-data, include=FALSE}
+load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
+readRDS(here::here("analysis", "data", "derived_data", "pre_bergm.rds"))
+readRDS(here::here("analysis", "data", "derived_data", "post_bergm.rds"))
 ```

### 2021-08-21 analysis/paper/paper.Rmd
knit and resubmit to JAS
edit the proof after acceptance

@@
 ---
-title: "A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period"
+title: "A Bayesian networks approach to infer social changes from burials in northeastern Taiwan during the European colonization period"
 author:
@@
 keywords: |
-  Bayesian modeling; network analysis; Monte-carlo simulation; Burials; European colonialism; social differentiation; Taiwan
+  Bayesian modeling; network analysis; Monte-carlo simulation; Burials; European colonialism; social inequality; Taiwan
 highlights: |
@@
 <!-- ERGMs in archaeology--> 
-Exponential random graph models (ERGMs) are statistical models for social networks that allows direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
+Exponential random graph models (ERGMs) are statistical models for social networks that allow direct modeling for the formation of edges, or ties, between nodes [@Robins2007].
 The application of ERGMs first appeared in archaeology with @Brughmans2014, who studied Iron age settlement patterns in Southern Spain by modeling inter-settlement visibility networks and visual control at 159 sites and at 190 sites for a later study [@Brughmans2015].
@@
 We assume that the use of imported goods and their values were amplified after the European arrival.
-Different from Heping dao and Tamsui in northern Taiwan where European forts are located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2005; @Cheng2017].
-This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that is still not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].
+Different from Heping dao and Tamsui in northern Taiwan where European forts are located, the remote geographical setting of northeastern Taiwan resulted in less interference from European colonial powers, but more frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2005; @Cheng2017].
+This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that remains not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].
 
 <!--theoretical framework: corporate/network strategies-->
-The involvement in long-distance exchange is often associated with an increased social inequality where ambitious individuals could gain power by building personal influence through the acquisition and distribution of foreign goods, leading to social change from corporate to network societies [@Blanton1996; @Feinman2000; @Klehm2017].
+Involvement in long-distance exchange is often associated with an increased social inequality where ambitious individuals could gain power by building personal influence through the acquisition and distribution of foreign goods, leading to social change from corporate to network societies [@Blanton1996; @Feinman2000; @Klehm2017].
 Corporate-network societies represent two distinct political-economic strategies for obtaining power, viewed as a horizontal axis crosscutting traditional vertical hierarchy [@Drennan2010; @Feinman2000].
@@
 We explore whether this change can be identified through the pattern of intercorrelations of foreign prestige goods among burials.
-We ask: (1) Do the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact?
+We ask: (1) Does the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact?
 and if so, (2) what are the major variables affecting or forming these two networks?
@@
 A total of 90 burials were unearthed from adjacent excavation squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)).
-The preservation of human remains at Kiwulan was generally poor, so we have incomplete data of age and sex (details in Supplementary Online Materials) and lack health status data.
+The preservation of human remains at Kiwulan was generally poor, so we have incomplete data for age and sex (details in Supplementary Online Materials) and lack health status data.
 Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses.
@@
 Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament; @LWandBM2020pot].
-We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size.
-For the two periods studied here, we evaluate differences in network size were using a vertex bootstrap method.
+We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size that corresponds to the population decline at Kiwulan in the Chinese period [@Chen2007].
+For the two periods studied here, we evaluated differences in network size using a vertex bootstrap method.
 We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology.
@@
 
-```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
+```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper.
+Seven burials with 60 and more (the maximum is 2936) glass beads are not shown in this plot."}
 knitr::include_graphics(here::here("analysis", "figures", "000-raincloud-beads.png"))
@@
              caption = "Network parameters in ERGMs used for model specifications with associated archaeological interpretation for burial relations.
-ERGM terminology is from the R package statnet (Handcock, 2008; Morris, 2008)")
+ERGM terminology is from the R package statnet (Handcock et al., 2008; Morris et al., 2008)")
 ```
@@
 <!-- add text about prior setting and anthropological assumptions of our prior values --> 
-Priors are a distinctive aspect of Bayesian methods that enable us combining our beliefs based on prior research with the newly observed data [@Gill2014].
+Priors are a distinctive aspect of Bayesian methods that enable us to combine our beliefs based on prior research with newly observed data [@Gill2014].
 The normal distribution N($\mu$, $\sigma^2$), where $\mu$ is the mean and $\sigma^2$ is the standard deviation, is typically used for priors when modelling network parameters because it is sensible and convenient, allowing adjustments for a wide range of prior information [@Caimo2017; @Normand1992].
@@
 
-To evaluate our anthropological model of increased social inequality, specifically a shift from a corporate to network society [@Drennan2010; @Feinman2000], we specified priors that are meaningful for social inequality, especially for the variables of transitivity and centralization.
+To evaluate our anthropological model of increased social inequality, specifically a shift from a corporate to network society [@Drennan2010; @Feinman2000], we specified priors that are meaningful for social inequality, especially for the variables of transitivity and centralization (see our supplementary information for more details on prior selection).
 For the network before European contact, we set the priors to model less social differentiation and emphasize the ritual element shared in corporate groups: higher transitivity (N(`r priors$pre_prior_mean[6]`, `r priors$pre_prior_sigma[6]`)), lower centralization (N(`r priors$pre_prior_mean[7]`, `r priors$pre_prior_sigma[7]`)), and a higher covariate effect of ritual activity (N(`r priors$pre_prior_mean[4]`, `r priors$pre_prior_sigma[4]`)).
@@
                values_to = "posterior") %>%
-  mutate(parameter= case_when(parameter == "edges" ~ "density",
-                              parameter == "nodematch.age" ~ "age-homophily",
-                              parameter == "nodematch.sex" ~ "sex-homophily",
-                              parameter == "nodematch.ritual" ~ "ritual-homophily",
-                              parameter == "nodematch.value" ~ "wealth-homophily",
-                              parameter == "gwesp" ~ "transitivity",
-                              parameter == "gwdeg" ~ "centralization",
-                              parameter == "dyadcov.distance" ~ "physical distance")) %>% 
+  mutate(parameter= case_when(parameter == "edges" ~ "Density",
+                              parameter == "nodematch.age" ~ "Age-homophily",
+                              parameter == "nodematch.sex" ~ "Sex-homophily",
+                              parameter == "nodematch.ritual" ~ "Ritual-homophily",
+                              parameter == "nodematch.value" ~ "Wealth-homophily",
+                              parameter == "gwesp" ~ "Transitivity",
+                              parameter == "gwdeg" ~ "Centralization",
+                              parameter == "dyadcov.distance" ~ "Physical distance")) %>% 
   mutate(parameter = factor(parameter,
-                            levels=c("density", "age-homophily", 
-                                     "sex-homophily", "ritual-homophily",
-                                     "wealth-homophily", "transitivity",	
-                                     "centralization", "physical distance"))) %>% 
+                            levels=c("Density", "Age-homophily", 
+                                     "Sex-homophily", "Ritual-homophily",
+                                     "Wealth-homophily", "Transitivity",	
+                                     "Centralization", "Physical distance"))) %>% 
   group_by(parameter, phase) %>% 
@@
 The pre-European group presents a long tailed distribution with larger values for the transitivity, but negative values for the centralization parameter compared to the post-European group.
-The edges and distance parameters for the pre-European group have larger values generally compared to pos-European one."}
+The edges and distance parameters for the pre-European group have larger values generally compared to the post-European one."}
 knitr::include_graphics(here::here("analysis", "figures", "006-posterior-distribution.png"))
@@
 We explored the sample size effect by removing nodes at certain percentages (5-40%) for both networks and comparing their confidence intervals.
-The results demonstrate a consistent difference between the two networks up to 35% node removals for the centralization variable as shown in Figure \@ref(fig:figure-bootstrap-CI).
+The results demonstrate a consistent difference between the two networks at up to 35% node removals for the centralization variable as shown in Figure \@ref(fig:figure-bootstrap-CI).
 This shows consistency with our earlier finding of negative centralization in the pre-European period and positive centralization in the post-European period, over several conditions of node removal.
@@
 Before European arrival, ritual behaviors could have been important factors that structured the interconnections of prestige goods in the burial contexts.
-This might imply that ritual was status indicator at the time before many foreign goods were introduced to Indigenous societies.
+This might imply that ritual was a status indicator at the time before many foreign goods were introduced to Indigenous societies.
 After the European presence was established, status indicators shifted from ritual to wealth differences, corresponding with the characteristics of corporate/network strategies.
@@
 <!-- trade dynamic during this period and future work for comparison -->
-We recognize that complex local and international trade dynamics between different Indigenous groups and other foreign presence, Han Chinese and Japanese, might have contributed to social changes at Kiwulan as well.
+We recognize that complex local and international trade dynamics between different Indigenous groups and other foreign groups, such as the Han Chinese and Japanese, might have contributed to social changes at Kiwulan as well.
 However, the current changes observed from burial networks and spatial distribution of trade items in the residential area at Kiwulan demonstrate consistent changes before and after the European arrival.
 The frequent trade activities stimulated by the presence of Europeans led to many trade items, including Anping jars, stonewares, Chinese porcelains, and beads found at archaeological sites with Dutch occupation at different parts of Taiwan, such as Fort Zeelandia.
-As a large village with continuous occupation since the 14th century in northeastern Taiwan, Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale.
+Although shortages mentioned in Spanish documents might indicate insufficient goods were brought into Taiwan [@Borao2009], we observed a large amount of trade goods at Kiwulan.
+Trade goods may not have been originally owned by Europeans if we consider the pre-existent regional trade with Chinese merchants.
+However, the large amount of foreign goods suggests frequent and active trade activities with local places involved in a global trade network.
+
+<!-- future work -->
+As a large village with continuous occupation since the 14th century in northeastern Taiwan,  the archaeological evidence at Kiwulan provides a special example to explore the reaction of Indigenous people living on the periphery to European colonial presence at a settlement scale that historical records rarely offer.
 Was its experience with the European presence unique or typical of the region?
 We need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui.
-Future work should focus on collecting data from Indigenous sites related to European contact for comparison with Kiwulan.
+The contemporary community of descendants of Kiwulan, the Kavalan people, shows features of a corporate society, such as a matrilineal system with age group administration and ritual practices [@Liu2021].
+This indicates that social change has been continuous in this area, from the more centralized network mode of social organization we observed in the burials after European arrival, to the current corporate model of social organization.
+Future work should focus on collecting data from Indigenous sites related to European contact and Chinese contact for comparison with Kiwulan.
 
@@
 We interpret the changes in formation mechanisms of networks, observed as a change from tie formation dominated by ritual to wealth, as an increase in wealth accumulation behaviors, stimulated by the European presence and its associated long-distance trade network.
-This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power (Feinman, 2000).
-Using burial data with historical documents, we can better understand indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
+This aligns with changes in horizontal hierarchy represented by corporate-network societies, that demonstrates different degrees of social inequality in relation to different strategies for achieving power [@Feinman2000].
+While ethnographic records may be ambiguous on the degree of social hierarchy at Kiwulan, the archaeological record allows us to see clear signals of changes in social organization.
+The archaeological record is valuable here because it gives us a better understanding of the indirect European impacts in understudied pericolonial contexts of East Asia [e.g.
 @Acabado2017; @Trabert2017].

### 2021-08-21 analysis/paper/supplementary-material.Rmd
edit the SI at the proof stage

@@
 ---
-title: "Supplementary Materials- A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period"
+title: "Supplementary Materials- A Bayesian networks approach to infer social changes from burials in northeastern Taiwan during the European colonization period"
 author:
@@
 Regarding burials, @Sosna2013 examined spatial patterns of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological similarity between burials.
-Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistically tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
+Recently, complex network modeling has been used to evaluate networks at both node and graph level through simulations of particular processes and statistical tests of the formation of network properties [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010].
 Such applications include simulations and testing of food exchange modes for Ancestral Pueblos on the aggregation of households in the American Southwest [@Crabtree2015], and exploring the diffusion of fired bricks across Hellenistic Europe by comparing similarity networks of sites with random networks [@Ostborn2015].
@@
 We conducted a sensitivity analysis to understand the effect of the prior specification on our two models.
-We set all the priors to be vague normal distribution (mean is 0 with standard deviation is 10) and a range of plausible priors (mean from 0 to 5 and standard deviation from 0 to 10).
+To explore the sensitivity, we ran the analysis with prior values to be vague normal distribution (mean is 0 with standard deviation is 10) and a range of plausible priors (mean ranging from 0 to 5 and standard deviation ranging from 0 to 10).
+We found very minor differences in the posterior distributions, but the directionality was consistent for all values.
+The results of these are not shown here.
+We choose the priors that have the best autocorrelation properties.
 The comparison demonstrates that our results are robust to different priors that preserve the differences between our two models, especially for density, transitivity, and centralization.
@@
 
-```{r figure-bgof-pre-E, fig.cap = "Goodness-of-fit diagnostics for the pre-European model and the post-European model (below).
+```{r figure-bgof-pre-E, fig.cap = "Goodness-of-fit diagnostics for the pre-European model.
 Boxplots represent distributions calculated on 100 network graphs simulated from the estimated posterior distribution.

### 2021-08-24 analysis/paper/paper.Rmd
knitting succeeded!

@@
 load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
-readRDS(here::here("analysis", "data", "derived_data", "pre_bergm.rds"))
-readRDS(here::here("analysis", "data", "derived_data", "post_bergm.rds"))
 ```

### 2021-08-24 analysis/code/006-bergm-distribution-stats.R

@@ -1,7 +1,7 @@
 # read saved models
 load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
-readRDS(here::here("analysis", "data", "derived_data", "pre_bergm.rds"))
-readRDS(here::here("analysis", "data", "derived_data", "post_bergm.rds"))
+
+library(cowplot)
 
 # get data from pre-European model
 pre_bergm_stats <- as.data.frame(pre_bergm$Theta)

### 2021-08-24 analysis/code/007-bgof-assessment-vis.R

@@ -5,6 +5,7 @@ library(Bergm)
 library(tidyverse)
 
 load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
+suppressWarnings(source(here::here("analysis", "code", "999-bgof-custom-function.R")))
 
 # Bayesian Model assessment
 #------------- Pre-E network

### 2021-08-24 analysis/paper/paper.Rmd
knit successfully after reordering the bgof2() and using lockfile pkgs by removing the updated ones

--- 2021-08-21 analysis/paper/paper.Rmd
@@
 Map data from naturalearthdata.com."}
-
 knitr::include_graphics(here::here("analysis", "figures", "000-KWL-map.png"))
@@
 # Methods {-}
-
-```{r custom-function, include=FALSE}
-suppressWarnings(source(here::here("analysis", "code", "999-bgof-custom-function.R")))
-```
 

### 2021-08-24 analysis/code/007-bgof-assessment-vis.R

@@ -1,9 +1,6 @@
 # read saved models
-library(igraph)
-library(Bergm)
-library(tidyverse)
-
 load(here::here("analysis", "data", "derived_data", "burial_bergm_model.RData"))
+suppressWarnings(source(here::here("analysis", "code", "999-bgof-custom-function.R")))
 
 # Bayesian Model assessment
 #------------- Pre-E network
@@ -18,16 +15,15 @@ bgof_pre <-
         n.deg     = 15,
         n.dist    = 15,
         n.esp     = 10)
-
 dev.off()
 
 # load library for measuring moments of distribution
 library(psych)
 
 # calculate moments for observed and simulated GOF distribution for pre-E model
-obs_degree_pre <- describe(bgof_pre$obs.degree * 0:(length(bgof_pre$obs.degree)-1))
-obs_dist_pre <- describe(bgof_pre$obs.dist * 1:length(bgof_pre$obs.dist))
-obs_esp_pre <- describe(bgof_pre$obs.esp * 0:(length(bgof_pre$obs.esp)-1))
+obs_degree_pre <- psych::describe(bgof_pre$obs.degree * 0:(length(bgof_pre$obs.degree)-1))
+obs_dist_pre <- psych::describe(bgof_pre$obs.dist * 1:length(bgof_pre$obs.dist))
+obs_esp_pre <- psych::describe(bgof_pre$obs.esp * 0:(length(bgof_pre$obs.esp)-1))
 
 sim_degree_pre <-
   psych::describe(bgof_pre$sim.degree * rep(0:(length(bgof_pre$obs.degree)-1),
@@ -36,12 +32,12 @@ sim_degree_pre <-
   dplyr::filter(skew != "NaN")
 
 sim_dist_pre <-
-  describe(bgof_pre$sim.dist * rep(1:length(bgof_pre$obs.dist),
+  psych::describe(bgof_pre$sim.dist * rep(1:length(bgof_pre$obs.dist),
                                    times = length(bgof_pre$sim.dist)
                                    /length(bgof_pre$obs.dist)))
 
 sim_esp_pre <-
-  describe(bgof_pre$sim.esp * rep(0:(length(bgof_pre$obs.esp)-1),
+  psych::describe(bgof_pre$sim.esp * rep(0:(length(bgof_pre$obs.esp)-1),
                                   times = length(bgof_pre$sim.esp)
                                   /length(bgof_pre$obs.esp)))
   #filter(skew != "NaN" & skew != 0.00)

### 2021-08-24 analysis/paper/paper.Rmd
fix a citation in the rmd and knit

@@
 We need more evidence from other sites in northeastern Taiwan to dive deeper into pericolonial effects, and compare evidence between sites on the periphery and colonial centers, such as Heping Dao and Tamsui.
-The contemporary community of descendants of Kiwulan, the Kavalan people, shows features of a corporate society, such as a matrilineal system with age group administration and ritual practices [@Liu2021].
+The contemporary community of descendants of Kiwulan, the Kavalan people, shows features of a corporate society, such as a matrilineal system with age group administration and ritual practices [@LiuPC2021].
 This indicates that social change has been continuous in this area, from the more centralized network mode of social organization we observed in the burials after European arrival, to the current corporate model of social organization.

### 2021-08-24 analysis/paper/supplementary-material.Rmd
note of code for fig 3 & 4 in the SI

@@
 ```{r pre-diag-plots, include= FALSE}
+# Figure 1 in the SI
 # get the first set of diagnostic plots
@@
 ```{r post-diag-plots, include= FALSE}
+# Figure 2 in the SI
 # get the first set of diagnostic plots
@@
 ```{r figure-pre-diag, fig.cap = "MCMC output of the pre-European model. In the order of edges, nodematch-age, nodematch-sex, nodematch-ritual, nodematch-value, gwesp, gwdeg, and dyadcov.distance."}
+# Figure 3 in the SI (code is in 007-bgof-assessment-vis.R)
 knitr::include_graphics(here::here("analysis", "figures", "008-pre-diag.png"))
@@
 ```{r figure-post-diag, fig.cap = "MCMC output of the post-European model. In the order of edges, nodematch-age, nodematch-sex, nodematch-ritual, nodematch-value, gwesp, gwdeg, and dyadcov.distance."}
+# Figure 4 in the SI (code is in 007-bgof-assessment-vis.R)
 knitr::include_graphics(here::here("analysis", "figures", "009-post-diag.png"))
